### 2016-01-28 paper/paper.tex
starting the paper

%Document Class
\documentclass{latex/emulateapj}
\usepackage{multirow,color,wrapfig,ulem}

%Packages and Commands
\include{latex/packages_commands}

\begin{document}

\title{Influence of galaxy rotation and outflows in the Lyman Alpha spectral line}
\shorttitle{Influence of galaxy rotation and outflows in the Lyman Alpha spectral line}

\shortauthors{Remolina-Gutierrez et al.}

\author{Maria Camila Remolina-Gutierrez, Jaime E. Forero-Romero} 
\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
\email{mc.remolina197@uniandes.edu.co}
\email{je.forero@uniandes.edu.co}

\keywords{Galaxies: high-redshift, Lyman Alpha Emission, Galaxy Rotation, Galaxy Outflows, Radiative Transfer}

\begin{abstract}
\noindent Young galaxies in the Universe have a strong \lya emission caused by the ionized Hydrogen atoms in their interstellar medium. When the spectrum of a galaxy has an intense peak around the \lya natural frequency ($2.46\times 10^{15}$ Hz) it is called a Lyman Alpha Emitter (LAE). Typical LAEs are very distant ($z \gtrsim 2$). This makes that all the data astronomers can obtain from them is their spectra, and from there all the physical information of the galaxy must be derived. Trying to solve this task requires the creation of a simplified and solid model. In this monograph I propose to consider LAEs as a spherical distribution of Hydrogen atoms that undergoes a solid body rotation and a radial expansion. I simulate the effect of rotational velocity, outflow velocity and optical depth of the LAE in the outgoing spectra. The main conclusion is that this new model reproduces LAEs observed features in a clear way and with consistent physical parameters. However, proper observational fits are left for future work. This monograph accomplishes the objective of extracting as much information as possible from a LAE's \lya line.
\end{abstract}

\section{Introduction}
\label{sec:intro}

Galaxies are key to understand our Universe. So far, human kind has been able to study in detail the closest ones. Distant galaxies are still challenging to detect. They contain however information of millions of stars and gas that could unravel mysteries like for example, how Milky Way type of galaxies formed. It is an open challenge to obtain as much information as possible from these young galaxies. One way to accomplish this is to look at their spectra. Astronomers noted that a that relevant fraction of these galaxies emitted a really strong line at a particular wavelength and named them Lyman Alpha Emitters (LAEs). The purpose of this monograph is to model and simulate LAEs in order to advance our knowledge about this galaxy population. \\

\section{The physics of Lyman Alpha Emitters}
\label{sec1:laes}
In the early Universe the most abundant elements were Hydrogen(H) and Helium(He), causing young galaxies to be rich in H atoms. Also, due to stellar activity in a galaxy, the surrounding gas can be ionized. These two facts leave the H atoms prone to excite and emit radiation at definite frequencies. The most basic of these frequencies corresponds to the \lya line. \cite{PartridgePeebles}  \\

When a Hydrogen atom is excited, the only electron it has moves to a higher energy level. However after some time, it is no longer able to maintain itself in this state. For this, the electron goes back to the ground level. Finally, as the energy obtained from this de-excitement has to be conserved, it is emitted as a photon with a wavelength of 1215.67 $\AA$ if levels are $2\rightarrow1$. This is the Lyman Alpha ( \lya) transition, discovered by Theodore Lyman in 1906. \cite{LymanBio} \\

\lya radiation is emitted from the H atoms in a galaxy. Due to the amount of atoms inside it, the whole body could become a strong \lya radiator. Astronomers can take spectra of a galaxy to detect emission at that wavelength. If the galaxy shows a strong \lya line, it classifies as a Lyman Alpha Emitter (LAE). (\cite{DjorgovskiThompson}, \cite{Rhoads00}, \cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08}, \cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12}, \cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013}, \cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014}, \cite{Fumagalli2015}) \\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/radiative_transfer}
	\end{center}
	\caption{\textbf{Radiative Transfer Sketch:} A \lya photon being absorbed and re-emitted by Hydrogen atoms. 
		\label{fig:radiative_transfer}}
\end{figure}

When a \lya photon is emitted inside the galaxy, it travels through its interstellar medium (ISM). During the photon's path it can be absorbed and re-emitted by an H atom. The new frequency of the photon is different that the initial one, in an observer frame of reference due to the atom's velocity. In LAEs, the state of the ISM gas, before and after a photon's re-emission is pretty much the same. This allows a random walk approximation and consider photon-atom encounters as scatterings. A photon's scattering can happen several times, as seen in Fig. \ref{fig:radiative_transfer}. This stops only until the photon is able to escape the galaxy. \\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.3\textwidth]{./figures/static}
	\end{center}
	\caption{\textbf{Figure A3 of Forero-Romero et al. paper:} CLARA's view on the escape fraction of Lyman-$\alpha$ photons in high redshift galaxies \cite{CLARA}. Reproduced with the permission of the author.
		\label{fig:static}}
\end{figure}

In a static galaxy, this random walk process produces a spectrum with two equal and symmetric peaks around the natural \lya wavelength. This can be seen in Figure A3 of Forero-Romero et al. paper \cite{CLARA} (Fig. \ref{fig:static}). If now the gas has a bulk velocity, the shape of the \lya profile changes. I explore these effects in this work by proposing a new model for a LAE. It consists of a spherical distribution of H atoms undergoing a solid body rotation and radial expansion(outflows). This new model can help observers to determine physical characteristics of a LAE by just looking at its \lya line. \\

The main goal is to measure the effect of the model's physical parameters on the outgoing \lya line. In this case, due to the resonant\footnote{Resonant is a common term used in radiative transfer. It means that the photons that create the line are absorbed and re-emitted several times before escaping the cloud of gas.} nature of the \lya line, analytical solutions can not be derived. It becomes necessary to run simulations to explore and test the model. \\

In this monograph I use a radiative transfer code that exists in the literature called CLARA (Code for Lyman Alpha Radiation Analysis). CLARA was written by Forero-Romero et al. \cite{CLARA}. It can simulate the \lya line of a spherical rotating LAE depending on its mass and velocity. In this work I modify CLARA to include outflows and explore the resulting consequences on the \lya line. \\

\section{A new LAE model}

\label{sec:newmodel}
Galaxies are ellipsoidal. However, for simplicity I model them as a sphere. This shape facilitates the interpretation of the results in our numerical experiments. Furthermore, this approximation is commonly used in the literature, as it explains a wide variety of observational features (\cite{Ahn03}, \cite{Verhamme06}, \cite{Dijkstra06}). \\

There are 3 parameters in my model that define a LAE: the rotational velocity (\vrot), the outflow velocity (\vout) and the optical depth (\tauh). \vout is due to material ejected from the galaxy, by supernovas(\cite{Verhamme06}, \cite{Orsi12}, \cite{Hashimoto2015}, \cite{Gronke2015}). \tauh roughly corresponds to number of H atoms found by a \lya photon if one traces a line from the center of the galaxy to its edge, and it resembles the mass of the LAE.\\

In this model, the LAE has a bulk velocity which is the superposition of the rotation and the outflow, as shown in Fig. \ref{fig:model}. The velocity components are written in Eqs. (\ref{eq:vx}) (\ref{eq:vy}) (\ref{eq:vz}). In these equations, $R$ is the radius of the sphere; $x$, $y$ and $z$ are the coordinates in a cartesian frame; and the $\mp$ signs in $v_x$ and $v_y$ indicate the direction of rotation, respectively. This rotation is a solid body rotation and its direction goes according to the right hand rule applied to the $\hat{k}$ unit vector. The outflow velocity is dependent on the position relative to the center of the galaxy, being it zero at the center and maximum at the edge of the sphere.\\

\begin{equation}
	v_{x}=\frac{x}{R}v_{\rm out}-\frac{y}{R}v_{\rm rot} 
	\label{eq:vx}
\end{equation}

\begin{equation}
	v_{y}=\frac{y}{R}v_{\rm out}+\frac{x}{R}v_{\rm rot} 
	\label{eq:vy}
\end{equation}

\begin{equation}
	v_{z}=\frac{z}{R}v_{\rm out}
	\label{eq:vz}
\end{equation}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.3\textwidth]{./figures/model}
	\end{center}
	\caption{\textbf{Model:} Spherical LAE with tangential and radial velocity due to rotation and outflows.
		\label{fig:model}}
\end{figure}

These 3 parameters, \vrot, \vout and \tauh, leave the idea of a model of LAE, that although simple, considers the main galaxy's dynamics. All of them have been previously proposed and used by different authors, but never combined together (\cite{Adams72}, \cite{Harrington73}, \cite{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12}, \cite{Martin2015}, \cite{Garavito14}, \cite{Neufeld91}, \cite{Laursen09}, \cite{Barnes11}, \cite{Verhamme12}, \cite{Yajima12}).\\

\subsection{A \lya Photon's Path in CLARA}
In several problems in science it is useful to define dimensionless variables. This new formulation has many advantages. The numerical equations do not depend on the unit choice. This tends to make the results' analysis straight forward. In the \lya radiative transfer problem, a commonly used dimensionless variable $x$ is used to describe a photon's frequency and is defined as:\\

\begin{equation}
	\label{eq:x_freq}
	x \equiv \frac{\nu -\nu_{\rm \alpha}}{\Delta\nu_{\rm D}},
\end{equation}

where $\nu$ is the photon's frequency and $\nu_{\rm \alpha} = 2.46\times 10^{15}$ Hz is the \lya natural frequency. The denominator $\Delta\nu_{\rm D}$ is defined in Eq. \ref{eq:delta_nuD}.

\begin{equation}
	\label{eq:delta_nuD}
	\Delta\nu_{\rm D} \equiv \nu_{\rm \alpha}\sqrt{\frac{2kT}{m_pc^2}} \equiv \nu_{\rm \alpha} \frac{v_{\rm th}}{c},
\end{equation}

where $\Delta\nu_{\rm D}$ is known as the Doppler broadening of the \lya line. It depends on the neutral gas temperature $T$ or equivalently the thermal velocity $v_{\rm th}$ of the atoms. In the model the temperature is kept constant at $T=10^4$K and the thermal velocity is $v_{\rm th}=12.8$\kms. \\

If $x < 0$, the final frequency $\nu < \nu_{\rm \alpha}$. This translates to the final wavelength being larger than the \lya natural one. The phenomenon is called a redshift in frequency. If, on the contrary, $x > 0$, the photon suffers a blueshift in frequency. \\

However, in the plots shown in this monograph I use instead velocity, $V$, units. This eases the comparison against observational data. The velocity units are defined by

\begin{equation}
	\label{eq:V}
	V = xv_{\rm th} = \frac{\nu -\nu_{\rm \alpha}}{\nu_{\rm \alpha}}c.
\end{equation}

The units of $V$ are usually \kms. In this case, the photon is redshifted in frequency when the velocity is greater than 0, and blueshifted in the opposite case. \\

The initial emission of photons is taken at the center of the sphere for practicality due to the fact that both, center and off-center emissions, give analogous results. From here, $100000$ photons are emitted with the natural \lya wavelength. Then, they start to behave as described in section \ref{sec1:laes} (See Fig. \ref{fig:radiative_transfer}). When each photon is re-emitted, its new wavelength depends on the H atom's velocity (both thermal and bulk) and direction (both initial and final). However the photon's new direction of propagation is random in the rest frame of the atom. \\

The individual scattering of all the photons is tracked through the complete 3D Hydrogen distribution. Once each photon escapes the galaxy, its final values are stored: position $\vec{r}$, direction of propagation $\hat{k}$, dimensionless frequency $x$, and number of scatterings $N$. To build the observed spectrum I make a histogram of the escape frequencies $x$. The number of scatterings $N$ tells how many steps the random walk requires to reach a distance to the center that is $\geq R$. In order to avoid situations in which the photon has not escaped after a long computational time, CLARA defines a number $N_{max}$ so the code stops. However, according to statistics, this last situation has low probabilities, so the photon is always most likely to exit the sphere. \\

The process that takes place from the moment the first photon is emitted to the moment the last photon escapes the sphere, is called a simulation. Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires. Depending on \tauh, I need a different number of processors in order to minimize the time demands. For \tauh $=10^5, 10^6, 10^7$ I use 6, 12, 24 processors, respectively. The running times go from 7 hours to 1 week approximately, depending on the case. \\

\subsection{Galaxy's Viewing Angle}
An observer located far away, only receives photons emitted along its line of sight. That means, only photons escaping in the direction of the observer must be counted in the spectrum. In the simulation I approximate this by taking into account only the photons with escaping direction angle $\theta$ respect to the rotation axis within the range $[\theta_{min}-\theta_{max}]$. I illustrate this in Fig. \ref{fig:viewing_angle_sketch}. \\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/viewing_angle_sketch}
	\end{center}
	\caption{\textbf{Viewing Angle Sketch:} The galaxy cut is in the $y-z$ plane perspective and the observer is located at an specific viewing angle of the sphere. Only photons with a direction that enters in this range of vision are taken into account to build the observed spectrum.
		\label{fig:viewing_angle_sketch}}
\end{figure}

Therefore, in principle the spectra depend on two new parameters, the azimuthal and the polar angle. However, the galaxy's motion is symmetrical respect to its rotation axis. This implies that the resulting spectrum is independent from the azimuthal angle. Taking into account this symmetry, I only select photons on their polar angle, regardless of their azimuthal angle. Regarding the polar angle, I build the spectra for observers located on 3 different positions, with $\theta$ intervals uniformly distributed in $\cos(\theta)$. \\

To summarize, the parameters influencing the spectra are \vrot, \vout, \tauh and $\theta$. In the next chapter I evaluate the impact each of these has on the resulting profile. \\

\section{Results}
\label{sec:results}

\subsection{Parameters' Initial Values}

In order to define the ranges of \tauh, \vrot and \vout, it is necessary to refer to observational constraints. The common values for typical LAEs are in Tab. \ref{tab:values}. I run CLARA for all the permutations of these 3 parameters. \\

\begin{table}[htbp]
	\centering
	\begin{tabular}{|c|c|c|}
		\hline
		$\tau_{\mathrm{H}}$ & $v_{rot}$ (\kms) & $v_{out}$ (\kms) \\
		\hline
		$10^5$, $10^6$, $10^7$ & 50, 100 & 5, 10, 15, 20, 25, 50, 75 \\
		\hline
	\end{tabular}
	\caption{\textbf{Parameters' Values:} All consistent with a LAE's typical properties}
	\label{tab:values}
\end{table}

The resulting sets of spectra are in Figs. \ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$. I do not include the spectra for \vout = 50, 75 \kms because they are similar to \vout = 25 \kms. For this reason, I limit in all the plots \vout to a maximum value of 25 \kms. Nevertheless, the figures for \vout = 50, 75 \kms are in the Appendix \ref{first_plots} in Figs. \ref{fig:2_tau10E5_phi83-90}, \ref{fig:2_tau10E6_phi83-90} and \ref{fig:2_tau10E7_phi83-90}.\\

\newpage

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/3_tau10E5_phi83-90}
	\end{center}
	\caption{\textbf{\lya profile for \tauh$=10^5$:} With \vrot ranging $50,100$ \kms and \vout ranging $5,10,15,20,25$ \kms. The intensity is in arbitrary units. The viewing angle $\theta \simeq 90^\circ$. The dashed vertical line indicates the \lya line's natural frequency. 
		\label{fig:3_tau10E5_phi83-90}}
\end{figure}

\newpage

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/3_tau10E6_phi83-90}
	\end{center}
	\caption{\textbf{\lya profile for \tauh$=10^6$:}  Same as Fig. \ref{fig:3_tau10E5_phi83-90} but for \tauh$=10^6$.
		\label{fig:3_tau10E6_phi83-90}}
\end{figure}

\newpage

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/3_tau10E7_phi83-90}
	\end{center}
	\caption{\textbf{\lya profile for \tauh$=10^7$:} Same as Fig. \ref{fig:3_tau10E5_phi83-90} but for \tauh$=10^7$.
		\label{fig:3_tau10E7_phi83-90}}
\end{figure}

\newpage

These last 3 plots show a clear creation of two asymmetric peaks around $V=0$ \kms with the tallest peak is always redshifted, with a strong dependence on the outflow velocity.\\

\subsection{Influence of the viewing angle $\theta$}
I take now into account the viewing angle of the galaxy to build the observed spectra. For all of the physical parameters' combinations, the effect of $\theta$ in the \lya line is always the same.\\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/influence_viewing_angle_5}
	\end{center}
	\caption{\textbf{\lya profile for different $\theta$:} With \tauh$=10^5$, \vrot$=50$ \kms and \vout$=20$ \kms.The intensity is in arbitrary units.
		\label{fig:influence_viewing_angle_5}}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/influence_viewing_angle_6}
	\end{center}
	\caption{\textbf{\lya profile for different $\theta$:} With \tauh$=10^6$, \vrot$=100$ \kms and \vout$=5$ \kms.The intensity is in arbitrary units.
		\label{fig:influence_viewing_angle_6}}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/influence_viewing_angle_7}
	\end{center}
	\caption{\textbf{\lya profile for different $\theta$:} With \tauh$=10^7$, \vrot$=100$ \kms and \vout$=15$ \kms.The intensity is in arbitrary units.
		\label{fig:influence_viewing_angle_7}}
\end{figure}

From Figs. \ref{fig:influence_viewing_angle_5}, \ref{fig:influence_viewing_angle_6} and \ref{fig:influence_viewing_angle_7} is clear that the intensity of the valley between the two peaks increases along with $\theta$. This causes an intensity decrease in the rest of the frequencies, thus a broadening of the line. The asymmetry also changes with the viewing angle.\\

In the next subsection, I show only the results of the angle at which an observer sees the galaxy's angular momentum vector perpendicular to the line of sight. The purpose of this is only to decrease the number of plots in the document, as there is an analogous behavior for all viewing angles. \\

\subsection{Morphology of \lya line}
To quantify the morphology of the \lya profile, I use three statistics: the standard deviation, the skewness and a factor sigma of asymmetry ($\sigma_A$). \\

\subsubsection{Standard Deviation}
The standard deviation measures the dispersion of frequencies in the \lya line. If it increases, the frequencies spread to wider ranges. \\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/std}
	\end{center}
	\caption{\textbf{Standard deviation plot for each \tauh.} The viewing angle $\theta \simeq 90^\circ$.
		\label{fig:std}}
\end{figure}

Fig. \ref{fig:std} shows that the standard deviation is inversely proportional to the outflow velocity. This implies that the greater \vout is, the less disperse is the \lya frequency distribution. This signals that the peaks start merging to one another. Also, the higher the \tauh, the more inclined the curves. The horizontal dotted line in the plot is the value of standard deviation for an observational LAE, in particular Q2206BX151 (at redshift $z\approx2.1974$). It is shown in Fig. \ref{fig:kulas} in the third column with second row. \\

\subsubsection{Skewness}
The skewness measures the asymmetry of the \lya line. If the value is greater than zero, there is more weight in the left tail of the line, if it is less than zero, there is more weight in the right one.\\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/skw}
	\end{center}
	\caption{\textbf{Skewness plot for each \tauh.} The viewing angle $\theta \simeq 90^\circ$. 
		\label{fig:skw}}
\end{figure}

Fig. \ref{fig:skw} show that the skewness is proportional to the outflow velocity, at a fixed rotational velocity. The greater \vout is, the more asymmetric is the \lya frequency distribution. The horizontal dotted line is the value of skewness for Q2206BX151. As seen, the line intersects with the skewness curves. This would imply that Q2206BX151, according to the model, has its parameters narrowed to: \vrot $\in [50,100]$ \kms, \vout $\in [15,20]$ \kms and \tauh $\in [10^6, 10^7]$. However, these values are not completely consistent with the standard deviation.\\

\subsubsection{Sigma of Asymmetry}
Sigma of asymmetry $\sigma_A$ is another measure for the asymmetry of the \lya line. It was defined by McLinden et al. \cite{McLinden2011} as follows. Each of the two peaks, the red one and the blue one, are fitted with a Gaussian curve. Their standard deviations $\sigma_{red}$ and $\sigma_{blue}$ are obtained. Then the factor $\sigma_A$ is:

\begin{equation}
\sigma_A = \frac{\sigma_{red}}{\sigma_{blue}}
\end{equation}

If the value is less than one, the left peak is thiner than the right peak. If it is greater than one, there left peak is wider than the right peak.\\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/sigma}
	\end{center}
	\caption{\textbf{Sigma of asymmetry $\sigma_A$.} 
		\label{fig:sigma}}
\end{figure}

Fig. \ref{fig:sigma} shows that this parameters has high fluctuations. I can not see a clear behavior or tendency that applies to all the curves. The only thing is that the values of $\sigma_A$ start close together and \vout makes them separate. Also, $\sigma_A$ is always greater than $1$, which means that there is always an asymmetry present in these distributions.\\

\subsection{Influences of the free parameters}

To summarize the influence of the 3 parameters on the \lya morphology is the following:

\begin{itemize}
	\item \tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
	\item \vout decreases the right peak's intensity. Higher \vout make the left peak smaller until it merges with the right one. \\
	\item \vrot broadens the line and decreases the maximum intensity. Higher \vrot implies a flatter spectrum. This effect has not been deeply studied in literature. Only Garavito et al. \cite{Garavito14} has simulated its effect. Our results are consistent with their conclusions. 
\end{itemize}

\section{Comparison against observations}
\label{sec:comparisonobservations}

There are several observations of LAEs in the literature. Kulas et al. \cite{Kulas12} observed distant LAEs ($z\sim2-3$) with high resolution. In the Figure 3 of their paper (Fig. \ref{fig:kulas}) they show the 18 \lya profiles. \\

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{./figures/figure3_kulas}
	\end{center}
	\caption{\textbf{Figure 3 of Kulas et al. paper:} The Kinematics of Multiple-peaked \lya Emission in Star-forming Galaxies at $z\sim2-3$ \cite{Kulas12}. This plot is reproduced with the authors permission.
		\label{fig:kulas}}
\end{figure}
%http://arohatgi.info/WebPlotDigitizer/app/

The observed spectra in Fig. \ref{fig:kulas} are modified by the redshift $z$ of the galaxies. This is, how much distance light had to travel to reach their telescopes. That causes a broadening of the line. \\

Taking into account this effect, I notice similitudes between them and the simulated spectra. The \lya profiles are mostly 2 peaks with an asymmetry between them, just as in the simulation. Also, the small peak (if existing) is always smaller than the tall one.\\

Many authors that only include the outflows effect are able to fit observational spectra but with a very large magnitude of \vout, on the order of $50-100$ \kms. However, \vout is caused mainly by ejection of material out the galaxy. It is then common that a galaxy rotates faster than it expands, not the opposite. Other authors that include the rotational effect are able to obtain the two peaks but not their asymmetry. Simulating LAEs without outflows does not replicate the majority of the cases. \\

In my model, the combination of rotation and outflows velocities not only makes a lot of sense but seems to be able to fit observations. This new model proposed could replicate \lya profiles with typical LAE's values. The main result of this monograph is that I presented a new LAE model roughly consistent with observations.\\

\section{Conclusions}
\label{sec:conclusions}

In this monograph, the objective was to analyze and measure the influence of galaxy rotation and outflows on the \lya line. The motivation for this is to be able to obtain physical information of a LAE by just looking at its \lya profile. In order to accomplish this objective, I propose a new model of a LAE consisting of a sphere of Hydrogen atoms that expands radially and rotates as a solid body. The program CLARA \cite{CLARA} is used to set the conditions and emulate the transfer code. \\

The conclusions obtained from this work are: \\

\begin{itemize}
	\item The outgoing spectra depend on the angle an external observer is viewing the galaxy from. The closer it is to the equator of the galaxy, the higher the central valley of the frequency distribution. \\

	\item The effects of \vrot, \vout and \tauh are consistent with the different authors that have used them. \vrot broadens the \lya line. \vout increases the peaks asymmetry and \tauh induces a redshift around the zero velocity.\\

	\item The \vrot effect is much larger than the \vout one. Only a little fraction of \vrot, in the radial direction, is necessary to obtain 2 asymmetric peaks. If it \vout approaches the value of \vrot, both peaks start merging until only the outflows effect is visible. \\

	\item The final spectra obtained are roughly consistent with LAEs observations.  \\

\end{itemize}

\subsection{Future work}

Due to the long time CLARA takes to run, it was not possible to fit an observational LAE and predict its parameters. However, the next step is to use tools as MCMC (Monte Carlo Markov Chain) to obtain the galaxy's \tauh, \vrot and \vout that would agree with my model. \\

All of this work is available online and free to use to anyone. The data, source code and instructions to replicate my monograph's results are in the GitHub repository:  \texttt{https://github.com/astroandes/CLARA\_RotationOutflows}. \\

\section*{Acknowledgments}

...\\

%------------------------REFERENCES----------------------------

\bibliographystyle{latex/apj}
\bibliography{references}

\newpage

%------------------------APPENDIX----------------------------

\appendix
\section{Additional Figures}

Additional figures of spectra with \vrot $= 50,100$ \kms and \vout $= 50,75$ \kms, for all the different \tauh.

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\textwidth]{./figures/appendix/2_tau10E5_phi83-90}
	\end{center}
	\caption{\textbf{\lya profile for \tauh$=10^5$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
		\label{fig:2_tau10E5_phi83-90}}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\textwidth]{./figures/appendix/2_tau10E6_phi83-90}
	\end{center}
	\caption{\textbf{\lya profile for \tauh$=10^6$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
		\label{fig:2_tau10E6_phi83-90}}
\end{figure}

\begin{figure}[h!]
	\begin{center}
		\includegraphics[width=1\textwidth]{./figures/appendix/2_tau10E7_phi83-90}
	\end{center}
	\caption{\textbf{\lya profile for \tauh$=10^7$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
		\label{fig:2_tau10E7_phi83-90}}
\end{figure}

\end{document}

### 2016-02-04 abstract/abstract.tex
new format

%\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
%\email{mc.remolina197@uniandes.edu.co}
%\email{je.forero@uniandes.edu.co}

\documentclass[11pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[latin1]{inputenc}    %% european characters can be used (Windows, old Linux)
\usepackage[T1]{fontenc}   %% get hyphenation and accented letters right

% do not change these lines:
\pagestyle{empty}                %% no page numbers!
\usepackage[left=35mm, right=35mm, top=15mm, bottom=20mm, noheadfoot]{geometry}

\begin{document}
\thispagestyle{empty}

\title{\textbf{Influence of galaxy rotation and outflows \\
			   in the Lyman Alpha spectral line}}

\author{Maria Camila Remolina-Gutierrez$^1$, Jaime E. Forero-Romero$^1$\\
	    mc.remolina197@uniandes.edu.co, \hspace{0.8mm} je.forero@uniandes.edu.co\\ 
		$^1$ Departamento de F\'{i}sica, Universidad de los Andes \\
		Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
\date{} % <--- leave date empty
\maketitle\thispagestyle{empty} %% <-- you need this for the first page
\textit{\textbf{Keywords - Galaxies: high-redshift, Lyman Alpha Emission, Galaxy Rotation, Galaxy Outflows, Radiative Transfer.}}\\

Young galaxies in the Universe have a strong Ly-$\alpha$ emission caused by the ionized Hydrogen atoms in their interstellar medium. When the spectrum of a galaxy has an intense peak around the Ly-$\alpha$ natural frequency ($2.46\times 10^{15}$ Hz) it is called a Lyman Alpha Emitter (LAE). Typical LAEs are very distant ($z \gtrsim 2$). This makes that all the data astronomers can obtain from them is their spectra, and from there all the physical information of the galaxy must be derived. Trying to solve this task requires the creation of a simplified and solid model. In this monograph I propose to consider LAEs as a spherical distribution of Hydrogen atoms that undergoes a solid body rotation and a radial expansion, as seen in Fig. \ref{model}.

\begin{figure}[h]
	%uncomment next line to include a graphic file
	\centerline{\includegraphics[width=4cm]{model.png}}
	\caption{Model of a Expanding and Rotating LAE.}
	\label{model}
\end{figure}

I simulate the effect of rotational velocity, outflow velocity and optical depth of the LAE in the outgoing spectra. The main conclusion is that this new model reproduces LAEs observed features in a clear way and with consistent physical parameters. However, proper observational fits are left for future work. This monograph accomplishes the objective of extracting as much information as possible from a LAE's Ly-$\alpha$ line.

%... a few optional references, such as \cite{Verhamme06}\\

%------------------------REFERENCES----------------------------

\bibliographystyle{latex/apj}
\bibliography{references}

\end{document}

### 2016-02-04 abstract/abstract.tex
spanish abstract updated

@@
 
-\author{Maria Camila Remolina-Gutierrez$^1$, Jaime E. Forero-Romero$^1$\\
+\author{Maria Camila Remolina-Gutierrez$^1$, Jaime E. Forero-Romero$^1$\\ \vspace{3mm}
 	    mc.remolina197@uniandes.edu.co, \hspace{0.8mm} je.forero@uniandes.edu.co\\ 
@@
 \maketitle\thispagestyle{empty} %% <-- you need this for the first page
-\textit{\textbf{Keywords - Galaxies: high-redshift, Lyman Alpha Emission, Galaxy Rotation, Galaxy Outflows, Radiative Transfer.}}\\
+\textit{Keywords - Galaxies: high-redshift, Lyman Alpha Emission, Galaxy Rotation, Galaxy Outflows, Radiative Transfer.}\\
 
@@
 
-\bibliographystyle{latex/apj}
+\bibliographystyle{apj}
 \bibliography{references}

### 2016-02-04 abstract/abstract_spanish.tex
spanish abstract updated

\documentclass[11pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage[latin1]{inputenc}    %% european characters can be used (Windows, old Linux)
\usepackage[T1]{fontenc}   %% get hyphenation and accented letters right

\pagestyle{empty} % no page numbers
\usepackage[left=35mm, right=35mm, top=15mm, bottom=20mm, noheadfoot]{geometry}

\begin{document}
\thispagestyle{empty}

\title{\textbf{Influencia de rotaci\'on y outflows\\
			    en la l\'inea espectral Lyman Alpha}}

\author{Maria Camila Remolina-Gutierrez$^1$, Jaime E. Forero-Romero$^1$\\ \vspace{3mm}
	    mc.remolina197@uniandes.edu.co, \hspace{0.8mm} je.forero@uniandes.edu.co\\ 
		$^1$ Departamento de F\'{i}sica, Universidad de los Andes, Bogot\'a, Colombia}
\date{} % <--- leave date empty
\maketitle\thispagestyle{empty} %% <-- you need this for the first page
\hyphenation{emi-si\'on ellas}

%Introducción y motivación
Las galaxias son un una estructura clave para entender el Universo. Hasta ahora, la humanidad ha podido estudiar en detalle las galaxias cercanas. Sin embargo, las distantes ($z \gtrsim 2$) presentan un reto observacional, tanto para ser detectadas, como para extraer información de ellas. \'Estas galaxias lejanas contienen características físicas de los principios del Universo que podrían responder misterios como por ejemplo, ¿cómo galaxias de tipo Vía Láctea fueron formadas? Es entonces un desafío abierto en la ciencia el obtener la mayor cantidad de información posible sobre ellas. Una manera de lograrlo es descomponer la luz que llega a los telescopios en diferentes longitudes de onda, obteniendo así su espectro electromagnético. Al hacer esto los astrónomos notaron que una fracción relevante de estas galaxias emit\'ia una fuerte l\'inea en la longitud de onda $\lambda = 1215.67$ \AA. \'Esta es debida a una presencia muy alta de \'atomos de Hidr\'ogeno que emiten la l\'inea espectral Lyman Alpha (Ly-$\alpha$). Por esta razón, estas galaxias se denominan Lyman Alpha Emitters (LAEs).\\

%Breve estado del arte
Al necesitar entonces que toda la información sea derivada del espectro de estas galaxias, es necesario crear un modelo s\'olido y simplificado que pueda describirlas. Varios autores las han modelado como cuerpos que se expanden debido a material proyectado hacía afuera de la galaxia (denominados outflows), causados principalmente por supernovas. Otros las han modelado como cuerpos que rotan. Lastimosamente, \'estas aproximaciones no permiten encontrar todas las posibles morfologías de la línea. Por esta razón, para mi trabajo de grado, propongo considerar las LAEs como una distribuci\'on esf\'erica de \'atomos de Hidr\'ogeno, que rota como un cuerpo r\'igido y se expande radialmente, tal como se ve en la Fig. \ref{model}:

\begin{figure}[h]
	%uncomment next line to include a graphic file
	\centerline{\includegraphics[width=4cm]{model.png}}
	\caption{Modelo de una LAE que se expande y rota.}
	\label{model}
\end{figure}

%Resultados principales y conclusión corta
Usando herramientas computacionales, simulo los efectos de velocidad rotacional, velocidad de expansión y masa de la LAE en el espectro de salida para diferentes combinaciones de \'estos tres parametros. La conclusi\'on principal es que este nuevo modelo reproduce las caracter\'isticas principales de LAEs observadas. Sin embargo, los ajustes observacionales se dejan para trabajo futuro. Este trabajo de grado, logra el objetivo de extraer la mayor informaci\'on posible de la l\'inea Ly-$\alpha$ de una LAE, con el fin de avanzar nuestro conocimiento sobre esta importante población de galaxias.

%------------------------REFERENCES----------------------------

\bibliographystyle{apj}
\bibliography{references}

\end{document}
%%https://www.modelica.org/events/modelica2014/authors-guide/example-abstract.tex/view

### 2016-02-04 abstract/abstract_spanish.tex
abstract espanol revisado

@@
 %Introducción y motivación
-Las galaxias son un una estructura clave para entender el Universo.
-Hasta ahora, la humanidad ha podido estudiar en detalle las galaxias cercanas.
-Sin embargo, las distantes ($z \gtrsim 2$) presentan un reto observacional, tanto para ser detectadas, como para extraer información de ellas.
-\'Estas galaxias lejanas contienen características físicas de los principios del Universo que podrían responder misterios como por ejemplo, ¿cómo galaxias de tipo Vía Láctea fueron formadas?
-Es entonces un desafío abierto en la ciencia el obtener la mayor cantidad de información posible sobre ellas.
-Una manera de lograrlo es descomponer la luz que llega a los telescopios en diferentes longitudes de onda, obteniendo así su espectro electromagnético.
-Al hacer esto los astrónomos notaron que una fracción relevante de estas galaxias emit\'ia una fuerte l\'inea en la longitud de onda $\lambda = 1215.67$ \AA.
-\'Esta es debida a una presencia muy alta de \'atomos de Hidr\'ogeno que emiten la l\'inea espectral Lyman Alpha (Ly-$\alpha$).
-Por esta razón, estas galaxias se denominan Lyman Alpha Emitters (LAEs).\\
+Las galaxias son la clave fundamental para entender el Universo.  Cuando observamos una galaxia distante estamos viendo el pasado. A partir de este esfuerzo de
+arqueolog\'ia astron\'omica podemos reconstruir la historia del Universo.
+
+Este proceso de reconstrucci\'on empez\'o en el Siglo XX. Los
+astr\'onomos estudiaron en detalle las galaxias m\'as cercanas,
+haciendo observaciones de un Universo ya maduro. 
+A finales del Siglo XX y comienzos del XXI se logran obtener se\~nales de
+galaxias mucho m\'as distantes; conjuntos de gas y estrellas que
+emitieron su luz cuando el Universo era j\'oven y ten\'ia menos
+de un quinto de su edad actual.  
+Estas galaxias son tan lejanas que que
+las im\'agenes aportan poco al momento de deducir las propiedades de
+una galaxia. 
+En estos casos mayor parte de la informaci\'on viene en los
+espectros, es decir, en la descomposici\'on de la luz en radiaci\'on de
+diferentes longitudes de onda. 
+De esta manera, el reto astrof\'isico se convierte en poder darle a estos
+espectros de galaxias distantes una  interpretaci\'on.
+
+En los primeros esfuerzos conjuntos de interpretaci\'on y
+observaci\'on, se hizo claro que varias galaxias
+j\'ovenes deben tener una alta cantidad de hidr\'ogeno y estrellas en
+proceso de formaci\'on. Bajo estas condiciones gran parte de su
+luz tiene la longitud de onda de la l\'inea de emisi\'on Lyman-$\alpha$,
+correspondiente a la transici\'on de niveles de energ\'ia m\'as probable
+del \'atomo de hidr\'ogeno.  Las galaxias que fueron observadas con
+esta fuerte l\'inea de emisi\'on se denominaron Lyman Alpha Emitters
+(LAEs). La deducci\'on de las propiedades f\'isicas de una galaxia a
+partir de esta l\'inea de emisi\'on se considera en gran parte como
+un problema abierto. Esto debido a su complejidad que requiere de
+t\'ecnicas computacionales para ser atacado.
 
 %Breve estado del arte
-Al necesitar entonces que toda la información sea derivada del espectro de estas galaxias, es necesario crear un modelo s\'olido y simplificado que pueda describirlas.
-Varios autores las han modelado como cuerpos que se expanden debido a material proyectado hacía afuera de la galaxia (denominados outflows), causados principalmente por supernovas.
-Otros las han modelado como cuerpos que rotan.
-Lastimosamente, \'estas aproximaciones no permiten encontrar todas las posibles morfologías de la línea.
-Por esta razón, para mi trabajo de grado, propongo considerar las LAEs como una distribuci\'on esf\'erica de \'atomos de Hidr\'ogeno, que rota como un cuerpo r\'igido y se expande radialmente, tal como se ve en la Fig.
-\ref{model}:
-
-\begin{figure}[h]
-	%uncomment next line to include a graphic file
-	\centerline{\includegraphics[width=4cm]{model.png}}
-	\caption{Modelo de una LAE que se expande y rota.}
-	\label{model}
-\end{figure}
+En esta tesis utilizo t\'ecnicas de transferencia radiativa
+computacional masivamente paralela para interpretar espectros de LAEs.
+Propongo considerar a las LAEs como una distribuci\'on esf\'erica de
+\'atomos de Hidr\'ogeno que tienen un movimiento de rotaci\'on
+superpuesto a movimiento radial de gas proyectado (outflow). Esto se
+motiva en que la rotaci\'on es un fen\'omeno com\'un a todas las
+galaxias y en que los outflows son una consecuencia natural de la
+explosi\'on de supernovas en una galaxia. 
+Este es un tipo de modelo que no hab\'ia sido considerado hasta la
+fecha para la interpretaci\'on de espectros de LAEs.
 
 %Resultados principales y conclusión corta
-Usando herramientas computacionales, simulo los efectos de velocidad rotacional, velocidad de expansión y masa de la LAE en el espectro de salida para diferentes combinaciones de \'estos tres parametros.
-La conclusi\'on principal es que este nuevo modelo reproduce las caracter\'isticas principales de LAEs observadas.
-Sin embargo, los ajustes observacionales se dejan para trabajo futuro.
-Este trabajo de grado, logra el objetivo de extraer la mayor informaci\'on posible de la l\'inea Ly-$\alpha$ de una LAE, con el fin de avanzar nuestro conocimiento sobre esta importante población de galaxias.
+Encuentro que la rotaci\'on sumada a una cantidad modesta
+de outflow es suficiente para reproducir  las caracter\'isticas
+principales de LAEs observadas. Esto se encuentra en constraste con
+la perspectiva m\'as aceptada en la comunidad que espera velocidades
+alt\'isimas de outflow para reproducir las observaciones, algo que
+desde el punto de vista energ\'etico es dificil de justificar para una
+galaxia j\'oven. Estos resultados dan pie para revisar las
+interpretaciones que se han hecho hasta ahora del estado f\'isico de
+galaxias j\'ovenes; haciendo un aporte a nuestro conocimiento para una
+reconstrucci\'on m\'as acertada de la historia de la evoluci\'on del
+Universo.
 

### 2016-02-25 abstract/abstract.tex
abstract v1

@@
 Typical LAEs are very distant ($z \gtrsim 2$).
-This makes that all the data astronomers can obtain from them is their spectra, and from there all the physical information of the galaxy must be derived.
+As a result all the data astronomers can obtain from them is their spectra, and from there all the physical information of the galaxy must be derived.
 Trying to solve this task requires the creation of a simplified and solid model.
-In this monograph I propose to consider LAEs as a spherical distribution of Hydrogen atoms that undergoes a solid body rotation and a radial expansion, as seen in Fig.
-\ref{model}.
+In this work we propose to consider LAEs as a spherical distribution of Hydrogen atoms that undergoes a solid body rotation and a radial expansion due to outflows.
+\\
 
-\begin{figure}[h]
-	%uncomment next line to include a graphic file
-	\centerline{\includegraphics[width=4cm]{model.png}}
-	\caption{Model of a Expanding and Rotating LAE.}
-	\label{model}
-\end{figure}
+We simulate the effect of rotational velocity ($v_{\mathrm{rot}}$), outflow velocity ($v_{\mathrm{out}}$) and optical depth ($\tau_{\mathrm{H}}$) of a LAE on its outgoing spectrum.
+This spectrum is obtained by considering the photons' path through the galaxy as a radiative transfer process.
+Then we used CLARA (Code for Lyman Alpha Radiation Analysis), a radiative transfer code that returns the resulting Ly-$\alpha$ line from the combination of the 3 parameters $v_{\mathrm{rot}}$, $v_{\mathrm{out}}$ and $\tau_{\mathrm{H}}$.\\
 
-I simulate the effect of rotational velocity, outflow velocity and optical depth of the LAE in the outgoing spectra.
 The main conclusion is that this new model reproduces LAEs observed features in a clear way and with consistent physical parameters.
-However, proper observational fits are left for future work.
-This monograph accomplishes the objective of extracting as much information as possible from a LAE's Ly-$\alpha$ line.
-
-%... a few optional references, such as \cite{Verhamme06}\\
+We find that rotation and a modest outflow quantity is enough to reproduce the primary characteristics of observed LAEs.
+This result goes against the most accepted community's perspective, which consists of having really high outflows in order to reproduce observations.
+However for such young galaxies these high outflows are energetically really hard to explain.
+We show that the outflows could be lowered by adding galaxy rotation to the model.
+The resulting spectra are roughly coherent with LAEs' observations and all the effects the parameters cause in the Ly-$\alpha$ profile, have been previously obtained by different authors.
+Nonetheless, proper observational fits are left for future work.
+This work accomplishes the objective of extracting as much information as possible from a LAE's Ly-$\alpha$ line.\\
 

### 2016-02-25 abstract/abstract.tex
nuevo abstract

--- 2016-02-04 abstract/abstract.tex
@@
 
-Young galaxies in the Universe have a strong Ly-$\alpha$ emission caused by the ionized Hydrogen atoms in their interstellar medium.
-When the spectrum of a galaxy has an intense peak around the Ly-$\alpha$ natural frequency ($2.46\times 10^{15}$ Hz) it is called a Lyman Alpha Emitter (LAE).
-Typical LAEs are very distant ($z \gtrsim 2$).
-This makes that all the data astronomers can obtain from them is their spectra, and from there all the physical information of the galaxy must be derived.
-Trying to solve this task requires the creation of a simplified and solid model.
-In this monograph I propose to consider LAEs as a spherical distribution of Hydrogen atoms that undergoes a solid body rotation and a radial expansion, as seen in Fig.
-\ref{model}.
-
-\begin{figure}[h]
-	%uncomment next line to include a graphic file
-	\centerline{\includegraphics[width=4cm]{model.png}}
-	\caption{Model of a Expanding and Rotating LAE.}
-	\label{model}
-\end{figure}
-
-I simulate the effect of rotational velocity, outflow velocity and optical depth of the LAE in the outgoing spectra.
-The main conclusion is that this new model reproduces LAEs observed features in a clear way and with consistent physical parameters.
-However, proper observational fits are left for future work.
-This monograph accomplishes the objective of extracting as much information as possible from a LAE's Ly-$\alpha$ line.
-
-%... a few optional references, such as \cite{Verhamme06}\\
+Radiative transfer MonteCarlo simulations are nowadays a required tool
+to interprete the Lyman-$\alpha$ line morphology. In this work we
+explore a new model that computes the join effect of outflows and
+bulk rotation. The main conclusion is that rotation and a modest
+outflow velocity is enough to reproduce the primary characteristics of
+observed LAEs. 
+We will show results of adjusting observational data for some selected
+objects to models with and without bulk rotation. We finalize by
+discussing the possible implications for these results in terms of the
+energetics required for supernova feedback and outflows in high
+redshift galaxies.
 

### 2016-02-26 abstract/abstract.tex
minor typos

@@
 Radiative transfer MonteCarlo simulations are nowadays a required tool
-to interprete the Lyman-$\alpha$ line morphology. In this work we
-explore a new model that computes the join effect of outflows and
+to interpret the Lyman-$\alpha$ line morphology. In this work we
+explore a new model that computes the joint effect of outflows and
 bulk rotation. The main conclusion is that rotation and a modest

### 2016-04-04 data/tau10E5/to_join_parts4.py
data for the 4th run

@@ -1,0 +1,23 @@
+import os
+
+logtau = 5
+vrots = [0]
+vouts = [5,25,50,100,200]
+
+for vrot in vrots:
+    for vout in vouts:
+
+        path = './vrot'+str(vrot)+'/vout'+str(vout)+'/'
+        filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+
+        #os.system('rm '+path+filename+'_out.ascii')
+        os.system('cat '+path+filename+'_out* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_out.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_out.ascii')
+        os.system('rm temp.txt')
+
+        #os.system('rm '+path+filename+'_in.ascii')
+        os.system('cat '+path+filename+'_in* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_in.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_in.ascii')
+        os.system('rm temp.txt')

### 2016-04-04 data/tau10E5/vrot0/vout25/tau10E5_vrot0_vout25.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E5_vrot0_vout25
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot0/vout25/tau10E5_vrot0_vout25.input

### 2016-04-04 data/tau10E5/vrot0/vout5/tau10E5_vrot0_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E5_vrot0_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot0/vout5/tau10E5_vrot0_vout5.input

### 2016-04-04 data/tau10E5/vrot0/vout50/tau10E5_vrot0_vout50.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E5_vrot0_vout50
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot0/vout50/tau10E5_vrot0_vout50.input

### 2016-04-04 data/tau10E6/to_join_parts4.py

@@ -1,0 +1,23 @@
+import os
+
+logtau = 6
+vrots = [0]
+vouts = [5,25,50,100,200]
+
+for vrot in vrots:
+    for vout in vouts:
+
+        path = './vrot'+str(vrot)+'/vout'+str(vout)+'/'
+        filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+
+        #os.system('rm '+path+filename+'_out.ascii')
+        os.system('cat '+path+filename+'_out* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_out.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_out.ascii')
+        os.system('rm temp.txt')
+
+        #os.system('rm '+path+filename+'_in.ascii')
+        os.system('cat '+path+filename+'_in* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_in.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_in.ascii')
+        os.system('rm temp.txt')

### 2016-04-04 data/tau10E6/vrot0/vout25/tau10E6_vrot0_vout25.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E6_vrot0_vout25
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=12
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout25/tau10E6_vrot0_vout25.input

### 2016-04-04 data/tau10E6/vrot0/vout50/tau10E6_vrot0_vout50.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E6_vrot0_vout50
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=12
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout50/tau10E6_vrot0_vout50.input

### 2016-04-04 data/tau10E6/vrot0/vout50/vout300/tau10E6_vrot0_vout300.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E6_vrot0_vout300
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=12
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout300/tau10E6_vrot0_vout300.input

### 2016-04-04 data/tau10E6/vrot0/vout50/vout5/tau10E6_vrot0_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E6_vrot0_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=12
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout5/tau10E6_vrot0_vout5.input

### 2016-04-04 data/tau10E7/to_join_parts4.py

@@ -1,0 +1,23 @@
+import os
+
+logtau = 7
+vrots = [0]
+vouts = [5,25,50,100,200]
+
+for vrot in vrots:
+    for vout in vouts:
+
+        path = './vrot'+str(vrot)+'/vout'+str(vout)+'/'
+        filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+
+        #os.system('rm '+path+filename+'_out.ascii')
+        os.system('cat '+path+filename+'_out* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_out.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_out.ascii')
+        os.system('rm temp.txt')
+
+        #os.system('rm '+path+filename+'_in.ascii')
+        os.system('cat '+path+filename+'_in* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_in.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_in.ascii')
+        os.system('rm temp.txt')

### 2016-04-04 data/tau10E7/vrot0/vout25/tau10E7_vrot0_vout25.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E7_vrot0_vout25
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=24
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 24 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E7/vrot0/vout25/tau10E7_vrot0_vout25.input

### 2016-04-04 data/tau10E7/vrot0/vout5/tau10E7_vrot0_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E7_vrot0_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=24
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 24 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E7/vrot0/vout5/tau10E7_vrot0_vout5.input

### 2016-04-04 data/tau10E7/vrot0/vout50/tau10E7_vrot0_vout50.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E7_vrot0_vout50
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=24
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 24 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E7/vrot0/vout50/tau10E7_vrot0_vout50.input

### 2016-04-04 script/4th_run/to_make_inputs.py

@@ -1,0 +1,81 @@
+#Example of directory: /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot0/vout100
+
+logtaus = [5,6,7]
+vrots = [0]
+vouts = [5,25,50,100,200]
+
+for logtau in logtaus:
+    for vrot in vrots:
+        for vout in vouts:
+
+            filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+            path = '/lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E'+str(logtau)+'/vrot'+str(vrot)+'/vout'+str(vout)+'/'
+            input_file = open(path+filename+'.input', 'w')
+
+            input_file.write('% input files and formats*/\n')
+            input_file.write('InputDir\t'+path+'\n')
+            input_file.write('CubeName\tgrid_object_100\n')
+            input_file.write('\n')
+            input_file.write('% output files\n')
+            input_file.write('OutputDir\t'+path+'\n')
+            input_file.write('OutputFile\t'+filename+'\n')
+            input_file.write('\n')
+            input_file.write('\n')
+            input_file.write('% define the problem/test to solve\n')
+            input_file.write('NeufeldSlab\t0\n')
+            input_file.write('NeufeldCube\t0\n')
+            input_file.write('ExpandingSphere\t0\n')
+            input_file.write('RotatingSphere\t0\n')
+            input_file.write('OutflowRotationSphere\t1\n')
+            input_file.write('TestParallelVel\t0\n')
+            input_file.write('TestParallelVelFast\t0\n')
+            input_file.write('TestFirstScatter\t0\n')
+            input_file.write('TestRND\t0\n')
+            input_file.write('TestPerpVel\t0\n')
+            input_file.write('SimulationCube\t0\n')
+            input_file.write('HomogeneousInit\t0\n')
+            input_file.write('\n')
+            input_file.write('% parameters for the tests\n')
+            input_file.write('Test_a\t0.00014\n')
+            input_file.write('Test_x\t2.0\n')
+            input_file.write('OutputTestFile\ttest_b\n')
+            input_file.write('\n')
+            input_file.write('% define some physical parameters of the problem to solve\n')
+            input_file.write('Temperature\t10000.0 % in  Kelvin\n')
+            input_file.write('Tau\t1.0e'+str(logtau)+'\n')
[…]

### 2016-04-04 script/4th_run/to_make_scripts.py

@@ -1,0 +1,32 @@
+#Example of directory: /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot0/vout100
+
+ppns = [6, 12, 24]
+logtaus = [5,6,7]
+vrots = [0]
+vouts = [5,25,50,100,200]
+
+for i in range(3):
+
+    logtau = logtaus[i]
+    ppn = ppns[i]
+
+    for vrot in vrots:
+        for vout in vouts:
+
+            filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+            path = '/lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E'+str(logtau)+'/vrot'+str(vrot)+'/vout'+str(vout)+'/'
+            script = open(path+filename+'.sh', 'w')
+
+            script.write('#!/bin/sh\n')
+            script.write('#PBS -q batch\n')
+            script.write('#PBS -N '+filename+'\n')
+            script.write('#PBS -l mem=128mb\n')
+            script.write('#PBS -l nodes=1:ppn='+str(ppn)+'\n')
+            script.write('#PBS -M mc.remolina197@uniandes.edu.co\n')
+            script.write('#PBS -m abe\n')
+            script.write('\n')
+            script.write('module load openmpi/1.8.5\n')
+            script.write('cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/\n')
+            script.write('mpiexec -n '+str(ppn)+' ./mine.x '+path+filename+'.input')
+
+            script.close()

### 2016-04-04 script/4th_run/to_run_scripts.py

@@ -1,0 +1,13 @@
+import os
+
+logtaus = [5,6,7]
+vrots = [0]
+vouts = [5,25,50,100,200]
+
+for logtau in logtaus:
+    for vrot in vrots:
+        for vout in vouts:
+            filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+            path = '/lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E'+str(logtau)+'/vrot'+str(vrot)+'/vout'+str(vout)+'/'
+            #os.system('qsub '+path+filename+'.sh')
+            print 'qsub '+path+filename+'.sh'

### 2016-04-04 data/tau10E6/vrot0/vout50/vout100/tau10E6_vrot0_vout100.sh
data for the 4th run

@@ -1,11 +1,0 @@
-#!/bin/sh
-#PBS -q batch
-#PBS -N tau10E6_vrot0_vout100
-#PBS -l mem=128mb
-#PBS -l nodes=1:ppn=12
-#PBS -M mc.remolina197@uniandes.edu.co
-#PBS -m abe
-
-module load openmpi/1.8.5
-cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
-mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout100/tau10E6_vrot0_vout100.input

### 2016-04-04 data/tau10E6/vrot0/vout50/vout200/tau10E6_vrot0_vout200.sh

@@ -1,11 +1,0 @@
-#!/bin/sh
-#PBS -q batch
-#PBS -N tau10E6_vrot0_vout200
-#PBS -l mem=128mb
-#PBS -l nodes=1:ppn=12
-#PBS -M mc.remolina197@uniandes.edu.co
-#PBS -m abe
-
-module load openmpi/1.8.5
-cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
-mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout200/tau10E6_vrot0_vout200.input

### 2016-04-04 data/tau10E6/vrot0/vout50/vout25/tau10E6_vrot0_vout25.sh

@@ -1,11 +1,0 @@
-#!/bin/sh
-#PBS -q batch
-#PBS -N tau10E6_vrot0_vout25
-#PBS -l mem=128mb
-#PBS -l nodes=1:ppn=12
-#PBS -M mc.remolina197@uniandes.edu.co
-#PBS -m abe
-
-module load openmpi/1.8.5
-cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
-mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout25/tau10E6_vrot0_vout25.input

### 2016-04-04 data/tau10E6/vrot0/vout50/vout300/tau10E6_vrot0_vout300.sh

@@ -1,11 +1,0 @@
-#!/bin/sh
-#PBS -q batch
-#PBS -N tau10E6_vrot0_vout300
-#PBS -l mem=128mb
-#PBS -l nodes=1:ppn=12
-#PBS -M mc.remolina197@uniandes.edu.co
-#PBS -m abe
-
-module load openmpi/1.8.5
-cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
-mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout300/tau10E6_vrot0_vout300.input

### 2016-04-04 data/tau10E6/vrot0/vout50/vout5/tau10E6_vrot0_vout5.sh

@@ -1,11 +1,0 @@
-#!/bin/sh
-#PBS -q batch
-#PBS -N tau10E6_vrot0_vout5
-#PBS -l mem=128mb
-#PBS -l nodes=1:ppn=12
-#PBS -M mc.remolina197@uniandes.edu.co
-#PBS -m abe
-
-module load openmpi/1.8.5
-cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
-mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot0/vout5/tau10E6_vrot0_vout5.input

### 2016-04-04 abstract/abstract_spanish.tex

@@
 
-\title{\textbf{Influencia de rotaci\'on y outflows\\
-			    en la l\'inea espectral Lyman Alpha}}
+\title{\textbf{Influencia de rotaci\'on y outflows en la\\
+			   l\'inea espectral Lyman Alpha}}
 
-\author{Maria Camila Remolina-Gutierrez$^1$, Jaime E. Forero-Romero$^1$\\ \vspace{3mm}
-	    mc.remolina197@uniandes.edu.co, \hspace{0.8mm} je.forero@uniandes.edu.co\\ 
-		$^1$ Departamento de F\'{i}sica, Universidad de los Andes, Bogot\'a, Colombia}
+\author{Autora:\\ \vspace{2mm}
+		Maria Camila Remolina-Gutierrez - mc.remolina197@uniandes.edu.co\\ 
+		Director: \\ \vspace{3mm}
+		Jaime E. Forero-Romero - je.forero@uniandes.edu.co\\ 
+		Departamento de F\'{i}sica, Universidad de los Andes, Bogot\'a, Colombia}
 \date{} % <--- leave date empty
 \maketitle\thispagestyle{empty} %% <-- you need this for the first page
-\hyphenation{emi-si\'on ellas}
+\hyphenation{emi-si\'on ellas es-tu-dia-ron}
 
 %Introducción y motivación
-Las galaxias son la clave fundamental para entender el Universo.  Cuando observamos una galaxia distante estamos viendo el pasado. A partir de este esfuerzo de
-arqueolog\'ia astron\'omica podemos reconstruir la historia del Universo.
+Las galaxias son la clave fundamental para entender el Universo.  Cuando 
+observamos una galaxia distante estamos viendo el pasado. A partir de este esfuerzo
+de arqueolog\'ia astron\'omica podemos reconstruir la historia del Universo.
 
-Este proceso de reconstrucci\'on empez\'o en el Siglo XX. Los
-astr\'onomos estudiaron en detalle las galaxias m\'as cercanas,
-haciendo observaciones de un Universo ya maduro. 
-A finales del Siglo XX y comienzos del XXI se logran obtener se\~nales de
-galaxias mucho m\'as distantes; conjuntos de gas y estrellas que
-emitieron su luz cuando el Universo era j\'oven y ten\'ia menos
-de un quinto de su edad actual.  
-Estas galaxias son tan lejanas que que
-las im\'agenes aportan poco al momento de deducir las propiedades de
-una galaxia. 
-En estos casos mayor parte de la informaci\'on viene en los
-espectros, es decir, en la descomposici\'on de la luz en radiaci\'on de
-diferentes longitudes de onda. 
-De esta manera, el reto astrof\'isico se convierte en poder darle a estos
-espectros de galaxias distantes una  interpretaci\'on.
+Este proceso de reconstrucci\'on empez\'o en el Siglo XX. Los astr\'onomos 
+estudiaron en detalle las galaxias m\'as cercanas, haciendo observaciones de
+un Universo ya maduro. A finales del Siglo XX y comienzos del XXI, se logran 
+obtener se\~nales de galaxias mucho m\'as distantes; conjuntos de gas y estrellas 
+que emitieron su luz cuando el Universo era joven y ten\'ia menos de un quinto 
+de su edad actual. Estas galaxias son tan lejanas, que las im\'agenes aportan 
+poco al momento de deducir las propiedades de una galaxia. En estos casos, mayor 
+parte de la informaci\'on viene en los espectros, es decir, en la descomposici\'on 
+de la luz en radiaci\'on de diferentes longitudes de onda. De esta manera, el reto 
+astrof\'isico se convierte en poder darle a estos espectros de galaxias distantes
+una  interpretaci\'on.
 
-En los primeros esfuerzos conjuntos de interpretaci\'on y
-observaci\'on, se hizo claro que varias galaxias
-j\'ovenes deben tener una alta cantidad de hidr\'ogeno y estrellas en
-proceso de formaci\'on. Bajo estas condiciones gran parte de su
-luz tiene la longitud de onda de la l\'inea de emisi\'on Lyman-$\alpha$,
-correspondiente a la transici\'on de niveles de energ\'ia m\'as probable
-del \'atomo de hidr\'ogeno.  Las galaxias que fueron observadas con
-esta fuerte l\'inea de emisi\'on se denominaron Lyman Alpha Emitters
-(LAEs). La deducci\'on de las propiedades f\'isicas de una galaxia a
-partir de esta l\'inea de emisi\'on se considera en gran parte como
-un problema abierto. Esto debido a su complejidad que requiere de
-t\'ecnicas computacionales para ser atacado.
+En los primeros esfuerzos conjuntos de interpretaci\'on y observaci\'on, se hizo
+claro que varias galaxias j\'ovenes deben tener una alta cantidad de hidr\'ogeno 
+y estrellas en proceso de formaci\'on. Bajo estas condiciones, gran parte de su luz
+tiene la longitud de onda de la l\'inea de emisi\'on Lyman-$\alpha$, correspondiente
+a la transici\'on de niveles de energ\'ia m\'as probable del \'atomo de hidr\'ogeno.
+Las galaxias que fueron observadas con esta fuerte l\'inea de emisi\'on se denominaron
+Lyman Alpha Emitters (LAEs). La deducci\'on de las propiedades f\'isicas de una 
+galaxia a partir de esta l\'inea de emisi\'on se considera, en gran parte, como un
+problema abierto. Esto debido a que su complejidad requiere de t\'ecnicas 
+computacionales para ser atacado.
 
 %Breve estado del arte
-En esta tesis utilizo t\'ecnicas de transferencia radiativa
-computacional masivamente paralela para interpretar espectros de LAEs.
-Propongo considerar a las LAEs como una distribuci\'on esf\'erica de
-\'atomos de Hidr\'ogeno que tienen un movimiento de rotaci\'on
-superpuesto a movimiento radial de gas proyectado (outflow). Esto se
-motiva en que la rotaci\'on es un fen\'omeno com\'un a todas las
-galaxias y en que los outflows son una consecuencia natural de la
-explosi\'on de supernovas en una galaxia. 
-Este es un tipo de modelo que no hab\'ia sido considerado hasta la
-fecha para la interpretaci\'on de espectros de LAEs.
+En esta tesis utilizo t\'ecnicas de transferencia radiativa computacional masivamente 
+paralela para interpretar espectros de LAEs. Propongo considerar a las LAEs como una 
+distribuci\'on esf\'erica de \'atomos de Hidr\'ogeno que tienen un movimiento de rotaci\'on
+superpuesto a movimiento radial de gas proyectado (outflow). Esto se motiva en que la 
+rotaci\'on es un fen\'omeno com\'un a todas las galaxias y en que los outflows son una 
+consecuencia natural de la explosi\'on de supernovas en una galaxia. Este es un tipo 
+de modelo que no hab\'ia sido considerado, hasta la fecha, para la interpretaci\'on de 
+espectros de LAEs.
 
 %Resultados principales y conclusión corta
-Encuentro que la rotaci\'on sumada a una cantidad modesta
-de outflow es suficiente para reproducir  las caracter\'isticas
-principales de LAEs observadas. Esto se encuentra en constraste con
-la perspectiva m\'as aceptada en la comunidad que espera velocidades
-alt\'isimas de outflow para reproducir las observaciones, algo que
-desde el punto de vista energ\'etico es dificil de justificar para una
-galaxia j\'oven. Estos resultados dan pie para revisar las
-interpretaciones que se han hecho hasta ahora del estado f\'isico de
-galaxias j\'ovenes; haciendo un aporte a nuestro conocimiento para una
-reconstrucci\'on m\'as acertada de la historia de la evoluci\'on del
-Universo.
+Encuentro que la rotaci\'on sumada a una cantidad modesta de outflow, es suficiente 
+para reproducir  las caracter\'isticas principales de LAEs observadas. Esto se encuentra
+en contraste con la perspectiva m\'as aceptada en la comunidad, que espera velocidades
+alt\'isimas de outflow para reproducir las observaciones, algo que desde el punto de 
+vista energ\'etico es dificil de justificar para una galaxia joven. Estos resultados 
+dan pie para revisar las interpretaciones que se han hecho hasta ahora sobre el estado 
+f\'isico de galaxias j\'ovenes; haciendo un aporte a nuestro conocimiento para una 
+reconstrucci\'on m\'as acertada de la historia de la evoluci\'on del Universo.
 

### 2016-04-13 fit/emcee_skewed_gaussian.py
emcee for gaussian fit

@@ -1,0 +1,139 @@
+from numpy import *
+from scipy.interpolate import interp1d
+from scipy.special import erf
+import emcee
+import os
+import corner
+import matplotlib.pyplot as plt
+#FUNCTIONS
+
+#Bifurcation index
+def find_bifurcation_index(intensities):
+    i_prev = 0
+    bif_index = -1
+
+    intensities_length = len(intensities)
+    start_index = 0.3*intensities_length
+    end_index = 0.7*intensities_length
+
+    intensities = intensities[start_index:end_index]
+
+    for i in range(len(intensities)-1):
+        i_curr = intensities[i]
+        i_next = intensities[i+1]
+        if(i_prev >= i_curr and i_curr<=i_next):
+            bif_index = i
+        i_prev = i_curr
+
+    if (bif_index==-1):
+        return intensities_length/2
+    else:
+        return bif_index+start_index
+
+#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
+def gaussian(x, amplitude, center, sigma):
+    """1 dimensional gaussian:
+    gaussian(x, amplitude, center, sigma)
+    """
+    return (amplitude/(sqrt(2*pi)*sigma)) * exp(-(1.0*x-center)**2 /(2*sigma**2))
+
+#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
+def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos):
+    """Gaussian, skewed with error function, equal to
+     gaussian(x, center, sigma)*(1+erf(beta*(x-center)))
+    with beta = gamma/(sigma*sqrt(2))
+    with  gamma < 0:  tail to low value of centroid
+          gamma > 0:  tail to high value of centroid
+    see http://en.wikipedia.org/wiki/Skew_normal_distribution
+    """
+    asym_neg = 1 + erf(gamma_neg*(x-center_neg)/(sqrt(2.0)*sigma_neg))
[…]

### 2016-04-13 paper/paper.tex

@@
 %Document Class
-\documentclass{latex/emulateapj}
-\usepackage{multirow,color,wrapfig,ulem}
+\documentclass[twocolappendix]{latex/emulateapj}
 
@@
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.3\textwidth]{./figures/model}
-	\end{center}
-	\caption{\textbf{Model:} Spherical LAE with tangential and radial velocity due to rotation and outflows.
-		\label{fig:model}}
-\end{figure}
-
 These 3 parameters, \vrot, \vout and \tauh, leave the idea of a model of LAE, that although simple, considers the main galaxy's dynamics.
@@
 I illustrate this in Fig.
-\ref{fig:viewing_angle_sketch}.
+\ref{fig:model}.
 \\
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/viewing_angle_sketch}
+		\includegraphics[width=0.48\textwidth]{./figures/model}
 	\end{center}
-	\caption{\textbf{Viewing Angle Sketch:} The galaxy cut is in the $y-z$ plane perspective and the observer is located at an specific viewing angle of the sphere.
+	\caption{\textbf{Model:} Spherical LAE with tangential and radial velocities due to rotation and outflows, respectively.
+The galaxy cut is in the $y-z$ plane perspective.
+The observer is located at an specific viewing angle of the sphere.
 Only photons with a direction that enters in this range of vision are taken into account to build the observed spectrum.
-		\label{fig:viewing_angle_sketch}}
+		\label{fig:model}}
 \end{figure}
@@
 The resulting sets of spectra are in Figs.
-\ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$.
+\ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$ available in the Appendix \ref{ap:results_figures}.
 I do not include the spectra for \vout = 50, 75 \kms because they are similar to \vout = 25 \kms.
 For this reason, I limit in all the plots \vout to a maximum value of 25 \kms.
-Nevertheless, the figures for \vout = 50, 75 \kms are in the Appendix \ref{first_plots} in Figs.
+Nevertheless, the figures for \vout = 50, 75 \kms are in the Appendix \ref{ap:additional_figures} in Figs.
 \ref{fig:2_tau10E5_phi83-90}, \ref{fig:2_tau10E6_phi83-90} and \ref{fig:2_tau10E7_phi83-90}.\\
-
-\newpage
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/3_tau10E5_phi83-90}
-	\end{center}
-	\caption{\textbf{\lya profile for \tauh$=10^5$:} With \vrot ranging $50,100$ \kms and \vout ranging $5,10,15,20,25$ \kms.
-The intensity is in arbitrary units.
-The viewing angle $\theta \simeq 90^\circ$.
-The dashed vertical line indicates the \lya line's natural frequency.
-		\label{fig:3_tau10E5_phi83-90}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/3_tau10E6_phi83-90}
-	\end{center}
-	\caption{\textbf{\lya profile for \tauh$=10^6$:}  Same as Fig. \ref{fig:3_tau10E5_phi83-90} but for \tauh$=10^6$.
-		\label{fig:3_tau10E6_phi83-90}}
-\end{figure}
-
-\newpage
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/3_tau10E7_phi83-90}
-	\end{center}
-	\caption{\textbf{\lya profile for \tauh$=10^7$:} Same as Fig. \ref{fig:3_tau10E5_phi83-90} but for \tauh$=10^7$.
-		\label{fig:3_tau10E7_phi83-90}}
-\end{figure}
-
-\newpage
 
@@
 All of this work is available online and free to use to anyone.
-The data, source code and instructions to replicate my monograph's results are in the GitHub repository:  \texttt{https://github.com/astroandes/CLARA\_RotationOutflows}.
+The data, source code and instructions to replicate my monograph's results are in the GitHub repository:  \texttt{https://github.com/astroandes/CLARA\_RotationOut flows}.
 \\
@@
 
-\newpage
-
 %------------------------APPENDIX----------------------------
@@
 \appendix
+
+\begin{figure}[!htbp]
+	\centering
+	\begin{minipage}[b]{0.45\textwidth}
+		\section{Results' Figures}
+		\label{ap:results_figures}
+
+		\includegraphics[width=\textwidth]{./figures/3_tau10E5_phi83-90}
+		\caption{\textbf{\lya profile for \tauh$=10^5$:} With \vrot ranging $50,100$ \kms and \vout ranging $5,10,15,20,25$ \kms.
+The intensity is in arbitrary units.
+The viewing angle $\theta \simeq 90^\circ$.
+The dashed vertical line indicates the \lya line's natural frequency.
+			\label{fig:3_tau10E5_phi83-90}}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{./figures/3_tau10E6_phi83-90}
+		\caption{\textbf{\lya profile for \tauh$=10^6$:}  Same as Fig. \ref{fig:3_tau10E5_phi83-90} but for \tauh$=10^6$. 
+			\label{fig:3_tau10E6_phi83-90}}
+		\vspace{1.5mm}
+	\end{minipage}
+\end{figure}
+
+\begin{figure}[!htbp]
+	\centering
+	\begin{minipage}[b]{0.45\textwidth}
+		\vspace{3mm}
+		\includegraphics[width=\textwidth]{./figures/3_tau10E7_phi83-90}
+		\caption{\textbf{\lya profile for \tauh$=10^7$:} Same as Fig. \ref{fig:3_tau10E5_phi83-90} but for \tauh$=10^7$.
+			\label{fig:3_tau10E7_phi83-90}}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.45\textwidth}
 		\section{Additional Figures}
+		\label{ap:additional_figures}
 
@@
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=1\textwidth]{./figures/appendix/2_tau10E5_phi83-90}
-	\end{center}
+		\vspace{2.7cm}
+
+		\includegraphics[width=\textwidth]{./figures/appendix/2_tau10E5_phi83-90}
 		\caption{\textbf{\lya profile for \tauh$=10^5$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
 			\label{fig:2_tau10E5_phi83-90}}
+	\end{minipage}
 \end{figure}
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=1\textwidth]{./figures/appendix/2_tau10E6_phi83-90}
-	\end{center}
+\begin{figure}[!htbp]
+	\centering
+	\begin{minipage}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{./figures/appendix/2_tau10E6_phi83-90}
 		\caption{\textbf{\lya profile for \tauh$=10^6$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
 			\label{fig:2_tau10E6_phi83-90}}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=1\textwidth]{./figures/appendix/2_tau10E7_phi83-90}
-	\end{center}
+	\end{minipage}
+	\hfill
+	\begin{minipage}[b]{0.45\textwidth}
+		\includegraphics[width=\textwidth]{./figures/appendix/2_tau10E7_phi83-90}
 		\caption{\textbf{\lya profile for \tauh$=10^7$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
 			\label{fig:2_tau10E7_phi83-90}}
+	\end{minipage}
 \end{figure}

### 2016-04-13 fit/emcee_skewed_gaussian.py
emcee hates me :( I have to try by hand

@@ -4,7 +4,7 @@ from scipy.special import erf
 import emcee
 import os
 import corner
-import matplotlib.pyplot as plt
+
 #FUNCTIONS
 
 #Bifurcation index
@@ -55,15 +55,35 @@ def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, am
     return g_neg+g_pos
 
 
+def lnprior(param):
 
-def lnprob(param, x_d, y_d):
     amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos = param
 
-    if (0.0<amplitude_neg<500.0) and (-50<center_neg<30) and (0<sigma_neg<50) and (-10<gamma_neg<10) and (0.0<amplitude_pos<500.0) and (-30<center_pos<50) and (0<sigma_pos<50) and (-10<gamma_pos<10):
-        y_m = sum_2_skewed_gaussian(x_d, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_neg, center_neg, sigma_neg, gamma_neg)
-        chi_squared = (1.0/2.0)*sum((y_d-y_m)**2)
-        return -chi_squared
+    if (0.0<amplitude_neg<200.0) and (-10<center_neg<0) and (0<sigma_neg<30) and (-15<gamma_neg<0) and (0.0<amplitude_pos<200.0) and (0<center_pos<10) and (0<sigma_pos<30) and (0<gamma_pos<15):
+        return 0.0
+
     return -inf
+
+
+def lnlike(param, x_d, y_d):
+
+    amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos = param
+
+    y_m = sum_2_skewed_gaussian(x_d, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_neg, center_neg, sigma_neg, gamma_neg)
+
+    chi_squared = (1.0/2.0)*sum((y_d-y_m)**2)
+
+    return -chi_squared
+
+
+def lnprob(param, x_d, y_d):
+
+    lp = lnprior(param)
+
+    if not isfinite(lp):
+        return -inf
+
+    return lp + lnlike(param, x_d, y_d)
 
 
 #emcee
[…]

### 2016-04-13 plots/3rd_run/mcmc_skewed_gaussian.py
mcmc fit for the 3rd run

@@ -1,0 +1,156 @@
+from numpy import *
+from scipy.special import erf
+
+#FUNCTIONS
+
+#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
+def gaussian(x, amplitude, center, sigma):
+    """1 dimensional gaussian:
+    gaussian(x, amplitude, center, sigma)
+    """
+    return (amplitude/(sqrt(2*pi)*sigma)) * exp(-(1.0*x-center)**2 /(2*sigma**2))
+
+#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
+def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos):
+    """Gaussian, skewed with error function, equal to
+     gaussian(x, center, sigma)*(1+erf(beta*(x-center)))
+    with beta = gamma/(sigma*sqrt(2))
+    with  gamma < 0:  tail to low value of centroid
+          gamma > 0:  tail to high value of centroid
+    see http://en.wikipedia.org/wiki/Skew_normal_distribution
+    """
+    asym_neg = 1 + erf(gamma_neg*(x-center_neg)/(sqrt(2.0)*sigma_neg))
+    g_neg = asym_neg * gaussian(x, amplitude_neg, center_neg, sigma_neg)
+
+    asym_pos = 1 + erf(gamma_pos*(x-center_pos)/(sqrt(2.0)*sigma_pos))
+    g_pos = asym_pos * gaussian(x, amplitude_pos, center_pos, sigma_pos)
+
+    return g_neg+g_pos
+
+def likelihood(y_obs, y_mod):
+    chi_squared_y = (1.0/2.0)*sum((y_obs - y_mod)**2)
+    return -chi_squared_y
+
+
+#MCMC
+def mcmc_skewed_gaussian(x_d, y_d):
+
+    #Initial guess
+    amplitude_neg0 = 9000
+    center_neg0 = -2
+    sigma_neg0 = 4
+    gamma_neg0 = -4
+    amplitude_pos0 = 4000
+    center_pos0 = 2
+    sigma_pos0 = 3
+    gamma_pos0 = 2
+
+    x_fit0 = x_d
[…]

### 2016-04-13 fit/mcmc_skewed_gaussian.py
fit obtained for all data, now time to compare them

@@ -1,0 +1,165 @@
+from numpy import *
+from scipy.special import erf
+
+#FUNCTIONS
+
+#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
+def gaussian(x, amplitude, center, sigma):
+    """1 dimensional gaussian:
+    gaussian(x, amplitude, center, sigma)
+    """
+    return (amplitude/(sqrt(2*pi)*sigma)) * exp(-(1.0*x-center)**2 /(2*sigma**2))
+
+#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
+def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos):
+    """Gaussian, skewed with error function, equal to
+     gaussian(x, center, sigma)*(1+erf(beta*(x-center)))
+    with beta = gamma/(sigma*sqrt(2))
+    with  gamma < 0:  tail to low value of centroid
+          gamma > 0:  tail to high value of centroid
+    see http://en.wikipedia.org/wiki/Skew_normal_distribution
+    """
+    asym_neg = 1 + erf(gamma_neg*(x-center_neg)/(sqrt(2.0)*sigma_neg))
+    g_neg = asym_neg * gaussian(x, amplitude_neg, center_neg, sigma_neg)
+
+    asym_pos = 1 + erf(gamma_pos*(x-center_pos)/(sqrt(2.0)*sigma_pos))
+    g_pos = asym_pos * gaussian(x, amplitude_pos, center_pos, sigma_pos)
+
+    return g_neg+g_pos
+
+def likelihood(y_obs, y_mod):
+    chi_squared_y = (1.0/2.0)*sum((y_obs - y_mod)**2)
+    return -chi_squared_y
+
+
+#MCMC
+def mcmc_skewed_gaussian(x_d, y_d):
+
+    #Initial guess
+    amplitude_neg0 = 65
+    center_neg0 = -4
+    sigma_neg0 = 9
+    gamma_neg0 = -4
+    amplitude_pos0 = 38
+    center_pos0 = 2
+    sigma_pos0 = 8
+    gamma_pos0 = 4
+
+    x_fit0 = x_d
[…]

### 2016-04-13 fit/mcmc_skewed_gaussian.py
removing some files

@@ -1,165 +1,0 @@
-from numpy import *
-from scipy.special import erf
-
-#FUNCTIONS
-
-#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
-def gaussian(x, amplitude, center, sigma):
-    """1 dimensional gaussian:
-    gaussian(x, amplitude, center, sigma)
-    """
-    return (amplitude/(sqrt(2*pi)*sigma)) * exp(-(1.0*x-center)**2 /(2*sigma**2))
-
-#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
-def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos):
-    """Gaussian, skewed with error function, equal to
-     gaussian(x, center, sigma)*(1+erf(beta*(x-center)))
-    with beta = gamma/(sigma*sqrt(2))
-    with  gamma < 0:  tail to low value of centroid
-          gamma > 0:  tail to high value of centroid
-    see http://en.wikipedia.org/wiki/Skew_normal_distribution
-    """
-    asym_neg = 1 + erf(gamma_neg*(x-center_neg)/(sqrt(2.0)*sigma_neg))
-    g_neg = asym_neg * gaussian(x, amplitude_neg, center_neg, sigma_neg)
-
-    asym_pos = 1 + erf(gamma_pos*(x-center_pos)/(sqrt(2.0)*sigma_pos))
-    g_pos = asym_pos * gaussian(x, amplitude_pos, center_pos, sigma_pos)
-
-    return g_neg+g_pos
-
-def likelihood(y_obs, y_mod):
-    chi_squared_y = (1.0/2.0)*sum((y_obs - y_mod)**2)
-    return -chi_squared_y
-
-
-#MCMC
-def mcmc_skewed_gaussian(x_d, y_d):
-
-    #Initial guess
-    amplitude_neg0 = 65
-    center_neg0 = -4
-    sigma_neg0 = 9
-    gamma_neg0 = -4
-    amplitude_pos0 = 38
-    center_pos0 = 2
-    sigma_pos0 = 8
-    gamma_pos0 = 4
-
-    x_fit0 = x_d
[…]

### 2016-04-14 plots/3rd_run/mcmc_skewed_gaussian.py
changes meanwhile I fix plots limits so they look pretty

@@ -153,4 +153,15 @@ def mcmc_skewed_gaussian(x_d, y_d):
     sigma_pos = sigma_pos_walk[-1]
     gamma_pos = gamma_pos_walk[-1]
 
-    return amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos
+    #Standard Deviations (discarding first 50 steps)
+
+    amplitude_neg_std = std(amplitude_neg_walk[50:])
+    center_neg_std = std(center_neg_walk[50:])
+    sigma_neg_std = std(sigma_neg_walk[50:])
+    gamma_neg_std = std(gamma_neg_walk[50:])
+    amplitude_pos_std = std(amplitude_pos_walk[50:])
+    center_pos_std = std(center_pos_walk[50:])
+    sigma_pos_std = std(sigma_pos_walk[50:])
+    gamma_pos_std = std(gamma_pos_walk[50:])
+
+    return amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos, amplitude_neg_std, center_neg_std, sigma_neg_std, gamma_neg_std, amplitude_pos_std, center_pos_std, sigma_pos_std, gamma_pos_std

### 2016-04-15 fit/emcee_skewed_gaussian.py
Fixed emcee function :)

@@ -1,9 +1,7 @@
-from numpy import *
+# Hala "import numpy as np"
+import numpy as np, emcee, os, corner
 from scipy.interpolate import interp1d
 from scipy.special import erf
-import emcee
-import os
-import corner
 
 #FUNCTIONS
 
@@ -35,7 +33,7 @@ def gaussian(x, amplitude, center, sigma):
     """1 dimensional gaussian:
     gaussian(x, amplitude, center, sigma)
     """
-    return (amplitude/(sqrt(2*pi)*sigma)) * exp(-(1.0*x-center)**2 /(2*sigma**2))
+    return (amplitude/(np.sqrt(2*np.pi)*sigma)) * np.exp(-(1.0*x-center)**2 /(2*sigma**2))
 
 #Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
 def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos):
@@ -46,44 +44,24 @@ def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, am
           gamma > 0:  tail to high value of centroid
     see http://en.wikipedia.org/wiki/Skew_normal_distribution
     """
-    asym_neg = 1 + erf(gamma_neg*(x-center_neg)/(sqrt(2.0)*sigma_neg))
+    asym_neg = 1 + erf(gamma_neg*(x-center_neg)/(np.sqrt(2.0)*sigma_neg))
     g_neg = asym_neg * gaussian(x, amplitude_neg, center_neg, sigma_neg)
 
-    asym_pos = 1 + erf(gamma_pos*(x-center_pos)/(sqrt(2.0)*sigma_pos))
+    asym_pos = 1 + erf(gamma_pos*(x-center_pos)/(np.sqrt(2.0)*sigma_pos))
     g_pos = asym_pos * gaussian(x, amplitude_pos, center_pos, sigma_pos)
 
     return g_neg+g_pos
 
-
-def lnprior(param):
-
-    amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos = param
-
[…]

### 2016-04-15 fit/emcee_skewed_gaussian.py
Added percentiles to triangle plot and organized code

@@ -96,7 +96,7 @@ def emcee_skewed_gaussian(x_d, y_d):
     #Running emcee
     ndim = 8
     nwalkers = 16
-    nsteps = 10000
+    nsteps = 5000
 
     pos = [first_guess+ 1e-3*np.random.randn(ndim) for i in range(nwalkers)]
 
@@ -106,6 +106,7 @@ def emcee_skewed_gaussian(x_d, y_d):
 
     # Saving results
     samples_fc = sampler.flatchain
+    logprob_fc = sampler.flatlnprobability
 
     np.savetxt('sampler_flatchain.dat', samples_fc, delimiter=',')
 
@@ -113,7 +114,8 @@ def emcee_skewed_gaussian(x_d, y_d):
     print("Mean acceptance fraction: {0:.3f} (Should be between 0.25 and 0.5 approximately)".format(np.mean(sampler.acceptance_fraction)))
 
     #Discard the initial 50 steps
-    samples = sampler.chain[:, 50:, :].reshape((-1, ndim))
+    samples = samples_fc[50:]
+    logprob = logprob_fc[50:]
 
     # Unpack the walk for each parameter
     amplitude_neg_walk, center_neg_walk, sigma_neg_walk, gamma_neg_walk, amplitude_pos_walk, center_pos_walk, sigma_pos_walk, gamma_pos_walk = np.transpose(samples)
@@ -128,6 +130,16 @@ def emcee_skewed_gaussian(x_d, y_d):
     sigma_pos_mcmc      = np.percentile(sigma_pos_walk, [16, 50, 84])
     gamma_pos_mcmc      = np.percentile(gamma_pos_walk, [16, 50, 84])
 
+    # Takes the best parameters as the 50 percentile
+    amplitude_neg_best = amplitude_neg_walk[1]
+    center_neg_best    = center_neg_walk[1]
+    sigma_neg_best     = sigma_neg_walk[1]
+    gamma_neg_best     = gamma_neg_walk[1]
+    amplitude_pos_best = amplitude_pos_walk[1]
+    center_pos_best    = center_pos_walk[1]
+    sigma_pos_best     = sigma_pos_walk[1]
+    gamma_pos_best     = gamma_pos_walk[1]
+
     # Prints them
     print('Parameter = [16 50 84]')
     print('amplitude_neg = ', amplitude_neg_mcmc)
[…]

### 2016-10-19 paper/paper.tex
still need to add a lot more content to new section of the paper

@@
 Trying to solve this task requires the creation of a simplified and solid model.
-In this monograph I propose to consider LAEs as a spherical distribution of Hydrogen atoms that undergoes a solid body rotation and a radial expansion.
-I simulate the effect of rotational velocity, outflow velocity and optical depth of the LAE in the outgoing spectra.
+In this paper we propose to consider LAEs as a spherical distribution of Hydrogen atoms that undergoes a solid body rotation and a radial expansion due to outflows.
+We use radiative transfer Monte Carlo simulations to interpret the Lyman-$\alpha$ line morphology.
 The main conclusion is that this new model reproduces LAEs observed features in a clear way and with consistent physical parameters.
-However, proper observational fits are left for future work.
-This monograph accomplishes the objective of extracting as much information as possible from a LAE's \lya line.
+We also show results of adjusting observational data for some selected objects to models with and without bulk rotation.
+We finalize by discussing the possible implications for these results in terms of the energetics required for supernova feedback and outflows in high redshift galaxies.
+\\
 \end{abstract}
@@
 Astronomers noted that a that relevant fraction of these galaxies emitted a really strong line at a particular wavelength and named them Lyman Alpha Emitters (LAEs).
-The purpose of this monograph is to model and simulate LAEs in order to advance our knowledge about this galaxy population.
+The purpose of this paper is to model and simulate LAEs in order to advance our knowledge about this galaxy population.
 \\
 
-\section{The physics of Lyman Alpha Emitters}
-\label{sec1:laes}
 In the early Universe the most abundant elements were Hydrogen(H) and Helium(He), causing young galaxies to be rich in H atoms.
@@
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/radiative_transfer}
-	\end{center}
-	\caption{\textbf{Radiative Transfer Sketch:} A \lya photon being absorbed and re-emitted by Hydrogen atoms. 
-		\label{fig:radiative_transfer}}
-\end{figure}
-
 When a \lya photon is emitted inside the galaxy, it travels through its interstellar medium (ISM).
@@
 	\begin{center}
-		\includegraphics[width=0.3\textwidth]{./figures/static}
+		\includegraphics[width=0.45\textwidth]{./figures/radiative_transfer}
+	\end{center}
+	\caption{\textbf{Radiative Transfer Sketch:} A \lya photon being absorbed and re-emitted by Hydrogen atoms. 
+		\label{fig:radiative_transfer}}
+\end{figure}
+
+In a static galaxy, this random walk process produces a spectrum with two equal and symmetric peaks around the natural \lya wavelength.
+This can be seen in Figure A3 of Forero-Romero et al.
+paper \cite{CLARA} (Fig.
+\ref{fig:static}).
+If now the gas has a bulk velocity, the shape of the \lya profile changes.
+We explore these effects in this work by proposing a new model for a LAE.
+It consists of a spherical distribution of H atoms undergoing a solid body rotation and radial expansion(outflows).
+This new model can help observers to determine physical characteristics of a LAE by just looking at its \lya line.
+\\
+
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=0.35\textwidth]{./figures/static}
 	\end{center}
@@
 
-In a static galaxy, this random walk process produces a spectrum with two equal and symmetric peaks around the natural \lya wavelength.
-This can be seen in Figure A3 of Forero-Romero et al.
-paper \cite{CLARA} (Fig.
-\ref{fig:static}).
-If now the gas has a bulk velocity, the shape of the \lya profile changes.
-I explore these effects in this work by proposing a new model for a LAE.
-It consists of a spherical distribution of H atoms undergoing a solid body rotation and radial expansion(outflows).
-This new model can help observers to determine physical characteristics of a LAE by just looking at its \lya line.
-\\
-
 The main goal is to measure the effect of the model's physical parameters on the outgoing \lya line.
@@
 
-In this monograph I use a radiative transfer code that exists in the literature called CLARA (Code for Lyman Alpha Radiation Analysis).
+In this paper we use a radiative transfer code that exists in the literature called CLARA (Code for Lyman Alpha Radiation Analysis).
 CLARA was written by Forero-Romero et al.
@@
 It can simulate the \lya line of a spherical rotating LAE depending on its mass and velocity.
-In this work I modify CLARA to include outflows and explore the resulting consequences on the \lya line.
+In this work we modify CLARA to include outflows and explore the resulting consequences on the \lya line.
 \\
@@
 \section{A new LAE model}
+\label{sec:newmodel}
 
-\label{sec:newmodel}
 Galaxies are ellipsoidal.
-However, for simplicity I model them as a sphere.
+However, for simplicity we model them as a sphere.
 This shape facilitates the interpretation of the results in our numerical experiments.
@@
 
-There are 3 parameters in my model that define a LAE: the rotational velocity (\vrot), the outflow velocity (\vout) and the optical depth (\tauh).
+There are 3 parameters in this model that define a LAE: the rotational velocity (\vrot), the outflow velocity (\vout) and the optical depth (\tauh).
 \vout is due to material ejected from the galaxy, by supernovas(\cite{Verhamme06}, \cite{Orsi12}, \cite{Hashimoto2015}, \cite{Gronke2015}).
@@
 \subsection{A \lya Photon's Path in CLARA}
-In several problems in science it is useful to define dimensionless variables.
-This new formulation has many advantages.
-The numerical equations do not depend on the unit choice.
-This tends to make the results' analysis straight forward.
+\label{subsec:photons_path_clara}
+
 In the \lya radiative transfer problem, a commonly used dimensionless variable $x$ is used to describe a photon's frequency and is defined as:\\
@@
 
-However, in the plots shown in this monograph I use instead velocity, $V$, units. This eases the comparison against observational data. The velocity units are defined by
+However, in the plots shown in this paper we use instead velocity, $V$, units. This eases the comparison against observational data. The velocity units are defined by
 
@@
 From here, $100000$ photons are emitted with the natural \lya wavelength.
-Then, they start to behave as described in section \ref{sec1:laes} (See Fig.
+Then, they start to behave as described in sub-section \ref{subsec:photons_path_clara} (See Fig.
 \ref{fig:radiative_transfer}).
@@
 Once each photon escapes the galaxy, its final values are stored: position $\vec{r}$, direction of propagation $\hat{k}$, dimensionless frequency $x$, and number of scatterings $N$.
-To build the observed spectrum I make a histogram of the escape frequencies $x$.
+To build the observed spectrum we make a histogram of the escape frequencies $x$.
 The number of scatterings $N$ tells how many steps the random walk requires to reach a distance to the center that is $\geq R$.
@@
 Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires.
-Depending on \tauh, I need a different number of processors in order to minimize the time demands.
-For \tauh $=10^5, 10^6, 10^7$ I use 6, 12, 24 processors, respectively.
+Depending on \tauh, we need a different number of processors in order to minimize the time demands.
+For \tauh $=10^5, 10^6, 10^7$ we use 6, 12, 24 processors, respectively.
 The running times go from 7 hours to 1 week approximately, depending on the case.
@@
 That means, only photons escaping in the direction of the observer must be counted in the spectrum.
-In the simulation I approximate this by taking into account only the photons with escaping direction angle $\theta$ respect to the rotation axis within the range $[\theta_{min}-\theta_{max}]$.
-I illustrate this in Fig.
+In the simulation we approximate this by taking into account only the photons with escaping direction angle $\theta$ respect to the rotation axis within the range $[\theta_{min}-\theta_{max}]$.
+We illustrate this in Fig.
 \ref{fig:model}.
@@
 This implies that the resulting spectrum is independent from the azimuthal angle.
-Taking into account this symmetry, I only select photons on their polar angle, regardless of their azimuthal angle.
-Regarding the polar angle, I build the spectra for observers located on 3 different positions, with $\theta$ intervals uniformly distributed in $\cos(\theta)$.
+Taking into account this symmetry, we only select photons on their polar angle, regardless of their azimuthal angle.
+Regarding the polar angle, we build the spectra for observers located on 3 different positions, with $\theta$ intervals uniformly distributed in $\cos(\theta)$.
 \\
 
-To summarize, the parameters influencing the spectra are \vrot, \vout, \tauh and $\theta$. In the next chapter I evaluate the impact each of these has on the resulting profile. \\
+To summarize, the parameters influencing the spectra are \vrot, \vout, \tauh and $\theta$. In the next section we evaluate the impact each of these has on the resulting profile. \\
 
@@
 \ref{tab:values}.
-I run CLARA for all the permutations of these 3 parameters.
+We run CLARA for all the permutations of these 3 parameters.
 \\
@@
 \ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$ available in the Appendix \ref{ap:results_figures}.
-I do not include the spectra for \vout = 50, 75 \kms because they are similar to \vout = 25 \kms.
-For this reason, I limit in all the plots \vout to a maximum value of 25 \kms.
+We do not include the spectra for \vout = 50, 75 \kms because they are similar to \vout = 25 \kms.
+For this reason, we limit in all the plots \vout to a maximum value of 25 \kms.
 Nevertheless, the figures for \vout = 50, 75 \kms are in the Appendix \ref{ap:additional_figures} in Figs.
@@
 \subsection{Influence of the viewing angle $\theta$}
-I take now into account the viewing angle of the galaxy to build the observed spectra. For all of the physical parameters' combinations, the effect of $\theta$ in the \lya line is always the same.\\
+We take now into account the viewing angle of the galaxy to build the observed spectra. For all of the physical parameters' combinations, the effect of $\theta$ in the \lya line is always the same.\\
 
@@
 
-In the next subsection, I show only the results of the angle at which an observer sees the galaxy's angular momentum vector perpendicular to the line of sight.
+In the next subsection, we show only the results of the angle at which an observer sees the galaxy's angular momentum vector perpendicular to the line of sight.
 The purpose of this is only to decrease the number of plots in the document, as there is an analogous behavior for all viewing angles.
@@
 \subsection{Morphology of \lya line}
-To quantify the morphology of the \lya profile, I use three statistics: the standard deviation, the skewness and a factor sigma of asymmetry ($\sigma_A$). \\
-
-\subsubsection{Standard Deviation}
-The standard deviation measures the dispersion of frequencies in the \lya line. If it increases, the frequencies spread to wider ranges. \\
+To quantify the morphology of the \lya profile, we use an asymmetric gaussian fit to the curve as seen in Fig.
+\ref{fig:asymmetric_gaussian_fit}.
+Each of the two peaks, the red one and the blue one, is fitted with a curve parametrized by: an amplitude $A_{\pm}$, a standard deviation $\sigma_{\pm}$, a center $c_{\pm}$ and a skewness factor $\gamma_{\pm}$.\\
 
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/std}
+		\includegraphics[width=0.3\textwidth]{./figures/asymmetric_gaussian_fit}
 	\end{center}
-	\caption{\textbf{Standard deviation plot for each \tauh.} The viewing angle $\theta \simeq 90^\circ$.
-		\label{fig:std}}
+	\caption{\textbf{Asymmetric gaussian fit for a \lya profile:} With \tauh$=10^5$, \vrot$=50$ \kms, \vout$=20$ \kms and $\theta \approx 0^\circ$.The intensity is in arbitrary units.
+		\label{fig:asymmetric_gaussian_fit}}
 \end{figure}
 
-Fig.
-\ref{fig:std} shows that the standard deviation is inversely proportional to the outflow velocity.
-This implies that the greater \vout is, the less disperse is the \lya frequency distribution.
-This signals that the peaks start merging to one another.
-Also, the higher the \tauh, the more inclined the curves.
-The horizontal dotted line in the plot is the value of standard deviation for an observational LAE, in particular Q2206BX151 (at redshift $z\approx2.1974$).
-It is shown in Fig.
-\ref{fig:kulas} in the third column with second row.
-\\
+TERMINAR...\\
 
-\subsubsection{Skewness}
-The skewness measures the asymmetry of the \lya line.
-If the value is greater than zero, there is more weight in the left tail of the line, if it is less than zero, there is more weight in the right one.\\
+\subsection{Influences of the free parameters}
+
+TERMINAR...\\
 
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/skw}
+		\includegraphics[width=0.3\textwidth]{./figures/rotation_doppler_outflow}
 	\end{center}
-	\caption{\textbf{Skewness plot for each \tauh.} The viewing angle $\theta \simeq 90^\circ$. 
-		\label{fig:skw}}
+	\caption{\textbf{Rotations induces a Doppler shift of the only-outflow spectra:} Each of the black lines is then weighted to obtain the red line.
+		\label{fig:rotation_doppler_outflow}}
 \end{figure}
-
-Fig.
-\ref{fig:skw} show that the skewness is proportional to the outflow velocity, at a fixed rotational velocity.
-The greater \vout is, the more asymmetric is the \lya frequency distribution.
-The horizontal dotted line is the value of skewness for Q2206BX151.
-As seen, the line intersects with the skewness curves.
-This would imply that Q2206BX151, according to the model, has its parameters narrowed to: \vrot $\in [50,100]$ \kms, \vout $\in [15,20]$ \kms and \tauh $\in [10^6, 10^7]$.
-However, these values are not completely consistent with the standard deviation.\\
-
-\subsubsection{Sigma of Asymmetry}
-Sigma of asymmetry $\sigma_A$ is another measure for the asymmetry of the \lya line.
-It was defined by McLinden et al.
-\cite{McLinden2011} as follows.
-Each of the two peaks, the red one and the blue one, are fitted with a Gaussian curve.
-Their standard deviations $\sigma_{red}$ and $\sigma_{blue}$ are obtained.
-Then the factor $\sigma_A$ is:
-
-\begin{equation}
-\sigma_A = \frac{\sigma_{red}}{\sigma_{blue}}
-\end{equation}
-
-If the value is less than one, the left peak is thiner than the right peak. If it is greater than one, there left peak is wider than the right peak.\\
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/sigma}
-	\end{center}
-	\caption{\textbf{Sigma of asymmetry $\sigma_A$.} 
-		\label{fig:sigma}}
-\end{figure}
-
-Fig.
-\ref{fig:sigma} shows that this parameters has high fluctuations.
-I can not see a clear behavior or tendency that applies to all the curves.
-The only thing is that the values of $\sigma_A$ start close together and \vout makes them separate.
-Also, $\sigma_A$ is always greater than $1$, which means that there is always an asymmetry present in these distributions.\\
-
-\subsection{Influences of the free parameters}
 
@@
 
-\section{Comparison against observations}
+\section{Observational Comparison}
 \label{sec:comparisonobservations}
@@
 In the Figure 3 of their paper (Fig.
-\ref{fig:kulas}) they show the 18 \lya profiles.
+\ref{fig:kulas}) they show 18 \lya profiles.
 \\
@@
 The observed spectra in Fig.
-\ref{fig:kulas} are modified by the redshift $z$ of the galaxies.
-This is, how much distance light had to travel to reach their telescopes.
-That causes a broadening of the line.
-\\
-
-Taking into account this effect, I notice similitudes between them and the simulated spectra.
+\ref{fig:kulas} are modified by the redshift $z$ of the galaxies, which causes a broadening of the line.
+Taking into account this effect, we notice similitudes between them and the simulated spectra.
 The \lya profiles are mostly 2 peaks with an asymmetry between them, just as in the simulation.
@@
 
-Many authors that only include the outflows effect are able to fit observational spectra but with a very large magnitude of \vout, on the order of $50-100$ \kms.
-However, \vout is caused mainly by ejection of material out the galaxy.
-It is then common that a galaxy rotates faster than it expands, not the opposite.
-Other authors that include the rotational effect are able to obtain the two peaks but not their asymmetry.
-Simulating LAEs without outflows does not replicate the majority of the cases.
-\\
+In this section we will focus in one particular observation from Fig.
+\ref{fig:kulas}: \textit{Westphal-BX154}, seen in more detail in Fig.
+\ref{fig:kulas_w}.
+We quantify the morphology of this observation by applying the asymmetric gaussian fit described previously.
+We take the obtaining values and select the simulations that fall in it within a reasonable range.
+This in order to filter all the spectra combinations.\\
 
-In my model, the combination of rotation and outflows velocities not only makes a lot of sense but seems to be able to fit observations.
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=0.3\textwidth]{./figures/kulas_w}
+	\end{center}
+	\caption{\textbf{Westphal-BX154:} The Kinematics of Multiple-peaked \lya Emission in Star-forming Galaxies at $z\sim2-3$ \cite{Kulas12}.
+This spectra is at $z=2.5954$.This plot is reproduced with the authors permission.
+		\label{fig:kulas_w}}
+\end{figure}
+
+TERMINAR...\\
+
+In this model, the combination of rotation and outflows velocities not only makes a lot of sense but seems to be able to fit observations.
 This new model proposed could replicate \lya profiles with typical LAE's values.
-The main result of this monograph is that I presented a new LAE model roughly consistent with observations.\\
+The main result of this paper is that we presented a new LAE model roughly consistent with observations.\\
 
@@
 
-In this monograph, the objective was to analyze and measure the influence of galaxy rotation and outflows on the \lya line.
+In this paper, the objective was to analyze and measure the influence of galaxy rotation and outflows on the \lya line.
 The motivation for this is to be able to obtain physical information of a LAE by just looking at its \lya profile.
-In order to accomplish this objective, I propose a new model of a LAE consisting of a sphere of Hydrogen atoms that expands radially and rotates as a solid body.
+In order to accomplish this objective, we propose a new model of a LAE consisting of a sphere of Hydrogen atoms that expands radially and rotates as a solid body.
 The program CLARA \cite{CLARA} is used to set the conditions and emulate the transfer code.
@@
 
-	\item The \vrot effect is much larger than the \vout one.
-Only a little fraction of \vrot, in the radial direction, is necessary to obtain 2 asymmetric peaks.
-If it \vout approaches the value of \vrot, both peaks start merging until only the outflows effect is visible.
-\\
-
 	\item The final spectra obtained are roughly consistent with LAEs observations.  \\
@@
 
+% Rotational effects should be clearly detected and characterized by MUSE.
+%Possible rotational velocity constraints using the Lyman-alpha line.
+
 Due to the long time CLARA takes to run, it was not possible to fit an observational LAE and predict its parameters.
-However, the next step is to use tools as MCMC (Monte Carlo Markov Chain) to obtain the galaxy's \tauh, \vrot and \vout that would agree with my model.
+However, the next step is to use tools as MCMC (Monte Carlo Markov Chain) to obtain the galaxy's \tauh, \vrot and \vout that would agree with this model.
 \\
 
+TERMINAR...\\
+
 All of this work is available online and free to use to anyone.
-The data, source code and instructions to replicate my monograph's results are in the GitHub repository:  \texttt{https://github.com/astroandes/CLARA\_RotationOut flows}.
+The data, source code and instructions to replicate this paper's results are in the GitHub repository:  \texttt{github.com/astroandes/CLARA\_RotationOutflows}.
 \\
@@
 %------------------------APPENDIX----------------------------
+
+\newpage
 

### 2017-02-09 paper/paper.tex
quick intro review

@@
 Galaxies are key to understand our Universe.
-So far, human kind has been able to study in detail the closest ones.
-Distant galaxies are still challenging to detect.
-They contain however information of millions of stars and gas that could unravel mysteries like for example, how Milky Way type of galaxies formed.
-It is an open challenge to obtain as much information as possible from these young galaxies.
-One way to accomplish this is to look at their spectra.
-Astronomers noted that a that relevant fraction of these galaxies emitted a really strong line at a particular wavelength and named them Lyman Alpha Emitters (LAEs).
-The purpose of this paper is to model and simulate LAEs in order to advance our knowledge about this galaxy population.
+However, distant ones are very challenging to detect.
+It is then an open challenge to obtain as much information as possible from them, as they represent the younger stages of galaxy evolution.
+Astronomers noted that a that relevant fraction of these galaxies emitted a really strong line at $1215.67\;\AA$ in their spectra at  and named them Lyman Alpha Emitters (LAEs).
+The purpose of this paper is to use this spectral line to model and simulate LAEs in order to narrow computationally their physical and kinematic properties.
 \\
@@
 In the early Universe the most abundant elements were Hydrogen(H) and Helium(He), causing young galaxies to be rich in H atoms.
-Also, due to stellar activity in a galaxy, the surrounding gas can be ionized.
-These two facts leave the H atoms prone to excite and emit radiation at definite frequencies.
-The most basic of these frequencies corresponds to the \lya line.
-\cite{PartridgePeebles}  \\
-
-When a Hydrogen atom is excited, the only electron it has moves to a higher energy level.
-However after some time, it is no longer able to maintain itself in this state.
-For this, the electron goes back to the ground level.
-Finally, as the energy obtained from this de-excitement has to be conserved, it is emitted as a photon with a wavelength of 1215.67 $\AA$ if levels are $2\rightarrow1$.
-This is the Lyman Alpha ( \lya) transition, discovered by Theodore Lyman in 1906.
+Due to stellar activity in a galaxy, the surrounding gas is ionized, so H atoms excite and emit radiation at definite frequencies.
+The most basic of these frequencies corresponds to the \lya line equal to $2.47 \times 10^{15}\;Hz$ \cite{PartridgePeebles}.
+This is equivalent to a wavelength of 1215.67 $\AA$ cause by the transition of the H atom electron from levels $2\rightarrow1$ dubbed Lyman Alpha ( \lya) transition, discovered by Theodore Lyman in 1906.
 \cite{LymanBio} \\
 
-\lya radiation is emitted from the H atoms in a galaxy.
-Due to the amount of atoms inside it, the whole body could become a strong \lya radiator.
-Astronomers can take spectra of a galaxy to detect emission at that wavelength.
-If the galaxy shows a strong \lya line, it classifies as a Lyman Alpha Emitter (LAE).
-(\cite{DjorgovskiThompson}, \cite{Rhoads00}, \cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08}, \cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12}, \cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013}, \cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014}, \cite{Fumagalli2015}) \\
-
-When a \lya photon is emitted inside the galaxy, it travels through its interstellar medium (ISM).
-During the photon's path it can be absorbed and re-emitted by an H atom.
+Due to the amount of H atoms inside a galaxy, the whole body becomes a strong \lya radiator.
+If the galaxy spectra shows a strong \lya line, it classifies as a Lyman Alpha Emitter (LAE).
+(\cite{DjorgovskiThompson}, \cite{Rhoads00}, \cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08}, \cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12}, \cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013}, \cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014}, \cite{Fumagalli2015}).
+However, when a \lya photon is emitted inside the galaxy, it travels through its interstellar medium (ISM).
+During the photon's path it can be absorbed and re-emitted by other H atoms.
 The new frequency of the photon is different that the initial one, in an observer frame of reference due to the atom's velocity.
 In LAEs, the state of the ISM gas, before and after a photon's re-emission is pretty much the same.
-This allows a random walk approximation and consider photon-atom encounters as scatterings.
+This allows a random walk approximation and lets us consider photon-atom encounters as scatterings.
 A photon's scattering can happen several times, as seen in Fig.
@@
 If now the gas has a bulk velocity, the shape of the \lya profile changes.
-We explore these effects in this work by proposing a new model for a LAE.
-It consists of a spherical distribution of H atoms undergoing a solid body rotation and radial expansion(outflows).
-This new model can help observers to determine physical characteristics of a LAE by just looking at its \lya line.
+We explore these effects in this paper by proposing a new model for a LAE.
+The model consists of a spherical distribution of H atoms undergoing a solid body rotation and radial expansion (outflows).
 \\
@@
 
-The main goal is to measure the effect of the model's physical parameters on the outgoing \lya line.
+Our main goal with this work is to measure the effect of the model's physical parameters on the outgoing \lya line.
 In this case, due to the resonant\footnote{Resonant is a common term used in radiative transfer.
 It means that the photons that create the line are absorbed and re-emitted several times before escaping the cloud of gas.} nature of the \lya line, analytical solutions can not be derived.
-It becomes necessary to run simulations to explore and test the model.
-\\
-
-In this paper we use a radiative transfer code that exists in the literature called CLARA (Code for Lyman Alpha Radiation Analysis).
-CLARA was written by Forero-Romero et al.
+Because of this, it becomes necessary to run simulations that explore and test the model.
+In this paper we use the radiative transfer code \textbf{CLARA} (Code for Lyman Alpha Radiation Analysis) written by Forero-Romero et al.
 \cite{CLARA}.
-It can simulate the \lya line of a spherical rotating LAE depending on its mass and velocity.
-In this work we modify CLARA to include outflows and explore the resulting consequences on the \lya line.
+CLARA can simulate the \lya line of a spherical rotating LAE depending on its mass and velocity, so we modify it to include outflows and then we explore the resulting consequences on the \lya line.
 \\

### 2017-02-17 paper/paper.tex
updates

@@
 
-Galaxies are ellipsoidal.
-However, for simplicity we model them as a sphere.
-This shape facilitates the interpretation of the results in our numerical experiments.
+We model spherical galaxy to facilitate the interpretation of the results, by deleting what would be another degree of freedom regarding a different geometry.
 Furthermore, this approximation is commonly used in the literature, as it explains a wide variety of observational features (\cite{Ahn03}, \cite{Verhamme06}, \cite{Dijkstra06}).
@@
 
-In this model, the LAE has a bulk velocity which is the superposition of the rotation and the outflow, as shown in Fig.
+In this model, the LAE has a bulk velocity corresponding to the superposition of rotation and outflows, as shown in Fig.
 \ref{fig:model}.
@@
 If $x < 0$, the final frequency $\nu < \nu_{\rm \alpha}$.
-This translates to the final wavelength being larger than the \lya natural one.
-The phenomenon is called a redshift in frequency.
+This translates to the final wavelength being larger than the \lya natural one, casing a redshift in frequency.
 If, on the contrary, $x > 0$, the photon suffers a blueshift in frequency.
@@
 The initial emission of photons is taken at the center of the sphere for practicality due to the fact that both, center and off-center emissions, give analogous results.
-From here, $100000$ photons are emitted with the natural \lya wavelength.
-Then, they start to behave as described in sub-section \ref{subsec:photons_path_clara} (See Fig.
+From here, $100000$ photons are emitted with the natural \lya wavelength starting to behave as described in sub-section \ref{subsec:photons_path_clara} (See Fig.
 \ref{fig:radiative_transfer}).
@@
 
-The process that takes place from the moment the first photon is emitted to the moment the last photon escapes the sphere, is called a simulation.
+%The process that takes place from the moment the first photon is emitted to the moment the last photon escapes the sphere, is called a simulation.
 Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires.
@@
 \subsection{Parameters' Initial Values}
-
 In order to define the ranges of \tauh, \vrot and \vout, it is necessary to refer to observational constraints.
@@
 \ref{tab:values}.
-We run CLARA for all the permutations of these 3 parameters.
+We run CLARA's modified version for all the permutations of these 3 parameters.
 \\
@@
 
-The resulting sets of spectra are in Figs.
-\ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$ available in the Appendix \ref{ap:results_figures}.
+The resulting sets of spectra are in the Appendix \ref{ap:results_figures} in Figs.
+\ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$.
 We do not include the spectra for \vout = 50, 75 \kms because they are similar to \vout = 25 \kms.
-For this reason, we limit in all the plots \vout to a maximum value of 25 \kms.
+For this reason, we limit \vout to a maximum value of 25 \kms in all the plots.
 Nevertheless, the figures for \vout = 50, 75 \kms are in the Appendix \ref{ap:additional_figures} in Figs.
@@
 In the next subsection, we show only the results of the angle at which an observer sees the galaxy's angular momentum vector perpendicular to the line of sight.
-The purpose of this is only to decrease the number of plots in the document, as there is an analogous behavior for all viewing angles.
+The purpose of this is only to decrease the number of plots as there is an analogous behavior for all viewing angles.
 \\

### 2017-02-24 paper/paper.tex
paper's updated version

@@
 \\
+
+TERMINAR...\\
 \end{abstract}
@@
 This allows a random walk approximation and lets us consider photon-atom encounters as scatterings.
-A photon's scattering can happen several times, as seen in Fig.
-\ref{fig:radiative_transfer}.
-This stops only until the photon is able to escape the galaxy.
+A photon's scattering can happen several times and stops only until the photon is able to escape the galaxy.
 \\
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.45\textwidth]{./figures/radiative_transfer}
-	\end{center}
-	\caption{\textbf{Radiative Transfer Sketch:} A \lya photon being absorbed and re-emitted by Hydrogen atoms. 
-		\label{fig:radiative_transfer}}
-\end{figure}
-
-In a static galaxy, this random walk process produces a spectrum with two equal and symmetric peaks around the natural \lya wavelength.
-This can be seen in Figure A3 of Forero-Romero et al.
-paper \cite{CLARA} (Fig.
-\ref{fig:static}).
+In a static galaxy, this random walk process produces a spectrum with two equal and symmetric peaks around the natural \lya wavelength and zero intensity at the center (Fig.
+A3 from \cite{CLARA}).
 If now the gas has a bulk velocity, the shape of the \lya profile changes.
@@
 \\
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.35\textwidth]{./figures/static}
-	\end{center}
-	\caption{\textbf{Figure A3 of Forero-Romero et al.
-paper:} CLARA's view on the escape fraction of Lyman-$\alpha$ photons in high redshift galaxies \cite{CLARA}.
-Reproduced with the permission of the author.
-		\label{fig:static}}
-\end{figure}
 
@@
 
-\subsection{A \lya Photon's Path in CLARA}
-\label{subsec:photons_path_clara}
-
-In the \lya radiative transfer problem, a commonly used dimensionless variable $x$ is used to describe a photon's frequency and is defined as:\\
-
-\begin{equation}
-	\label{eq:x_freq}
-	x \equiv \frac{\nu -\nu_{\rm \alpha}}{\Delta\nu_{\rm D}},
-\end{equation}
-
-where $\nu$ is the photon's frequency and $\nu_{\rm \alpha} = 2.46\times 10^{15}$ Hz is the \lya natural frequency. The denominator $\Delta\nu_{\rm D}$ is defined in Eq. \ref{eq:delta_nuD}.
-
-\begin{equation}
-	\label{eq:delta_nuD}
-	\Delta\nu_{\rm D} \equiv \nu_{\rm \alpha}\sqrt{\frac{2kT}{m_pc^2}} \equiv \nu_{\rm \alpha} \frac{v_{\rm th}}{c},
-\end{equation}
-
-where $\Delta\nu_{\rm D}$ is known as the Doppler broadening of the \lya line.
-It depends on the neutral gas temperature $T$ or equivalently the thermal velocity $v_{\rm th}$ of the atoms.
-In the model the temperature is kept constant at $T=10^4$K and the thermal velocity is $v_{\rm th}=12.8$\kms.
-\\
-
-If $x < 0$, the final frequency $\nu < \nu_{\rm \alpha}$.
-This translates to the final wavelength being larger than the \lya natural one, casing a redshift in frequency.
-If, on the contrary, $x > 0$, the photon suffers a blueshift in frequency.
-\\
-
+%mirar donde pongo esto
 However, in the plots shown in this paper we use instead velocity, $V$, units. This eases the comparison against observational data. The velocity units are defined by
-
-\begin{equation}
-	\label{eq:V}
-	V = xv_{\rm th} = \frac{\nu -\nu_{\rm \alpha}}{\nu_{\rm \alpha}}c.
-\end{equation}
-
 The units of $V$ are usually \kms. In this case, the photon is redshifted in frequency when the velocity is greater than 0, and blueshifted in the opposite case. \\
@@
 The initial emission of photons is taken at the center of the sphere for practicality due to the fact that both, center and off-center emissions, give analogous results.
-From here, $100000$ photons are emitted with the natural \lya wavelength starting to behave as described in sub-section \ref{subsec:photons_path_clara} (See Fig.
-\ref{fig:radiative_transfer}).
+From here, $100000$ photons are emitted with the natural \lya wavelength and start to scatter.
 When each photon is re-emitted, its new wavelength depends on the H atom's velocity (both thermal and bulk) and direction (both initial and final).
@@
 
-The resulting sets of spectra are in the Appendix \ref{ap:results_figures} in Figs.
-\ref{fig:3_tau10E5_phi83-90}, \ref{fig:3_tau10E7_phi83-90} and \ref{fig:3_tau10E6_phi83-90} for $\theta \simeq 90^\circ$.
-We do not include the spectra for \vout = 50, 75 \kms because they are similar to \vout = 25 \kms.
-For this reason, we limit \vout to a maximum value of 25 \kms in all the plots.
-Nevertheless, the figures for \vout = 50, 75 \kms are in the Appendix \ref{ap:additional_figures} in Figs.
-\ref{fig:2_tau10E5_phi83-90}, \ref{fig:2_tau10E6_phi83-90} and \ref{fig:2_tau10E7_phi83-90}.\\
+The behavior of the resulting sets of spectra can be summarized by Fig.
+\ref{fig:summary}.
+We concluded from the simulations a clear creation of two asymmetric peaks around $V=0$ \kms with the tallest peak always redshifted.
+We detected a strong dependence on the outflow velocity that induces the peak asymmetry.
+We also detected that the rotation velocity broadens the line horizontally.
+Both these effects have been previously reported separately in the literature, so these results can valid their superposition to some extend.
+\\
 
-These last 3 plots show a clear creation of two asymmetric peaks around $V=0$ \kms with the tallest peak is always redshifted, with a strong dependence on the outflow velocity.\\
+\begin{figure}[h!]
+	\begin{center}
+		\includegraphics[width=0.5\textwidth]{./figures/summary}
+	\end{center}
+	\caption{\textbf{4 different \lya profiles:} With \tauh$=10^5$ and $\theta \simeq 90^\circ$.
+The rotational velocity \vrot increases to the right and the outflows velocity \vout increases downwards.
+The intensity is in arbitrary units.
+		\label{fig:summary}}
+\end{figure}
 
@@
 
-In the next subsection, we show only the results of the angle at which an observer sees the galaxy's angular momentum vector perpendicular to the line of sight.
-The purpose of this is only to decrease the number of plots as there is an analogous behavior for all viewing angles.
-\\
+\subsection{Morphology of \lya line}
 
-\subsection{Morphology of \lya line}
-To quantify the morphology of the \lya profile, we use an asymmetric gaussian fit to the curve as seen in Fig.
-\ref{fig:asymmetric_gaussian_fit}.
-Each of the two peaks, the red one and the blue one, is fitted with a curve parametrized by: an amplitude $A_{\pm}$, a standard deviation $\sigma_{\pm}$, a center $c_{\pm}$ and a skewness factor $\gamma_{\pm}$.\\
+To summarize, the influence of the 4 parameters on the \lya morphology is the following:
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.3\textwidth]{./figures/asymmetric_gaussian_fit}
-	\end{center}
-	\caption{\textbf{Asymmetric gaussian fit for a \lya profile:} With \tauh$=10^5$, \vrot$=50$ \kms, \vout$=20$ \kms and $\theta \approx 0^\circ$.The intensity is in arbitrary units.
-		\label{fig:asymmetric_gaussian_fit}}
-\end{figure}
+\begin{itemize}
+	\item \tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
+	\item \vout decreases the right peak's intensity. Higher \vout makes the left peak smaller until it merges with the right one. \\
+	\item \vrot broadens the line and decreases the maximum intensity.
+Higher \vrot implies a flatter spectrum.
+This effect has not been deeply studied in literature.
+Only Garavito et al.
+\cite{Garavito14} has simulated its effect.
+Our results are consistent with their conclusions.
+	\item $\theta$ increases the central valley of the line. Higher $\theta$ makes more emission at \lya natural frequency. \\ 
+\end{itemize}
 
-TERMINAR...\\
+\subsection{Doppler shift in frequency}
 
-\subsection{Influences of the free parameters}
-
-TERMINAR...\\
+Rotations induces a Doppler shift of the only-outflow spectra.\\
 
@@
 
-To summarize the influence of the 3 parameters on the \lya morphology is the following:
+TERMINAR...\\
+PONER GRAFICA REAL...\\
 
-\begin{itemize}
-	\item \tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
-	\item \vout decreases the right peak's intensity. Higher \vout make the left peak smaller until it merges with the right one. \\
-	\item \vrot broadens the line and decreases the maximum intensity.
-Higher \vrot implies a flatter spectrum.
-This effect has not been deeply studied in literature.
-Only Garavito et al.
-\cite{Garavito14} has simulated its effect.
-Our results are consistent with their conclusions.
-\end{itemize}
+\section{Observational Implications}
+\label{sec:observationalimplications}
 
-\section{Observational Comparison}
-\label{sec:comparisonobservations}
+** Rotational effects should be clearly detected and characterized by MUSE.\\
 
-There are several observations of LAEs in the literature.
-Kulas et al.
-\cite{Kulas12} observed distant LAEs ($z\sim2-3$) with high resolution.
-In the Figure 3 of their paper (Fig.
-\ref{fig:kulas}) they show 18 \lya profiles.
-\\
+** The V=0 emission is not necessary because of radiation escaping without scattering, but that the viewing angle can control that. \\
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/figure3_kulas}
-	\end{center}
-	\caption{\textbf{Figure 3 of Kulas et al.
-paper:} The Kinematics of Multiple-peaked \lya Emission in Star-forming Galaxies at $z\sim2-3$ \cite{Kulas12}.
-This plot is reproduced with the authors permission.
-		\label{fig:kulas}}
-\end{figure}
-%http://arohatgi.info/WebPlotDigitizer/app/
-
-The observed spectra in Fig.
-\ref{fig:kulas} are modified by the redshift $z$ of the galaxies, which causes a broadening of the line.
-Taking into account this effect, we notice similitudes between them and the simulated spectra.
-The \lya profiles are mostly 2 peaks with an asymmetry between them, just as in the simulation.
-Also, the small peak (if existing) is always smaller than the tall one.\\
-
-In this section we will focus in one particular observation from Fig.
-\ref{fig:kulas}: \textit{Westphal-BX154}, seen in more detail in Fig.
-\ref{fig:kulas_w}.
-We quantify the morphology of this observation by applying the asymmetric gaussian fit described previously.
-We take the obtaining values and select the simulations that fall in it within a reasonable range.
-This in order to filter all the spectra combinations.\\
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.3\textwidth]{./figures/kulas_w}
-	\end{center}
-	\caption{\textbf{Westphal-BX154:} The Kinematics of Multiple-peaked \lya Emission in Star-forming Galaxies at $z\sim2-3$ \cite{Kulas12}.
-This spectra is at $z=2.5954$.This plot is reproduced with the authors permission.
-		\label{fig:kulas_w}}
-\end{figure}
+** This new LAE model is roughly consistent with observations.\\
 
 TERMINAR...\\
-
-In this model, the combination of rotation and outflows velocities not only makes a lot of sense but seems to be able to fit observations.
-This new model proposed could replicate \lya profiles with typical LAE's values.
-The main result of this paper is that we presented a new LAE model roughly consistent with observations.\\
 
@@
 In order to accomplish this objective, we propose a new model of a LAE consisting of a sphere of Hydrogen atoms that expands radially and rotates as a solid body.
-The program CLARA \cite{CLARA} is used to set the conditions and emulate the transfer code.
+A modified version of the program CLARA \cite{CLARA} is used to set the conditions and emulate the radiative transfer process inside the galaxy.
 \\
@@
 
+	\item Rotations induces a Doppler shift of the only-outflow spectra.\\
+
+	\item TERMINAR...\\
 \end{itemize}
@@
 
-% Rotational effects should be clearly detected and characterized by MUSE.
-%Possible rotational velocity constraints using the Lyman-alpha line.
-
-Due to the long time CLARA takes to run, it was not possible to fit an observational LAE and predict its parameters.
-However, the next step is to use tools as MCMC (Monte Carlo Markov Chain) to obtain the galaxy's \tauh, \vrot and \vout that would agree with this model.
-\\
+** Differential rotation instead of solid body. \\
+** Fit and observation's parameters with MCMC. \\
 
@@
 
-...\\
+** Cluster Uniandes?
+
+TERMINAR...\\
 
@@
 
-%------------------------APPENDIX----------------------------
-
-\newpage
-
-\appendix
-
-\begin{figure}[!htbp]
-	\centering
-	\begin{minipage}[b]{0.45\textwidth}
-		\section{Results' Figures}
-		\label{ap:results_figures}
-
-		\includegraphics[width=\textwidth]{./figures/3_tau10E5_phi83-90}
-		\caption{\textbf{\lya profile for \tauh$=10^5$:} With \vrot ranging $50,100$ \kms and \vout ranging $5,10,15,20,25$ \kms.
-The intensity is in arbitrary units.
-The viewing angle $\theta \simeq 90^\circ$.
-The dashed vertical line indicates the \lya line's natural frequency.
-			\label{fig:3_tau10E5_phi83-90}}
-	\end{minipage}
-	\hfill
-	\begin{minipage}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{./figures/3_tau10E6_phi83-90}
-		\caption{\textbf{\lya profile for \tauh$=10^6$:}  Same as Fig. \ref{fig:3_tau10E5_phi83-90} but for \tauh$=10^6$. 
-			\label{fig:3_tau10E6_phi83-90}}
-		\vspace{1.5mm}
-	\end{minipage}
-\end{figure}
-
-\begin{figure}[!htbp]
-	\centering
-	\begin{minipage}[b]{0.45\textwidth}
-		\vspace{3mm}
-		\includegraphics[width=\textwidth]{./figures/3_tau10E7_phi83-90}
-		\caption{\textbf{\lya profile for \tauh$=10^7$:} Same as Fig. \ref{fig:3_tau10E5_phi83-90} but for \tauh$=10^7$.
-			\label{fig:3_tau10E7_phi83-90}}
-	\end{minipage}
-	\hfill
-	\begin{minipage}[b]{0.45\textwidth}
-		\section{Additional Figures}
-		\label{ap:additional_figures}
-
-		Additional figures of spectra with \vrot $= 50,100$ \kms and \vout $= 50,75$ \kms, for all the different \tauh.
-
-		\vspace{2.7cm}
-
-		\includegraphics[width=\textwidth]{./figures/appendix/2_tau10E5_phi83-90}
-		\caption{\textbf{\lya profile for \tauh$=10^5$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
-			\label{fig:2_tau10E5_phi83-90}}
-	\end{minipage}
-\end{figure}
-
-\begin{figure}[!htbp]
-	\centering
-	\begin{minipage}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{./figures/appendix/2_tau10E6_phi83-90}
-		\caption{\textbf{\lya profile for \tauh$=10^6$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
-			\label{fig:2_tau10E6_phi83-90}}
-	\end{minipage}
-	\hfill
-	\begin{minipage}[b]{0.45\textwidth}
-		\includegraphics[width=\textwidth]{./figures/appendix/2_tau10E7_phi83-90}
-		\caption{\textbf{\lya profile for \tauh$=10^7$:} With \vrot ranging $50,100$ \kms and \vout ranging $50,75$ \kms.
-			\label{fig:2_tau10E7_phi83-90}}
-	\end{minipage}
-\end{figure}
-
 \end{document}

### 2017-03-02 paper/paper.tex
only changes in results left

@@
 
-** Rotational effects should be clearly detected and characterized by MUSE.\\
+Among the observational implications that this paper has over observations we select three principal ones.
+First, the rotational effects of the galaxy should be clearly detected and characterized by MUSE's high resolution data.
+The Doppler redshift is already visible in other observations, such as the one seen in Fig.
+7 of Prescott et al.
+\cite{Prescott14}.\\
 
-** The V=0 emission is not necessary because of radiation escaping without scattering, but that the viewing angle can control that. \\
+Second, the central ($V=0$) emission of the spectra seen in the results is a consequence of the viewing angle of the galaxy and can be controlled by it.
+This means that the reason is not necessary that radiation is escaping without scattering, as several authors have suggested (CITAS...).\\
 
-** This new LAE model is roughly consistent with observations.\\
-
-TERMINAR...\\
+Finally, all the spectra produced from this new LAE model is roughly consistent with observations.
+The consideration of adding the new rotation parameter to the standard only-outflow model found in literature is reasonable and very powerful.
+\vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\
 
@@
 
+	\item Rotations induces a Doppler shift in frequency of the only-outflow spectra.\\
+
 	\item The final spectra obtained are roughly consistent with LAEs observations.  \\
 
-	\item Rotations induces a Doppler shift of the only-outflow spectra.\\
-
-	\item TERMINAR...\\
 \end{itemize}
@@
 
-** Differential rotation instead of solid body. \\
-** Fit and observation's parameters with MCMC. \\
+For future work regarding the model, we would like to implement a differential rotation for the LAE instead of solid body.
+Regarding result analysis, we would like to compare against observations in two ways.
+Firstly by using the kinematic observation obtained from newly observed LAEs.
+We could do a consistency check between their real spectra and the simulated one that our model would produce with the respective parameters.
+Secondly we would like to take a specific observed spectrum and fit it using MCMC in order to predict ranges for its parameters.
+\\
 
-TERMINAR...\\
+\subsection{Reproducibility}
 
@@
 \section*{Acknowledgments}
+The authors would like to acknowledge the high performance computational cluster (HPC) hosted by Universidad de los Andes, where we ran the simulations for CLARA. \\
 
-** Cluster Uniandes?
-
-TERMINAR...\\
+We would also like to acknowledge the participants of the meeting \textit{Escape of Lyman radiation from galactic labyrinths} at Crete, Greece, for their valuable comments on this work, specially to Max Gronke and Christian Herenz.
+\\
 

### 2017-03-02 paper/paper.tex
now i need to update plots

@@
 
-\subsection{Doppler shift in frequency}
+\subsection{Doppler shift by rotation}
 
-Rotations induces a Doppler shift of the only-outflow spectra.\\
+The rotation effect induces a Doppler shift in frequency and displaces the only-outflow spectra.
+As seen in Fig.
+\ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with \vrot$=0$ and displaces it from its central location.
+Then it weights the lines and they all merge into the resulting red \lya line.\\
 
@@
 
-TERMINAR...\\
 PONER GRAFICA REAL...\\
@@
 
-The conclusions obtained from this work are: \\
+The conclusions obtained from this work are:
 
@@
 The closer it is to the equator of the galaxy, the higher the central valley of the frequency distribution.
-\\
 
@@
 \vrot broadens the \lya line.
-\vout increases the peaks asymmetry and \tauh induces a redshift around the zero velocity.\\
+\vout increases the peaks asymmetry and \tauh induces a redshift around the zero velocity.
 
-	\item Rotations induces a Doppler shift in frequency of the only-outflow spectra.\\
+	\item Rotations induces a Doppler shift in frequency of the only-outflow spectra.
 

### 2017-03-02 data/tau10E5/to_join_parts_p.py
new data in git

@@ -1,0 +1,25 @@
+import os
+
+logtau = 5
+vrots = [50]
+#vrots = [50,100]
+vouts = [20]
+#vouts = [5,25]
+
+for vrot in vrots:
+    for vout in vouts:
+
+        path = './vrot'+str(vrot)+'/vout'+str(vout)+'/'
+        filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+
+        #os.system('rm '+path+filename+'_out.ascii')
+        os.system('cat '+path+filename+'_out* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_out.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_out.ascii')
+        os.system('rm temp.txt')
+
+        #os.system('rm '+path+filename+'_in.ascii')
+        os.system('cat '+path+filename+'_in* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_in.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_in.ascii')
+        os.system('rm temp.txt')

### 2017-03-02 data/tau10E6/to_join_parts_p.py

@@ -1,0 +1,23 @@
+import os
+
+logtau = 6
+vrots = [100]
+vouts = [5]
+
+for vrot in vrots:
+    for vout in vouts:
+
+        path = './vrot'+str(vrot)+'/vout'+str(vout)+'/'
+        filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+
+        #os.system('rm '+path+filename+'_out.ascii')
+        os.system('cat '+path+filename+'_out* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_out.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_out.ascii')
+        os.system('rm temp.txt')
+
+        #os.system('rm '+path+filename+'_in.ascii')
+        os.system('cat '+path+filename+'_in* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_in.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_in.ascii')
+        os.system('rm temp.txt')

### 2017-03-02 data/tau10E7/to_join_parts_p.py

@@ -1,0 +1,23 @@
+import os
+
+logtau = 7
+vrots = [100]
+vouts = [15]
+
+for vrot in vrots:
+    for vout in vouts:
+
+        path = './vrot'+str(vrot)+'/vout'+str(vout)+'/'
+        filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+
+        #os.system('rm '+path+filename+'_out.ascii')
+        os.system('cat '+path+filename+'_out* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_out.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_out.ascii')
+        os.system('rm temp.txt')
+
+        #os.system('rm '+path+filename+'_in.ascii')
+        os.system('cat '+path+filename+'_in* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_in.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_in.ascii')
+        os.system('rm temp.txt')

### 2017-03-02 paper/paper.tex
one plot missing

@@
 
-TERMINAR...\\
+REVISAR AL FINAL...\\
 \end{abstract}
@@
 
-\subsection{Parameters' Initial Values}
+\subsection{Resulting simulated spectra}
 In order to define the ranges of \tauh, \vrot and \vout, it is necessary to refer to observational constraints.
-The common values for typical LAEs are in Tab.
+The common values for typical LAEs that were set as parameters are in Tab.
 \ref{tab:values}.
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/summary}
+		\includegraphics[width=0.48\textwidth]{./figures/summary}
 	\end{center}
@@
 	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/influence_viewing_angle_5}
+		\includegraphics[width=0.5\textwidth]{./figures/angles}
 	\end{center}
-	\caption{\textbf{\lya profile for different $\theta$:} With \tauh$=10^5$, \vrot$=50$ \kms and \vout$=20$ \kms.The intensity is in arbitrary units.
-		\label{fig:influence_viewing_angle_5}}
+	\caption{\textbf{\lya profile for different $\theta$.} For \tauh$=10^5$, \vrot$=50$ \kms and \vout$=20$ \kms.
+For \tauh$=10^6$, \vrot$=100$ \kms and \vout$=5$ \kms.
+For \tauh$=10^7$, \vrot$=100$ \kms and \vout$=15$ \kms.
+The intensity is in arbitrary units.
+		\label{fig:angles}}
 \end{figure}
 
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/influence_viewing_angle_6}
-	\end{center}
-	\caption{\textbf{\lya profile for different $\theta$:} With \tauh$=10^6$, \vrot$=100$ \kms and \vout$=5$ \kms.The intensity is in arbitrary units.
-		\label{fig:influence_viewing_angle_6}}
-\end{figure}
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/influence_viewing_angle_7}
-	\end{center}
-	\caption{\textbf{\lya profile for different $\theta$:} With \tauh$=10^7$, \vrot$=100$ \kms and \vout$=15$ \kms.The intensity is in arbitrary units.
-		\label{fig:influence_viewing_angle_7}}
-\end{figure}
-
-From Figs.
-\ref{fig:influence_viewing_angle_5}, \ref{fig:influence_viewing_angle_6} and \ref{fig:influence_viewing_angle_7} is clear that the intensity of the valley between the two peaks increases along with $\theta$.
+From Fig.
+\ref{fig:angles} it is clear that the intensity of the valley between the two peaks increases along with $\theta$.
 This causes an intensity decrease in the rest of the frequencies, thus a broadening of the line.

### 2017-03-02 paper/paper.tex
rewriting intro

@@
 
-\title{Influence of galaxy rotation and outflows in the Lyman Alpha spectral line}
-\shorttitle{Influence of galaxy rotation and outflows in the Lyman Alpha spectral line}
+\title{Influence of galaxy rotation and outflows in the Lyman-$\alpha$
+  spectral line}
+\shorttitle{Rotation and outflows in the Lyman-$\alpha$ lines}
 
-\shortauthors{Remolina-Gutierrez et al.}
+\shortauthors{Remolina-Gutierrez and Forero-Romero.}
 
 \author{Maria Camila Remolina-Gutierrez, Jaime E. Forero-Romero} 
-\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
+\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
+  No. 18A-10, Edificio Ip, Bogot\'a, Colombia} 
 \email{mc.remolina197@uniandes.edu.co}
@@
 
-\keywords{Galaxies: high-redshift, Lyman Alpha Emission, Galaxy Rotation, Galaxy Outflows, Radiative Transfer}
+\keywords{Galaxies: high-redshift, Lyman Alpha Emission, Galaxy
+  Rotation, Galaxy Outflows, Radiative Transfer}
 
 \begin{abstract}
-\noindent Young galaxies in the Universe have a strong \lya emission caused by the ionized Hydrogen atoms in their interstellar medium.
-When the spectrum of a galaxy has an intense peak around the \lya natural frequency ($2.46\times 10^{15}$ Hz) it is called a Lyman Alpha Emitter (LAE).
-Typical LAEs are very distant ($z \gtrsim 2$).
-This makes that all the data astronomers can obtain from them is their spectra, and from there all the physical information of the galaxy must be derived.
-Trying to solve this task requires the creation of a simplified and solid model.
-In this paper we propose to consider LAEs as a spherical distribution of Hydrogen atoms that undergoes a solid body rotation and a radial expansion due to outflows.
-We use radiative transfer Monte Carlo simulations to interpret the Lyman-$\alpha$ line morphology.
-The main conclusion is that this new model reproduces LAEs observed features in a clear way and with consistent physical parameters.
-We also show results of adjusting observational data for some selected objects to models with and without bulk rotation.
-We finalize by discussing the possible implications for these results in terms of the energetics required for supernova feedback and outflows in high redshift galaxies.
-\\
+\noindent Young galaxies in the Universe have a strong \lya emission
+caused by the ionized Hydrogen atoms in their interstellar
+medium. When the spectrum of a galaxy has an intense peak around the
+\lya natural frequency ($2.46\times 10^{15}$ Hz) it is called a Lyman
+Alpha Emitter (LAE). Typical LAEs are very distant ($z \gtrsim
+2$). This makes that all the data astronomers can obtain from them is
+their spectra, and from there all the physical information of the
+galaxy must be derived. Trying to solve this task requires the
+creation of a simplified and solid model. In this paper we propose to
+consider LAEs as a spherical distribution of Hydrogen atoms that
+undergoes a solid body rotation and a radial expansion due to
+outflows. We use radiative transfer Monte Carlo simulations to
+interpret the Lyman-$\alpha$ line morphology. The main conclusion is
+that this new model reproduces LAEs observed features in a clear way
+and with consistent physical parameters. We also show results of
+adjusting observational data for some selected objects to models with
+and without bulk rotation. We finalize by discussing the possible
+implications for these results in terms of the energetics required for
+supernova feedback and outflows in high redshift galaxies. \\
 
@@
 
-Galaxies are key to understand our Universe.
-However, distant ones are very challenging to detect.
-It is then an open challenge to obtain as much information as possible from them, as they represent the younger stages of galaxy evolution.
-Astronomers noted that a that relevant fraction of these galaxies emitted a really strong line at $1215.67\;\AA$ in their spectra at  and named them Lyman Alpha Emitters (LAEs).
-The purpose of this paper is to use this spectral line to model and simulate LAEs in order to narrow computationally their physical and kinematic properties.
-\\
+Distant galaxies are key to understand early evolutionary stages of
+our Universe.
+However, those galaxies are challenging to detect.
+Physical conditions in those galaxies allows the emergence of
+Lyman-$\alpha$ line emission at $1216$ \AA \cite{PartrigePeebles}.
+Galaxies detected through its Lyman-$\alpha$ emission receive the name of
+spectra at  and named Lyman Alpha Emitters (LAEs).
 
-In the early Universe the most abundant elements were Hydrogen(H) and Helium(He), causing young galaxies to be rich in H atoms.
-Due to stellar activity in a galaxy, the surrounding gas is ionized, so H atoms excite and emit radiation at definite frequencies.
-The most basic of these frequencies corresponds to the \lya line equal to $2.47 \times 10^{15}\;Hz$ \cite{PartridgePeebles}.
-This is equivalent to a wavelength of 1215.67 $\AA$ cause by the transition of the H atom electron from levels $2\rightarrow1$ dubbed Lyman Alpha ( \lya) transition, discovered by Theodore Lyman in 1906.
-\cite{LymanBio} \\
+In the early Universe the most abundant elements were Hydrogen(H) and
+Helium(He), causing young galaxies to be rich in H atoms. Due to
+stellar activity in a galaxy, the surrounding gas is ionized, so H
+atoms excite and emit radiation at definite frequencies. The most
+basic of these frequencies corresponds to the \lya line equal to $2.47
+\times 10^{15}\;Hz$ \cite{PartridgePeebles}. This is equivalent to a
+wavelength of 1215.67 $\AA$ cause by the transition of the H atom
+electron from levels $2\rightarrow1$ dubbed Lyman Alpha ( \lya)
+transition, discovered by Theodore Lyman in 1906. \cite{LymanBio} \\
 
-Due to the amount of H atoms inside a galaxy, the whole body becomes a strong \lya radiator.
-If the galaxy spectra shows a strong \lya line, it classifies as a Lyman Alpha Emitter (LAE).
-(\cite{DjorgovskiThompson}, \cite{Rhoads00}, \cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08}, \cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12}, \cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013}, \cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014}, \cite{Fumagalli2015}).
-However, when a \lya photon is emitted inside the galaxy, it travels through its interstellar medium (ISM).
-During the photon's path it can be absorbed and re-emitted by other H atoms.
-The new frequency of the photon is different that the initial one, in an observer frame of reference due to the atom's velocity.
-In LAEs, the state of the ISM gas, before and after a photon's re-emission is pretty much the same.
-This allows a random walk approximation and lets us consider photon-atom encounters as scatterings.
-A photon's scattering can happen several times and stops only until the photon is able to escape the galaxy.
-\\
+Due to the amount of H atoms inside a galaxy, the whole body becomes a
+strong \lya radiator. If the galaxy spectra shows a strong \lya line,
+it classifies as a Lyman Alpha Emitter
+(LAE). (\cite{DjorgovskiThompson}, \cite{Rhoads00},
+\cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08},
+\cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12},
+\cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013},
+\cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014},
+\cite{Fumagalli2015}). However, when a \lya photon is emitted inside
+the galaxy, it travels through its interstellar medium (ISM). During
+the photon's path it can be absorbed and re-emitted by other H
+atoms. The new frequency of the photon is different that the initial
+one, in an observer frame of reference due to the atom's velocity. In
+LAEs, the state of the ISM gas, before and after a photon's
+re-emission is pretty much the same. This allows a random walk
+approximation and lets us consider photon-atom encounters as
+scatterings. A photon's scattering can happen several times and stops
+only until the photon is able to escape the galaxy. \\
 
-In a static galaxy, this random walk process produces a spectrum with two equal and symmetric peaks around the natural \lya wavelength and zero intensity at the center (Fig.
-A3 from \cite{CLARA}).
-If now the gas has a bulk velocity, the shape of the \lya profile changes.
-We explore these effects in this paper by proposing a new model for a LAE.
-The model consists of a spherical distribution of H atoms undergoing a solid body rotation and radial expansion (outflows).
-\\
+In a static galaxy, this random walk process produces a spectrum with
+two equal and symmetric peaks around the natural \lya wavelength and
+zero intensity at the center (Fig. A3 from \cite{CLARA}). If now the
+gas has a bulk velocity, the shape of the \lya profile changes. We
+explore these effects in this paper by proposing a new model for a
+LAE. The model consists of a spherical distribution of H atoms
+undergoing a solid body rotation and radial expansion (outflows). \\
 
-Our main goal with this work is to measure the effect of the model's physical parameters on the outgoing \lya line.
-In this case, due to the resonant\footnote{Resonant is a common term used in radiative transfer.
-It means that the photons that create the line are absorbed and re-emitted several times before escaping the cloud of gas.} nature of the \lya line, analytical solutions can not be derived.
-Because of this, it becomes necessary to run simulations that explore and test the model.
-In this paper we use the radiative transfer code \textbf{CLARA} (Code for Lyman Alpha Radiation Analysis) written by Forero-Romero et al.
-\cite{CLARA}.
-CLARA can simulate the \lya line of a spherical rotating LAE depending on its mass and velocity, so we modify it to include outflows and then we explore the resulting consequences on the \lya line.
-\\
+Our main goal with this work is to measure the effect of the model's
+physical parameters on the outgoing \lya line. In this case, due to
+the resonant\footnote{Resonant is a common term used in radiative
+  transfer. It means that the photons that create the line are
+  absorbed and re-emitted several times before escaping the cloud of
+  gas.} nature of the \lya line, analytical solutions can not be
+derived. Because of this, it becomes necessary to run simulations that
+explore and test the model. In this paper we use the radiative
+transfer code \textbf{CLARA} (Code for Lyman Alpha Radiation Analysis)
+written by Forero-Romero et al. \cite{CLARA}. CLARA can simulate the
+\lya line of a spherical rotating LAE depending on its mass and
+velocity, so we modify it to include outflows and then we explore the
+resulting consequences on the \lya line. \\
 

### 2017-03-02 paper/paper.tex
title, intro

@@
 
-\title{Influence of galaxy rotation and outflows in the Lyman-$\alpha$
-  spectral line}
+\title{A simple model to describe how outflows and rotation shape the Lyman-$\alpha$
+  emission line in galaxies}
 \shorttitle{Rotation and outflows in the Lyman-$\alpha$ lines}
@@
 
-In the early Universe the most abundant elements were Hydrogen(H) and
-Helium(He), causing young galaxies to be rich in H atoms. Due to
-stellar activity in a galaxy, the surrounding gas is ionized, so H
-atoms excite and emit radiation at definite frequencies. The most
-basic of these frequencies corresponds to the \lya line equal to $2.47
-\times 10^{15}\;Hz$ \cite{PartridgePeebles}. This is equivalent to a
-wavelength of 1215.67 $\AA$ cause by the transition of the H atom
-electron from levels $2\rightarrow1$ dubbed Lyman Alpha ( \lya)
-transition, discovered by Theodore Lyman in 1906. \cite{LymanBio} \\
+Currently LAEs are commonly targetted in wide area galaxy surveys.
+They have been effectively used to study galaxy evolution,
+cosmology and the thermal history of the Universe.
+This has been able through the study of their spatial distribution and
+their global line shape.
+The recent release of MUSE has caused an switch in the kind of studies
+that can be performed on LAEs.
 
-Due to the amount of H atoms inside a galaxy, the whole body becomes a
-strong \lya radiator. If the galaxy spectra shows a strong \lya line,
-it classifies as a Lyman Alpha Emitter
-(LAE). (\cite{DjorgovskiThompson}, \cite{Rhoads00},
+Thanks to MUSE's ability to spatially resolve a galaxy, we are now able
+to infer detailed kinematic maps for nearby galaxies.
+The study of these maps would allow us to build data-driven models to
+interpret the \lya spectra of unresolved galaxies, helping us to
+constrain the physical conditions of the interstellar medium (ISM)
+processing the \lya radiation.
+
+On the ISM's features that plays an important role in shaping the \lya
+is HI kinematics. 
+In a static HI medium the \lya line has  two equal and symmetric peaks
+around the natural \lya wavelength and zero intensity at the line's
+center.
+If the ISM has an outflow velocity profile, the line becomes
+asymmetrical with a more pronounced read peak.
+If the galaxy rotates, the line shows different amounts of Doppler
+shifts that would also modify the profile.
+gas has a bulk velocity, the shape of the \lya profile changes. We
+explore these effects in this paper by proposing a new model for a
+LAE. The model consists of a spherical distribution of H atoms
+undergoing a solid body rotation and radial expansion (outflows). \\
+
+\cite{DjorgovskiThompson}, \cite{Rhoads00},
 \cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08},
@@
 \cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014},
-\cite{Fumagalli2015}). However, when a \lya photon is emitted inside
-the galaxy, it travels through its interstellar medium (ISM). During
-the photon's path it can be absorbed and re-emitted by other H
-atoms. The new frequency of the photon is different that the initial
-one, in an observer frame of reference due to the atom's velocity. In
-LAEs, the state of the ISM gas, before and after a photon's
-re-emission is pretty much the same. This allows a random walk
-approximation and lets us consider photon-atom encounters as
-scatterings. A photon's scattering can happen several times and stops
-only until the photon is able to escape the galaxy. \\
-
-In a static galaxy, this random walk process produces a spectrum with
-two equal and symmetric peaks around the natural \lya wavelength and
-zero intensity at the center (Fig. A3 from \cite{CLARA}). If now the
-gas has a bulk velocity, the shape of the \lya profile changes. We
-explore these effects in this paper by proposing a new model for a
-LAE. The model consists of a spherical distribution of H atoms
-undergoing a solid body rotation and radial expansion (outflows). \\
+\cite{Fumagalli2015}.
 

### 2017-03-28 paper/paper.tex
intro

@@
 our Universe.
-However, those galaxies are challenging to detect.
 Physical conditions in those galaxies allows the emergence of
@@
 This has been able through the study of their spatial distribution and
-their global line shape.
-The recent release of MUSE has caused an switch in the kind of studies
-that can be performed on LAEs.
+the shape of the \lya emission line.
 
-Thanks to MUSE's ability to spatially resolve a galaxy, we are now able
-to infer detailed kinematic maps for nearby galaxies.
+Recent imptovements in instrumentation have revolutionized the kind of
+studies that can be performed on LAEs.  
+It is now possible to infer detailed kinematic maps for nearby galaxies.
 The study of these maps would allow us to build data-driven models to
@@
 center.
-If the ISM has an outflow velocity profile, the line becomes
+For an outflowing ISM, the line becomes
 asymmetrical with a more pronounced read peak.
 If the galaxy rotates, the line shows different amounts of Doppler
-shifts that would also modify the profile.
-gas has a bulk velocity, the shape of the \lya profile changes. We
-explore these effects in this paper by proposing a new model for a
-LAE. The model consists of a spherical distribution of H atoms
-undergoing a solid body rotation and radial expansion (outflows). \\
+shifts modifying the overall line profile \cite{Garavito14}.
+
+In this paper we present for the first time a study of the joint
+effects of galaxy outflows and rotation.
+We study a simplified geometrical configuration corresponding to an
+spherical gas cloud with symmetrical radial outflows and a rotation
+profile corresponding to a solid body.
 
@@
 Our main goal with this work is to measure the effect of the model's
-physical parameters on the outgoing \lya line. In this case, due to
+physical parameters on the outgoing \lya line.
+In this case, due to
 the resonant\footnote{Resonant is a common term used in radiative

### 2017-03-28 paper/paper.tex
intro + mnras style

@@
-%Document Class
-\documentclass[twocolappendix]{latex/emulateapj}
+\documentclass[a4,useAMS,usenatbib,usegraphicx]{mn2e} 
+%=========================================================================
+\usepackage{amsmath} 
+\usepackage{amssymb} 
+\usepackage {graphicx}
+\usepackage[dvips]{epsfig}
+\usepackage{epsfig}  
+\usepackage{color}
+\usepackage[normalem]{ulem}
+\usepackage{hyperref}
+\usepackage{caption}
+%Non reposionated tables
+\usepackage{float}
+\restylefloat{table}
+%Multiple columns support for tables
+\usepackage{array}
+\usepackage{booktabs}
+\setlength{\heavyrulewidth}{1.5pt}
+\setlength{\abovetopsep}{4pt}
 
-%Packages and Commands
-\include{latex/packages_commands}
+%=========================================================================
+%		INTERNAL MACROS
+%=========================================================================
+\def\be{\begin{equation}}
+\def\ee{\end{equation}}
+\def\ba{\begin{eqnarray}}
+\def\ea{\end{eqnarray}}
+
+% To highlight comments 
+\definecolor{red}{rgb}{1,0.0,0.0}
+\newcommand{\red}{\color{red}}
+\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
+\newcommand{\SRK}[1]{\textcolor{darkgreen}{\bf SRK: \textit{#1}}}
+\newcommand{\SRKED}[1]{\textcolor{darkgreen}{\bf #1}}
+\newcommand{\before}[1]{\textcolor{red}{ #1}}
+\newcommand{\after}[1]{\textcolor{darkgreen}{ #1}}
+\newcommand{\tol}{Tololo 1214-277}
+\newcommand{\lya}{Ly$\alpha$}
+\newcommand{\hb}{H$\beta$}
+\newcommand{\ha}{H$\alpha$}
+\newcommand{\oiii}{[OIII]}
+\newcommand{\oii}{[OII]}
+\newcommand{\nii}{[NII]}
+\newcommand{\esca}{erg cm$^{-2}$ s$^{-1}$ \AA$^{-1}$}
+\newcommand{\esc}{erg cm$^{-2}$ s$^{-1}$}
+\newcommand{\es}{erg s$^{-1}$}
+\newcommand{\esa}{erg s$^{-1}$}
+\newcommand{\kms}{km s$^{-1}$}
+\newcommand{\apj}{ApJ}  
+\newcommand{\jcap}{JCAP}  
+\newcommand{\apjs}{ApJS}  
+\newcommand{\pasa}{PASA}  
+\newcommand{\apjl}{ApJL}  
+\newcommand{\aj}{AJ}  
+\newcommand{\mnras}{MNRAS}  
+\newcommand{\mnrassub}{MNRAS accepted}  
+\newcommand{\aap}{A\&A}  
+\newcommand{\aaps}{A\&AS}  
+\newcommand{\araa}{ARA\&A}  
+\newcommand{\nat}{Nature}  
+\newcommand{\physrep}{PhR}
+\newcommand{\pasp}{PASP}    
+\newcommand{\pasj}{PASJ}    
+\newcommand{\sigmaclump}{$54.3\pm 0.6$ km s$^{-1}$}
+\newcommand{\inftyclump}{$54.3\pm 5.1$ km s$^{-1}$}
+\newcommand{\probaclump}{$0.96\pm 0.01$}
+\def\simgt{\lower.5ex\hbox{\gtsima}}
+\def\simlt{\lower.5ex\hbox{\ltsima}}
 
@@
 
-\title{A simple model to describe how outflows and rotation shape the Lyman-$\alpha$
-  emission line in galaxies}
-\shorttitle{Rotation and outflows in the Lyman-$\alpha$ lines}
+%=========================================================================
+%		FRONT MATTER
+%=========================================================================
+\title[An atypical \lya\ dwarf galaxy]{
+Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy}
+\author[J.E. Forero-Romero et al.]{
+  Maria Camila Remolina-Guti\'errez$^1$ \&
+  Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co}\\
+  %%
+  $^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
+  No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
+}
 
-\shortauthors{Remolina-Gutierrez and Forero-Romero.}
+       %\vspace*{6pt}\\
 
-\author{Maria Camila Remolina-Gutierrez, Jaime E. Forero-Romero} 
-\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
-  No. 18A-10, Edificio Ip, Bogot\'a, Colombia} 
-\email{mc.remolina197@uniandes.edu.co}
-\email{je.forero@uniandes.edu.co}
-
-\keywords{Galaxies: high-redshift, Lyman Alpha Emission, Galaxy
-  Rotation, Galaxy Outflows, Radiative Transfer}
+       %}
+       \maketitle
 
 \begin{abstract}
-\noindent Young galaxies in the Universe have a strong \lya emission
-caused by the ionized Hydrogen atoms in their interstellar
-medium. When the spectrum of a galaxy has an intense peak around the
-\lya natural frequency ($2.46\times 10^{15}$ Hz) it is called a Lyman
-Alpha Emitter (LAE). Typical LAEs are very distant ($z \gtrsim
-2$). This makes that all the data astronomers can obtain from them is
-their spectra, and from there all the physical information of the
-galaxy must be derived. Trying to solve this task requires the
-creation of a simplified and solid model. In this paper we propose to
-consider LAEs as a spherical distribution of Hydrogen atoms that
-undergoes a solid body rotation and a radial expansion due to
-outflows. We use radiative transfer Monte Carlo simulations to
-interpret the Lyman-$\alpha$ line morphology. The main conclusion is
-that this new model reproduces LAEs observed features in a clear way
-and with consistent physical parameters. We also show results of
-adjusting observational data for some selected objects to models with
-and without bulk rotation. We finalize by discussing the possible
-implications for these results in terms of the energetics required for
-supernova feedback and outflows in high redshift galaxies. \\
+  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
+  pristine conditions of the first galaxies in the Universe and
+  are the best systems to test models on primordial
+galaxy formation and evolution.    
+Here we report on one of such CDGs,
 
-REVISAR AL FINAL...\\
 \end{abstract}
+
+\begin{keywords}
+galaxies: dwarf --- galaxies: individual:\tol\ --- radiative transfer --- Methods: numerical 
+\end{keywords}
+
+%=========================================================================
+%		PAPER CONTENT
+%=========================================================================
+
+%*************************************************************************
 
@@
 
+\bibliographystyle{mn2e}
+\bibliography{references}
+
+\end{document}
+
 \section{A new LAE model}
@@
 
-There are 3 parameters in this model that define a LAE: the rotational velocity (\vrot), the outflow velocity (\vout) and the optical depth (\tauh).
-\vout is due to material ejected from the galaxy, by supernovas(\cite{Verhamme06}, \cite{Orsi12}, \cite{Hashimoto2015}, \cite{Gronke2015}).
-\tauh roughly corresponds to number of H atoms found by a \lya photon if one traces a line from the center of the galaxy to its edge, and it resembles the mass of the LAE.\\
+There are 3 parameters in this model that define a LAE: the rotational velocity (vrot), the outflow velocity (vout) and the optical depth (tauh).
+vout is due to material ejected from the galaxy, by supernovas(\cite{Verhamme06}, \cite{Orsi12}, \cite{Hashimoto2015}, \cite{Gronke2015}).
+tauh roughly corresponds to number of H atoms found by a \lya photon if one traces a line from the center of the galaxy to its edge, and it resembles the mass of the LAE.\\
 
@@
 
-These 3 parameters, \vrot, \vout and \tauh, leave the idea of a model of LAE, that although simple, considers the main galaxy's dynamics.
+These 3 parameters, vrot, vout and tauh, leave the idea of a model of LAE, that although simple, considers the main galaxy's dynamics.
 All of them have been previously proposed and used by different authors, but never combined together (\cite{Adams72}, \cite{Harrington73}, \cite{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12}, \cite{Martin2015}, \cite{Garavito14}, \cite{Neufeld91}, \cite{Laursen09}, \cite{Barnes11}, \cite{Verhamme12}, \cite{Yajima12}).\\
@@
 Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires.
-Depending on \tauh, we need a different number of processors in order to minimize the time demands.
-For \tauh $=10^5, 10^6, 10^7$ we use 6, 12, 24 processors, respectively.
+Depending on tauh, we need a different number of processors in order to minimize the time demands.
+For tauh $=10^5, 10^6, 10^7$ we use 6, 12, 24 processors, respectively.
 The running times go from 7 hours to 1 week approximately, depending on the case.
@@
 
-To summarize, the parameters influencing the spectra are \vrot, \vout, \tauh and $\theta$. In the next section we evaluate the impact each of these has on the resulting profile. \\
+To summarize, the parameters influencing the spectra are vrot, vout, tauh and $\theta$. In the next section we evaluate the impact each of these has on the resulting profile. \\
 
@@
 \subsection{Resulting simulated spectra}
-In order to define the ranges of \tauh, \vrot and \vout, it is necessary to refer to observational constraints.
+In order to define the ranges of tauh, vrot and vout, it is necessary to refer to observational constraints.
 The common values for typical LAEs that were set as parameters are in Tab.
@@
 	\end{center}
-	\caption{\textbf{4 different \lya profiles:} With \tauh$=10^5$ and $\theta \simeq 90^\circ$.
-The rotational velocity \vrot increases to the right and the outflows velocity \vout increases downwards.
+	\caption{\textbf{4 different \lya profiles:} With tauh$=10^5$ and $\theta \simeq 90^\circ$.
+The rotational velocity vrot increases to the right and the outflows velocity vout increases downwards.
 The intensity is in arbitrary units.
@@
 	\end{center}
-	\caption{\textbf{\lya profile for different $\theta$.} For \tauh$=10^5$, \vrot$=50$ \kms and \vout$=20$ \kms.
-For \tauh$=10^6$, \vrot$=100$ \kms and \vout$=5$ \kms.
-For \tauh$=10^7$, \vrot$=100$ \kms and \vout$=15$ \kms.
+	\caption{\textbf{\lya profile for different $\theta$.} For tauh$=10^5$, vrot$=50$ \kms and vout$=20$ \kms.
+For tauh$=10^6$, vrot$=100$ \kms and vout$=5$ \kms.
+For tauh$=10^7$, vrot$=100$ \kms and vout$=15$ \kms.
 The intensity is in arbitrary units.
@@
 \begin{itemize}
-	\item \tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
-	\item \vout decreases the right peak's intensity. Higher \vout makes the left peak smaller until it merges with the right one. \\
-	\item \vrot broadens the line and decreases the maximum intensity.
-Higher \vrot implies a flatter spectrum.
+	\item tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
+	\item vout decreases the right peak's intensity. Higher vout makes the left peak smaller until it merges with the right one. \\
+	\item vrot broadens the line and decreases the maximum intensity.
+Higher vrot implies a flatter spectrum.
 This effect has not been deeply studied in literature.
@@
 As seen in Fig.
-\ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with \vrot$=0$ and displaces it from its central location.
+\ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with vrot$=0$ and displaces it from its central location.
 Then it weights the lines and they all merge into the resulting red \lya line.\\
@@
 The consideration of adding the new rotation parameter to the standard only-outflow model found in literature is reasonable and very powerful.
-\vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\
+vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\
 
@@
 
-	\item The effects of \vrot, \vout and \tauh are consistent with the different authors that have used them.
-\vrot broadens the \lya line.
-\vout increases the peaks asymmetry and \tauh induces a redshift around the zero velocity.
+	\item The effects of vrot, vout and tauh are consistent with the different authors that have used them.
+vrot broadens the \lya line.
+vout increases the peaks asymmetry and tauh induces a redshift around the zero velocity.
 
@@
 %------------------------REFERENCES----------------------------
-
-\bibliographystyle{latex/apj}
-\bibliography{references}
-
-\end{document}

### 2017-03-28 paper/paper.tex
new sections

@@
 %=========================================================================
-\title[An atypical \lya\ dwarf galaxy]{
-Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy}
-\author[J.E. Forero-Romero et al.]{
+\title[Outflows and rotation in LAEs]{
+Outflows and rotation in Lyman alpha galaxies}
+\author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
   Maria Camila Remolina-Guti\'errez$^1$ \&
@@
 
-       %\vspace*{6pt}\\
-
-       %}
 \maketitle
@@
 profile corresponding to a solid body.
+We base our modelling on a Monte-Carlo radiative transfer code called
+CLARA (Code for Lyman Alpha Radiation Analysis) presented for the
+first time in \cite{CLARA}.
+Besides modelling the impact of joint rotation and outflows, we also
+want to check to what extent the analytical model presented by
+\cite{Garavito14} to explain the effects of rotation can also be
+applied in our case.
+
+In this paper we introduce first our theoretical tools and assumptions
+in Section \ref{sec:theory}, then we present the numerical results and
+comparisons against the analytical solution in \ref{sec:results}.
+In Section \ref{sec:discussion} we discuss our results and their
+possible implications for observational analysis to finally present
+our conclusions in Section \ref{sec:conclusions}.
+
+\section{Theoretical Models}
+\label{sec:theory}
+
+\subsection{Monte-Carlo Radiative Transfer Model}
+
+\subsection{Analytical Model for Bulk Rotation}
+
+\section{Results}
+\label{sec:results}
+
+\subsection{Monte-Carlo Radiative Transfer Model}
+
+\subsection{Analytical Model for Bulk Rotation}
+
+\section{Discussion}
+\label{sec:discussion}
+
+\subsection{Theoretical Insights}
+
+\subsection{Observational Perspectives}
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+\bibliographystyle{mn2e}
+\bibliography{references}
+
+\end{document}
 
@@
 \cite{Fumagalli2015}.
-
-Our main goal with this work is to measure the effect of the model's
-physical parameters on the outgoing \lya line.
-In this case, due to
-the resonant\footnote{Resonant is a common term used in radiative
-  transfer. It means that the photons that create the line are
-  absorbed and re-emitted several times before escaping the cloud of
-  gas.} nature of the \lya line, analytical solutions can not be
-derived. Because of this, it becomes necessary to run simulations that
-explore and test the model. In this paper we use the radiative
-transfer code \textbf{CLARA} (Code for Lyman Alpha Radiation Analysis)
-written by Forero-Romero et al. \cite{CLARA}. CLARA can simulate the
-\lya line of a spherical rotating LAE depending on its mass and
-velocity, so we modify it to include outflows and then we explore the
-resulting consequences on the \lya line. \\
-
-\bibliographystyle{mn2e}
-\bibliography{references}
-
-\end{document}
 

### 2017-03-30 paper/paper.tex
section 2.1

@@
 
+CLARA follows the propagation of individual photons through a neutral
+Hydrogen medium characterized by its temperature, velocity field and
+global optical depth.
+The code assumes an homogeneous density throughout the simulated
+volume.
+In the current implementation we neglect the influence of dust.
+Our basic models is an spherical distribution of neutral hydrogen,
+an approximation commonly used in the literature, as it explains a
+wide variety of observational features \citep{Ahn03,Verhamme06,Dijkstra06}.
+
+The central element in this paper is the velocity field that captures
+outflows and rotation.
+Outflows are captured by a Hubble-like radial velocity profile with
+the velocity magnitude increasing linearly with the radial
+coordinate; the outflow model is fully characterized by $V_{\rm out}$, the
+velocity at the sphere's surface.
+Rotation follows a solid body rotation profile, which is fully
+characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.
+
+The total velocity field corresponds to the superposition of rotation and
+outflows.
+The cartesian components take the form
+
+\begin{equation}
+	v_{x}=\frac{x}{R}V_{\rm out}-\frac{y}{R}V_{\rm rot} ,
+	\label{eq:vx}
+\end{equation}
+
+\begin{equation}
+	v_{y}=\frac{y}{R}V_{\rm out}+\frac{x}{R}V_{\rm rot} ,
+	\label{eq:vy}
+\end{equation}
+
+\begin{equation}
+	v_{z}=\frac{z}{R}V_{\rm out},
+	\label{eq:vz}
+\end{equation}
+%
+where $x$, $y$ and $z$ are the cartesian position coordinates with the
+origin at the sphere's center $R$ is the radius of the sphere and the
+direction of the angular velocity vector corresponds to the $\hat{k}$
+unit vector.
+
+For each model setup we follow $10^4$ individual photons generated at
+the center of the sphere at the \lya\ line's center as they propagate
+through the spere and finally scape.
+We store the final frequency and propagation direction for each photon
+at its last scattering.
+
 \subsection{Analytical Model for Bulk Rotation}
@@
 \label{sec:newmodel}
-
-We model spherical galaxy to facilitate the interpretation of the results, by deleting what would be another degree of freedom regarding a different geometry.
-Furthermore, this approximation is commonly used in the literature, as it explains a wide variety of observational features (\cite{Ahn03}, \cite{Verhamme06}, \cite{Dijkstra06}).
-\\
-
-There are 3 parameters in this model that define a LAE: the rotational velocity (vrot), the outflow velocity (vout) and the optical depth (tauh).
-vout is due to material ejected from the galaxy, by supernovas(\cite{Verhamme06}, \cite{Orsi12}, \cite{Hashimoto2015}, \cite{Gronke2015}).
-tauh roughly corresponds to number of H atoms found by a \lya photon if one traces a line from the center of the galaxy to its edge, and it resembles the mass of the LAE.\\
-
-In this model, the LAE has a bulk velocity corresponding to the superposition of rotation and outflows, as shown in Fig.
-\ref{fig:model}.
-The velocity components are written in Eqs.
-(\ref{eq:vx}) (\ref{eq:vy}) (\ref{eq:vz}).
-In these equations, $R$ is the radius of the sphere; $x$, $y$ and $z$ are the coordinates in a cartesian frame; and the $\mp$ signs in $v_x$ and $v_y$ indicate the direction of rotation, respectively.
-This rotation is a solid body rotation and its direction goes according to the right hand rule applied to the $\hat{k}$ unit vector.
-The outflow velocity is dependent on the position relative to the center of the galaxy, being it zero at the center and maximum at the edge of the sphere.\\
-
-\begin{equation}
-	v_{x}=\frac{x}{R}v_{\rm out}-\frac{y}{R}v_{\rm rot} 
-	\label{eq:vx}
-\end{equation}
-
-\begin{equation}
-	v_{y}=\frac{y}{R}v_{\rm out}+\frac{x}{R}v_{\rm rot} 
-	\label{eq:vy}
-\end{equation}
-
-\begin{equation}
-	v_{z}=\frac{z}{R}v_{\rm out}
-	\label{eq:vz}
-\end{equation}
 

### 2017-04-03 plots/paper/doppler_shift.c
started doppler shift + gitignore

@@ -1,0 +1,29 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <math.h>
+
+int main() {
+
+  //Number of lines in file
+  static const int num_xs = 11169;
+
+  //The info with v_rot = 0
+  FILE * file;
+  file = fopen("vrot0_vout25.dat" , "r");
+
+  //Pointer to save xs
+  double *xs;
+  xs = malloc(num_xs*sizeof(double));
+
+  if (file) {
+    int i = 0;
+    while (fscanf(file, "%lf", &xs[i]) != EOF) {
+      i++;
+    }
+    fclose(file);
+  }
+
+  printf("%f\n", xs[0]);
+
+  return 0;
+}

### 2017-04-04 data/tau10E5/vrot100/vout5/tau10E5_vrot100_vout5.sh
adding scripts and inputs

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E5_vrot100_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot100/vout5/tau10E5_vrot100_vout5.input

### 2017-04-04 data/tau10E5/vrot50/vout20/tau10E5_vrot50_vout20.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E5_vrot50_vout20
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot50/vout20/tau10E5_vrot50_vout20.input

### 2017-04-04 data/tau10E5/vrot50/vout5/tau10E5_vrot50_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E5_vrot50_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot50/vout5/tau10E5_vrot50_vout5.input

### 2017-04-04 data/tau10E6/vrot100/vout5/tau10E6_vrot100_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E6_vrot100_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=12
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot100/vout5/tau10E6_vrot100_vout5.input

### 2017-04-04 data/tau10E7/vrot100/vout15/tau10E7_vrot100_vout15.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E7_vrot100_vout15
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=24
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 24 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E7/vrot100/vout15/tau10E7_vrot100_vout15.input

### 2017-04-04 paper/paper.tex
title

@@
 \title[Outflows and rotation in LAEs]{
-Outflows and rotation in Lyman alpha galaxies}
+Outflows and rotation in Lyman alpha emitting galaxies}
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
@@
 
+Rotation induces two main effects on the \lya line morphology.
+First. Break of spherical symmetry due to the 
+  preferential direction introduced by the rotation axis.
+  This symmetry break is reflected in the observed spectra as a
+  dependency on the viewing angle, $\theta$, between the line of sight
+  of a distant observer and the rotaional velocity.    
+Second. Line blurring as the rotational velocity increases.
+  At a fixed viewing angle, faster rotation induces makes the observed
+  line wider and increases the flux around the line's center.
+
+\cite{Garavito14} presented in the Appendix an analytical model that
+accounts for these two features.
+The basic assumption of the analytical model is that each
+differential surface element on the sphere Doppler shifts the photons
+that it emmits.
+
 \section{Results}
@@
 
-These 3 parameters, vrot, vout and tauh, leave the idea of a model of LAE, that although simple, considers the main galaxy's dynamics.
-All of them have been previously proposed and used by different authors, but never combined together (\cite{Adams72}, \cite{Harrington73}, \cite{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12}, \cite{Martin2015}, \cite{Garavito14}, \cite{Neufeld91}, \cite{Laursen09}, \cite{Barnes11}, \cite{Verhamme12}, \cite{Yajima12}).\\
+These 3 parameters, vrot, vout and tauh, leave the idea of a model of
+LAE, that although simple, considers the main galaxy's dynamics. All
+of them have been previously proposed and used by different authors,
+but never combined together (\cite{Adams72}, \cite{Harrington73}, \cit
+e{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12},
+\cite{Martin2015}, \cite{Garavito14}, \cite{Neufeld91},
+\cite{Laursen09}, \cite{Barnes11}, \cite{Verhamme12},
+\cite{Yajima12}).\\
 
 %mirar donde pongo esto
-However, in the plots shown in this paper we use instead velocity, $V$, units. This eases the comparison against observational data. The velocity units are defined by
-The units of $V$ are usually \kms. In this case, the photon is redshifted in frequency when the velocity is greater than 0, and blueshifted in the opposite case. \\
+However, in the plots shown in this paper we use instead velocity,
+$V$, units. This eases the comparison against observational data. The
+velocity units are defined by 
+The units of $V$ are usually \kms. In this case, the photon is
+redshifted in frequency when the velocity is greater than 0, and
+blueshifted in the opposite case. \\
 
-The initial emission of photons is taken at the center of the sphere for practicality due to the fact that both, center and off-center emissions, give analogous results.
-From here, $100000$ photons are emitted with the natural \lya wavelength and start to scatter.
-When each photon is re-emitted, its new wavelength depends on the H atom's velocity (both thermal and bulk) and direction (both initial and final).
-However the photon's new direction of propagation is random in the rest frame of the atom.
-\\
+The initial emission of photons is taken at the center of the sphere
+for practicality due to the fact that both, center and off-center
+emissions, give analogous results. From here, $100000$ photons are
+emitted with the natural \lya wavelength and start to scatter. When
+each photon is re-emitted, its new wavelength depends on the H atom's
+velocity (both thermal and bulk) and direction (both initial and
+final). However the photon's new direction of propagation is random in
+the rest frame of the atom. \\
 
-The individual scattering of all the photons is tracked through the complete 3D Hydrogen distribution.
-Once each photon escapes the galaxy, its final values are stored: position $\vec{r}$, direction of propagation $\hat{k}$, dimensionless frequency $x$, and number of scatterings $N$.
-To build the observed spectrum we make a histogram of the escape frequencies $x$.
-The number of scatterings $N$ tells how many steps the random walk requires to reach a distance to the center that is $\geq R$.
-In order to avoid situations in which the photon has not escaped after a long computational time, CLARA defines a number $N_{max}$ so the code stops.
-However, according to statistics, this last situation has low probabilities, so the photon is always most likely to exit the sphere.
-\\
+The individual scattering of all the photons is tracked through the
+complete 3D Hydrogen distribution. Once each photon escapes the
+galaxy, its final values are stored: position $\vec{r}$, direction of
+propagation $\hat{k}$, dimensionless frequency $x$, and number of
+scatterings $N$. To build the observed spectrum we make a histogram of
+the escape frequencies $x$. The number of scatterings $N$ tells how
+many steps the random walk requires to reach a distance to the center
+that is $\geq R$. In order to avoid situations in which the photon has
+not escaped after a long computational time, CLARA defines a number
+$N_{max}$ so the code stops. However, according to statistics, this
+last situation has low probabilities, so the photon is always most
+likely to exit the sphere. \\
 
@@
 \subsection{Galaxy's Viewing Angle}
-An observer located far away, only receives photons emitted along its line of sight.
-That means, only photons escaping in the direction of the observer must be counted in the spectrum.
-In the simulation we approximate this by taking into account only the photons with escaping direction angle $\theta$ respect to the rotation axis within the range $[\theta_{min}-\theta_{max}]$.
-We illustrate this in Fig.
-\ref{fig:model}.
-\\
+An observer located far away, only receives photons emitted along its
+line of sight. That means, only photons escaping in the direction of
+the observer must be counted in the spectrum. In the simulation we
+approximate this by taking into account only the photons with escaping
+direction angle $\theta$ respect to the rotation axis within the range
+$[\theta_{min}-\theta_{max}]$.
+
+We illustrate this in
+Fig. \ref{fig:model}. \\
 

### 2017-04-06 paper/paper.tex
list of models

@@
 
-For each model setup we follow $10^4$ individual photons generated at
+For each model setup we follow $10^5$ individual photons generated at
 the center of the sphere at the \lya\ line's center as they propagate
@@
 at its last scattering.
+
+In Table \ref{tab:values} we list the combination of $\tau_{\rm{H}}$,
+$v_{rot}$ and $v_{out}$ values used in this paper.
+The range of values have some overlap with the expectations from a dwarf
+galaxy with a total neutral hydrogen mass of $10^8$-$10^9$ $M_{\odot}$.
+We run a total of $21$ different models.
+
+\begin{table}[htbp]
+  \begin{center}
+    \begin{tabular}{|c|c|c|}
+      \hline
+      $\tau_{\mathrm{H}}$ & $v_{rot}$ (\kms) & $v_{out}$ (\kms) \\
+      \hline
+      $10^5$  & \{0, 50, 100\} & \{25, 50\} \\
+      $10^6$ & \{0, 50, 100\} & \{25, 50\}\\
+      $10^7$ & \{0, 50\} & \{25, 50\}\\
+      $10^7$ & \{0, 100\} & \{50, 100, 200\}\\
+      \hline    
+    \end{tabular}
+  \end{center}
+  \caption{\textbf{Parameters' Values.} List of values that were used
+    to construct the radiative transfer models. Values in braces
+    for a row indicate that all possible combinations in of
+    $\tau_{\rm{H}}$,
+    $v_{rot}$ and $v_{out}$ are used.}
+  \label{tab:values}
+\end{table}
 
@@
 that it emmits.
+In this paper we introduce this ansatz by postprocessing the results
+of the outflow simulations without rotation.
+This allows us to produce new \lya spectra and compare them with the
+full radiative transfer solution including both outflows and
+rotation.
 
@@
 \\
-
-\begin{table}[htbp]
-	\centering
-	\begin{tabular}{|c|c|c|}
-		\hline
-		$\tau_{\mathrm{H}}$ & $v_{rot}$ (\kms) & $v_{out}$ (\kms) \\
-		\hline
-		$10^5$, $10^6$, $10^7$ & 50, 100 & 5, 10, 15, 20, 25, 50, 75 \\
-		\hline
-	\end{tabular}
-	\caption{\textbf{Parameters' Values:} All consistent with a LAE's typical properties}
-	\label{tab:values}
-\end{table}
 

### 2017-04-07 paper/paper.tex
estructura de los resultados

@@
 galaxy with a total neutral hydrogen mass of $10^8$-$10^9$ $M_{\odot}$.
-We run a total of $21$ different models.
+We run a total of $24$ different models.
 
@@
       \hline
-      $10^5$  & \{0, 50, 100\} & \{25, 50\} \\
-      $10^6$ & \{0, 50, 100\} & \{25, 50\}\\
-      $10^7$ & \{0, 50\} & \{25, 50\}\\
-      $10^7$ & \{0, 100\} & \{50, 100, 200\}\\
+      $10^5$  & \{0\} & \{5, 25, 50\} \\
+      $10^6$ & \{0\} & \{5, 25, 50\}\\
+      $10^7$ & \{0\} & \{5, 25, 50\}\\
+      $10^5$  & \{50, 100\} & \{25, 50, 75\} \\
+      $10^6$ & \{50, 100\} & \{25, 50,75\}\\
+      $10^7$ & \{50, 100\} & \{25, 50,75\}\\
       \hline    
@@
 
+We evaluate first the effects of increasing rotational
+velocity at fixed values of the optical depth and outflow velocity.
+Figure \ref{fig:varying_rotation} summarizes the information.
+All plots correspond to the same outflow velocity of $v_{out}=50$\kms,
+rows have constant optical depth and columns have constant rotational
+velocity.
+Each panel shows different morphologies corresponding to varying
+line-of-sight viewing angle.
+
+% Esta figura deberia tener 3 x 3 vinetas. vout=50.
+% cada panel tiene una curva para theta=0, 45, 90.
+% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100, 
+% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100, 
+% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100,
+
+We now evaluate the effects of increasing outflow velocity at fixed
+values for the optical depth and rotational velocity.
+Figure \ref{fig:varying_outflow} summarizes the information following
+the same patters as Figure \ref{fig:varying_rotation}, only that this
+time all plots correspond to the same rotation velocity of
+$v_{rot}=50$\kms,
+
+% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
+% cada panel tiene una curva para theta=0, 45, 90.
+% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
+% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
+% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75,
+
 \subsection{Analytical Model for Bulk Rotation}
@@
 
-These 3 parameters, vrot, vout and tauh, leave the idea of a model of
-LAE, that although simple, considers the main galaxy's dynamics. All
-of them have been previously proposed and used by different authors,
-but never combined together (\cite{Adams72}, \cite{Harrington73}, \cit
+(\cite{Adams72}, \cite{Harrington73}, \cit
 e{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12},
@@
 
-%mirar donde pongo esto
-However, in the plots shown in this paper we use instead velocity,
-$V$, units. This eases the comparison against observational data. The
-velocity units are defined by 
-The units of $V$ are usually \kms. In this case, the photon is
-redshifted in frequency when the velocity is greater than 0, and
-blueshifted in the opposite case. \\
-
-The initial emission of photons is taken at the center of the sphere
-for practicality due to the fact that both, center and off-center
-emissions, give analogous results. From here, $100000$ photons are
-emitted with the natural \lya wavelength and start to scatter. When
-each photon is re-emitted, its new wavelength depends on the H atom's
-velocity (both thermal and bulk) and direction (both initial and
-final). However the photon's new direction of propagation is random in
-the rest frame of the atom. \\
-
-The individual scattering of all the photons is tracked through the
-complete 3D Hydrogen distribution. Once each photon escapes the
-galaxy, its final values are stored: position $\vec{r}$, direction of
-propagation $\hat{k}$, dimensionless frequency $x$, and number of
-scatterings $N$. To build the observed spectrum we make a histogram of
-the escape frequencies $x$. The number of scatterings $N$ tells how
-many steps the random walk requires to reach a distance to the center
-that is $\geq R$. In order to avoid situations in which the photon has
-not escaped after a long computational time, CLARA defines a number
-$N_{max}$ so the code stops. However, according to statistics, this
-last situation has low probabilities, so the photon is always most
-likely to exit the sphere. \\
-
-%The process that takes place from the moment the first photon is emitted to the moment the last photon escapes the sphere, is called a simulation.
-Each simulation runs in the scientific computational cluster of Universidad de los Andes due to the computing resources it requires.
-Depending on tauh, we need a different number of processors in order to minimize the time demands.
-For tauh $=10^5, 10^6, 10^7$ we use 6, 12, 24 processors, respectively.
-The running times go from 7 hours to 1 week approximately, depending on the case.
-\\
-
 \subsection{Galaxy's Viewing Angle}
-An observer located far away, only receives photons emitted along its
-line of sight. That means, only photons escaping in the direction of
-the observer must be counted in the spectrum. In the simulation we
-approximate this by taking into account only the photons with escaping
-direction angle $\theta$ respect to the rotation axis within the range
-$[\theta_{min}-\theta_{max}]$.
 
@@
 
-Therefore, in principle the spectra depend on two new parameters, the azimuthal and the polar angle.
-However, the galaxy's motion is symmetrical respect to its rotation axis.
-This implies that the resulting spectrum is independent from the azimuthal angle.
-Taking into account this symmetry, we only select photons on their polar angle, regardless of their azimuthal angle.
-Regarding the polar angle, we build the spectra for observers located on 3 different positions, with $\theta$ intervals uniformly distributed in $\cos(\theta)$.
-\\
-
-To summarize, the parameters influencing the spectra are vrot, vout, tauh and $\theta$. In the next section we evaluate the impact each of these has on the resulting profile. \\
-
 \section{Results}
@@
 \subsection{Resulting simulated spectra}
-In order to define the ranges of tauh, vrot and vout, it is necessary to refer to observational constraints.
-The common values for typical LAEs that were set as parameters are in Tab.
-\ref{tab:values}.
-We run CLARA's modified version for all the permutations of these 3 parameters.
-\\
+In order to define the ranges of tauh, vrot and vout, it is necessary
+to refer to observational constraints. The common values for typical
+LAEs that were set as parameters are in Tab. \ref{tab:values}. We run
+CLARA's modified version for all the permutations of these 3
+parameters. \\
 
-The behavior of the resulting sets of spectra can be summarized by Fig.
-\ref{fig:summary}.
-We concluded from the simulations a clear creation of two asymmetric peaks around $V=0$ \kms with the tallest peak always redshifted.
-We detected a strong dependence on the outflow velocity that induces the peak asymmetry.
-We also detected that the rotation velocity broadens the line horizontally.
-Both these effects have been previously reported separately in the literature, so these results can valid their superposition to some extend.
-\\
+The behavior of the resulting sets of spectra can be summarized by
+Fig. \ref{fig:summary}. We concluded from the simulations a clear
+creation of two asymmetric peaks around $V=0$ \kms with the tallest
+peak always redshifted. We detected a strong dependence on the outflow
+velocity that induces the peak asymmetry. We also detected that the
+rotation velocity broadens the line horizontally. Both these effects
+have been previously reported separately in the literature, so these
+results can valid their superposition to some extend. \\
 

### 2017-09-06 paper/paper.tex
updated mnras style
rewriting the abstract

@@
-\documentclass[a4,useAMS,usenatbib,usegraphicx]{mn2e} 
+\documentclass[a4paper,fleqn,usenatbib]{mnras}
 %=========================================================================
@@
 \usepackage{amssymb} 
-\usepackage {graphicx}
+\usepackage{graphicx}
+\usepackage{grffile}
 \usepackage[dvips]{epsfig}
@@
 \usepackage{color}
-\usepackage[normalem]{ulem}
+\usepackage{caption}
 \usepackage{hyperref}
-\usepackage{caption}
+\usepackage{bm}
 %Non reposionated tables
-\usepackage{float}
-\restylefloat{table}
-%Multiple columns support for tables
-\usepackage{array}
-\usepackage{booktabs}
-\setlength{\heavyrulewidth}{1.5pt}
-\setlength{\abovetopsep}{4pt}
-
-%=========================================================================
-%		INTERNAL MACROS
-%=========================================================================
-\def\be{\begin{equation}}
-\def\ee{\end{equation}}
-\def\ba{\begin{eqnarray}}
-\def\ea{\end{eqnarray}}
-
-% To highlight comments 
-\definecolor{red}{rgb}{1,0.0,0.0}
-\newcommand{\red}{\color{red}}
-\definecolor{darkgreen}{rgb}{0.0,0.5,0.0}
-\newcommand{\SRK}[1]{\textcolor{darkgreen}{\bf SRK: \textit{#1}}}
-\newcommand{\SRKED}[1]{\textcolor{darkgreen}{\bf #1}}
-\newcommand{\before}[1]{\textcolor{red}{ #1}}
-\newcommand{\after}[1]{\textcolor{darkgreen}{ #1}}
-\newcommand{\tol}{Tololo 1214-277}
-\newcommand{\lya}{Ly$\alpha$}
+\newcommand{\HI}{{\text{H\MakeUppercase{\romannumeral 1}}} }
+\newcommand{\HII}{{\text{H\MakeUppercase{\romannumeral 2}}} }
+\newcommand{\lya}{\ifmmode{{\rm Ly}\alpha}\else Ly$\alpha$\ \fi}
+\newcommand{\cm}{\ifmmode{{\rm cm}}\else cm\fi}
+\newcommand{\ccm}{\,\mathrm{cm}^{-3}}
+\newcommand{\ergps}{\,{\rm erg}\,{\rm s}\ifmmode{}^{-1}\else ${}^{-1}$\fi}
+\newcommand{\Mpch}{\,{\rm Mpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
+%\newcommand{\kms}{\,\mathrm{km}\,\mathrm{s}^{-1}}
+\newcommand{\dd}{\mathrm{d}}
+\newcommand{\vek}[1]{\bm{#1}}
+%\newcommand{\lya}{Ly$\alpha$}
 \newcommand{\hb}{H$\beta$}
@@
 \newcommand{\esa}{erg s$^{-1}$}
-\newcommand{\kms}{km s$^{-1}$}
-\newcommand{\apj}{ApJ}  
-\newcommand{\jcap}{JCAP}  
-\newcommand{\apjs}{ApJS}  
-\newcommand{\pasa}{PASA}  
-\newcommand{\apjl}{ApJL}  
-\newcommand{\aj}{AJ}  
-\newcommand{\mnras}{MNRAS}  
-\newcommand{\mnrassub}{MNRAS accepted}  
-\newcommand{\aap}{A\&A}  
-\newcommand{\aaps}{A\&AS}  
-\newcommand{\araa}{ARA\&A}  
-\newcommand{\nat}{Nature}  
-\newcommand{\physrep}{PhR}
-\newcommand{\pasp}{PASP}    
-\newcommand{\pasj}{PASJ}    
+\newcommand{\kms}{\ifmmode\mathrm{km\ s}^{-1}\else km s$^{-1}$\fi}
 \newcommand{\sigmaclump}{$54.3\pm 0.6$ km s$^{-1}$}
@@
 \newcommand{\probaclump}{$0.96\pm 0.01$}
-\def\simgt{\lower.5ex\hbox{\gtsima}}
-\def\simlt{\lower.5ex\hbox{\ltsima}}
 
@@
 \title[Outflows and rotation in LAEs]{
-Outflows and rotation in Lyman alpha emitting galaxies}
+Lyman-alpha spectra from rotating outflows}
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
@@
 \begin{abstract}
-  Star-forming Compact Dwarf Galaxies (CDGs) resemble the expected
-  pristine conditions of the first galaxies in the Universe and
-  are the best systems to test models on primordial
-galaxy formation and evolution.    
-Here we report on one of such CDGs,
+We perform Monte Carlo radiative transfer simulations of outflowing
+gas with additional solid body rotation.
+We reproduce the main features of the outcoming spectra with a simple
+semi-analytic model that modifies the spectra of outflows without
+rotation by adding a Doppler shift with the rotational information.
 
@@
 \begin{keywords}
-galaxies: dwarf --- galaxies: individual:\tol\ --- radiative transfer --- Methods: numerical 
+galaxies: dwarf --- radiative transfer --- Methods: numerical 
 \end{keywords}
@@
 
-\subsection{Analytical Model for Bulk Rotation}
+\subsection{Doppler Shifts to Include Rotational Effects}
 
@@
 
-\subsection{Observational Perspectives}
+\subsection{Application to observational data}
 
@@
 
-\bibliographystyle{mn2e}
+\bibliographystyle{mnras}
 \bibliography{references}

### 2017-10-13 paper/paper.tex
adding spectra ensemble and typos in paper

@@
 \newcommand{\HI}{{\text{H\MakeUppercase{\romannumeral 1}}} }
+\newcommand{\lya}{\ifmmode{{\rm Ly}\alpha}\else Ly$\alpha$\ \fi}
+\newcommand{\kms}{\ifmmode\mathrm{km\ s}^{-1}\else km s$^{-1}$\fi}
+\newcommand{\vrot}{\ifmmode\mathrm v_{\mathrm{rot}}\else $v_{\mathrm{rot}}$~\fi}
+\newcommand{\vout}{\ifmmode\mathrm v_{\mathrm{out}}\else $v_{\mathrm{out}}$~\fi}
+\newcommand{\tauh}{$\tau_{\mathrm{H}}$}
+
 \newcommand{\HII}{{\text{H\MakeUppercase{\romannumeral 2}}} }
-\newcommand{\lya}{\ifmmode{{\rm Ly}\alpha}\else Ly$\alpha$\ \fi}
 \newcommand{\cm}{\ifmmode{{\rm cm}}\else cm\fi}
@@
 \newcommand{\Mpch}{\,{\rm Mpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
-%\newcommand{\kms}{\,\mathrm{km}\,\mathrm{s}^{-1}}
 \newcommand{\dd}{\mathrm{d}}
 \newcommand{\vek}[1]{\bm{#1}}
-%\newcommand{\lya}{Ly$\alpha$}
 \newcommand{\hb}{H$\beta$}
@@
 \newcommand{\esa}{erg s$^{-1}$}
-\newcommand{\kms}{\ifmmode\mathrm{km\ s}^{-1}\else km s$^{-1}$\fi}
 \newcommand{\sigmaclump}{$54.3\pm 0.6$ km s$^{-1}$}
@@
 %=========================================================================
-\title[Outflows and rotation in LAEs]{
-Lyman-alpha spectra from rotating outflows}
+\title[Outflows and rotation in LAEs]{Lyman-alpha spectra from rotating outflows}
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
-  Maria Camila Remolina-Guti\'errez$^1$ \&
-  Jaime E. Forero-Romero $^{1}$ \thanks{je.forero@uniandes.edu.co}\\
+  Maria Camila Remolina-Guti\'errez$^{1}$ 
+  \thanks{mc.remolina197@uniandes.edu.co} \&
+  Jaime E. Forero-Romero $^{1}$ 
+  \thanks{je.forero@uniandes.edu.co}\\
   %%
-  $^1$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
+  $^{1}$ Departamento de F\'isica, Universidad de los Andes, Cra. 1
   No. 18A-10 Edificio Ip, CP 111711, Bogot\'a, Colombia \\
@@
 our Universe.
-Physical conditions in those galaxies allows the emergence of
-Lyman-$\alpha$ line emission at $1216$ \AA \cite{PartrigePeebles}.
+Physical conditions in those galaxies allow the emergence of a
+Lyman-$\alpha$ line emission at $1216$ \AA \citep{PartridgePeebles}.
 Galaxies detected through its Lyman-$\alpha$ emission receive the name of
-spectra at  and named Lyman Alpha Emitters (LAEs).
+Lyman Alpha Emitters (LAEs).
 
-Currently LAEs are commonly targetted in wide area galaxy surveys.
+Currently LAEs are commonly targeted in wide area galaxy surveys.
 They have been effectively used to study galaxy evolution,
 cosmology and the thermal history of the Universe.
-This has been able through the study of their spatial distribution and
-the shape of the \lya emission line.
+This has been possible through the study of their spatial distribution and
+the shape of their \lya emission line.
 
-Recent imptovements in instrumentation have revolutionized the kind of
+Recent improvements in instrumentation have revolutionized the kind of
 studies that can be performed on LAEs.  
@@
 
-On the ISM's features that plays an important role in shaping the \lya
-is HI kinematics. 
-In a static HI medium the \lya line has  two equal and symmetric peaks
+One of the ISM's features that plays an important role in shaping the \lya
+is \HI kinematics. 
+In a static \HI medium the \lya line ha  two equal and symmetric peaks
 around the natural \lya wavelength and zero intensity at the line's
@@
 If the galaxy rotates, the line shows different amounts of Doppler
-shifts modifying the overall line profile \cite{Garavito14}.
+shifts modifying the overall line profile \citep{Garavito14}.
 
@@
 effects of galaxy outflows and rotation.
-We study a simplified geometrical configuration corresponding to an
+We study a simplified geometrical configuration corresponding to a
 spherical gas cloud with symmetrical radial outflows and a rotation
 profile corresponding to a solid body.
-We base our modelling on a Monte-Carlo radiative transfer code called
+We base our modeling on a Monte-Carlo radiative transfer code called
 CLARA (Code for Lyman Alpha Radiation Analysis) presented for the
-first time in \cite{CLARA}.
-Besides modelling the impact of joint rotation and outflows, we also
+first time by \cite{CLARA}.
+Besides modeling the impact of joint rotation and outflows, we also
 want to check to what extent the analytical model presented by
@@
 outflows.
-The cartesian components take the form
+The cartesian components take the following form:
 
@@
 where $x$, $y$ and $z$ are the cartesian position coordinates with the
-origin at the sphere's center $R$ is the radius of the sphere and the
+origin at the sphere's center, $R$ is the radius of the sphere and the
 direction of the angular velocity vector corresponds to the $\hat{k}$
@@
 the center of the sphere at the \lya\ line's center as they propagate
-through the spere and finally scape.
+through the volume and finally scape.
 We store the final frequency and propagation direction for each photon
@@
 
-In Table \ref{tab:values} we list the combination of $\tau_{\rm{H}}$,
-$v_{rot}$ and $v_{out}$ values used in this paper.
+In Table \ref{tab:values} we list the combination of \tauh,
+\vrot and \vrot values used in this paper.
 The range of values have some overlap with the expectations from a dwarf
@@
 
-\begin{table}[htbp]
+\begin{table}
   \begin{center}
@@
       \hline
-      $\tau_{\mathrm{H}}$ & $v_{rot}$ (\kms) & $v_{out}$ (\kms) \\
+      \tauh & \vrot (\kms) & \vout (\kms) \\
       \hline
@@
       $10^5$ & \{50, 100\} & \{25, 50, 75\} \\
-      $10^6$ & \{50, 100\} & \{25, 50,75\}\\
-      $10^7$ & \{50, 100\} & \{25, 50,75\}\\
+      $10^6$ & \{50, 100\} & \{25, 50, 75\}\\
+      $10^7$ & \{50, 100\} & \{25, 50, 75\}\\
       \hline    
@@
     for a row indicate that all possible combinations in of
-    $\tau_{\rm{H}}$,
-    $v_{rot}$ and $v_{out}$ are used.}
+    \tauh, \vrot and \vout are used.}
   \label{tab:values}
@@
 dependency on the viewing angle, $\theta$, between the line of sight
-  of a distant observer and the rotaional velocity.    
+of a distant observer and the rotational velocity.    
 Second. Line blurring as the rotational velocity increases.
@@
 differential surface element on the sphere Doppler shifts the photons
-that it emmits.
-In this paper we introduce this ansatz by postprocessing the results
+that it emits
+In this paper we introduce this ansatz by post-processing the results
 of the outflow simulations without rotation.
@@
 Figure \ref{fig:varying_rotation} summarizes the information.
-All plots correspond to the same outflow velocity of $v_{out}=50$\kms,
+All plots correspond to the same outflow velocity of $\vout=50$\kms,
 rows have constant optical depth and columns have constant rotational
@@
 % |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100, 
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.48\textwidth]{./figures/results/varying_rotation}
+	\end{center}
+	\caption{\textbf{Caption:} Caption.
+	\label{fig:varying_rotation}}
+\end{figure}
 
@@
 time all plots correspond to the same rotation velocity of
-$v_{rot}=50$\kms,
+$\vrot=50$\kms.
 
@@
 % |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75, 
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.48\textwidth]{./figures/results/varying_outflow}
+	\end{center}
+	\caption{\textbf{Caption:} Caption.
+		\label{fig:varying_outflow}}
+\end{figure}
 
@@
 \section{A new LAE model}
-\label{sec:newmodel}
 
@@
 \section{Results}
-\label{sec:results}
 
@@
 \section{Observational Implications}
-\label{sec:observationalimplications}
 
@@
 \section{Conclusions}
-\label{sec:conclusions}
 

### 2017-10-13 paper/paper.tex
making a grid plot of doppler shift

@@
 \newcommand{\tauh}{$\tau_{\mathrm{H}}$}
+\newcommand{\vth}{\ifmmode\mathrm v_{\mathrm{th}}\else $v_{\mathrm{th}}$~\fi}
 
@@
 
+We know that rotation alone does have an impact on the \lya line morphology
+and we get the same influence in it that previous works. Together with the 
+outflows, we found that the resulting spectra is equivalent to a non-rotating
+galaxy with outflows that undergoes a Doppler Shift in velocity, weighted by 
+amount of photons that if rotating would project their escape in $\hat{k}$'s
+direction. The formulas to calculate this shift are:
+
+\begin{equation}
+V' = V + \vec{v}_{\rm rot}\cdot\hat{k} = V + V_{\rm rot} (x k_{x} - y k_{y})
+\label{eq:shift_x}
+\end{equation}
+
+where $V'$ is the new Doppler shifted velocity, $V$ is the velocity caused by outflows and $\vec{v}_{\rm rot} = V_{\rm rot} (x\hat{i} - y\hat{j})$.
+
+Figure \ref{fig:doppler_shift} displays this Doppler shift for four different configurations with a fixed optical depth \tauh.
+In each cell of the grid there is a \lya spectrum for a galaxy with only outflows and two \lya spectra for the galaxy with outflows and rotation, one obtained from the radiative transfer and the other by the Doppler shifting of the non-rotating galaxy.
+
+% Esta figura deberia tener 3 x 2 viñetas. tau=10E5.
+% cada panel tiene una curva para no rotation, rotation (radiative transfer)
+% y rotation (doppler)
+% |vrot=50, vout=5,  | vrot=100, vout=5,  |
+% |vrot=50, vout=25, | vrot=100, vout=25, |
+% |vrot=50, vout=50, | vrot=100, vout=50, |
+\begin{figure}
+	\begin{center}
+		\includegraphics[width=0.48\textwidth]{./figures/results/doppler_shift}
+	\end{center}
+	\caption{\textbf{Caption:} Caption.
+		\label{fig:doppler_shift}}
+\end{figure}
+
 \section{Discussion}

### 2017-10-14 plots/paper/emcee_skewed_gaussian.py
emcee

@@ -1,0 +1,159 @@
+# Hala "import numpy as np"
+import numpy as np, emcee, os, corner
+from scipy.interpolate import interp1d
+from scipy.special import erf
+
+#FUNCTIONS
+
+#Bifurcation index
+def find_bifurcation_index(intensities):
+    i_prev = 0
+    bif_index = -1
+
+    intensities_length = len(intensities)
+    start_index = 0.3*intensities_length
+    end_index = 0.7*intensities_length
+
+    intensities = intensities[start_index:end_index]
+
+    for i in range(len(intensities)-1):
+        i_curr = intensities[i]
+        i_next = intensities[i+1]
+        if(i_prev >= i_curr and i_curr<=i_next):
+            bif_index = i
+        i_prev = i_curr
+
+    if (bif_index==-1):
+        return intensities_length/2
+    else:
+        return bif_index+start_index
+
+#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
+def gaussian(x, amplitude, center, sigma):
+    """1 dimensional gaussian:
+    gaussian(x, amplitude, center, sigma)
+    """
+    return (amplitude/(np.sqrt(2*np.pi)*sigma)) * np.exp(-(1.0*x-center)**2 /(2*sigma**2))
+
+#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
+def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos):
+    """Gaussian, skewed with error function, equal to
+     gaussian(x, center, sigma)*(1+erf(beta*(x-center)))
+    with beta = gamma/(sigma*sqrt(2))
+    with  gamma < 0:  tail to low value of centroid
+          gamma > 0:  tail to high value of centroid
+    see http://en.wikipedia.org/wiki/Skew_normal_distribution
+    """
+    asym_neg = 1 + erf(gamma_neg*(x-center_neg)/(np.sqrt(2.0)*sigma_neg))
+    g_neg = asym_neg * gaussian(x, amplitude_neg, center_neg, sigma_neg)
+
[…]

### 2017-10-16 paper/paper.tex
change figsize
minor updates

@@
 In this paper we introduce first our theoretical tools and assumptions
-in Section \ref{sec:theory}, then we present the numerical results and
-comparisons against the analytical solution in \ref{sec:results}.
+in Section \ref{sec:theory}. We continue in Section  \ref{sec:results}
+with the results from the Monte-Carlo simulation, the comparison
+against the semi-analytical approximation which we use to make a
+thorough exploration of the effect of rotation.
 In Section \ref{sec:discussion} we discuss our results and their
@@
 % |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100, 
-\begin{figure}
+\begin{figure*}
 	\begin{center}
-		\includegraphics[width=0.48\textwidth]{./figures/results/varying_rotation}
+		\includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation}
 	\end{center}
@@
 	\label{fig:varying_rotation}}
-\end{figure}
+\end{figure*}
 
@@
 % |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75, 
-\begin{figure}
+\begin{figure*}
 	\begin{center}
-		\includegraphics[width=0.48\textwidth]{./figures/results/varying_outflow}
+		\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow}
 	\end{center}
@@
 		\label{fig:varying_outflow}}
-\end{figure}
+\end{figure*}
 
@@
 % |vrot=50, vout=50, | vrot=100, vout=50, |
-\begin{figure}
+\begin{figure*}
   \begin{center}
-		\includegraphics[width=0.48\textwidth]{./figures/results/doppler_shift}
+    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift}
   \end{center}
@@
     \label{fig:doppler_shift}}
-\end{figure}
+\end{figure*}
 

### 2017-10-19 data/tau10E6/to_join_parts_p.py
data for tau 6 and 7 and vout 5

@@ -1,7 +1,7 @@
 import os
 
 logtau = 6
-vrots = [100]
+vrots = [50, 100]
 vouts = [5]
 
 for vrot in vrots:

### 2017-10-19 data/tau10E6/vrot50/vout5/tau10E6_vrot50_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E6_vrot50_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=12
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 12 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E6/vrot50/vout5/tau10E6_vrot50_vout5.input

### 2017-10-19 data/tau10E7/vrot100/vout5/tau10E7_vrot100_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E7_vrot100_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=24
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 24 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E7/vrot100/vout5/tau10E7_vrot100_vout5.input

### 2017-10-19 data/tau10E7/vrot50/vout5/tau10E7_vrot50_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E7_vrot50_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=24
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /lustre/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 24 ./mine.x /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E7/vrot50/vout5/tau10E7_vrot50_vout5.input

### 2017-10-19 paper/paper.tex
plots for line characterization
doppler line stats in paper

@@
 
-where $V'$ is the new Doppler shifted velocity, $V$ is the velocity caused by outflows and $\vec{v}_{\rm rot} = V_{\rm rot} (x\hat{i} - y\hat{j})$.
+where $V'$ is the new Doppler shifted velocity, $V$ is the velocity caused by 
+outflows and $\vec{v}_{\rm rot} = V_{\rm rot} (x\hat{i} - y\hat{j})$.
 
-Figure \ref{fig:doppler_shift} displays this Doppler shift for four different configurations with a fixed optical depth \tauh.
-In each cell of the grid there is a \lya spectrum for a galaxy with only outflows and two \lya spectra for the galaxy with outflows and rotation, one obtained from the radiative transfer and the other by the Doppler shifting of the non-rotating galaxy.
+Figure \ref{fig:doppler_shift} displays this Doppler shift for four different 
+configurations with a fixed optical depth \tauh. In each cell of the grid 
+there is a \lya spectrum for a galaxy with only outflows and two \lya spectra 
+for the galaxy with outflows and rotation, one obtained from the radiative 
+transfer and the other by the Doppler shifting of the non-rotating galaxy.
 
-% Esta figura deberia tener 3 x 2 viñetas. tau=10E5.
+% Esta figura deberia tener 3 x 2 viñetas. tau=10E6.
 % cada panel tiene una curva para no rotation, rotation (radiative transfer)
@@
   \begin{center}
-    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift}
+    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
   \end{center}
-  \caption{\textbf{Caption:} Caption.
+  \caption{\textbf{Caption:} Caption. Tau 10E6.
     \label{fig:doppler_shift}}
 \end{figure*}
+
+In order to compare the line creative with the radiative transfer process that 
+involves the rotation (RT) and the line with only outflows that is Doppler 
+shifted (DS), we calculate three main characteristics for each spectrum. We 
+measure the line's mean velocity, standard deviation and skewness for both 
+RT and DS approaches. We plot the results for each value in Figures 
+\ref{fig:mean_line}, \ref{fig:standard_deviation} and \ref{fig:skewness}, 
+respectively for the same \tauh as Figure \ref{fig:doppler_shift}.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/results/mean_line_6}
+	\end{center}
+	\caption{\textbf{Caption:} The dots represent the Radiative Transfer and
+		the lines represent the Doppler Shift. Tau 10E6.
+		\label{fig:mean_line}}
+\end{figure*}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/results/standard_deviation_6}
+	\end{center}
+	\caption{\textbf{Caption:} The dots represent the Radiative Transfer and 
+		the lines represent the Doppler Shift. Tau 10E6.
+		\label{fig:standard_deviation}}
+\end{figure*}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/results/skewness_6}
+	\end{center}
+	\caption{\textbf{Caption:} The dots represent the Radiative Transfer and 
+		the lines represent the Doppler Shift. Tau 10E6.
+		\label{fig:skewness}}
+\end{figure*}
+
+This shows that the DS model can reproduce the main characteristics of the 
+RT one. Because of this, we can introduce rotation effects in any spectra 
+without running the radiative transfer simulations, that usually take a 
+significant amount of time and computational resources.
 

### 2017-11-02 paper/paper.tex
changing style of present plots, about to add new

@@
 \newcommand{\vout}{\ifmmode\mathrm v_{\mathrm{out}}\else $v_{\mathrm{out}}$~\fi}
-\newcommand{\tauh}{$\tau_{\mathrm{H}}$}
+\newcommand{\tauh}{\ifmmode\mathrm \tau_{\mathrm{H}}\else $\tau_{\mathrm{H}}$~\fi}
 \newcommand{\vth}{\ifmmode\mathrm v_{\mathrm{th}}\else $v_{\mathrm{th}}$~\fi}
@@
 velocity at fixed values of the optical depth and outflow velocity.
-Figure \ref{fig:varying_rotation} summarizes the information.
+Figure \ref{fig:varying_rotation_small} summarizes the information.
 All plots correspond to the same outflow velocity of $\vout=50$\kms,
@@
 
-% Esta figura deberia tener 3 x 3 vinetas. vout=50.
+% Esta figura deberia tener 2 x 2 vinetas. vout=50.
 % cada panel tiene una curva para theta=0, 45, 90.
-% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100, 
-% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100, 
-% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100, 
+% | tau=1E5, vrot=50, | tau=1E5, vrot=100 |
+% | tau=1E6, vrot=50, | tau=1E6, vrot=100 | 
 \begin{figure*}
 	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation}
+		\includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
 	\end{center}
-	\caption{\textbf{Caption:} Caption.
-	\label{fig:varying_rotation}}
+	\caption{\textbf{Spectra for fixed outflow velocity:} We fixed $\vout=50$\kms. 
+			 We vary \vrot increasing from left to right, \tauh increasing from top to 
+			 bottom, and $\theta$ according to the legend.
+	\label{fig:varying_rotation_small}}
 \end{figure*}
@@
 
-% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
+% Esta figura deberia tener 2 x 2 vinetas. vrot=50.
 % cada panel tiene una curva para theta=0, 45, 90.
-% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
-% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
-% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75, 
+% | tau=1E5, vout=25, | tau=1E5, vout=75 |
+% | tau=1E6, vout=25, | tau=1E6, vout=75 |
 \begin{figure*}
 	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow}
+		\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow_small}
 	\end{center}
-	\caption{\textbf{Caption:} Caption.
-		\label{fig:varying_outflow}}
+	\caption{\textbf{Spectra for fixed rotational velocity:} We fixed $\vrot=50$\kms. 
+			 We vary \vout increasing from left to right, \tauh increasing from top to 
+			 bottom, and $\theta$ according to the legend.
+		\label{fig:varying_outflow_small}}
 \end{figure*}
@@
   \end{center}
-  \caption{\textbf{Caption:} Caption. Tau 10E6.
+  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=?$. 
+  	We vary \vrot increasing from left to right and \vout increasing from top to 
+  	bottom. The legend represents the \lya line obtained if there is no rotation 
+  	(in thin line), if there is a radiative transfer of rotation and outflows 
+  	(thick and clear line), and if there is a radiative transfer of only outflows,
+  	but also a Doppler shift from the rotational velocity (thick and dark line). 
 	    \label{fig:doppler_shift}}
@@
 
-In order to compare the line creative with the radiative transfer process that 
+In order to compare the line created with the radiative transfer process that 
 involves the rotation (RT) and the line with only outflows that is Doppler 
@@
 measure the line's mean velocity, standard deviation and skewness for both 
-RT and DS approaches. We plot the results for each value in Figures 
+RT's and DS's approaches. We plot the results for each value in Figures 
 \ref{fig:mean_line}, \ref{fig:standard_deviation} and \ref{fig:skewness}, 
@@
 	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/results/mean_line_6}
+		\includegraphics[width=0.65\textwidth]{./figures/results/mean_line_7}
 	\end{center}
-	\caption{\textbf{Caption:} The dots represent the Radiative Transfer and
-		the lines represent the Doppler Shift. Tau 10E6.
+	\caption{\textbf{Mean line for RT and DS:} The dots represent the Radiative 
+		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
 		\label{fig:mean_line}}
@@
 	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/results/standard_deviation_6}
+		\includegraphics[width=0.65\textwidth]{./figures/results/standard_deviation_7}
 	\end{center}
-	\caption{\textbf{Caption:} The dots represent the Radiative Transfer and 
-		the lines represent the Doppler Shift. Tau 10E6.
+	\caption{\textbf{Standard Deviation for RT and DS:} The dots represent the Radiative 
+		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
 		\label{fig:standard_deviation}}
@@
 	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/results/skewness_6}
+		\includegraphics[width=0.65\textwidth]{./figures/results/skewness_7}
 	\end{center}
-	\caption{\textbf{Caption:} The dots represent the Radiative Transfer and 
-		the lines represent the Doppler Shift. Tau 10E6.
+	\caption{\textbf{Skewness for RT and DS:} The dots represent the Radiative 
+		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
 		\label{fig:skewness}}
@@
 
+In addition to this, we also evaluate the effects of the viewing angle
+for fixed rotational and outflows velocity. We measure the intensity of the 
+valley for each $\theta$ and \tauh. Figure \ref{fig:valley_intensity} 
+summarizes the information with all plots corresponding to the same outflow
+velocity of $\vout=75$\kms and rotation velocity of $\vrot=50$\kms.
+
+% Esta figura deberia tener en el eje x theta y en el eje y intensity of valley
+% cada panel tiene una curva para logtau = 5, 6, 7.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.65\textwidth]{./figures/results/valley_intensity}
+	\end{center}
+	\caption{\textbf{Caption:} We fixed $\vout=75$\kms and $\vrot=50$\kms.
+		\label{fig:valley_intensity}}
+\end{figure*}
+
 \section{Discussion}
@@
 \subsection{Theoretical Insights}
+Talk about radiative transfer...
+
+Regarding de Doppler shift... By this logic then, the quality and accuracy of 
+the spectrum, obtained by a Doppler shift, increases with the optical depth \tauh. 
+This because the more \tauh, the more anisotropic is the photons' distribution 
+at the galaxy's border.
+
+The introduction of rotation in the model implies that it will only affect 
+galaxies with its viewing angle
 
 \subsection{Application to observational data}
+
+Kulas...
+
+MUSE...
 
@@
 \bibliography{references}
+
+\appendix
+
+\section{Additional figures}
+\label{sec:appendix}
+
+In this appendix we show Figures \ref{fig:varying_rotation} and \ref{fig:varying_outflow} with more combinations of \vrot and \vout.
+
+% Esta figura deberia tener 3 x 3 vinetas. vout=50.
+% cada panel tiene una curva para theta=0, 45, 90.
+% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100, 
+% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100, 
+% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100, 
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation}
+	\end{center}
+	\caption{\textbf{Caption:} $\vout=50$\kms.
+		\label{fig:varying_rotation}}
+\end{figure*}
+
+% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
+% cada panel tiene una curva para theta=0, 45, 90.
+% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
+% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
+% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75, 
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow}
+	\end{center}
+	\caption{\textbf{Caption:} $\vrot=50$\kms.
+		\label{fig:varying_outflow}}
+\end{figure*}
 

### 2017-11-03 paper/paper.tex
new plot theta

@@
 for fixed rotational and outflows velocity. We measure the intensity of the 
-valley for each $\theta$ and \tauh. Figure \ref{fig:valley_intensity} 
+valley between peaks, for each $\theta$ and \tauh. Figure \ref{fig:valley_intensity} 
 summarizes the information with all plots corresponding to the same outflow
@@
 % cada panel tiene una curva para logtau = 5, 6, 7.
-
 \begin{figure*}
@@
 
-Regarding de Doppler shift... By this logic then, the quality and accuracy of 
+Regarding de Doppler shift... Then, the quality and accuracy of 
 the spectrum, obtained by a Doppler shift, increases with the optical depth \tauh. 
-This because the more \tauh, the more anisotropic is the photons' distribution 
+This is because the more \tauh, the more anisotropic is the photons' distribution 
 at the galaxy's border.
@@
 The introduction of rotation in the model implies that it will only affect 
-galaxies with its viewing angle
+galaxies with its viewing angle...
 

### 2017-11-08 paper/paper.tex
adding reference, missing text
missing kulas, doppler explanation and more conclusions

@@
 \ref{fig:mean_line}, \ref{fig:standard_deviation} and \ref{fig:skewness}, 
-respectively for the same \tauh as Figure \ref{fig:doppler_shift}.
+respectively for $\tauh = 10^7$. We highlight that these 3 plots increase
+their fit tendency with a higher optical depth. The reasons why will be 
+discussed in section \ref{sec:discussion}.
 
@@
 \subsection{Theoretical Insights}
-Talk about radiative transfer...
+\color{red}
+Why Doppler Shift works. Talk about radiative transfer...
+\color{black}
 
-Regarding de Doppler shift... Then, the quality and accuracy of 
-the spectrum, obtained by a Doppler shift, increases with the optical depth \tauh. 
-This is because the more \tauh, the more anisotropic is the photons' distribution 
-at the galaxy's border.
+A first approach to an analytical expression that returns the \lya spectrum 
+from a rotating galaxy is presented by \cite{Garavito14}. This derivation is 
+based on the assumption that the distribution of photons' propagation directions
+at the edge of the galaxy is anisotropic and depends of \tauh only. So this 
+approximation becomes more accurate with higher optical depth and it is also the 
+reason why the Doppler Shift technique to induce rotational effects to an outflow
+\lya spectrum works.
 
-The introduction of rotation in the model implies that it will only affect 
-galaxies with its viewing angle...
+On the other hand, regarding the parameters \vrot, \vout, and $\theta$ for
+which this model would work, we have the following. The effect of the outflows 
+velocity is much more sensitive on the spectrum than the rotational one; meaning 
+that for a high \vout, the \vrot value would have to be equal or greater in order 
+to be able to glimpse the shift in frequency that rotation induces. Regarding $\theta$, 
+the introduction of rotation in the model implies that it will only affect galaxies 
+with its viewing angle $\theta \neq 90^\circ$. This can be observed in the cartesian 
+components of the velocity of the atoms (Eq. \ref{eq:vx}, \ref{eq:vy}, \ref{eq:vz})
+where the only velocity factor affecting $v_z$ is the outflows one.
 
@@
 
+We find three different approaches how this model can have observational
+implications. Firstly, as the resolution of astronomical instruments is 
+increasing, new images from LAEs can be spatially resolved in a way that 
+rotation velocities of the galaxy can be determined. New spectrographs like 
+MUSE should help obtain kinematic information from LAEs by using our model.  
+As for now, previous works in the literatures have already found useful data
+to support this idea. Fig 7 of Prescott et al. \cite{Prescott14} shows the
+presence of Doppler shift and \cite{Herenz2016} creates 2D line of sight 
+velocity maps of several LARS (Lyman Alpha Reference Sample) galaxies.
+
+Another application of this new model to the current interpretation of \lya
+spectra, is that the central ($V=0$) emission of the spectra seen in the 
+\lya line, is a consequence of the viewing angle of the galaxy and can be 
+controlled by it. Several authors have suggested that this central emission 
+is caused by radiation that escapes the galaxy without scattering. So we 
+present an alternative that solves this issue. In addition, the effect of
+outflows velocities \vrot in current models can modify the asymmetry of peaks
+present in the \lya line. However the intensity of the valley and the width 
+of the line were not easily reproduced. In this work we provide a tool to
+provide these effects with values of \vout and \vrot that do not exceed the
+typical LAEs' numbers found in the literature.
+
+\color{red}
 Kulas...
-
-MUSE...
+\color{black}
+Finally, we present a reconstruction of a selected \lya spectrum presented
+by \cite{Kulas12} and reproduced with the permission of the author. We create
+a fit of the \lya line and predict its kinematic properties.
 
@@
 \label{sec:conclusions}
+
+In this paper we create, for the first time, a computational model for LAEs 
+that emulate the kinematic effects of galaxy rotation and outflows in the \lya
+spectral line.
+
+\color{red}
+Complete conclusions...
+\color{black}
 
@@
 
-In this appendix we show Figures \ref{fig:varying_rotation} and \ref{fig:varying_outflow} with more combinations of \vrot and \vout.
+In this appendix we show more \lya spectra in the Figures 
+\ref{fig:varying_rotation} and \ref{fig:varying_outflow} 
+with more combinations of \vrot and \vout.
 

### 2017-11-08 plots/paper/emcee_skewed_gaussian.py
removing 2 files

@@ -1,159 +1,0 @@
-# Hala "import numpy as np"
-import numpy as np, emcee, os, corner
-from scipy.interpolate import interp1d
-from scipy.special import erf
-
-#FUNCTIONS
-
-#Bifurcation index
-def find_bifurcation_index(intensities):
-    i_prev = 0
-    bif_index = -1
-
-    intensities_length = len(intensities)
-    start_index = 0.3*intensities_length
-    end_index = 0.7*intensities_length
-
-    intensities = intensities[start_index:end_index]
-
-    for i in range(len(intensities)-1):
-        i_curr = intensities[i]
-        i_next = intensities[i+1]
-        if(i_prev >= i_curr and i_curr<=i_next):
-            bif_index = i
-        i_prev = i_curr
-
-    if (bif_index==-1):
-        return intensities_length/2
-    else:
-        return bif_index+start_index
-
-#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
-def gaussian(x, amplitude, center, sigma):
-    """1 dimensional gaussian:
-    gaussian(x, amplitude, center, sigma)
-    """
-    return (amplitude/(np.sqrt(2*np.pi)*sigma)) * np.exp(-(1.0*x-center)**2 /(2*sigma**2))
-
-#Based on the definition in: https://github.com/lmfit/lmfit-py/blob/master/lmfit/lineshapes.py
-def sum_2_skewed_gaussian(x, amplitude_neg, center_neg, sigma_neg, gamma_neg, amplitude_pos, center_pos, sigma_pos, gamma_pos):
-    """Gaussian, skewed with error function, equal to
-     gaussian(x, center, sigma)*(1+erf(beta*(x-center)))
-    with beta = gamma/(sigma*sqrt(2))
-    with  gamma < 0:  tail to low value of centroid
-          gamma > 0:  tail to high value of centroid
-    see http://en.wikipedia.org/wiki/Skew_normal_distribution
-    """
-    asym_neg = 1 + erf(gamma_neg*(x-center_neg)/(np.sqrt(2.0)*sigma_neg))
-    g_neg = asym_neg * gaussian(x, amplitude_neg, center_neg, sigma_neg)
-
[…]

### 2017-11-08 paper/paper.tex
paper updated

@@
 
-\newcommand{\HII}{{\text{H\MakeUppercase{\romannumeral 2}}} }
-\newcommand{\cm}{\ifmmode{{\rm cm}}\else cm\fi}
-\newcommand{\ccm}{\,\mathrm{cm}^{-3}}
-\newcommand{\ergps}{\,{\rm erg}\,{\rm s}\ifmmode{}^{-1}\else ${}^{-1}$\fi}
-\newcommand{\Mpch}{\,{\rm Mpc}\,\ifmmode h^{-1}\else $h^{-1}$\fi}
-\newcommand{\dd}{\mathrm{d}}
-\newcommand{\vek}[1]{\bm{#1}}
-\newcommand{\hb}{H$\beta$}
-\newcommand{\ha}{H$\alpha$}
-\newcommand{\oiii}{[OIII]}
-\newcommand{\oii}{[OII]}
-\newcommand{\nii}{[NII]}
-\newcommand{\esca}{erg cm$^{-2}$ s$^{-1}$ \AA$^{-1}$}
-\newcommand{\esc}{erg cm$^{-2}$ s$^{-1}$}
-\newcommand{\es}{erg s$^{-1}$}
-\newcommand{\esa}{erg s$^{-1}$}
-\newcommand{\sigmaclump}{$54.3\pm 0.6$ km s$^{-1}$}
-\newcommand{\inftyclump}{$54.3\pm 5.1$ km s$^{-1}$}
-\newcommand{\probaclump}{$0.96\pm 0.01$}
-
 \begin{document}
@@
 We perform Monte Carlo radiative transfer simulations of outflowing
-gas with additional solid body rotation.
-We reproduce the main features of the outcoming spectra with a simple
-semi-analytic model that modifies the spectra of outflows without
-rotation by adding a Doppler shift with the rotational information.
+gas with additional solid body rotation. We reproduce the main features 
+of the outcoming \lya line spectra with a simple semi-analytic model 
+that modifies the spectra of outflows without rotation by adding a 
+Doppler shift with the rotational information.
 

### 2017-11-14 paper/paper.tex
some edits in the itro

@@
 Currently LAEs are commonly targeted in wide area galaxy surveys.
+The conditions that allow for the emergence of the lYa line is that
+these galaxies are star forming with a low dust content. 
 They have been effectively used to study galaxy evolution,
@@
 
-One of the ISM's features that plays an important role in shaping the \lya
-is \HI kinematics. 
-In a static \HI medium the \lya line ha  two equal and symmetric peaks
-around the natural \lya wavelength and zero intensity at the line's
-center.
-For an outflowing ISM, the line becomes
-asymmetrical with a more pronounced read peak.
-If the galaxy rotates, the line shows different amounts of Doppler
-shifts modifying the overall line profile \citep{Garavito14}.
+These kinematic maps show that lyman alpha emitting galaxies have two 
+overimposed velocity patterns: coherent rotation plus and a random
+motions around the mean value provided by rotation. 
+Depending on the system under study, one of the two features can be
+more important, or they can be found in equal measure. 
+These kinematic features play an important role in shaping the \lya.
 
-In this paper we present for the first time a study of the joint
-effects of galaxy outflows and rotation.
+Models of the \lya have also shown that there is clear signal for
+outflowing gas most probably produced by star formation activity.
+The characteristic signal for these outflows is the asymmetrical \lya
+linea with a more pronounced read peak.
+
+The effects of either outflows or rotation on the \lya line have been
+already explored.
+Here e present for the first time a study of the joint effects of
+galaxy outflows and rotation, two of the three major kinematic
+components expected in LAEs.
+
 We study a simplified geometrical configuration corresponding to a
@@
 first time by \cite{CLARA}.
+
 Besides modeling the impact of joint rotation and outflows, we also
-want to check to what extent the analytical model presented by
+check to what extent the analytical model presented by
 \cite{Garavito14} to explain the effects of rotation can also be
 applied in our case.
+We show that solid body rotation effects can be included
+in the results of other radiative transfer model in the case of large
+optical depths. In this case it is a good approximation to Doppler
+boost the results of the model without rotation.
 
-In this paper we introduce first our theoretical tools and assumptions
+The structure of the paper is the following. 
+We introduce first our theoretical tools and assumptions
 in Section \ref{sec:theory}. We continue in Section  \ref{sec:results}

### 2017-11-21 paper/paper.tex
doppler shift in discussion and rough fit to kulas

@@
 \color{black}
+In Fig.
+\ref{fig:doppler} we present the spectrum pf a LAE taken from to different sides of the galaxy.
+As the LAE is rotating, one side is being redshifted while the other is blueshifted.
+We see that the combined spectrum is a weighted line, in black, that combines these past two.
+This shows that there is in effect a Doppler Shift and justifies our choice to apply it after the outflows spectrum is obtained.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.55\textwidth]{./figures/discussion/doppler}
+	\end{center}
+	\caption{\textbf{Caption:} $\vout=25$\kms, $\vrot=100$\kms, $\tauh=10^5$.
+		\label{fig:doppler}}
+\end{figure*}
 
@@
 
-\color{red}
-Complete conclusions...
-\color{black}
+Our main conclusions are:
+
+\begin{itemize}
+	\item The effect of rotation \vrot in a LAE spectrum is to broaden the 
+	\lya line.
+
+	\item The effect of outflows \vout in a LAE spectrum is to increase the peak 
+	asymmetry of the \lya line.
+
+	\item The effect of optical depth \tauh in a LAE spectrum is to induces a 
+	redshift around the zero velocity in the \lya line.
+
+	\item Rotation induces a dependency in viewing angle on the \lya line. The 
+	closer the observer is to the equator of the galaxy, the higher is the central 
+	valley of the frequency distribution.
+
+	\item Rotations induces a Doppler Shift in frequency of the only-outflow 
+	spectrum of the LAE.
+
+	\item The final spectra obtained are roughly consistent with LAEs 
+	observations. 
+\end{itemize}
 

### 2018-02-14 paper/paper.tex
edit

@@
 
-Distant galaxies are key to understand early evolutionary stages of
-our Universe.
-Physical conditions in those galaxies allow the emergence of a
-Lyman-$\alpha$ line emission at $1216$ \AA \citep{PartridgePeebles}.
-Galaxies detected through its Lyman-$\alpha$ emission receive the name of
-Lyman Alpha Emitters (LAEs).
-
-Currently LAEs are commonly targeted in wide area galaxy surveys.
-The conditions that allow for the emergence of the lYa line is that
-these galaxies are star forming with a low dust content. 
-They have been effectively used to study galaxy evolution,
-cosmology and the thermal history of the Universe.
-This has been possible through the study of their spatial distribution and
-the shape of their \lya emission line.
-
+The interpretation of the Lyman-$\alpha$ emission in galaxies is
+key to understand their evolutionary processes \citep{PartridgePeebles}.
 Recent improvements in instrumentation have revolutionized the kind of
-studies that can be performed on LAEs.  
-It is now possible to infer detailed kinematic maps for nearby galaxies.
+studies that can be performed on Lyman-$\alpha$ emitting galaxies
+(LAEs.)
+For instance, it is now possible to infer detailed kinematic maps for
+nearby galaxies. 
 The study of these maps would allow us to build data-driven models to

### 2018-02-14 paper/paper.tex
edit

@@
 %=========================================================================
-\title[Outflows and rotation in LAEs]{Lyman-alpha spectra from rotating outflows}
+\title[Outflows and rotation in LAEs]{Lyman-alpha spectra from
+  rotating outflows} 
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
@@
 \begin{abstract}
+Outflows and rotation are two ubiquitous kinematic features in the gas
+distributions of galaxies.
 We perform Monte Carlo radiative transfer simulations of outflowing
-gas with additional solid body rotation. We reproduce the main features 
-of the outcoming \lya line spectra with a simple semi-analytic model 
-that modifies the spectra of outflows without rotation by adding a 
-Doppler shift with the rotational information.
-
+gas with additional solid body rotation to understand how these two
+features impact the morphology of the Lyman-$\alpha$ emission line.
+One of the most notorious results is that rotation increases the
+detected flux at the line's center in comparison with the
+non-rotational case. 
+Rotation also induces a line broadening and a dependency with the
+viewing angle.
+We present a semi-analytic model that modifies the spectra of
+non-rotational that manages to reproduce the main features of these
+Monte Carlo simulations. 
 \end{abstract}

### 2018-03-09 paper/paper.tex
edit

@@
 distributions of galaxies.
-We perform Monte Carlo radiative transfer simulations of outflowing
+Here we perform Monte Carlo radiative transfer simulations of outflowing
 gas with additional solid body rotation to understand how these two
@@
 Monte Carlo simulations. 
+As an application example we contraint the rotation parameters to the
+observed Lyman-$\alpha$ profile of a high redshift LAE, providing the
+first a constraint on the rotational velocity of a high redshift
+galaxy based on its integrated \lya emission.
 \end{abstract}
@@
 The interpretation of the Lyman-$\alpha$ emission in galaxies is
-key to understand their evolutionary processes \citep{PartridgePeebles}.
+key to understand their evolutionary processes, specially in
+star-forming, low-dust galaxies \citep{PartridgePeebles}.
 Recent improvements in instrumentation have revolutionized the kind of
@@
 
-These kinematic maps show that lyman alpha emitting galaxies have two 
-overimposed velocity patterns: coherent rotation plus and a random
-motions around the mean value provided by rotation. 
+These kinematic maps show that LAEs have two  overimposed velocity
+patterns: coherent rotation plus and a random motions around the mean
+value provided by rotation.  
 Depending on the system under study, one of the two features can be
@@
 
-We know that rotation alone does have an impact on the \lya line morphology
-and we get the same influence in it that previous works. Together with the 
-outflows, we found that the resulting spectra is equivalent to a non-rotating
-galaxy with outflows that undergoes a Doppler Shift in velocity, weighted by 
-amount of photons that if rotating would project their escape in $\hat{k}$'s
-direction. The formulas to calculate this shift are:
+Rotation alone does have an impact on the \lya line morphology
+and we get the same influence in it that previous works. 
+Together with the  outflows, we found that the resulting spectra is
+equivalent to a non-rotating galaxy with outflows that undergoes a
+Doppler Shift in velocity, weighted by  amount of photons that if
+rotating would project their escape in $\hat{k}$'s direction:
 
 \begin{equation}
-V' = V + \vec{v}_{\rm rot}\cdot\hat{k} = V + V_{\rm rot} (x k_{x} - y k_{y})
+V' = V + \vec{v}_{\rm rot}\cdot\hat{k} 
 \label{eq:shift_x}
 \end{equation}
-
+%
 where $V'$ is the new Doppler shifted velocity, $V$ is the velocity caused by 
-outflows and $\vec{v}_{\rm rot} = V_{\rm rot} (x\hat{i} - y\hat{j})$.
+outflows.
 

### 2018-03-13 paper/paper.tex
fixing paper plots

@@
 velocity at fixed values of the optical depth and outflow velocity.
-Figure \ref{fig:varying_rotation_small} summarizes the information.
+Figure \ref{fig:varying_rotation_small} summarizes its behavior 
+(for largest range see Figure \ref{fig:varying_rotation}).
 All plots correspond to the same outflow velocity of $\vout=50$\kms,
@@
 values for the optical depth and rotational velocity.
-Figure \ref{fig:varying_outflow} summarizes the information following
-the same patters as Figure \ref{fig:varying_rotation}, only that this
-time all plots correspond to the same rotation velocity of
+Figure \ref{fig:varying_outflow_small} summarizes its behavior
+(for largest range see Figure \ref{fig:varying_outflow}).
+All plots correspond to the same rotation velocity of
 $\vrot=50$\kms.
@@
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.65\textwidth]{./figures/results/mean_line_7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_mn_vout5_logtau7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_mn_vout25_logtau7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_mn_vout50_logtau7}
 	\end{center}
@@
 		\label{fig:mean_line}}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.65\textwidth]{./figures/results/standard_deviation_7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_std_vout5_logtau7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_std_vout25_logtau7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_std_vout50_logtau7}
 	\end{center}
@@
 		\label{fig:standard_deviation}}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.65\textwidth]{./figures/results/skewness_7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
+		\includegraphics[width=0.15\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
 	\end{center}
@@
 		\label{fig:skewness}}
-\end{figure*}
+\end{figure}
 

### 2018-03-13 plots/kulas/parameters_emcee.py

@@ -1,34 +1,41 @@
 import emcee, corner
 import numpy as np
 import matplotlib.pyplot as plt
-from model import read_data, doppler_shift, filter_by_theta, get_spectra
+from model import read_data, doppler_shift, filter_by_theta, get_spectra, bins_to_x
 
 # CONSTANTS
 
 VROT_MIN = 0.0
-VROT_MAX = 300.0
-DELTA_THETA = 0.05
+VROT_MAX = 600.0
 THETA_MIN = 0.0
-THETA_MAX = np.pi/2.0 - DELTA_THETA
-
+THETA_MAX = np.pi/2.0
+NUM_THETAS = 9
+THETAS_RANGE = np.arccos(np.linspace(1, 0, NUM_THETAS))
 
 # FUNCTIONS
 
-def bins_to_x(b):
-    x = []
-    for i in range(len(b)-1):
-        x.append( (b[i]+b[i+1])/2.0 )
-    x = np.array(x)
-    return x
+def get_theta_range(theta):
+    '''Returns the range of theta in a cosine linear space.'''
+    min_theta = THETAS_RANGE[0]
+    max_theta = THETAS_RANGE[1]
+    for i in range(NUM_THETAS-1):
+        min_theta = THETAS_RANGE[i]
+        max_theta = THETAS_RANGE[i+1]
+        if (theta >= min_theta and theta <= max_theta):
+            break
+    return min_theta, max_theta
 
 def model(data_doppler, vrot, theta):
+    '''Model.'''
     doppler = doppler_shift(data_doppler, vrot=vrot)
-    doppler_theta = filter_by_theta(doppler, min_theta=theta,
-                                    max_theta=theta+DELTA_THETA)
+    min_theta, max_theta = get_theta_range(theta)
+    doppler_theta = filter_by_theta(doppler, min_theta=min_theta,
+                                    max_theta=max_theta)
     b_m, n_m = get_spectra(doppler_theta)
     return b_m, n_m
 
 def get_difference(b_m, n_m, x_d, y_d):
+    '''Use the data resolution to subtract model and data.'''
     # kulas is wider than model
     diff = []
     ind = 0
@@ -46,12 +53,14 @@ def get_difference(b_m, n_m, x_d, y_d):
     return diff
 
 def ln_prior(param):
+    '''Log of prior.'''
     vrot, theta = param
[…]

### 2018-03-13 plots/paper/doppler_shift.c

@@ -1,29 +1,0 @@
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-
-int main() {
-
-  //Number of lines in file
-  static const int num_xs = 11169;
-
-  //The info with v_rot = 0
-  FILE * file;
-  file = fopen("vrot0_vout25.dat" , "r");
-
-  //Pointer to save xs
-  double *xs;
-  xs = malloc(num_xs*sizeof(double));
-
-  if (file) {
-    int i = 0;
-    while (fscanf(file, "%lf", &xs[i]) != EOF) {
-      i++;
-    }
-    fclose(file);
-  }
-
-  printf("%f\n", xs[0]);
-
-  return 0;
-}

### 2018-03-13 plots/paper/model.py

@@ -65,37 +65,33 @@ def bins_to_x(bins):
 def get_spectra(data, v_th=12.86, as_hist=True):
     '''Returns the spectrum in velocity units.'''
     velocity = -data['x_frec']*v_th
-    n, b = np.histogram(velocity, normed=True, bins=50)
+    n, b = np.histogram(velocity, bins=50)
+    n = n / n.sum()
     if (as_hist):
         return b, n
     else:
         return bins_to_x(b), n
 
-def plot_spectra(x, y, alpha=1.0, lw=2, label='label'):
+def plot_spectra(x, y, ax, alpha=1.0, lw=2, label='label'):
     '''Basic plot of spectra.'''
-    plt.figure()
-    plt.hist(x, weights=y, histtype='step', fill=False, normed=True,
-             color='black', linewidth=lw, bins=50, alpha=alpha, label=label)
-    plt.show()
-
-def main():
+    ax.hist(x, weights=y, histtype='step', fill=False, 
+            normed=True, color='black', linewidth=lw, bins=50, 
+            alpha=alpha, label=label)
+    
+def main(vrot, vout, logtau, min_theta, max_theta):
     '''Main. Shows an example of its use.'''
     # Input
     vrot0 = 0
-    vrot = 150
-    vout = 5
-    logtau = 6
-    min_theta = np.pi/2.0 - 0.15
-    max_theta = np.pi/2.0
 
     data = read_data(vrot0, vout, logtau)
     doppler = doppler_shift(data, vrot=vrot)
     doppler_theta = filter_by_theta(doppler, min_theta, max_theta)
-    xs, ys = get_spectra(doppler_theta, as_hist=False)
-
-    plot_spectra(xs, ys)
+    x, y = get_spectra(doppler_theta, as_hist=False)
+    
+    return x, y
 
 
 # MAIN
 if __name__ == "__main__":
-    main()
+    main(vrot=150, vout=5, logtau=6, min_theta=np.pi/2.0-0.15, 
+         max_theta=np.pi/2.0)

### 2018-03-13 plots/kulas/model.py
everything should use model.py

@@ -1,101 +1,0 @@
-import os
-import numpy as np
-import matplotlib.pyplot as plt
-
-def get_unit_r_sphere(data):
-    '''Re-scale positions on the sphere at which the photons escaped.'''
-    r_norm = np.sqrt(data['x']**2 + data['y']**2 + data['z']**2)
-    x_sphere = data['x'] / r_norm
-    y_sphere = data['y'] / r_norm
-    z_sphere = data['z'] / r_norm
-    return x_sphere, y_sphere, z_sphere
-
-def get_atom_velocities(x_unitary, y_unitary, z_unitary, vrot):
-    '''Defines new atom velocities based on vrot (vr).'''
-    v_x = - y_unitary * vrot
-    v_y = x_unitary * vrot
-    v_z = np.zeros(len(z_unitary))
-    return {'x':v_x, 'y':v_y, 'z':v_z}
-
-def doppler_shift(data, vrot=0.0, v_th=12.86):
-    '''Updates old_data adding doppler shifts. the result is stored in new_data.'''
-    new_data = data.copy()
-    new_data['x'], new_data['y'], new_data['z'] = get_unit_r_sphere(data)
-
-    atom_velocities = get_atom_velocities(new_data['x'], new_data['y'], new_data['z'], vrot=vrot)
-
-    new_data['x_frec'] = data['x_frec']
-    new_data['x_frec'] += atom_velocities['x']*data['k_x']/v_th
-    new_data['x_frec'] += atom_velocities['y']*data['k_y']/v_th
-    new_data['x_frec'] += atom_velocities['z']*data['k_z']/v_th
-
-    return new_data
-
-def read_data(vrot=0, vout=5, logtau=6, input_dir="../../data/"):
-    '''Reads the data.'''
-    tau_name = 'tau10E' + str(logtau)
-    vrot_name = 'vrot' + str(vrot)
-    vout_name = 'vout' + str(vout)
-    filename = tau_name + '_' + vrot_name + '_' + vout_name + '_out.ascii'
-    fname = os.path.join(input_dir, tau_name, vrot_name, vout_name, filename)
-
-    dtype = [('x', 'f8'), ('y', 'f8'), ('z', 'f8'),
-             ('k_x', 'f8'), ('k_y', 'f8'), ('k_z', 'f8'),
-             ('x_frec', 'f8'), ('escaped', 'i8'), ('n_scattering', 'i8')]
-
-    df = np.loadtxt(fname, skiprows=1, dtype=dtype)
[…]

### 2018-03-26 paper/paper.tex
ready line stats and muse

--- 2018-03-09 paper/paper.tex
@@
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.65\textwidth]{./figures/results/mean_line_7}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout5_logtau7}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout25_logtau7}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout50_logtau7}
 	\end{center}
-	\caption{\textbf{Mean line for RT and DS:} The dots represent the Radiative 
+	\caption{\textbf{Mean line for RT and DS:} The triangles represent the Radiative 
 		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
 		\label{fig:mean_line}}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.65\textwidth]{./figures/results/standard_deviation_7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout5_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout25_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout50_logtau7}
 	\end{center}
-	\caption{\textbf{Standard Deviation for RT and DS:} The dots represent the Radiative 
+	\caption{\textbf{Standard Deviation for RT and DS:} The triangles represent the Radiative 
 		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
 		\label{fig:standard_deviation}}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.65\textwidth]{./figures/results/skewness_7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout5_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout25_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout50_logtau7}
 	\end{center}
-	\caption{\textbf{Skewness for RT and DS:} The dots represent the Radiative 
+	\caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative 
 		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
 		\label{fig:skewness}}
-\end{figure*}
+\end{figure}
 
@@
 summarizes the information with all plots corresponding to the same outflow
-velocity of $\vout=75$\kms and rotation velocity of $\vrot=50$\kms.
+velocity of $\vout=5$\kms and rotation velocity of $\vrot=100$\kms.
 
@@
 % cada panel tiene una curva para logtau = 5, 6, 7.
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.65\textwidth]{./figures/results/valley_intensity}
+		\includegraphics[height=0.12\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
+		\includegraphics[height=0.12\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
+		\includegraphics[height=0.12\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
 	\end{center}
-	\caption{\textbf{Caption:} We fixed $\vout=75$\kms and $\vrot=50$\kms.
+	\caption{\textbf{Valley Intensity:} We show for each \tauh the dependency that
+		the viewing angle $\theta$ has on the line's the valley intensity. The larger 
+		$\cos{\theta}$ the lower is the valley intensity, so the higher $\theta$, the 
+		higher the valley intensity. We fixed $\vout=5$\kms and $\vrot=100$\kms. 
 		\label{fig:valley_intensity}}
-\end{figure*}
+\end{figure}
 
@@
 \color{black}
-In Fig.
-\ref{fig:doppler} we present the spectrum pf a LAE taken from to different sides of the galaxy.
-As the LAE is rotating, one side is being redshifted while the other is blueshifted.
-We see that the combined spectrum is a weighted line, in black, that combines these past two.
-This shows that there is in effect a Doppler Shift and justifies our choice to apply it after the outflows spectrum is obtained.
+In Fig. \ref{fig:doppler} we present the spectrum of a LAE taken from to different 
+sides of the galaxy. As the LAE is rotating, one side is being redshifted while 
+the other is blueshifted. We see that the combined spectrum is a weighted line, 
+in solid black, that combines these past two. This shows that there is in effect a 
+Doppler Shift and justifies our choice to apply it after the outflows spectrum is 
+obtained.
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-		\includegraphics[width=0.55\textwidth]{./figures/discussion/doppler}
+		\includegraphics[width=0.45\textwidth]{./figures/discussion/doppler}
 	\end{center}
-	\caption{\textbf{Caption:} $\vout=25$\kms, $\vrot=100$\kms, $\tauh=10^5$.
+	\caption{\textbf{Doppler Shift:} We fix $\vout=25$\kms, $\vrot=100$\kms, $\tauh=10^5$.
 		\label{fig:doppler}}
-\end{figure*}
+\end{figure}
 

### 2018-03-26 plots/paper/model.py

@@ -95,7 +95,7 @@ def get_spectra(data, v_th=12.86, as_hist=True):
 def plot_spectra(x, y, ax, alpha=1.0, lw=2, ls='-', c='k', label='label'):
     '''Basic plot of spectra.'''
     ax.hist(x, weights=y, histtype='step', fill=False, 
-            normed=True, color=c, linewidth=lw, bins=50, 
+            color=c, linewidth=lw, bins=50, 
             alpha=alpha, linestyle=ls, label=label)
     
 def main(vrot, vout, logtau, min_theta, max_theta):

### 2018-03-26 paper/paper.tex
minor typos

@@
 already explored.
-Here e present for the first time a study of the joint effects of
+Here we present for the first time a study of the joint effects of
 galaxy outflows and rotation, two of the three major kinematic
@@
 In the current implementation we neglect the influence of dust.
-Our basic models is an spherical distribution of neutral hydrogen,
+Our basic model is an spherical distribution of neutral hydrogen,
 an approximation commonly used in the literature, as it explains a
@@
 the velocity magnitude increasing linearly with the radial
-coordinate; the outflow model is fully characterized by $V_{\rm out}$, the
+coordinate; the outflows model is fully characterized by $V_{\rm out}$, the
 velocity at the sphere's surface.
@@
 galaxy with a total neutral hydrogen mass of $10^8$-$10^9$ $M_{\odot}$.
-We run a total of $24$ different models.
+We run a total of $27$ different models.
 
@@
 In this paper we introduce this ansatz by post-processing the results
-of the outflow simulations without rotation.
+of the outflows simulations without rotation.
 This allows us to produce new \lya spectra and compare them with the
@@
 
-We evaluate first the effects of increasing rotational
-velocity at fixed values of the optical depth and outflow velocity.
+We evaluate the effects of increasing rotational
+velocity at fixed values of the optical depth and outflows velocity.
 Figure \ref{fig:varying_rotation_small} summarizes its behavior 
-(for largest range see Figure \ref{fig:varying_rotation}).
-All plots correspond to the same outflow velocity of $\vout=50$\kms,
+(for a largest range see Figure \ref{fig:varying_rotation}).
+All plots correspond to the same outflows velocity of $\vout=50$\kms,
 rows have constant optical depth and columns have constant rotational
@@
 	\end{center}
-	\caption{\textbf{Spectra for fixed outflow velocity:} We fixed $\vout=50$\kms. 
+	\caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms. 
 			 We vary \vrot increasing from left to right, \tauh increasing from top to 
@@
 
-We now evaluate the effects of increasing outflow velocity at fixed
+We evaluate the effects of increasing outflows velocity at fixed
 values for the optical depth and rotational velocity.
 Figure \ref{fig:varying_outflow_small} summarizes its behavior
-(for largest range see Figure \ref{fig:varying_outflow}).
+(for a largest range see Figure \ref{fig:varying_outflow}).
 All plots correspond to the same rotation velocity of
-$\vrot=50$\kms.
+$\vrot=50$\kms, rows have constant optical depth and columns 
+have constant outflows velocity.
 
@@
 Rotation alone does have an impact on the \lya line morphology
-and we get the same influence in it that previous works. 
+and we obtain the same influence as in the literature.
 Together with the outflows, we found that the resulting spectra is
 equivalent to a non-rotating galaxy with outflows that undergoes a
-Doppler Shift in velocity, weighted by  amount of photons that if
+Doppler Shift in velocity, weighted by the amount of photons that if
 rotating would project their escape in $\hat{k}$'s direction:
@@
 
-Figure \ref{fig:doppler_shift} displays this Doppler shift for four different 
+Figure \ref{fig:doppler_shift} displays this Doppler shift for six different 
 configurations with a fixed optical depth \tauh. In each cell of the grid 
 there is a \lya spectrum for a galaxy with only outflows and two \lya spectra 
-for the galaxy with outflows and rotation, one obtained from the radiative 
-transfer and the other by the Doppler shifting of the non-rotating galaxy.
+for the galaxy with outflows and rotation: one obtained from the radiative 
+transfer simulation and the other obtained from the Doppler shifting of the 
+non-rotating galaxy.
 
@@
   \end{center}
-  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=?$. 
+  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$. 
   	We vary \vrot increasing from left to right and \vout increasing from top to 
@@
 
-In order to compare the line created with the radiative transfer process that 
-involves the rotation (RT) and the line with only outflows that is Doppler 
+In order to compare the line created with the radiative transfer (RT) process that 
+involves the rotation and the line with only outflows that is Doppler 
 shifted (DS), we calculate three main characteristics for each spectrum. We 
@@
 the other is blueshifted. We see that the combined spectrum is a weighted line, 
-in solid black, that combines these past two. This shows that there is in effect a 
-Doppler Shift and justifies our choice to apply it after the outflows spectrum is 
+in solid black, that combines these past two. This shows that there is in fact a 
+Doppler Shift and then it justifies our choice to apply it after the outflows spectrum is 
 obtained.
@@
 provide these effects with values of \vout and \vrot that do not exceed the
-typical LAEs' numbers found in the literature.
+typical LAEs' values found in the literature.
 
@@
 
-	\item The effect of optical depth \tauh in a LAE spectrum is to induces a 
+	\item The effect of optical depth \tauh in a LAE spectrum is to induce a 
 	redshift around the zero velocity in the \lya line.
@@
 
-	\item Rotations induces a Doppler Shift in frequency of the only-outflow 
+	\item Rotations induces a Doppler Shift in frequency of the only-outflows 
 	spectrum of the LAE.
@@
 \ref{fig:varying_rotation} and \ref{fig:varying_outflow} 
-with more combinations of \vrot and \vout.
+with the complete range of combinations of \vrot and \vout.
 
@@
 
-The rotation effect induces a Doppler shift in frequency and displaces the only-outflow spectra.
+The rotation effect induces a Doppler shift in frequency and displaces the only-outflows spectra.
 As seen in Fig.
@@
 	\end{center}
-	\caption{\textbf{Rotations induces a Doppler shift of the only-outflow spectra:} Each of the black lines is then weighted to obtain the red line.
+	\caption{\textbf{Rotations induces a Doppler shift of the only-outflows spectra:} Each of the black lines is then weighted to obtain the red line.
 		\label{fig:rotation_doppler_outflow}}
@@
 Finally, all the spectra produced from this new LAE model is roughly consistent with observations.
-The consideration of adding the new rotation parameter to the standard only-outflow model found in literature is reasonable and very powerful.
+The consideration of adding the new rotation parameter to the standard only-outflows model found in literature is reasonable and very powerful.
 vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\
@@
 
-	\item Rotations induces a Doppler shift in frequency of the only-outflow spectra.
+	\item Rotations induces a Doppler shift in frequency of the only-outflows spectra.
 

### 2018-03-26 paper/paper.tex
figuras

@@
 viewing angle.
-We present a semi-analytic model that modifies the spectra of
+We also present a semi-analytic model that modifies the spectra of
 non-rotational that manages to reproduce the main features of these
@@
 These kinematic maps show that LAEs have two  overimposed velocity
-patterns: coherent rotation plus and a random motions around the mean
-value provided by rotation.  
+patterns: coherent rotation plus and a random motions.
 Depending on the system under study, one of the two features can be
@@
 These kinematic features play an important role in shaping the \lya. 
-
 Models of the \lya have also shown that there is clear signal for
@@
 linea with a more pronounced read peak.
-
 The effects of either outflows or rotation on the \lya line have been
 already explored.
+
 Here we present for the first time a study of the joint effects of
@@
 components expected in LAEs. 
-
 We study a simplified geometrical configuration corresponding to a
@@
 applied in our case.
-We show that solid body rotation effects can be included
+We show how solid body rotation effects can be modelled by
+postprocessing the results of outflow only simulation.
+
 in the results of other radiative transfer model in the case of large
-optical depths. In this case it is a good approximation to Doppler
+optical depths. 
+In this case it is a good approximation to Doppler
 boost the results of the model without rotation.
@@
 
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
+  \end{center}
+  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$. 
+  	We vary \vrot increasing from left to right and \vout increasing from top to 
+  	bottom. The legend represents the \lya line obtained if there is no rotation 
+  	(in thin line), if there is a radiative transfer of rotation and outflows 
+  	(thick and clear line), and if there is a radiative transfer of only outflows,
+  	but also a Doppler shift from the rotational velocity (thick and dark line). 
+	    \label{fig:doppler_shift}}
+\end{figure*}
+
 \subsection{Monte-Carlo Radiative Transfer Model}
@@
 % | tau=1E6, vrot=50, | tau=1E6, vrot=100 | 
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
-	\end{center}
-	\caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms. 
-			 We vary \vrot increasing from left to right, \tauh increasing from top to 
-			 bottom, and $\theta$ according to the legend.
-	\label{fig:varying_rotation_small}}
-\end{figure*}
+%\begin{figure*}
+%  \begin{center}
+%    \includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
+%  \end{center}
+%  \caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms. 
+%    We vary \vrot increasing from left to right, \tauh increasing from top to 
+%    bottom, and $\theta$ according to the legend.
+%    \label{fig:varying_rotation_small}}
+%\end{figure*}
 
@@
 % | tau=1E6, vout=25, | tau=1E6, vout=75 |
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow_small}
-	\end{center}
-	\caption{\textbf{Spectra for fixed rotational velocity:} We fixed $\vrot=50$\kms. 
-			 We vary \vout increasing from left to right, \tauh increasing from top to 
+%\begin{figure*}
+%\begin{center}
+%\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow_small}
+%\end{center}
+%\caption{\textbf{Spectra for fixed rotational velocity:} We fixed $\vrot=50$\kms.  We vary \vout in%creasing from left to right, \tauh increasing from top to 
 			 bottom, and $\theta$ according to the legend.
-		\label{fig:varying_outflow_small}}
-\end{figure*}
+%\label{fig:varying_outflow_small}}
+%\end{figure*}
 
@@
 % |vrot=50, vout=50, | vrot=100, vout=50, |
-\begin{figure*}
-  \begin{center}
-    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
-  \end{center}
-  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$. 
-  	We vary \vrot increasing from left to right and \vout increasing from top to 
-  	bottom. The legend represents the \lya line obtained if there is no rotation 
-  	(in thin line), if there is a radiative transfer of rotation and outflows 
-  	(thick and clear line), and if there is a radiative transfer of only outflows,
-  	but also a Doppler shift from the rotational velocity (thick and dark line). 
-	    \label{fig:doppler_shift}}
-\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout5_logtau7}
-		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout25_logtau7}
-		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout50_logtau7}
+ \includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_mn_vout5_logtau7}
+ \includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_mn_vout25_logtau7}
+ \includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_mn_vout50_logtau7}
 \end{center}
@@
     \label{fig:mean_line}}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
 \begin{center}
-		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout5_logtau7}
-		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout25_logtau7}
-		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout50_logtau7}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau7}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau7}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau7}
 \end{center}
@@
   \label{fig:standard_deviation}}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
+\begin{figure*}
   \begin{center}
@@
     \label{fig:skewness}}
-\end{figure}
+\end{figure*}
 

### 2018-03-26 paper/paper.tex
figura

@@
 \begin{center}
-    \includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout5_logtau7}
-    \includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout25_logtau7}
-    \includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout50_logtau7}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
   \end{center}

### 2018-03-26 paper/paper.tex
figuras

@@
 % cada panel tiene una curva para logtau = 5, 6, 7.
-\begin{figure}
+\begin{figure*}
 \begin{center}
-		\includegraphics[height=0.12\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
-		\includegraphics[height=0.12\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
-		\includegraphics[height=0.12\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
 \end{center}
@@
 		\label{fig:valley_intensity}}
-\end{figure}
+\end{figure*}
 

### 2018-03-26 paper/paper.tex
figura

@@
 discussed in section \ref{sec:discussion}.
-
-\begin{figure*}
-\begin{center}
- \includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_mn_vout5_logtau7}
- \includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_mn_vout25_logtau7}
- \includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_mn_vout50_logtau7}
-\end{center}
-  \caption{\textbf{Mean line for RT and DS:} The triangles represent the Radiative 
-    Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
-    \label{fig:mean_line}}
-\end{figure*}
 

### 2018-03-26 paper/paper.tex
making paper plots in pdf format

@@
 viewing angle.
-We also present a semi-analytic model that modifies the spectra of
+We present a semi-analytic model that modifies the spectra of
 non-rotational that manages to reproduce the main features of these
@@
 These kinematic maps show that LAEs have two  overimposed velocity
-patterns: coherent rotation plus and a random motions.
+patterns: coherent rotation plus and a random motions around the mean
+value provided by rotation.  
 Depending on the system under study, one of the two features can be
@@
 These kinematic features play an important role in shaping the \lya.
+
 Models of the \lya have also shown that there is clear signal for
@@
 linea with a more pronounced read peak.
+
 The effects of either outflows or rotation on the \lya line have been
 already explored.
-
 Here we present for the first time a study of the joint effects of
@@
 components expected in LAEs.
+
 We study a simplified geometrical configuration corresponding to a
@@
 applied in our case.
-We show how solid body rotation effects can be modelled by
-postprocessing the results of outflow only simulation.
-
+We show that solid body rotation effects can be included
 in the results of other radiative transfer model in the case of large
-optical depths. 
-In this case it is a good approximation to Doppler
+optical depths. In this case it is a good approximation to Doppler
 boost the results of the model without rotation.
@@
 
-\begin{figure*}
-  \begin{center}
-    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
-  \end{center}
-  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$. 
-  	We vary \vrot increasing from left to right and \vout increasing from top to 
-  	bottom. The legend represents the \lya line obtained if there is no rotation 
-  	(in thin line), if there is a radiative transfer of rotation and outflows 
-  	(thick and clear line), and if there is a radiative transfer of only outflows,
-  	but also a Doppler shift from the rotational velocity (thick and dark line). 
-	    \label{fig:doppler_shift}}
-\end{figure*}
-
 \subsection{Monte-Carlo Radiative Transfer Model}
@@
 % | tau=1E6, vrot=50, | tau=1E6, vrot=100 | 
-%\begin{figure*}
-%  \begin{center}
-%    \includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
-%  \end{center}
-%  \caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms. 
-%    We vary \vrot increasing from left to right, \tauh increasing from top to 
-%    bottom, and $\theta$ according to the legend.
-%    \label{fig:varying_rotation_small}}
-%\end{figure*}
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
+	\end{center}
+	\caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms. 
+			 We vary \vrot increasing from left to right, \tauh increasing from top to 
+			 bottom, and $\theta$ according to the legend.
+	\label{fig:varying_rotation_small}}
+\end{figure*}
 
@@
 % | tau=1E6, vout=25, | tau=1E6, vout=75 |
-%\begin{figure*}
-%\begin{center}
-%\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow_small}
-%\end{center}
-%\caption{\textbf{Spectra for fixed rotational velocity:} We fixed $\vrot=50$\kms.  We vary \vout in%creasing from left to right, \tauh increasing from top to 
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow_small}
+	\end{center}
+	\caption{\textbf{Spectra for fixed rotational velocity:} We fixed $\vrot=50$\kms. 
+			 We vary \vout increasing from left to right, \tauh increasing from top to 
 			 bottom, and $\theta$ according to the legend.
-%\label{fig:varying_outflow_small}}
-%\end{figure*}
+		\label{fig:varying_outflow_small}}
+\end{figure*}
 
@@
 % |vrot=50, vout=50, | vrot=100, vout=50, |
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
+  \end{center}
+  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$. 
+  	We vary \vrot increasing from left to right and \vout increasing from top to 
+  	bottom. The legend represents the \lya line obtained if there is no rotation 
+  	(in thin line), if there is a radiative transfer of rotation and outflows 
+  	(thick and clear line), and if there is a radiative transfer of only outflows,
+  	but also a Doppler shift from the rotational velocity (thick and dark line). 
+	    \label{fig:doppler_shift}}
+\end{figure*}
 
@@
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau7}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau7}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau7}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout5_logtau7}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout25_logtau7}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_mn_vout50_logtau7}
+	\end{center}
+	\caption{\textbf{Mean line for RT and DS:} The triangles represent the Radiative 
+		Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
+		\label{fig:mean_line}}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout5_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout25_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_std_vout50_logtau7}
 	\end{center}
@@
 		\label{fig:standard_deviation}}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout5_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout25_logtau7}
+		\includegraphics[height=0.11\textheight]{./figures/results/line_characterization_skw_vout50_logtau7}
 	\end{center}
@@
 % cada panel tiene una curva para logtau = 5, 6, 7.
-\begin{figure*}
+\begin{figure}
 	\begin{center}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
+		\includegraphics[height=0.1\textheight]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
 	\end{center}
@@
 		\label{fig:valley_intensity}}
-\end{figure*}
+\end{figure}
 

### 2018-03-26 plots/paper/model.py

@@ -86,7 +86,8 @@ def get_spectra(data, v_th=12.86, as_hist=True):
     '''Returns the spectrum in velocity units.'''
     velocity = -data['x_frec']*v_th
     n, b = np.histogram(velocity, bins=50)
-    n = n / n.sum()
+    area = np.trapz(n, b[:-1])
+    n = n / area
     if (as_hist):
         return b, n
     else:

### 2018-03-26 paper/paper.tex

--- 2018-03-26 paper/paper.tex
@@
 \begin{center}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
   \end{center}
@@
 \begin{center}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
 \end{center}

### 2018-03-27 paper/paper.tex
plots

@@
 	\begin{center}
-		\includegraphics[width=0.45\textwidth]{./figures/discussion/doppler}
+		\includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
 	\end{center}
@@
 
-In this appendix we show more \lya spectra in the Figures 
-\ref{fig:varying_rotation} and \ref{fig:varying_outflow} 
-with the complete range of combinations of \vrot and \vout.
+In this appendix we show \lya spectra with fixed \vout (Fig. \ref{fig:varying_rotation_small})
+and fixed \vrot (Fig. \ref{fig:varying_outflow_small}). For a larger set of spectra with the
+same fixed \vout and \vrot refer to Fig. \ref{fig:varying_rotation} and Fig. \ref{fig:varying_outflow} 
+respectively.
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation_small}
+	\end{center}
+	\caption{\textbf{Caption:} $\vout=50$\kms.
+		\label{fig:varying_rotation_small}}
+\end{figure*}
+
+\begin{figure*}
+	\begin{center}
+		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow_small}
+	\end{center}
+	\caption{\textbf{Caption:} $\vrot=50$\kms.
+		\label{fig:varying_outflow_small}}
+\end{figure*}
 

### 2018-03-27 paper/paper.tex
edit caption
conclusions

@@
 Outflows and rotation are two ubiquitous kinematic features in the gas
-distributions of galaxies.
+kinematics of galaxies.
 Here we perform Monte Carlo radiative transfer simulations of outflowing
-gas with additional solid body rotation to understand how these two
+gas with additional solid body rotation to understand how these kinematic
 features impact the morphology of the Lyman-$\alpha$ emission line.
-One of the most notorious results is that rotation increases the
-detected flux at the line's center in comparison with the
-non-rotational case. 
-Rotation also induces a line broadening and a dependency with the
-viewing angle.
-We also present a semi-analytic model that modifies the spectra of
-non-rotational that manages to reproduce the main features of these
+We find three important consequences of rotation.
+First, it increases the detected flux at the line's center in comparison with the
+non-rotational case, second it produces a line broadening and third it introduces
+a dependency with the  viewing angle.
+We also develp a semi-analytic model that modifies the spectra of
+pure outflow simulations that manages to reproduce the main features of the
 Monte Carlo simulations.
-As an application example we contraint the rotation parameters to the
-observed Lyman-$\alpha$ profile of a high redshift LAE, providing the
-first a constraint on the rotational velocity of a high redshift
-galaxy based on its integrated \lya emission.
+As an application we constraint the rotation parameters of a set of observed LAE
+emission spectra.
 \end{abstract}
@@
 
-In this paper we create, for the first time, a computational model for LAEs 
-that emulate the kinematic effects of galaxy rotation and outflows in the \lya
-spectral line.
+In this paper we explore, for the first time in the literature,
+the results of a model for the emergent \lya\ line from rotating outflows.
+The main results for the model are computed from a Monte-Carlo radiative transfer
+simulation and confronted with a simple semi-analytic ansatz that adds the effects
+of rotation onto results of pure outflow kinematics.
 
-Our main conclusions are:
+The main effects of rotation on the \lya\ line morphology are:
 
 \begin{itemize}
-	\item The effect of rotation \vrot in a LAE spectrum is to broaden the 
-	\lya line.
+  \item Broadening the line.
+	\item Increasing the intensity at the line's center.
+	\item Inducing a dependency on the viewing angle,
+  The closer the observer is to the pole of the galaxy (defined by the rotation axis),
+  the smaller are the effects from rotation.
+\end{itemize}
 
-	\item The effect of outflows \vout in a LAE spectrum is to increase the peak 
-	asymmetry of the \lya line.
-
-	\item The effect of optical depth \tauh in a LAE spectrum is to induce a 
-	redshift around the zero velocity in the \lya line.
-
-	\item Rotation induces a dependency in viewing angle on the \lya line. The 
-	closer the observer is to the equator of the galaxy, the higher is the central 
-	valley of the frequency distribution.
-
-	\item Rotations induces a Doppler Shift in frequency of the only-outflows 
-	spectrum of the LAE.
-
-	\item The final spectra obtained are roughly consistent with LAEs 
-	observations. 
-\end{itemize}
+We found that in the case of solid body rotation these effects can be quantitatively
+explained by a Doppler boost, where the Doppler factor can be computed as the
+ producet of quantities at the surface of last scattering, namely $\vec{v}_{\rm rot}\cdot \hat{k}$,
+ where $\vec{v}_{\rm rot}$ is the velocity due to rotation and $\hat{k}$ is the direction
+ of the photon's propagation.
+These conclusions trenghten the evidence reported by \cite{Garavito14} when they modeled
+the effect of pure rotation on \lya\ spectra.
 
@@
 	\end{center}
-	\caption{\textbf{Caption:} $\vout=50$\kms.
+	\caption{\textbf{Spectra with fixed outflows velocity:} We fix $\vout=50$\kms.
 		\label{fig:varying_rotation_small}}
@@
 	\end{center}
-	\caption{\textbf{Caption:} $\vrot=50$\kms.
+	\caption{\textbf{Spectra with fixed rotation velocity:} We fix $\vrot=50$\kms.
 		\label{fig:varying_outflow_small}}
@@
 	\end{center}
-	\caption{\textbf{Caption:} $\vout=50$\kms.
+	\caption{\textbf{Spectra with fixed outflows velocity:} We fix $\vout=50$\kms.
 		\label{fig:varying_rotation}}
@@
 	\end{center}
-	\caption{\textbf{Caption:} $\vrot=50$\kms.
+	\caption{\textbf{Spectra with fixed rotation velocity:} We fix $\vrot=50$\kms.
 		\label{fig:varying_outflow}}

### 2018-03-27 paper/paper.tex
new references

@@
 
-These kinematic maps show that LAEs have two  overimposed velocity
-patterns: coherent rotation plus and a random motions.
-Depending on the system under study, one of the two features can be
-more important, or they can be found in equal measure.
-These kinematic features play an important role in shaping the \lya.
-Models of the \lya have also shown that there is clear signal for
-outflowing gas most probably produced by star formation activity.
-The characteristic signal for these outflows is the asymmetrical \lya
-linea with a more pronounced read peak.
-The effects of either outflows or rotation on the \lya line have been
-already explored.
+Dwarf galaxies show signs of coherent rotation
+\citep{2009A&A...493..871S}.
+Some of them are expected to have high neutral gas contents and thus
+show \lya\ emission with rotation imprints
+\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
+Some Compact Dwarf Galaxies actually show gas kinematics that range
+from pure rotation to high velocity dispersion without a clear
+rotation pattern.
+\citep{2015A&A...577A..21C,2017arXiv170809407C,2017A&A...600A.125C}
+
+The case for outflows as traced in the \lya\ line is much bettwe
+stablished.
+There are abundance cases \lya line profiles with a single peak
+redwards from the line's center. 
+In other cases there is a double peak but the peak on the red side is
+stronger \citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.  
+This has been readily explained as the consequence of multiple \lya photon
+scatterings through an homogeneous outflowing shell of neutral Hydrogen
+\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.
 
@@
 of the photon's propagation.
-These conclusions trenghten the evidence reported by \cite{Garavito14} when they modeled
-the effect of pure rotation on \lya\ spectra.
+These conclusions, specially the confirmation of the Doppler boost as
+a good proxy for solid body rotation, strenghten the evidence reported
+by \cite{Garavito14} when they modeled the effect of pure rotation on
+\lya\ spectra.
+
+As an application to observational data we find that recent results
+that take the spectra of two different sides of a galaxy can detect 
+the effect of approaching/receeding gas motions. 
+However, the distances between the peaks of these two spectra 
+correspond to $\approx v_{\rm rot} \cos\theta$ due to the weights of the
+emitting regions.
 

### 2018-03-27 paper/paper.tex
referencias
edit

@@
 Monte Carlo simulations.
-As an application we constraint the rotation parameters of a set of observed LAE
-emission spectra.
+As an application to observational data we constraint the rotation
+parameters of a set of observed LAE emission spectra.    
 \end{abstract}
@@
 
-Dwarf galaxies show signs of coherent rotation
+For instances, dwarf galaxies show signs of coherent rotation
 \citep{2009A&A...493..871S}.
@@
 rotation pattern.
-\citep{2015A&A...577A..21C,2017arXiv170809407C,2017A&A...600A.125C}
+\citep{2015A&A...577A..21C,2017A&A...600A.125C}
 
@@
 
-\subsection{Monte-Carlo Radiative Transfer Model}
-
-CLARA follows the propagation of individual photons through a neutral
-Hydrogen medium characterized by its temperature, velocity field and
-global optical depth.
+The Monte Carlo code we use (CLARA)  follows the propagation of
+individual photons through a neutral Hydrogen medium characterized by
+its temperature, velocity field and global optical depth.
 The code assumes an homogeneous density throughout the simulated
@@
 an approximation commonly used in the literature, as it explains a
-wide variety of observational features \citep{Ahn03,Verhamme06,Dijkstra06}.
+wide variety of observational features
+\citep{Ahn03,Verhamme06,Dijkstra06}.
 
-The central element in this paper is the velocity field that captures
-outflows and rotation.
-Outflows are captured by a Hubble-like radial velocity profile with
+The velocity field we use captures both outflows and rotation.
+Outflows are described by a Hubble-like radial velocity profile with
 the velocity magnitude increasing linearly with the radial
-coordinate; the outflows model is fully characterized by $V_{\rm out}$, the
+coordinate; the outflows model is fully characterized by $v_{\rm out}$, the
 velocity at the sphere's surface.
 Rotation follows a solid body rotation profile, which is fully
-characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.
+characterized by $v_{\rm rot}$, the linear velocity at the sphere's surface.
 
@@
 
-\subsection{Analytical Model for Bulk Rotation}
-
-Rotation induces two main effects on the \lya line morphology.
-First. Break of spherical symmetry due to the
-preferential direction introduced by the rotation axis.
-This symmetry break is reflected in the observed spectra as a
-dependency on the viewing angle, $\theta$, between the line of sight
-of a distant observer and the rotational velocity.
-Second. Line blurring as the rotational velocity increases.
-At a fixed viewing angle, faster rotation induces makes the observed
-line wider and increases the flux around the line's center.
-
 \cite{Garavito14} presented in the Appendix an analytical model that
-accounts for these two features.
-The basic assumption of the analytical model is that each
+accounts for the effects of pure rotation on the
+\lya\ lines 'morphology. 
+The basic assumption of their analytical model is that each
 differential surface element on the sphere Doppler shifts the photons
 that it emits
+
 In this paper we introduce this ansatz by post-processing the results
 of the outflows simulations without rotation.
+We introduce a Doppler shift as follows
+\begin{equation}
+x' = x + \frac{\vec{v}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
+\label{eq:shift_x}
+\end{equation}
+%
+where $x'$ is the photon's new frequency, $x$ is the photon's
+frequency after being procesed by the outflow, $v_{\rm rot}$ is the 
+rotational velocity at the point of escape of the photon, $\hat{k}$ is
+the photon's direction of propagation and $v_{\rm th}$ is the thermal
+velocity of the sphere.
+
 This allows us to produce new \lya spectra and compare them with the
@@
 
+Figure \ref{fig:doppler_shift} displays this Doppler shift for six different
+configurations with a fixed optical depth \tauh. In each cell of the grid
+there is a \lya spectrum for a galaxy with only outflows and two \lya spectra
+for the galaxy with outflows and rotation: one obtained from the radiative
+transfer simulation and the other obtained from the Doppler shifting of the
+non-rotating galaxy.
+
 % Esta figura deberia tener 2 x 2 vinetas. vrot=50.
@@
 \subsection{Doppler Shifts to Include Rotational Effects}
-
-Rotation alone does have an impact on the \lya line morphology
-and we obtain the same influence as in the literature.
-Together with the outflows, we found that the resulting spectra is
-equivalent to a non-rotating galaxy with outflows that undergoes a
-Doppler Shift in velocity, weighted by the amount of photons that if
-rotating would project their escape in $\hat{k}$'s direction:
-
-\begin{equation}
-V' = V + \vec{v}_{\rm rot}\cdot\hat{k}
-\label{eq:shift_x}
-\end{equation}
-%
-where $V'$ is the new Doppler shifted velocity, $V$ is the velocity caused by
-outflows.
-
-Figure \ref{fig:doppler_shift} displays this Doppler shift for six different
-configurations with a fixed optical depth \tauh. In each cell of the grid
-there is a \lya spectrum for a galaxy with only outflows and two \lya spectra
-for the galaxy with outflows and rotation: one obtained from the radiative
-transfer simulation and the other obtained from the Doppler shifting of the
-non-rotating galaxy.
 
@@
 
-In this paper, the objective was to analyze and measure the influence of galaxy rotation and outflows on the \lya line.
-The motivation for this is to be able to obtain physical information of a LAE by just looking at its \lya profile.
-In order to accomplish this objective, we propose a new model of a LAE consisting of a sphere of Hydrogen atoms that expands radially and rotates as a solid body.
-A modified version of the program CLARA \cite{CLARA} is used to set the conditions and emulate the radiative transfer process inside the galaxy.
-\\
+In this paper, the objective was to analyze and measure the influence
+of galaxy rotation and outflows on the \lya line. The motivation for
+this is to be able to obtain physical information of a LAE by just
+looking at its \lya profile. In order to accomplish this objective, we
+propose a new model of a LAE consisting of a sphere of Hydrogen atoms
+that expands radially and rotates as a solid body. A modified version
+of the program CLARA \cite{CLARA} is used to set the conditions and
+emulate the radiative transfer process inside the galaxy. \\
 

### 2018-03-27 paper/paper.tex
edit

@@
 Rotation follows a solid body rotation profile, which is fully
-characterized by $v_{\rm rot}$, the linear velocity at the sphere's surface.
+characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.
 
@@
 
-\cite{Garavito14} presented in the Appendix an analytical model that
+\cite{Garavito14} presented an analytical model that
 accounts for the effects of pure rotation on the

### 2018-03-27 paper/paper.tex
edit

@@
 accounts for the effects of pure rotation on the
-\lya\ lines 'morphology. 
+\lya\ line morphology. 
 The basic assumption of their analytical model is that each
 differential surface element on the sphere Doppler shifts the photons
-that it emits
-
+that it emits.
 In this paper we introduce this ansatz by post-processing the results
 of the outflows simulations without rotation.
-We introduce a Doppler shift as follows
+The frequency of each photon is Doppler shifted as follows
+
 \begin{equation}
@@
 \label{sec:results}
+
+Figure \ref{fig:doppler_shift} summarizes the most important trends of
+our simulations.
 

### 2018-03-27 paper/paper.tex
edit
edita

@@
 \newcommand{\kms}{\ifmmode\mathrm{km\ s}^{-1}\else km s$^{-1}$\fi}
-\newcommand{\vrot}{\ifmmode\mathrm v_{\mathrm{rot}}\else $v_{\mathrm{rot}}$~\fi}
-\newcommand{\vout}{\ifmmode\mathrm v_{\mathrm{out}}\else $v_{\mathrm{out}}$~\fi}
+\newcommand{\vrot}{\ifmmode\mathrm V_{\mathrm{rot}}\else $V_{\mathrm{rot}}$~\fi}
+\newcommand{\vout}{\ifmmode\mathrm V_{\mathrm{out}}\else $V_{\mathrm{out}}$~\fi}
 \newcommand{\tauh}{\ifmmode\mathrm \tau_{\mathrm{H}}\else $\tau_{\mathrm{H}}$~\fi}
@@
 %=========================================================================
-\title[Outflows and rotation in LAEs]{Lyman-alpha spectra from
-  rotating outflows}
+\title[Outflows and rotation in LAEs]{Lyman-$\alpha$ photons through rotating outflows}
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
@@
       \hline
-      $10^5$ & \{0\} & \{5, 25, 50\} \\
-      $10^6$ & \{0\} & \{5, 25, 50\}\\
-      $10^7$ & \{0\} & \{5, 25, 50\}\\
-      $10^5$ & \{50, 100\} & \{25, 50, 75\} \\
-      $10^6$ & \{50, 100\} & \{25, 50, 75\}\\
-      $10^7$ & \{50, 100\} & \{25, 50, 75\}\\
+      \{$10^5$, $10^6$, $10^7$\}  & \{0, 50, 100\} & \{5, 25, 50\} \\
       \hline
@@
 \begin{equation}
-x' = x + \frac{\vec{v}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
+x' = x + \frac{\vec{V}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
 \label{eq:shift_x}
@@
 
+\subsection{Qualitative Trends}
+\label{sec:qualitative}
 Figure \ref{fig:doppler_shift} summarizes the most important trends of
 our simulations. 
+The six panels correspond to $\tau=10^6$ and a viewing angle of
+$\theta =90^{\deg}$, that is perpendicular to the rotation axis of the
+galaxy. 
+In every panel the thin black line corresponds to the pure outflow
+solution (i.e. without rotation). 
+From top to bottom we see the effect of increasing the outflow
+velocity, which is the expected increasing asymmetry towards the red
+peak.
+
+The thick black line corresponds to the solution that includes both
+outflows and rotation.
+Comparing the left and right columns (lower versus higher rotational
+velocity) we can see two immediate effects.
+First, the line broadens and second, the intensity at the line's
+center increases.
+
+Finally, the thick gray line correspods to the pure outflow solution
+with the Doppler boost added to model rotation's influence.
+In all cases the Doppler boost does a good job at capturing the broad
+morphological features introduced by rotation: the angle dependence,
+the broadening and the intensity increase at the line's center.
+
+In Figure \ref{fig:angle_influence} we show the same results as in
+Figure \ref{fig:doppler_shift} but for a viewing angle of $\theta =
+90^{\deg}$, that is parallel to the rotation axis. 
+In this case we see that the effect of rotation is very hard to see. 
+As \ref{Garavito14} showed, pure rotation introduces a strong
+dependence with viewing angle, a trend that we find also holds for
+rotation mixed with ouflows.
 
@@
   \end{center}
-  \caption{\textbf{Doppler shift effect:} We fixed $\tauh=10^6$ and $\theta=90^\circ$.
+  \caption{\textbf{Qualitative trends of changing outflow and
+      rotational velocity}. 
+    Here we fix $\tauh=10^6$ and $\theta=90^\circ$.
     We vary \vrot increasing from left to right and \vout increasing from top to
-  	bottom. The legend represents the \lya line obtained if there is no rotation
+    bottom. 
+    The thin black line corresponds to the \lya line obtained with
+    CLARA without any rotation and the indicated outflow velocity.
+    The thick black line corresponds to the results including both
+    outflows and rotation.
+    The thick gray line shows the results of modifyin the pure outflow
+    solution by the Doppler shift presented in Equation \ref{eq:shift_x}
     (in thin line), if there is a radiative transfer of rotation and outflows
@@
 
-\subsection{Monte-Carlo Radiative Transfer Model}
+\subsection{Quantitative trends}
+\label{sec:quantitative}
 
-We evaluate the effects of increasing rotational
-velocity at fixed values of the optical depth and outflows velocity.
-Figure \ref{fig:varying_rotation_small} summarizes its behavior
-(for a largest range see Figure \ref{fig:varying_rotation}).
-All plots correspond to the same outflows velocity of $\vout=50$\kms,
-rows have constant optical depth and columns have constant rotational
-velocity.
-Each panel shows different morphologies corresponding to varying
-line-of-sight viewing angle.
-
-% Esta figura deberia tener 2 x 2 vinetas. vout=50.
-% cada panel tiene una curva para theta=0, 45, 90.
-% | tau=1E5, vrot=50, | tau=1E5, vrot=100 |
-% | tau=1E6, vrot=50, | tau=1E6, vrot=100 |
-%\begin{figure*}
-%  \begin{center}
-%    \includegraphics[width=0.90\textwidth]{./figures/results/varying_rotation_small}
-%  \end{center}
-%  \caption{\textbf{Spectra for fixed outflows velocity:} We fixed $\vout=50$\kms.
-%    We vary \vrot increasing from left to right, \tauh increasing from top to
-%    bottom, and $\theta$ according to the legend.
-%    \label{fig:varying_rotation_small}}
-%\end{figure*}
+After finding the qualitative influence of the different parameters.
 

### 2018-03-28 paper/paper.tex
fig 1 with theta 0 and 90
graficas

@@
   \begin{center}
-    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_6}
+    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_logtau6_theta90}
   \end{center}
   \caption{\textbf{Qualitative trends of changing outflow and
-      rotational velocity}. 
+      rotational velocity viewed perpendicular to the rotation axis}. 
     Here we fix $\tauh=10^6$ and $\theta=90^\circ$.
@@
     \label{fig:doppler_shift}}
+\end{figure*}
+
+\begin{figure*}
+  \begin{center}
+    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_logtau6_theta0}
+  \end{center}
+  \caption{\textbf{Qualitative trends of changing outflow and
+      rotational velocity viewed parallel to the rotation axis}. 
+    Here we fix $\tauh=10^6$ and $\theta=0^\circ$.
+    \label{fig:doppler_shift_theta_0}}
 \end{figure*}
@@
 \begin{center}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau7}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau7}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau7}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
+\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
 \end{center}
@@
 \begin{center}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
 \includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}

### 2018-03-28 paper/paper.tex
adding bimodality plots

@@
 shifted (DS), we calculate three main characteristics for each spectrum. We
-measure the line's mean velocity, standard deviation and skewness for both
-RT's and DS's approaches. We plot the results for each value in Figures
-\ref{fig:mean_line}, \ref{fig:standard_deviation} and \ref{fig:skewness},
-respectively for $\tauh = 10^7$. We highlight that these 3 plots increase
-their fit tendency with a higher optical depth. The reasons why will be
-discussed in section \ref{sec:discussion}.
+measure the line's standard deviation (STD), skewness (SKW) and bimodality (BI) 
+for both RT's and DS's approaches. We use the definitions in Eqs. \ref{eq:std}, 
+\ref{eq:skw}, \ref{eq:bi} for each value \citep{kokoska1999}.
+
+\begin{equation}
+\label{eq:std}
+\mathrm{STD} = \sqrt{m_2},
+\end{equation}
+
+\begin{equation}
+\label{eq:skw}
+\mathrm{SKW} = \frac{m_3}{m_2^{3/2}},
+\end{equation}
+
+\begin{equation}
+\label{eq:bi}
+\mathrm{BI} = \mathrm{KURTOSIS} - \mathrm{SKW}^2 = \frac{m_4}{m_2^{2}} - \frac{m_3^2}{m_2^{3}},
+\end{equation}
+
+where each $m_i$ is the i-th moment about the mean. The STD has velocity units and 
+quantifies the dispersion of the line's velocity: the larger it is, the wider the line. 
+The SKW is adimensional and quantifies the peaks' asymmetry: if it is $>0$ the left peak is 
+taller and if it is $<0$ the right peak is taller. The BI has velocity$^{-1}$ units and
+quantifies if the line has 1 or 2 peaks: it is always $\geq 1$ and the closer it is to 1 the more
+bimodal is the line (i.e. has 2 peaks).
+
+We plot the results for each 
+value in Figures \ref{fig:standard_deviation}, \ref{fig:skewness} and \ref{fig:bimodality},
+respectively for $\tauh = 10^7$. We highlight that these 3 plots increase their fit 
+tendency with a higher optical depth. The reasons why will be discussed in section 
+\ref{sec:discussion}.
 
@@
 \caption{\textbf{Standard Deviation for RT and DS:} The triangles represent the Radiative
-  Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
+  Transfer and the lines represent the Doppler Shift.
   \label{fig:standard_deviation}}
@@
 \caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative
-    Transfer and the lines represent the Doppler Shift. We fixed $\tauh=10^7$.
+    Transfer and the lines represent the Doppler Shift.
     \label{fig:skewness}}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
+\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
+\end{center}
+\caption{\textbf{Bimodality for RT and DS:} The triangles represent the Radiative
+	Transfer and the lines represent the Doppler Shift.
+	\label{fig:bimodality}}
 \end{figure*}
@@
 \includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau5}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau5}
 \includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau6}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau6}
 \includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau7}
+\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau7}
 \end{center}
@@
 		$\cos{\theta}$ the lower is the valley intensity, so the higher $\theta$, the
-		higher the valley intensity. We fixed $\vout=5$\kms and $\vrot=100$\kms.
+		higher the valley intensity. Also, the higher \tauh, the better is the fit RT-DS.
+		We fixed $\vrot=100$\kms.
 		\label{fig:valley_intensity}}

### 2018-03-28 paper/paper.tex
paper descrip

@@
 quantifies the dispersion of the line's velocity: the larger it is, the wider the line. 
-The SKW is adimensional and quantifies the peaks' asymmetry: if it is $>0$ the left peak is 
+The SKW has velocity$^{-1/2}$ units and quantifies the peaks' asymmetry: if it is $>0$ the left peak is 
 taller and if it is $<0$ the right peak is taller. The BI has velocity$^{-1}$ units and
-quantifies if the line has 1 or 2 peaks: it is always $\geq 1$ and the closer it is to 1 the more
-bimodal is the line (i.e. has 2 peaks).
+quantifies if the line has 1 or 2 peaks.
 
-We plot the results for each 
-value in Figures \ref{fig:standard_deviation}, \ref{fig:skewness} and \ref{fig:bimodality},
-respectively for $\tauh = 10^7$. We highlight that these 3 plots increase their fit 
-tendency with a higher optical depth. The reasons why will be discussed in section 
-\ref{sec:discussion}.
+We plot the results for each value in Figures \ref{fig:standard_deviation}, 
+\ref{fig:skewness} and \ref{fig:bimodality}, respectively for $\tauh = 10^7$. 
+We highlight that these 3 plots increase their fit tendency with a higher optical 
+depth. The reasons why will be discussed in section \ref{sec:discussion}.
 

### 2018-03-28 paper/paper.tex
fixed bi and scatter plots

@@
 quantifies the dispersion of the line's velocity: the larger it is, the wider the line. 
-The SKW has velocity$^{-1/2}$ units and quantifies the peaks' asymmetry: if it is $>0$ the left peak is 
-taller and if it is $<0$ the right peak is taller. The BI has velocity$^{-1}$ units and
-quantifies if the line has 1 or 2 peaks.
+The SKW is adimensional and quantifies the peaks' asymmetry: if it is $>0$ the left peak is 
+taller and if it is $<0$ the right peak is taller. The BI is adimensional and
+quantifies if the line has 1 or 2 peaks: it is  always $\geq 1$ and the closer it is to 1, the 
+more bimodal is the line (i.e. has 2 peaks).
 

### 2018-03-28 paper/paper.tex
muse in appendix temporarly and cite to theorem

@@
 taller and if it is $<0$ the right peak is taller. The BI is adimensional and
-quantifies if the line has 1 or 2 peaks: it is  always $\geq 1$ and the closer it is to 1, the 
-more bimodal is the line (i.e. has 2 peaks).
+quantifies if the line has 1 or 2 peaks: it is  always $\geq 1$ \citep{Pearson1929} and the closer it is to 1, the more bimodal is the line (i.e. has 2 peaks).
 
@@
 
+% MUSE
+\begin{figure*}
+	\begin{center}
+		\includegraphics[height=0.3\textheight]{./figures/appendix/muse_vs_simulations.pdf}
+		\includegraphics[height=0.3\textheight]{./figures/appendix/museall_vs_muserange.pdf}
+		\includegraphics[height=0.3\textheight]{./figures/appendix/muserange_vs_simulations.pdf}
+	\end{center}
+	\caption{\textbf{Caption:} Caption.
+		\label{fig:varying_outflow}}
+\end{figure*}
+
 \end{document}

### 2018-03-29 paper/paper.tex
edit + fig size

@@
 
-\begin{figure*}
+\begin{figure}
   \begin{center}
-    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_logtau6_theta90}
+    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau6_theta90}
   \end{center}
@@
     Here we fix $\tauh=10^6$ and $\theta=90^\circ$.
-    We vary \vrot increasing from left to right and \vout increasing from top to
-    bottom. 
+    We vary \vrot increasing from left to right and \vout increasing
+    from top to bottom. 
     The thin black line corresponds to the \lya line obtained with
@@
     (in thin line), if there is a radiative transfer of rotation and outflows
-    (thick and clear line), and if there is a radiative transfer of only outflows,
-    but also a Doppler shift from the rotational velocity (thick and dark line).
+    (thick and clear line), and if there is a radiative transfer of
+    only outflows, but also a Doppler shift from the rotational
+    velocity (thick and dark line).  
     \label{fig:doppler_shift}}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
   \begin{center}
-    \includegraphics[width=0.90\textwidth]{./figures/results/doppler_shift_logtau6_theta0}
+    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau6_theta0}
   \end{center}
@@
       rotational velocity viewed parallel to the rotation axis}.     
-    Here we fix $\tauh=10^6$ and $\theta=0^\circ$.
+    Same layout as Figure \ref{fig:doppler_shift},
+    this time the viewing angle corresponds to $\theta=0^\circ$.
+    The effects of rotation on
+    the  \lya\ line morphology are not visible when viewed parallel to
+    the rotation axis.
     \label{fig:doppler_shift_theta_0}}
-\end{figure*}
+\end{figure}
 
@@
 \begin{center}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
-\includegraphics[height=0.3\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
 \end{center}

### 2018-03-29 paper/paper.tex
edit

@@
 \newcommand{\kms}{\ifmmode\mathrm{km\ s}^{-1}\else km s$^{-1}$\fi}
-\newcommand{\vrot}{\ifmmode\mathrm V_{\mathrm{rot}}\else $V_{\mathrm{rot}}$~\fi}
-\newcommand{\vout}{\ifmmode\mathrm V_{\mathrm{out}}\else $V_{\mathrm{out}}$~\fi}
+\newcommand{\vrot}{\ifmmode\mathrm v_{\mathrm{rot}}\else $v_{\mathrm{rot}}$~\fi}
+\newcommand{\vout}{\ifmmode\mathrm v_{\mathrm{out}}\else $v_{\mathrm{out}}$~\fi}
 \newcommand{\tauh}{\ifmmode\mathrm \tau_{\mathrm{H}}\else $\tau_{\mathrm{H}}$~\fi}
@@
 
-The case for outflows as traced in the \lya\ line is much bettwe
+The case for outflows as traced in the \lya\ line is much better
 stablished.
-There are abundance cases \lya line profiles with a single peak
-redwards from the line's center. 
-In other cases there is a double peak but the peak on the red side is
-stronger \citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.  
-This has been readily explained as the consequence of multiple \lya photon
-scatterings through an homogeneous outflowing shell of neutral Hydrogen
+In many cases the \lya line profiles has a single peak
+redwards from the line's center, in other cases there is a double peak
+but the peak on the red side is stronger
+\citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
+These features been readily explained as the consequence of multiple
+\lya photon scatterings through an homogeneous outflowing shell of
+neutral Hydrogen
 \citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.
@@
 The six panels correspond to $\tau=10^6$ and a viewing angle of
-$\theta =90^{\deg}$, that is perpendicular to the rotation axis of the
+$\theta =90^{\circ}$, that is perpendicular to the rotation axis of the
 galaxy. 
@@
 
-In Figure \ref{fig:angle_influence} we show the same results as in
-Figure \ref{fig:doppler_shift} but for a viewing angle of $\theta =
-90^{\deg}$, that is parallel to the rotation axis. 
-In this case we see that the effect of rotation is very hard to see. 
-As \ref{Garavito14} showed, pure rotation introduces a strong
-dependence with viewing angle, a trend that we find also holds for
-rotation mixed with ouflows.
+In Figure \ref{fig:doppler_shift_theta_0} we show the same results as
+in Figure \ref{fig:doppler_shift} but for a viewing angle of $\theta =
+90^{\circ}$, that is parallel to the rotation axis. 
+In this case we confirme the result presented by \ref{Garavito14}
+showed, namely that pure rotation introduces a strong dependence with
+viewing angle, a trend that we find also holds for rotation mixed with
+ouflows.
 
@@
 
-After finding the qualitative influence of the different parameters.
-
-We evaluate the effects of increasing outflows velocity at fixed
-values for the optical depth and rotational velocity.
-Figure \ref{fig:varying_outflow_small} summarizes its behavior
-(for a largest range see Figure \ref{fig:varying_outflow}).
-All plots correspond to the same rotation velocity of
-$\vrot=50$\kms, rows have constant optical depth and columns
-have constant outflows velocity.
-
-Figure \ref{fig:doppler_shift} displays this Doppler shift for six different
-configurations with a fixed optical depth \tauh. In each cell of the grid
-there is a \lya spectrum for a galaxy with only outflows and two \lya spectra
-for the galaxy with outflows and rotation: one obtained from the radiative
-transfer simulation and the other obtained from the Doppler shifting of the
-non-rotating galaxy.
-
-% Esta figura deberia tener 2 x 2 vinetas. vrot=50.
-% cada panel tiene una curva para theta=0, 45, 90.
-% | tau=1E5, vout=25, | tau=1E5, vout=75 |
-% | tau=1E6, vout=25, | tau=1E6, vout=75 |
-%\begin{figure*}
-%\begin{center}
-%\includegraphics[width=0.90\textwidth]{./figures/results/varying_outflow_small}
-%\end{center}
-%\caption{\textbf{Spectra for fixed rotational velocity:} We fixed $\vrot=50$\kms.  We vary \vout in%creasing from left to right, \tauh increasing from top to
-			 bottom, and $\theta$ according to the legend.
-%\label{fig:varying_outflow_small}}
-%\end{figure*}
-
-\subsection{Doppler Shifts to Include Rotational Effects}
-
-% Esta figura deberia tener 3 x 2 viñetas. tau=10E6.
-% cada panel tiene una curva para no rotation, rotation (radiative transfer)
-% y rotation (doppler)
-% |vrot=50, vout=5,  | vrot=100, vout=5,  |
-% |vrot=50, vout=25, | vrot=100, vout=25, |
-% |vrot=50, vout=50, | vrot=100, vout=50, |
-
-In order to compare the line created with the radiative transfer (RT) process that
-involves the rotation and the line with only outflows that is Doppler
-shifted (DS), we calculate three main characteristics for each spectrum. We
-measure the line's standard deviation (STD), skewness (SKW) and bimodality (BI) 
-for both RT's and DS's approaches. We use the definitions in Eqs. \ref{eq:std}, 
-\ref{eq:skw}, \ref{eq:bi} for each value \citep{kokoska1999}.
+After finding the quantitative influence of the different parameters we
+move onto a qualilitative study.
+To do this we summarize the line morphology by four different
+characteristics: standard deviation (STD), skewness (SKW), bimodality
+(BI) and valley/peak ration.
+These quantities are defined by the following equations \citep{kokoska1999}.:
 
@@
 \end{equation}
-
-where each $m_i$ is the i-th moment about the mean. The STD has velocity units and 
-quantifies the dispersion of the line's velocity: the larger it is, the wider the line. 
-The SKW is adimensional and quantifies the peaks' asymmetry: if it is $>0$ the left peak is 
-taller and if it is $<0$ the right peak is taller. The BI is adimensional and
-quantifies if the line has 1 or 2 peaks: it is  always $\geq 1$ \citep{Pearson1929} and the closer it is to 1, the more bimodal is the line (i.e. has 2 peaks).
+%
+where each $m_i$ is the i-th moment about the mean. 
+The STD has velocity units and quantifies the dispersion of the line's
+velocity: the larger it is, the wider the line.  
+The SKW is adimensional and quantifies the peaks' asymmetry. 
+In the case of a bimodal distributio 
+$SKW>0$ the blue peak is taller and for $SWK<0$ the right peak is
+taller. The BI is adimensional and quantifies if the line has 1 or 2 peaks: it is  always $\geq 1$ \citep{Pearson1929} and the closer it is to 1, the more bimodal is the line (i.e. has 2 peaks).
 

### 2018-03-29 paper/paper.tex
edit

@@
 \label{sec:qualitative}
-Figure \ref{fig:doppler_shift} summarizes the most important trends of
-our simulations. 
+Figure \ref{fig:doppler_shift} summarizes the most important trends.
 The six panels correspond to $\tau=10^6$ and a viewing angle of
-$\theta =90^{\circ}$, that is perpendicular to the rotation axis of the
+$\theta =90^{\circ}$, that is, perpendicular to the rotation axis of the
 galaxy. 
 In every panel the thin black line corresponds to the pure outflow
-solution (i.e. without rotation). 
+solution, i.e. without rotation. 
 From top to bottom we see the effect of increasing the outflow
@@
 with the Doppler boost added to model rotation's influence.
-In all cases the Doppler boost does a good job at capturing the broad
+At $\tauh=10^6$ the Doppler boost does a good job at capturing the broad
 morphological features introduced by rotation: the angle dependence,
@@
 90^{\circ}$, that is parallel to the rotation axis. 
-In this case we confirme the result presented by \ref{Garavito14}
-showed, namely that pure rotation introduces a strong dependence with
+In this case we confirmethe result presented by \cite{Garavito14},
+namely that pure rotation introduces a strong dependence with 
 viewing angle, a trend that we find also holds for rotation mixed with
 ouflows.
+
+The quality of the results from the Doppler boos improves for higher
+\tauh values. 
+In the Appendix we show the same plots as Figures
+\ref{fig:doppler_shift} and \ref{fig:doppler_shift}, there it is
+evident that for $\tauh=10^5$ the results are not as good as they are
+for $\tauh=10^6$, and that for $\tauh=10^7$ the Doppler boos
+provides a remarkable good approximation.
 
@@
 After finding the quantitative influence of the different parameters we
-move onto a qualilitative study.
+move onto a qualitative study.
 To do this we summarize the line morphology by four different
@@
 \begin{center}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
 \end{center}
@@
 
-In this appendix we show \lya spectra with fixed \vout (Fig. \ref{fig:varying_rotation_small})
-and fixed \vrot (Fig. \ref{fig:varying_outflow_small}). For a larger set of spectra with the
-same fixed \vout and \vrot refer to Fig. \ref{fig:varying_rotation} and Fig. \ref{fig:varying_outflow}
-respectively.
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation_small}
-	\end{center}
-	\caption{\textbf{Spectra with fixed outflows velocity:} We fix $\vout=50$\kms.
-		\label{fig:varying_rotation_small}}
-\end{figure*}
-
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow_small}
-	\end{center}
-	\caption{\textbf{Spectra with fixed rotation velocity:} We fix $\vrot=50$\kms.
-		\label{fig:varying_outflow_small}}
-\end{figure*}
-
-% Esta figura deberia tener 3 x 3 vinetas. vout=50.
-% cada panel tiene una curva para theta=0, 45, 90.
-% |tau=1E5, vrot=0, | tau=1E5, vrot=50, | tau=1E5, vrot=100,
-% |tau=1E6, vrot=0, | tau=1E6, vrot=50, | tau=1E6, vrot=100,
-% |tau=1E7, vrot=0, | tau=1E7, vrot=50, | tau=1E7, vrot=100,
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_rotation}
-	\end{center}
-	\caption{\textbf{Spectra with fixed outflows velocity:} We fix $\vout=50$\kms.
-		\label{fig:varying_rotation}}
-\end{figure*}
-
-% Esta figura deberia tener 3 x 3 vinetas. vrot=50.
-% cada panel tiene una curva para theta=0, 45, 90.
-% |tau=1E5, vout=25, | tau=1E5, vout=50, | tau=1E5, vout=75,
-% |tau=1E6, vout=25, | tau=1E6, vout=50, | tau=1E6, vout=75,
-% |tau=1E7, vout=25, | tau=1E7, vout=50, | tau=1E7, vout=75,
-\begin{figure*}
-	\begin{center}
-		\includegraphics[width=0.90\textwidth]{./figures/appendix/varying_outflow}
-	\end{center}
-	\caption{\textbf{Spectra with fixed rotation velocity:} We fix $\vrot=50$\kms.
-		\label{fig:varying_outflow}}
-\end{figure*}
-
-% MUSE
-\begin{figure*}
-	\begin{center}
-		\includegraphics[height=0.3\textheight]{./figures/appendix/muse_vs_simulations.pdf}
-		\includegraphics[height=0.3\textheight]{./figures/appendix/museall_vs_muserange.pdf}
-		\includegraphics[height=0.3\textheight]{./figures/appendix/muserange_vs_simulations.pdf}
-	\end{center}
-	\caption{\textbf{Caption:} Caption.
-		\label{fig:varying_outflow}}
-\end{figure*}
-
 \end{document}
-
-\cite{DjorgovskiThompson}, \cite{Rhoads00},
-\cite{Gawiser2007}, \cite{Koehler2007}, \cite{Ouchi08},
-\cite{Yamada2012}, \cite{Schenker2012}, \cite{Kulas12},
-\cite{Yamada2012}, \cite{Chonis2013}, \cite{Finkelstein2013},
-\cite{Ostlin14}, \cite{Hayes2014}, \cite{Faisst2014},
-\cite{Fumagalli2015}.
-
-\section{A new LAE model}
-
-(\cite{Adams72}, \cite{Harrington73}, \cit
-e{Neufeld90}, \cite{Dijkstra06}, \cite{Verhamme06}, \cite{Forero12},
-\cite{Martin2015}, \cite{Garavito14}, \cite{Neufeld91},
-\cite{Laursen09}, \cite{Barnes11}, \cite{Verhamme12},
-\cite{Yajima12}).\\
-
-\subsection{Galaxy's Viewing Angle}
-
-We illustrate this in
-Fig. \ref{fig:model}. \\
-
-\begin{figure}[h!]
-  \begin{center}
-    \includegraphics[width=0.48\textwidth]{./figures/model}
-  \end{center}
-	\caption{\textbf{Model:} Spherical LAE with tangential and radial velocities due to rotation and outflows, respectively.
-The galaxy cut is in the $y-z$ plane perspective.
-The observer is located at an specific viewing angle of the sphere.
-Only photons with a direction that enters in this range of vision are taken into account to build the observed spectrum.
-		\label{fig:model}}
-\end{figure}
-
-\section{Results}
-
-\subsection{Resulting simulated spectra}
-In order to define the ranges of tauh, vrot and vout, it is necessary
-to refer to observational constraints. The common values for typical
-LAEs that were set as parameters are in Tab. \ref{tab:values}. We run
-CLARA's modified version for all the permutations of these 3
-parameters. \\
-
-The behavior of the resulting sets of spectra can be summarized by
-Fig. \ref{fig:summary}. We concluded from the simulations a clear
-creation of two asymmetric peaks around $V=0$ \kms with the tallest
-peak always redshifted. We detected a strong dependence on the outflow
-velocity that induces the peak asymmetry. We also detected that the
-rotation velocity broadens the line horizontally. Both these effects
-have been previously reported separately in the literature, so these
-results can valid their superposition to some extend. \\
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.48\textwidth]{./figures/summary}
-	\end{center}
-	\caption{\textbf{4 different \lya profiles:} With tauh$=10^5$ and $\theta \simeq 90^\circ$.
-The rotational velocity vrot increases to the right and the outflows velocity vout increases downwards.
-The intensity is in arbitrary units.
-		\label{fig:summary}}
-\end{figure}
-
-\subsection{Influence of the viewing angle $\theta$}
-We take now into account the viewing angle of the galaxy to build the observed spectra. For all of the physical parameters' combinations, the effect of $\theta$ in the \lya line is always the same.\\
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.5\textwidth]{./figures/angles}
-	\end{center}
-	\caption{\textbf{\lya profile for different $\theta$.} For tauh$=10^5$, vrot$=50$ \kms and vout$=20$ \kms.
-For tauh$=10^6$, vrot$=100$ \kms and vout$=5$ \kms.
-For tauh$=10^7$, vrot$=100$ \kms and vout$=15$ \kms.
-The intensity is in arbitrary units.
-		\label{fig:angles}}
-\end{figure}
-
-From Fig.
-\ref{fig:angles} it is clear that the intensity of the valley between the two peaks increases along with $\theta$.
-This causes an intensity decrease in the rest of the frequencies, thus a broadening of the line.
-The asymmetry also changes with the viewing angle.\\
-
-\subsection{Morphology of \lya line}
-
-To summarize, the influence of the 4 parameters on the \lya morphology is the following:
-
-\begin{itemize}
-	\item tauh induces a redshift. Increasing the optical depth separates the line of the zero velocity line. \\
-	\item vout decreases the right peak's intensity. Higher vout makes the left peak smaller until it merges with the right one. \\
-	\item vrot broadens the line and decreases the maximum intensity.
-Higher vrot implies a flatter spectrum.
-This effect has not been deeply studied in literature.
-Only Garavito et al.
-\cite{Garavito14} has simulated its effect.
-Our results are consistent with their conclusions.
-	\item $\theta$ increases the central valley of the line. Higher $\theta$ makes more emission at \lya natural frequency. \\
-\end{itemize}
-
-\subsection{Doppler shift by rotation}
-
-The rotation effect induces a Doppler shift in frequency and displaces the only-outflows spectra.
-As seen in Fig.
-\ref{fig:rotation_doppler_outflow} rotation takes the resulting spectrum with vrot$=0$ and displaces it from its central location.
-Then it weights the lines and they all merge into the resulting red \lya line.\\
-
-\begin{figure}[h!]
-	\begin{center}
-		\includegraphics[width=0.3\textwidth]{./figures/rotation_doppler_outflow}
-	\end{center}
-	\caption{\textbf{Rotations induces a Doppler shift of the only-outflows spectra:} Each of the black lines is then weighted to obtain the red line.
-		\label{fig:rotation_doppler_outflow}}
-\end{figure}
-
-PONER GRAFICA REAL...\\
-
-\section{Observational Implications}
-
-Among the observational implications that this paper has over observations we select three principal ones.
-First, the rotational effects of the galaxy should be clearly detected and characterized by MUSE's high resolution data.
-The Doppler redshift is already visible in other observations, such as the one seen in Fig.
-7 of Prescott et al.
-\cite{Prescott14}.\\
-
-Second, the central ($V=0$) emission of the spectra seen in the results is a consequence of the viewing angle of the galaxy and can be controlled by it.
-This means that the reason is not necessary that radiation is escaping without scattering, as several authors have suggested (CITAS...).\\
-
-Finally, all the spectra produced from this new LAE model is roughly consistent with observations.
-The consideration of adding the new rotation parameter to the standard only-outflows model found in literature is reasonable and very powerful.
-vrot and its consequent viewing angle $\theta$ can modify the morphology of the \lya line in new different ways by keeping the model simplified.\\
-
-\section{Conclusions}
-
-In this paper, the objective was to analyze and measure the influence
-of galaxy rotation and outflows on the \lya line. The motivation for
-this is to be able to obtain physical information of a LAE by just
-looking at its \lya profile. In order to accomplish this objective, we
-propose a new model of a LAE consisting of a sphere of Hydrogen atoms
-that expands radially and rotates as a solid body. A modified version
-of the program CLARA \cite{CLARA} is used to set the conditions and
-emulate the radiative transfer process inside the galaxy. \\
-
-The conclusions obtained from this work are:
-
-\begin{itemize}
-	\item The outgoing spectra depend on the angle an external observer is viewing the galaxy from.
-The closer it is to the equator of the galaxy, the higher the central valley of the frequency distribution.
-
-	\item The effects of vrot, vout and tauh are consistent with the different authors that have used them.
-vrot broadens the \lya line.
-vout increases the peaks asymmetry and tauh induces a redshift around the zero velocity.
-
-	\item Rotations induces a Doppler shift in frequency of the only-outflows spectra.
-
-	\item The final spectra obtained are roughly consistent with LAEs observations. \\
-
-\end{itemize}
-
-\subsection{Future work}
-
-For future work regarding the model, we would like to implement a differential rotation for the LAE instead of solid body.
-Regarding result analysis, we would like to compare against observations in two ways.
-Firstly by using the kinematic observation obtained from newly observed LAEs.
-We could do a consistency check between their real spectra and the simulated one that our model would produce with the respective parameters.
-Secondly we would like to take a specific observed spectrum and fit it using MCMC in order to predict ranges for its parameters.
-\\
-
-\subsection{Reproducibility}
-
-All of this work is available online and free to use to anyone.
-The data, source code and instructions to replicate this paper's results are in the GitHub repository:  \texttt{github.com/astroandes/CLARA\_RotationOutflows}.
-\\
-
-\section*{Acknowledgments}
-The authors would like to acknowledge the high performance computational cluster (HPC) hosted by Universidad de los Andes, where we ran the simulations for CLARA. \\
-
-We would also like to acknowledge the participants of the meeting \textit{Escape of Lyman radiation from galactic labyrinths} at Crete, Greece, for their valuable comments on this work, specially to Max Gronke and Christian Herenz.
-\\
-
-%------------------------REFERENCES----------------------------

### 2018-03-29 paper/paper.tex
new plots

@@
 \begin{center}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
-\includegraphics[height=0.31\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau5}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau5}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau6}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau6}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau7}
-\includegraphics[height=0.29\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau7}
 \end{center}
@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau5_theta90}
+  \end{center}
+  \caption{\textbf{Qualitative trends of changing outflow and
+      rotational velocity.}
+    Same layout as Figure \ref{fig:doppler_shift},
+    this time  $\tauh=10^5$ and $\theta=90^\circ$.}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau7_theta90}
+  \end{center}
+  \caption{\textbf{Qualitative trends of changing outflow and
+      rotational velocity.}
+    Same layout as Figure \ref{fig:doppler_shift},
+    this time  $\tauh=10^7$ and $\theta=90^\circ$.}
+\end{figure}
+
 \end{document}

### 2018-03-29 paper/paper.tex
edit

@@
 
-The quality of the results from the Doppler boos improves for higher
+The quality of the results from the Doppler boost improves for higher
 \tauh values. 
@@
 where each $m_i$ is the i-th moment about the mean. 
-The STD has velocity units and quantifies the dispersion of the line's
-velocity: the larger it is, the wider the line.  
+The STD has velocity units and quantifies the line's width.
 The SKW is adimensional and quantifies the peaks' asymmetry. 
-In the case of a bimodal distributio 
-$SKW>0$ the blue peak is taller and for $SWK<0$ the right peak is
-taller. The BI is adimensional and quantifies if the line has 1 or 2 peaks: it is  always $\geq 1$ \citep{Pearson1929} and the closer it is to 1, the more bimodal is the line (i.e. has 2 peaks).
+In the case of a bimodal distribution, $SKW>0$ means that the blue
+peak is taller and for $SWK<0$ the red peak is taller. 
+The BI is adimensional and quantifies whether the line has 1 or 2
+peaks: it is  always $\geq 1$ \citep{Pearson1929} and the 
+to 1, the more bimodal is the line (i.e. has 2 similar peaks). 
+We found by visual inspection of our spectra that BI$=2.5$ marks the
+transition between two peaks (however imbalanced) and a dominang
+single peak.
 

### 2018-03-29 paper/paper.tex
edit

@@
 
+Figure \ref{fig:standard_deviation} summarizes the results for STD.
+Each panel shows the STD as a function of \vrot.
+The black triangles correspond to the full RT solution and the line to
+the Doppler Shift approximation. 
+The optical depth increases from top to bottom and the outflow
+velocity from left to right.
+This quantitative plot confirms that the line width increases with
+rotational velocity, all else being equal.
+The Doppler Shift sucessfuly reproduces this trend, albeit a consistent
+underestimation of the RT values; this underestimation has as an upper
+bound of $\sim 7\%$, $3\%$ and $\sim 2\%$ for $\tauh=10^5$, $10^6$ and
+$10^7$, respectively.
+Tee Doppler Shift approximation also improves as the outflow velocity
+decreases.
+
+The mismatch between RT and DS decreases for higher
+
 We plot the results for each value in Figures \ref{fig:standard_deviation}, 

### 2018-03-29 paper/paper.tex
edit

@@
 features impact the morphology of the Lyman-$\alpha$ emission line.
+We explore a range of Hydrogen optical depth of
+$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
+outflow velocity $0\leq \vout/\kms\leq 50$.  
 We find three important consequences of rotation.
@@
 
-Figure \ref{fig:standard_deviation} summarizes the results for STD.
+Figure \ref{fig:standard_deviation} summarizes the STD results for all
+our models.
 Each panel shows the STD as a function of \vrot.

### 2018-03-30 paper/paper.tex
edit

--- 2018-03-29 paper/paper.tex
@@
 features impact the morphology of the Lyman-$\alpha$ emission line.
+We explore a range of Hydrogen optical depth of
+$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
+outflow velocity $0\leq \vout/\kms\leq 50$.  
 We find three important consequences of rotation.
-First, it increases the detected flux at the line's center in comparison with the
-non-rotational case, second it produces a line broadening and third it introduces
-a dependency with the  viewing angle.
+First, it introduces a dependency with viewing angle; second it
+produces a line broadening and third it increases the flux at the
+line's center.
 We also develp a semi-analytic model that modifies the spectra of
-pure outflow simulations that manages to reproduce the main features of the
-Monte Carlo simulations.
-As an application to observational data we constraint the rotation
-parameters of a set of observed LAE emission spectra.    
+pure outflow simulations that manages to reproduce the line broadening
+and flux change at the line's center within a $7\%$ and $30\%$
+precision, respectively.
 \end{abstract}
@@
 
-We plot the results for each value in Figures \ref{fig:standard_deviation}, 
-\ref{fig:skewness} and \ref{fig:bimodality}, respectively for $\tauh = 10^7$. 
-We highlight that these 3 plots increase their fit tendency with a higher optical 
-depth. The reasons why will be discussed in section \ref{sec:discussion}.
+Figure \ref{fig:standard_deviation} summarizes the standard deviation
+results for all our models.
+Each panel shows the STD as a function of \vrot.
+The black triangles correspond to the full RT solution and the line to
+the DS approximation. 
+The optical depth increases from top to bottom and the outflow
+velocity from left to right.
+This quantitative plot confirms that the line width increases with
+rotational velocity, all else being equal.
+The DS successfuly reproduces all trends with the optical depth,
+rotational velocity and outflow velocity.
+However, the DS consistently underestimates the STD. 
+The difference between the RT and DS increses with the outflow
+velocity and the rotational velocity, and decreases with increasing
+optical depth.
+In the range of parameter space explore, this difference has as an
+upper bound of $\sim 7\%$, $3\%$ and $\sim 2\%$ for $\tauh=10^5$,
+$10^6$ and $10^7$, respectively.
 
@@
 \end{center}
-\caption{\textbf{Standard Deviation for RT and DS:} The triangles represent the Radiative
-  Transfer and the lines represent the Doppler Shift.
+\caption{\textbf{Standard Deviation trends.} Results for all the
+  Radiative Transfer simulations (in triangles) compares against the
+  Doppler Shift model (lines).
   \label{fig:standard_deviation}}

### 2018-03-30 paper/paper.tex
edit

@@
 
-Finally, the thick gray line correspods to the pure outflow solution
+The thick gray line correspods to the pure outflow solution
 with the Doppler boost added to model rotation's influence.
@@
 results for all our models.
-Each panel shows the STD as a function of \vrot.
-The black triangles correspond to the full RT solution and the line to
-the DS approximation. 
+Each panel shows the STD as a function of \vrot. The black triangles
+correspond to the full RT solution and the line to the DS
+approximation.  
 The optical depth increases from top to bottom and the outflow
@@
 $10^6$ and $10^7$, respectively.
+
+Figure \ref{fig:skewness} presente the skwewness results for all the
+models together with the DS comparison.
 
@@
 
-This shows that the DS model can reproduce the main characteristics of the
-RT one. Because of this, we can introduce rotation effects in any spectra
-without running the radiative transfer simulations, that usually take a
-significant amount of time and computational resources.
-
-In addition to this, we also evaluate the effects of the viewing angle
-for fixed rotational and outflows velocity. We measure the intensity of the
-valley between peaks, for each $\theta$ and \tauh. Figure \ref{fig:valley_intensity}
-summarizes the information with all plots corresponding to the same outflow
-velocity of $\vout=5$\kms and rotation velocity of $\vrot=100$\kms.
-
-% Esta figura deberia tener en el eje x theta y en el eje y intensity of valley
-% cada panel tiene una curva para logtau = 5, 6, 7.
 \begin{figure*}

### 2018-03-30 paper/paper.tex
edit skewness

@@
 This quantitative plot confirms that the line width increases with
-rotational velocity, all else being equal.
+rotational velocity and  optical depth.
 The DS successfuly reproduces all trends with the optical depth,
@@
 However, the DS consistently underestimates the STD. 
-The difference between the RT and DS increses with the outflow
+The difference between the RT and DS increases with the outflow
 velocity and the rotational velocity, and decreases with increasing
@@
 In the range of parameter space explore, this difference has as an
-upper bound of $\sim 7\%$, $3\%$ and $\sim 2\%$ for $\tauh=10^5$,
-$10^6$ and $10^7$, respectively.
+upper bound of $\sim 7\%$, $3\%$ and $\sim 2\%$ for
+ $\tauh=10^5$, $10^6$ and $10^7$, respectively.
 
-Figure \ref{fig:skewness} presente the skwewness results for all the
-models together with the DS comparison.
+Figure \ref{fig:skewness} presents the skwewness results for all the
+models together with the DS comparison following the same layout as
+Figure \ref{fig:standard_deviation}.
+In all cases the skewness is negative showing that all the lines
+are unbalanced towards the red side of the spectrum.
+Skewness increases with rotational velocity and decreases with
+optical depth; rotation tries to smoothes the line diminishing the
+asymetries while a higher optical depth reinforces the line asymetries.
+The skewness does not have a monotonous trend with outflow velocity because
+there is a transition between double and single peak line; for low
+outflow velocities the skewness signals the balance between the two
+existing peaks while for high outflow velocities it quantifies the
+asymmetry of the already dominant read peak.
+The DS reproduces the main trends, again with an underestimation that
+decreases at higher optical depths and increases with larger values of
+the rotational velocity and outflow velocity.
+In this case the differences between RT and DS have an upper bound of
+$85\%$, $35\%$ and $5\%$ for  $\tauh=10^5$, $10^6$ and $10^7$, respectively.
 

### 2018-03-30 paper/paper.tex
edit

@@
 \vrot and \vrot values used in this paper.
-The range of values have some overlap with the expectations from a dwarf
-galaxy with a total neutral hydrogen mass of $10^8$-$10^9$ $M_{\odot}$.
-We run a total of $27$ different models.
-
-\begin{table}
-  \begin{center}
-    \begin{tabular}{|c|c|c|}
-      \hline
-      \tauh & \vrot (\kms) & \vout (\kms) \\
-      \hline
-      \{$10^5$, $10^6$, $10^7$\}  & \{0, 50, 100\} & \{5, 25, 50\} \\
-      \hline
-    \end{tabular}
-  \end{center}
-  \caption{\textbf{Parameters' Values.} List of values that were used
-    to construct the radiative transfer models. Values in braces
-    for a row indicate that all possible combinations in of
-    \tauh, \vrot and \vout are used.}
-  \label{tab:values}
-\end{table}
+The range of values have some overlap with the expectations from a
+galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
+$M_{\odot}$. We run a total of $27$ different models.
 
@@
 \section{Results}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau6_theta90}
+  \end{center}
+  \caption{\textbf{Qualitative trends of changing outflow and
+      rotational velocity viewed perpendicular to the rotation axis}. 
+    Here we fix $\tauh=10^6$ and $\theta=90^\circ$.
+    We vary \vrot increasing from left to right and \vout increasing
+    from top to bottom. 
+    The thin black line corresponds to the \lya line obtained with
+    CLARA without any rotation and the indicated outflow velocity.
+    The thick black line corresponds to the results including both
+    outflows and rotation.
+    The thick gray line shows the results of modifyin the pure outflow
+    solution by the Doppler shift presented in Equation \ref{eq:shift_x}
+    (in thin line), if there is a radiative transfer of rotation and outflows
+    (thick and clear line), and if there is a radiative transfer of
+    only outflows, but also a Doppler shift from the rotational
+    velocity (thick and dark line).  
+    \label{fig:doppler_shift}}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau6_theta0}
+  \end{center}
+  \caption{\textbf{Qualitative trends of changing outflow and
+      rotational velocity viewed parallel to the rotation axis}.     
+    Same layout as Figure \ref{fig:doppler_shift},
+    this time the viewing angle corresponds to $\theta=0^\circ$.
+    The effects of rotation on
+    the  \lya\ line morphology are not visible when viewed parallel to
+    the rotation axis.
+    \label{fig:doppler_shift_theta_0}}
+\end{figure}
+
 \label{sec:results}
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau6_theta90}
-  \end{center}
-  \caption{\textbf{Qualitative trends of changing outflow and
-      rotational velocity viewed perpendicular to the rotation axis}. 
-    Here we fix $\tauh=10^6$ and $\theta=90^\circ$.
-    We vary \vrot increasing from left to right and \vout increasing
-    from top to bottom. 
-    The thin black line corresponds to the \lya line obtained with
-    CLARA without any rotation and the indicated outflow velocity.
-    The thick black line corresponds to the results including both
-    outflows and rotation.
-    The thick gray line shows the results of modifyin the pure outflow
-    solution by the Doppler shift presented in Equation \ref{eq:shift_x}
-    (in thin line), if there is a radiative transfer of rotation and outflows
-    (thick and clear line), and if there is a radiative transfer of
-    only outflows, but also a Doppler shift from the rotational
-    velocity (thick and dark line).  
-    \label{fig:doppler_shift}}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau6_theta0}
-  \end{center}
-  \caption{\textbf{Qualitative trends of changing outflow and
-      rotational velocity viewed parallel to the rotation axis}.     
-    Same layout as Figure \ref{fig:doppler_shift},
-    this time the viewing angle corresponds to $\theta=0^\circ$.
-    The effects of rotation on
-    the  \lya\ line morphology are not visible when viewed parallel to
-    the rotation axis.
-    \label{fig:doppler_shift_theta_0}}
-\end{figure}
-
 \subsection{Quantitative trends}
@@
 results for all our models.
-Each panel shows the STD as a function of \vrot. The black triangles
+Each panel shows the STD as a function of \vrot.
+All panels were computed using a viewing angle of $\theta =
+90^{\circ}$ (perpendicular to the rotation axis), which has the most
+extreme ifluence from rotation.
+ The black triangles
 correspond to the full RT solution and the line to the DS
@@
 In this case the differences between RT and DS have an upper bound of
-$85\%$, $35\%$ and $5\%$ for  $\tauh=10^5$, $10^6$ and $10^7$, respectively.
+$85\%$, $35\%$ and $5\%$ for  $\tauh=10^5$, $10^6$ and $10^7$,
+respectively.
+
+Figure \ref{fig:bimodality} shows the results for the bimodality using
+the same layout as in the two previous Figures.
 

### 2018-03-30 paper/paper.tex
edit

@@
 \newcommand{\kms}{\ifmmode\mathrm{km\ s}^{-1}\else km s$^{-1}$\fi}
-\newcommand{\vrot}{\ifmmode\mathrm v_{\mathrm{rot}}\else $v_{\mathrm{rot}}$~\fi}
-\newcommand{\vout}{\ifmmode\mathrm v_{\mathrm{out}}\else $v_{\mathrm{out}}$~\fi}
-\newcommand{\tauh}{\ifmmode\mathrm \tau_{\mathrm{H}}\else $\tau_{\mathrm{H}}$~\fi}
-\newcommand{\vth}{\ifmmode\mathrm v_{\mathrm{th}}\else $v_{\mathrm{th}}$~\fi}
+\newcommand{\vrot}{\ifmmode v_{\mathrm{rot}}\else $v_{\mathrm{rot}}$~\fi}
+\newcommand{\vout}{\ifmmode v_{\mathrm{out}}\else $v_{\mathrm{out}}$~\fi}
+\newcommand{\tauh}{\ifmmode \tau_{\mathrm{H}}\else $\tau_{\mathrm{H}}$~\fi}
+\newcommand{\vth}{\ifmmode v_{\mathrm{th}}\else $v_{\mathrm{th}}$~\fi}
 
@@
 
+\begin{figure*}
+\begin{center}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
+\end{center}
+\caption{\textbf{Standard Deviation trends.} Results for all the
+  Radiative Transfer simulations (in triangles) compares against the
+  Doppler Shift model (lines).
+  \label{fig:standard_deviation}}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
+\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
+\end{center}
+\caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative
+    Transfer and the lines represent the Doppler Shift.
+    \label{fig:skewness}}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
+\end{center}
+\caption{\textbf{Bimodality for RT and DS:} The triangles represent the Radiative
+	Transfer and the lines represent the Doppler Shift.
+	\label{fig:bimodality}}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau5}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau6}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau7}
+\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau7}
+\end{center}
+\caption{\textbf{Valley Intensity:} We show for each \tauh the dependency that
+		the viewing angle $\theta$ has on the line's the valley intensity. The larger
+		$\cos{\theta}$ the lower is the valley intensity, so the higher $\theta$, the
+		higher the valley intensity. Also, the higher \tauh, the better is the fit RT-DS.
+		We fixed $\vrot=100$\kms.
+		\label{fig:valley_intensity}}
+\end{figure*}
+
 \label{sec:results}
@@
 
+\subsubsection{Standard Deviation}
 Figure \ref{fig:standard_deviation} summarizes the standard deviation
@@
 
+\subsubsection{Skewness}
+
 Figure \ref{fig:skewness} presents the skwewness results for all the
@@
 In this case the differences between RT and DS have an upper bound of
-$85\%$, $35\%$ and $5\%$ for  $\tauh=10^5$, $10^6$ and $10^7$,
+$85\%$, $35\%$ and $5\%$ 
+for  $\tauh=10^5$, $10^6$ and $10^7$,
 respectively.
 
+\subsection{Bimodality}
 Figure \ref{fig:bimodality} shows the results for the bimodality using
 the same layout as in the two previous Figures.
-
-\begin{figure*}
-\begin{center}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
-\end{center}
-\caption{\textbf{Standard Deviation trends.} Results for all the
-  Radiative Transfer simulations (in triangles) compares against the
-  Doppler Shift model (lines).
-  \label{fig:standard_deviation}}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
-\end{center}
-\caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative
-    Transfer and the lines represent the Doppler Shift.
-    \label{fig:skewness}}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
-\end{center}
-\caption{\textbf{Bimodality for RT and DS:} The triangles represent the Radiative
-	Transfer and the lines represent the Doppler Shift.
-	\label{fig:bimodality}}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau7}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau7}
-\end{center}
-\caption{\textbf{Valley Intensity:} We show for each \tauh the dependency that
-		the viewing angle $\theta$ has on the line's the valley intensity. The larger
-		$\cos{\theta}$ the lower is the valley intensity, so the higher $\theta$, the
-		higher the valley intensity. Also, the higher \tauh, the better is the fit RT-DS.
-		We fixed $\vrot=100$\kms.
-		\label{fig:valley_intensity}}
-\end{figure*}
+Following the reasoning about the skewness, we observe that
+increasing the outflow velocity increases the value of bimodality,
+that is, it transitions to a more pronounced single peak. 
+The trend as a function of the rotational velocity and the optical
+depth are not monotonous.
+When the outflow velocity is low (\vout$<50\ \kms$), an increasing
+rotational velocity smears the two asymmetrical peaks pushing the line
+morphology towards a single peaks, making the bimodality statistics
+increase. 
+On other situations (\vout$=50\ \kms$ and \tauh$\geq 10^6$) higher
+rotational velocities the bimodality statistics decreases, which means
+that it manages to slightly enlarge the already dominant red peak.
+The DS continous to reproduce the main trends, still underestimating
+the bimodality statistics.
+As expected from the previous results the difference betwen RT and DS
+decreases at higher optical depths and increases with increasing
+values of the rotational and outflow velocities.
+In this case the differences have an upper bound of $4\%$, $2\%$ and
+$1\%$ for  $\tauh=10^5$, $10^6$ and $10^7$, respectively.
 

### 2018-03-31 paper/paper.tex
edit

@@
 
-\subsection{Bimodality}
+\subsubsection{Bimodality}
+
 Figure \ref{fig:bimodality} shows the results for the bimodality using
@@
 $1\%$ for  $\tauh=10^5$, $10^6$ and $10^7$, respectively.
+
+\subsection{Intensity at line's center}
+
+In Figure \ref{fig:valley_intensity} we quantify how the intensity at
+the line's center (i.e. the valley) changes with the viewing angle,
+the outflow velocity and the optical depth.
+These results correspond to a fixed rotational velocity of
+\vrot$=100\ \kms$.
+The triangles correspond to the RT simulations and the line represents
+the DS results. 
+The valley intensity is expresed as a fraction of the maximum peak
+intensty in the line, as such the valley/peak ratio is always less
+than one. 
+In every panel we see that the valley/peak ratio decreases as the
+observer moves from a line  of sight perpendicular to the rotation
+axis onto a parallel line of sight. 
+This is a clear demonstration of the viewing angle dependency
+introduced by rotation.
 

### 2018-03-31 paper/paper.tex
edit

@@
 introduced by rotation.
+The valley/peak ratio at $\cos\theta=1$ matches results
+without rotation, this shows that for increasing rotational velocity
+the valley/peak ratio increases.
+In turn, for increasing optical depth or outflowing velocity this
+ratio decreases.
+Again, the DS results correctly follow the trends for the full RT
+simulations.
 

### 2018-03-31 paper/paper.tex
edit

@@
 We also develp a semi-analytic model that modifies the spectra of
-pure outflow simulations that manages to reproduce the line broadening
-and flux change at the line's center within a $7\%$ and $30\%$
+pure outflow simulations that manages to reproduce the line widths
+and flux change at the line's center within a $7\%$ and $50\%$
 precision, respectively.
@@
   Doppler Shift model (lines).
+  All panels correspond to a viewing angle of $\theta = 90^{\circ}$
+  (perpendicular to the rotation axis). 
+  The optical depth increases from top to bottom and the outflow
+  velocity from left to right.
   \label{fig:standard_deviation}}
@@
 \end{center}
-\caption{\textbf{Skewness for RT and DS:} The triangles represent the Radiative
-    Transfer and the lines represent the Doppler Shift.
+\caption{\textbf{Skewness trends.} Results for all the
+  Radiative Transfer simulations (in triangles) compares against the
+  Doppler Shift model (lines).
+  Follows the same layout as Figure \ref{fig:standard_deviation}. 
   \label{fig:skewness}}
@@
 \end{center}
-\caption{\textbf{Bimodality for RT and DS:} The triangles represent the Radiative
-	Transfer and the lines represent the Doppler Shift.
+\caption{\textbf{Bimodality trends.} Results for all the
+  Radiative Transfer simulations (in triangles) compares against the
+  Doppler Shift model (lines). 
+  Follows the same layout as Figure \ref{fig:standard_deviation}. 
   \label{fig:bimodality}}
@@
 rotational velocity and optical depth.
+These trends are expected; higher rotational velocities can be seen as
+an addition of different Doppler shifts that smear out the line, while
+a higher optical depth translates into a larger number of scatterings
+that increase the probability of a photon to diffuse in frequency
+resulting in a broader line.
+
 The DS successfuly reproduces all trends with the optical depth,
@@
 asymmetry of the already dominant read peak.
+
 The DS reproduces the main trends, again with an underestimation that
@@
 that it manages to slightly enlarge the already dominant red peak.
-The DS continous to reproduce the main trends, still underestimating
-the bimodality statistics.
+
+The DS reproduces the main trends while underestimating the bimodality
+statistics. 
 As expected from the previous results the difference betwen RT and DS
@@
 
-\subsection{Intensity at line's center}
+\subsubsection{Intensity at line's center}
 
@@
 introduced by rotation.
+
 The valley/peak ratio at $\cos\theta=1$ matches results
@@
 ratio decreases.
-Again, the DS results correctly follow the trends for the full RT
+Once again, the DS results correctly follow the trends for the full RT
 simulations.  
+This time the differences have an upper bound of $55\%$, $2\%$ and
+$1\%$ for  $\tauh=10^5$, $10^6$ and $10^7$, respectively.
 
@@
 \subsection{Theoretical Insights}
-\color{red}
-Why Doppler Shift works. Talk about radiative transfer...
-\color{black}
+
 In Fig. \ref{fig:doppler} we present the spectrum of a LAE taken from to different

### 2018-03-31 paper/paper.tex
edit

@@
 \usepackage{grffile}
+\usepackage{float}
 \usepackage[dvips]{epsfig}
 \usepackage{epsfig}
+\usepackage{dblfloatfix}
 \usepackage{color}
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau6_theta90}
-  \end{center}
+\begin{figure*}
+\centering
+    \includegraphics[width=0.48\textwidth]{./figures/results/doppler_shift_logtau6_theta90}
+    \includegraphics[width=0.48\textwidth]{./figures/results/doppler_shift_logtau6_theta0}
   \caption{\textbf{Qualitative trends of changing outflow and
-      rotational velocity viewed perpendicular to the rotation axis}. 
-    Here we fix $\tauh=10^6$ and $\theta=90^\circ$.
+      rotational velocity viewed perpendicular/paralell to the
+      rotation axis}.  
+    Here we fix $\tauh=10^6$. 
+    The six panels on the left correspond to $\theta=90^\circ$ and the
+    panels on the right to $\theta=0^{\circ}$
     We vary \vrot increasing from left to right and \vout increasing
@@
     \label{fig:doppler_shift}}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau6_theta0}
-  \end{center}
-  \caption{\textbf{Qualitative trends of changing outflow and
-      rotational velocity viewed parallel to the rotation axis}.     
-    Same layout as Figure \ref{fig:doppler_shift},
-    this time the viewing angle corresponds to $\theta=0^\circ$.
-    The effects of rotation on
-    the  \lya\ line morphology are not visible when viewed parallel to
-    the rotation axis.
-    \label{fig:doppler_shift_theta_0}}
-\end{figure}
+\end{figure*}
 
@@
 
-In Fig. \ref{fig:doppler} we present the spectrum of a LAE taken from to different
-sides of the galaxy. As the LAE is rotating, one side is being redshifted while
-the other is blueshifted. We see that the combined spectrum is a weighted line,
-in solid black, that combines these past two. This shows that there is in fact a
-Doppler Shift and then it justifies our choice to apply it after the outflows spectrum is
-obtained.
+\subsection{Links to observations}
+
+There are at least three inmmediate ways the work we have presented
+here can be useful to observational efforts.
+
+\subsubsection{Kinematic Maps}
+The first is related to the capability of spatially resolving the
+extent of a LAE.
+For instance \cite{Prescott14} presented observational results of a
+Doppler shift when taking spectra at two oposite sides of a LAE.
+
+In Figure \ref{fig:doppler} we present a toy model for the spectrum of
+a LAE taken from to different sides of the galaxy. 
+As the LAE is rotating, one side is being redshifted while the other
+is blueshifted. 
+We see that the full spectrum is a weighted line, in solid black,
+that is found between these two.
+In this toy model the rotational velocity is \vrot$=100\ \kms$. 
+We highlight that the distance between the maxima of the blue and red
+spectra is not twice the rotational velocity as it could be naively
+expected.
+Although the spectra can be modeled by Doppler shifts, the peak of the
+spectra is also weighted by the amount of mass that has a given
+velocity. 
+In this toy model
+
+New spectrographs like MUSE should help obtain kinematic information
+from LAEs to build velocity maps in \lya. 
+This could be a natural extension of the work reported by
+\cite{Herenz2016} on the velocity maps of several LARS (Lyman Alpha
+Reference Sample) galaxies.
+
+Another application of this new model to the current interpretation of \lya
+spectra, is that the central ($V=0$) emission of the spectra seen in the
+\lya line, is a consequence of the viewing angle of the galaxy and can be
+controlled by it. Several authors have suggested that this central emission
+is caused by radiation that escapes the galaxy without scattering. So we
+present an alternative that solves this issue. In addition, the effect of
+outflows velocities \vrot in current models can modify the asymmetry of peaks
+present in the \lya line. However the intensity of the valley and the width
+of the line were not easily reproduced. In this work we provide a tool to
+provide these effects with values of \vout and \vrot that do not exceed the
+typical LAEs' values found in the literature.
 
@@
 from a rotating galaxy is presented by \cite{Garavito14}. This derivation is
-based on the assumption that the distribution of photons' propagation directions
-at the edge of the galaxy is anisotropic and depends of \tauh only. So this
-approximation becomes more accurate with higher optical depth and it is also the
+based on the assumption that the distribution of photons' propagation
+directions at the edge of the galaxy is anisotropic and depends of
+\tauh only. So this approximation becomes more accurate with higher optical depth and it is also the
 reason why the Doppler Shift technique to induce rotational effects to an outflow
@@
 We find three different approaches how this model can have observational
-implications. Firstly, as the resolution of astronomical instruments is
-increasing, new images from LAEs can be spatially resolved in a way that
-rotation velocities of the galaxy can be determined. New spectrographs like
-MUSE should help obtain kinematic information from LAEs by using our model.
-As for now, previous works in the literatures have already found useful data
-to support this idea. Fig 7 of Prescott et al. \cite{Prescott14} shows the
-presence of Doppler shift and \cite{Herenz2016} creates 2D line of sight
-velocity maps of several LARS (Lyman Alpha Reference Sample) galaxies.
-
-Another application of this new model to the current interpretation of \lya
-spectra, is that the central ($V=0$) emission of the spectra seen in the
-\lya line, is a consequence of the viewing angle of the galaxy and can be
-controlled by it. Several authors have suggested that this central emission
-is caused by radiation that escapes the galaxy without scattering. So we
-present an alternative that solves this issue. In addition, the effect of
-outflows velocities \vrot in current models can modify the asymmetry of peaks
-present in the \lya line. However the intensity of the valley and the width
-of the line were not easily reproduced. In this work we provide a tool to
-provide these effects with values of \vout and \vrot that do not exceed the
-typical LAEs' values found in the literature.
-
-\color{red}
-Kulas...
-\color{black}
-Finally, we present a reconstruction of a selected \lya spectrum presented
-by \cite{Kulas12} and reproduced with the permission of the author. We create
-a fit of the \lya line and predict its kinematic properties.
+implications. Firstly,
 

### 2018-03-31 paper/paper.tex
edit

@@
 
-\section{Theoretical Models}
-\label{sec:theory}
-
-The Monte Carlo code we use (CLARA)  follows the propagation of
-individual photons through a neutral Hydrogen medium characterized by
-its temperature, velocity field and global optical depth.
-The code assumes an homogeneous density throughout the simulated
-volume.
-In the current implementation we neglect the influence of dust.
-Our basic model is an spherical distribution of neutral hydrogen,
-an approximation commonly used in the literature, as it explains a
-wide variety of observational features
-\citep{Ahn03,Verhamme06,Dijkstra06}.
-
-The velocity field we use captures both outflows and rotation.
-Outflows are described by a Hubble-like radial velocity profile with
-the velocity magnitude increasing linearly with the radial
-coordinate; the outflows model is fully characterized by $v_{\rm out}$, the
-velocity at the sphere's surface.
-Rotation follows a solid body rotation profile, which is fully
-characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.
-
-The total velocity field corresponds to the superposition of rotation and
-outflows.
-The cartesian components take the following form:
-
-\begin{equation}
-	v_{x}=\frac{x}{R}V_{\rm out}-\frac{y}{R}V_{\rm rot} ,
-	\label{eq:vx}
-\end{equation}
-
-\begin{equation}
-	v_{y}=\frac{y}{R}V_{\rm out}+\frac{x}{R}V_{\rm rot} ,
-	\label{eq:vy}
-\end{equation}
-
-\begin{equation}
-	v_{z}=\frac{z}{R}V_{\rm out},
-	\label{eq:vz}
-\end{equation}
-%
-where $x$, $y$ and $z$ are the cartesian position coordinates with the
-origin at the sphere's center, $R$ is the radius of the sphere and the
-direction of the angular velocity vector corresponds to the $\hat{k}$
-unit vector.
-
-For each model setup we follow $10^5$ individual photons generated at
-the center of the sphere at the \lya\ line's center as they propagate
-through the volume and finally scape.
-We store the final frequency and propagation direction for each photon
-at its last scattering.
-
-In Table \ref{tab:values} we list the combination of \tauh,
-\vrot and \vrot values used in this paper.
-The range of values have some overlap with the expectations from a
-galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
-$M_{\odot}$. We run a total of $27$ different models.
-
-\cite{Garavito14} presented an analytical model that
-accounts for the effects of pure rotation on the
-\lya\ line morphology. 
-The basic assumption of their analytical model is that each
-differential surface element on the sphere Doppler shifts the photons
-that it emits.
-In this paper we introduce this ansatz by post-processing the results
-of the outflows simulations without rotation.
-The frequency of each photon is Doppler shifted as follows
-
-\begin{equation}
-x' = x + \frac{\vec{V}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
-\label{eq:shift_x}
-\end{equation}
-%
-where $x'$ is the photon's new frequency, $x$ is the photon's
-frequency after being procesed by the outflow, $v_{\rm rot}$ is the 
-rotational velocity at the point of escape of the photon, $\hat{k}$ is
-the photon's direction of propagation and $v_{\rm th}$ is the thermal
-velocity of the sphere.
-
-This allows us to produce new \lya spectra and compare them with the
-full radiative transfer solution including both outflows and
-rotation.
-
-\section{Results}
-
 \begin{figure*}
@@
 \end{figure*}
+
+\section{Theoretical Models}
+\label{sec:theory}
+
+The Monte Carlo code we use (CLARA)  follows the propagation of
+individual photons through a neutral Hydrogen medium characterized by
+its temperature, velocity field and global optical depth.
+The code assumes an homogeneous density throughout the simulated
+volume.
+In the current implementation we neglect the influence of dust.
+Our basic model is an spherical distribution of neutral hydrogen,
+an approximation commonly used in the literature, as it explains a
+wide variety of observational features
+\citep{Ahn03,Verhamme06,Dijkstra06}.
+
+The velocity field we use captures both outflows and rotation.
+Outflows are described by a Hubble-like radial velocity profile with
+the velocity magnitude increasing linearly with the radial
+coordinate; the outflows model is fully characterized by $v_{\rm out}$, the
+velocity at the sphere's surface.
+Rotation follows a solid body rotation profile, which is fully
+characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.
+
+The total velocity field corresponds to the superposition of rotation and
+outflows.
+The cartesian components take the following form:
+
+\begin{equation}
+	v_{x}=\frac{x}{R}V_{\rm out}-\frac{y}{R}V_{\rm rot} ,
+	\label{eq:vx}
+\end{equation}
+
+\begin{equation}
+	v_{y}=\frac{y}{R}V_{\rm out}+\frac{x}{R}V_{\rm rot} ,
+	\label{eq:vy}
+\end{equation}
+
+\begin{equation}
+	v_{z}=\frac{z}{R}V_{\rm out},
+	\label{eq:vz}
+\end{equation}
+%
+where $x$, $y$ and $z$ are the cartesian position coordinates with the
+origin at the sphere's center, $R$ is the radius of the sphere and the
+direction of the angular velocity vector corresponds to the $\hat{k}$
+unit vector.
+
+For each model setup we follow $10^5$ individual photons generated at
+the center of the sphere at the \lya\ line's center as they propagate
+through the volume and finally scape.
+We store the final frequency and propagation direction for each photon
+at its last scattering.
+
+In Table \ref{tab:values} we list the combination of \tauh,
+\vrot and \vrot values used in this paper.
+The range of values have some overlap with the expectations from a
+galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
+$M_{\odot}$. We run a total of $27$ different models.
+
+\cite{Garavito14} presented an analytical model that
+accounts for the effects of pure rotation on the
+\lya\ line morphology. 
+The basic assumption of their analytical model is that each
+differential surface element on the sphere Doppler shifts the photons
+that it emits.
+In this paper we introduce this ansatz by post-processing the results
+of the outflows simulations without rotation.
+The frequency of each photon is Doppler shifted as follows
+
+\begin{equation}
+x' = x + \frac{\vec{V}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
+\label{eq:shift_x}
+\end{equation}
+%
+where $x'$ is the photon's new frequency, $x$ is the photon's
+frequency after being procesed by the outflow, $v_{\rm rot}$ is the 
+rotational velocity at the point of escape of the photon, $\hat{k}$ is
+the photon's direction of propagation and $v_{\rm th}$ is the thermal
+velocity of the sphere.
+
+This allows us to produce new \lya spectra and compare them with the
+full radiative transfer solution including both outflows and
+rotation.
+
+\section{Results}
 
@@
 
-In Figure \ref{fig:doppler} we present a toy model for the spectrum of
+In Figure \ref{fig:doppler} we present a toy model (\vrot$=50\ \kms$,
+\vout$=25\ \kms$ and \tauh$=10^5$) for the spectrum of
 a LAE taken from to different sides of the galaxy. 
@@
 that is found between these two.
-In this toy model the rotational velocity is \vrot$=100\ \kms$. 
-We highlight that the distance between the maxima of the blue and red
+We notice that the distance between the maxima of the blue and red
 spectra is not twice the rotational velocity as it could be naively
 expected.
-Although the spectra can be modeled by Doppler shifts, the peak of the
-spectra is also weighted by the amount of mass that has a given
-velocity. 
-In this toy model
+Although is a good approximation to model the rotating spectra by a
+sum of Doppler shifts, the peak of the spectra is also weighted by the
+amount of mass with a given velocity. 
+In this toy model the distance between the peaks of the
+receeding/approaching spectra is close to $\sim 20\ \kms$, which is
+less than half of the naively expected value of $2\vrot=100\ \kms$
 
@@
   \end{center}
-  \caption{\textbf{Doppler Shift:} We fix $\vout=25$\kms, $\vrot=100$\kms, $\tauh=10^5$.
+  \caption{\textbf{Spectra from receeding/approaching sides of a toy
+      model LAE.}  These results correspond to the RT simulation with
+    $\vout=25$\kms, $\vrot=50$\kms, $\tauh=10^5$. 
+    The spectra were computed for a viewing angle of
+    $\theta=90^{\circ}$. 
     \label{fig:doppler}}
@@
 where the only velocity factor affecting $v_z$ is the outflows one.
-
-\subsection{Application to observational data}
-
-We find three different approaches how this model can have observational
-implications. Firstly,
 

### 2018-03-31 paper/paper.tex
edit

@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
+  \end{center}
+  \caption{\textbf{Spectra from receeding/approaching sides of a toy
+      model LAE.}  These results correspond to the RT simulation with
+    $\vout=25$\kms, $\vrot=50$\kms, $\tauh=10^5$. 
+    The spectra were computed for a viewing angle of
+    $\theta=90^{\circ}$. 
+    \label{fig:doppler}}
+\end{figure}
+
 \subsection{Theoretical Insights}
+
+A first approach to an analytical expression that returns the \lya spectrum
+from a rotating galaxy is presented by \cite{Garavito14}.
+This derivation is based on the assumption that the distribution of
+photons' propagation directions at the edge of the galaxy is
+anisotropic and depends of \tauh only. 
+So this approximation becomes more accurate with higher optical depth
+and it is also the reason why the Doppler Shift technique to induce
+rotational effects to an outflow \lya spectrum works.
 
@@
 
-\subsubsection{Kinematic Maps}
+\subsubsection{\lya Kinematic Maps}
 The first is related to the capability of spatially resolving the
@@
 expected.
-Although is a good approximation to model the rotating spectra by a
+
+Although it is a good approximation to think the rotating spectra by a
 sum of Doppler shifts, the peak of the spectra is also weighted by the
@@
 receeding/approaching spectra is close to $\sim 20\ \kms$, which is
-less than half of the naively expected value of $2\vrot=100\ \kms$
+less than half of the naively expected value of $2\vrot=100\ \kms$,
+due to the fact that only a small fraction of the photons are emitted
+at the extreme of the galaxy having the maximum rotational velocity of
+$50\ \kms$.
 
@@
 Reference Sample) galaxies.  
+The interpretation of such data should take into account the insights
+and trends we have presented in this paper.
+
+\subsubsection{\lya continuum escape or rotation?}
 
@@
 typical LAEs' values found in the literature.
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
-  \end{center}
-  \caption{\textbf{Spectra from receeding/approaching sides of a toy
-      model LAE.}  These results correspond to the RT simulation with
-    $\vout=25$\kms, $\vrot=50$\kms, $\tauh=10^5$. 
-    The spectra were computed for a viewing angle of
-    $\theta=90^{\circ}$. 
-    \label{fig:doppler}}
-\end{figure}
-
-A first approach to an analytical expression that returns the \lya spectrum
-from a rotating galaxy is presented by \cite{Garavito14}. This derivation is
-based on the assumption that the distribution of photons' propagation
-directions at the edge of the galaxy is anisotropic and depends of
-\tauh only. So this approximation becomes more accurate with higher optical depth and it is also the
-reason why the Doppler Shift technique to induce rotational effects to an outflow
-\lya spectrum works.
-
-On the other hand, regarding the parameters \vrot, \vout, and $\theta$ for
-which this model would work, we have the following. The effect of the outflows
-velocity is much more sensitive on the spectrum than the rotational one; meaning
-that for a high \vout, the \vrot value would have to be equal or greater in order
-to be able to glimpse the shift in frequency that rotation induces. Regarding $\theta$,
-the introduction of rotation in the model implies that it will only affect galaxies
-with its viewing angle $\theta \neq 90^\circ$. This can be observed in the cartesian
-components of the velocity of the atoms (Eq. \ref{eq:vx}, \ref{eq:vy}, \ref{eq:vz})
-where the only velocity factor affecting $v_z$ is the outflows one.
 

### 2018-03-31 paper/paper.tex
edit

@@
 
-\subsubsection{\lya continuum escape or rotation?}
+\subsubsection{\lya continuum escaping or rotation?}
 
@@
 \lya line, is a consequence of the viewing angle of the galaxy and can be
-controlled by it. Several authors have suggested that this central emission
-is caused by radiation that escapes the galaxy without scattering. So we
-present an alternative that solves this issue. In addition, the effect of
-outflows velocities \vrot in current models can modify the asymmetry of peaks
-present in the \lya line. However the intensity of the valley and the width
-of the line were not easily reproduced. In this work we provide a tool to
-provide these effects with values of \vout and \vrot that do not exceed the
-typical LAEs' values found in the literature.
+controlled by it. 
+Several authors have suggested that this central emission is caused by
+radiation that escapes the galaxy without scattering. 
+Rotation is an alternative that solves this issue.
+
+We note that the relative intensity of the two peaks in the outflowing
+spectra is not modified by rotation, that asymmetry is solely
+controled by the outflow velocity.
+This means that results that have already derive a typical outflowing
+velocity as to match the observational constraints do not have to be
+completely revised. 
+Only an additional exploration of rotational velocities and viewing
+angles need to be explored.
 

### 2018-03-31 paper/paper.tex
edit

@@
 
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.48\textwidth]{./figures/results/muse_vs_simulations.pdf}
+  \end{center}
+  \caption{\textbf{Comparison between observations and simulations.}  
+    \label{fig:muse}}
+\end{figure}
+
 \subsection{Theoretical Insights}
@@
 
+\subsubsection{\lya continuum leakage or rotation?}
+
+Another application of this new model to the current interpretation of \lya
+spectra, is that the central ($V=0$) emission of the spectra seen in the
+\lya line, is a consequence of the viewing angle of the galaxy and can be
+controlled by it. 
+Several authors have suggested that this central emission is caused by
+radiation that escapes the galaxy without scattering. 
+Rotation is an alternative that solves this issue.
+
+We note that the relative intensity of the two peaks in the outflowing
+spectra is not modified by rotation, that asymmetry is solely
+controled by the outflow velocity.
+This means that results that have already derive a typical outflowing
+velocity as to match the observational constraints do not have to be
+completely revised. 
+Only an additional exploration of rotational velocities and viewing
+angles need to be explored.
+
 \subsubsection{\lya Kinematic Maps}
-The first is related to the capability of spatially resolving the
-extent of a LAE.
+
+Current observational facilities have the capability of spatially
+resolving the extent of a LAE.
 For instance \cite{Prescott14} presented observational results of a
@@
 
-\subsubsection{\lya continuum escaping or rotation?}
+\subsubsection{Match to \lya observations by MUSE}
 
-Another application of this new model to the current interpretation of \lya
-spectra, is that the central ($V=0$) emission of the spectra seen in the
-\lya line, is a consequence of the viewing angle of the galaxy and can be
-controlled by it. 
-Several authors have suggested that this central emission is caused by
-radiation that escapes the galaxy without scattering. 
-Rotation is an alternative that solves this issue.
-
-We note that the relative intensity of the two peaks in the outflowing
-spectra is not modified by rotation, that asymmetry is solely
-controled by the outflow velocity.
-This means that results that have already derive a typical outflowing
-velocity as to match the observational constraints do not have to be
-completely revised. 
-Only an additional exploration of rotational velocities and viewing
-angles need to be explored.
+In order to give the reader a feeling of the ranges of observational
+data that can be explained by the models presented here we perform a
+comparison against observational data from MUSE.
 

### 2018-03-31 paper/paper.tex
edit

@@
 We also develp a semi-analytic model that modifies the spectra of
-pure outflow simulations that manages to reproduce the line widths
+pure outflow simulations and manages to reproduce the line widths
 and flux change at the line's center within a $7\%$ and $50\%$
@@
 
-The interpretation of the Lyman-$\alpha$ emission in galaxies is
-key to understand their evolutionary processes, specially in
-star-forming, low-dust galaxies \citep{PartridgePeebles}.
 Recent improvements in instrumentation have revolutionized the kind of
-studies that can be performed on Lyman-$\alpha$ emitting galaxies
-(LAEs.)
-For instance, it is now possible to infer detailed kinematic maps for
-nearby galaxies.
-The study of these maps would allow us to build data-driven models to
-interpret the \lya spectra of unresolved galaxies, helping us to
-constrain the physical conditions of the interstellar medium (ISM)
-processing the \lya radiation.
-
-For instances, dwarf galaxies show signs of coherent rotation
-\citep{2009A&A...493..871S}.
-Some of them are expected to have high neutral gas contents and thus
-show \lya\ emission with rotation imprints
-\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
+kinematic studies that can be performed on galaxies with small angular
+sizes on the sky.
+For instance, spatially resolved spectra on dwarf galaxies show signs
+of coherent rotation \citep{2009A&A...493..871S}.
 Some Compact Dwarf Galaxies actually show gas kinematics that range
@@
 \citep{2015A&A...577A..21C,2017A&A...600A.125C}
+These studies are now being extended to distant galaxies.
 
+Some of those galaxies are expected to have high neutral gas contents
+and thus show \lya\ emission with rotation imprints
+\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
+The computational modeling of systems showing rotation has only been
+studied until recently by \cite{Garavito14}.
+In that work the authors studied the influence of pure solid body
+rotation on the \lya line's morpholgy.
+They found that rotation indeed introduces changes, the most
+noticeable change is the dependence of the spectra with the viewing
+angle with respect to the rotation axis.
+
+\cite{Garavito14} also presented a simple semi-analytical
+approximation that accounted for the main features of the \lya spectra
+from a rotation sphere.
+Recently, this semi-analytic solution was used to perform a Markov
+Chain Monte Carlo exploration to fit the observed spectra Compact
+Dwarf Galaxy \citep{tololo}, that presented atypical features that
+could be explained by pure rotation.
+
+However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies is
+more complex than pure rotation.
 The case for outflows as traced in the \lya\ line is much better
 stablished.
-In many cases the \lya line profiles has a single peak
+In many observations the \lya line profiles has a single peak
 redwards from the line's center, in other cases there is a double peak
@@
 
-Here we present for the first time a study of the joint effects of
-galaxy outflows and rotation, two of the three major kinematic
-components expected in LAEs.
+An study of the combined effects of outflows and rotation has not been
+presented in the literature.
+Here we present for the first time such as study. 
 We study a simplified geometrical configuration corresponding to a
-spherical gas cloud with symmetrical radial outflows and a rotation
-profile corresponding to a solid body.
-We base our modeling on a Monte-Carlo radiative transfer code called
-CLARA (Code for Lyman Alpha Radiation Analysis) presented for the
-first time by \cite{CLARA}.
-
+spherical gas cloud with symmetrical radial outflows and solid body
+rotation. 
+We base our modeling on a Monte-Carlo radiative transfer code.
 Besides modeling the impact of joint rotation and outflows, we also
@@
 applied in our case.
-We show how solid body rotation effects can be modelled by
-postprocessing the results of outflow only simulation.
-
-in the results of other radiative transfer model in the case of large
-optical depths.
-In this case it is a good approximation to Doppler
-boost the results of the model without rotation.
 

### 2018-03-31 paper/paper.tex
results

@@
 \citep{2015A&A...577A..21C,2017A&A...600A.125C}
-These studies are now being extended to distant galaxies.
+Some of those star forming galaxies are expected to have high neutral
+gas contents and low dust contents and thus show \lya\ emission with
+rotation imprints
+\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
 
-Some of those galaxies are expected to have high neutral gas contents
-and thus show \lya\ emission with rotation imprints
-\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
-The computational modeling of systems showing rotation has only been
-studied until recently by \cite{Garavito14}.
+The computational \lya modeling of systems showing rotation has only
+been studied until recently by \cite{Garavito14}.
 In that work the authors studied the influence of pure solid body
@@
 Chain Monte Carlo exploration to fit the observed spectra Compact
-Dwarf Galaxy \citep{tololo}, that presented atypical features that
-could be explained by pure rotation.
+Dwarf Galaxy \citep{tololo} with atypical features that could be
+explained by pure rotation.
 
@@
 
-In Table \ref{tab:values} we list the combination of \tauh,
-\vrot and \vrot values used in this paper.
+As input parameters we use \tauh$=\{10^5, 10^6, 10^7\}$,
+\vout$=\{5,25,50\}\kms$ and \vrot$=\{0,50,100\}\kms$, for a total of
+$27$ models with all the possible parameter combinations.
 The range of values have some overlap with the expectations from a
 galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
-$M_{\odot}$. We run a total of $27$ different models.
+$M_{\odot}$.
 
@@
 
-\subsubsection{Match to \lya observations by MUSE}
+\subsubsection{Relationship to MUSE observations}
 
@@
 data that can be explained by the models presented here we perform a
-comparison against observational data from MUSE.
+comparison against observational data from MUSE-Wide survey (MWS)
+\citep{2017A&A...606A..12H}.
+
+MWS provides public access to 831 spectra out of which 237 contain a
+detected \lya line as dominante. 
+The public dataset reports as the redshift of the galaxy the position
+of the dominant peak.
+We use this redshift to convert the spectra to rest-frame and compute
+the standard deviation and skewness statitstics. 
+Given the uncertainty on the galaxy redshift we do not compute the
+valley/peak ratio. 
+We do not try to perform a full fit to all the spectra but only to
+explore whether the region of parameter space explored in our RT
+simulations shows any crossing with the results from the MWS.
 
@@
 the results of a model for the emergent \lya\ line from rotating outflows.
-The main results for the model are computed from a Monte-Carlo radiative transfer
-simulation and confronted with a simple semi-analytic ansatz that adds the effects
-of rotation onto results of pure outflow kinematics.
+The main results for the model are computed from a Monte-Carlo
+radiative transfer simulation and confronted with a simple
+semi-analytic ansatz that adds the effects of rotation onto results of
+pure outflow kinematics.
 
-The main effects of rotation on the \lya\ line morphology are:
+We find that the main effects of rotation on the \lya\ line morphology are:
 
 \begin{itemize}
+  \item Inducing a dependency on the viewing angle.
   \item Broadening the line.
   \item Increasing the intensity at the line's center.
-	\item Inducing a dependency on the viewing angle,
-  The closer the observer is to the pole of the galaxy (defined by the rotation axis),
-  the smaller are the effects from rotation.
 \end{itemize}
 
-We found that in the case of solid body rotation these effects can be quantitatively
-explained by a Doppler boost, where the Doppler factor can be computed as the
- producet of quantities at the surface of last scattering, namely $\vec{v}_{\rm rot}\cdot \hat{k}$,
- where $\vec{v}_{\rm rot}$ is the velocity due to rotation and $\hat{k}$ is the direction
-of the photon's propagation.
-These conclusions, specially the confirmation of the Doppler boost as
-a good proxy for solid body rotation, strenghten the evidence reported
-by \cite{Garavito14} when they modeled the effect of pure rotation on
-\lya\ spectra.
+We find that these effects can be quantitatively explained by a
+Doppler shift computed as product of quantities at the surface of last
+scattering, namely  $\vec{v}_{\rm rot}\cdot \hat{k}$,  where
+$\vec{v}_{\rm rot}$ is the velocity due to rotation and $\hat{k}$ is
+the direction of the photon's propagation.  
+These conclusions, specially the confirmation of the Doppler Boost as
+a good approximation to model solid body rotation, strenghten the
+semi-analytic model reported by \cite{Garavito14} when they modeled
+the effect of pure rotation on \lya\ spectra.
 
-As an application to observational data we find that recent results
-that take the spectra of two different sides of a galaxy can detect 
-the effect of approaching/receeding gas motions. 
-However, the distances between the peaks of these two spectra 
-correspond to $\approx v_{\rm rot} \cos\theta$ due to the weights of the
-emitting regions.
+As an application to observational data our results suggest that
+rotation should also be considered as an explanation of the
+non-vanishing \lya intensity at the line's center.
+We also show how the \lya spectra from two different sides of a galaxy
+can detect approaching/receeding gas motions We show that the
+distances between the peaks of these two spectra is smaller than the
+expected value of $2\times\vrot$ due to the differente weights at the
+surface of the emitting regions.  
+Finally we also use data from MUSE to demonstrate to what extent the
+outputs of the RT results presented here can be expected to match
+hundreds of observed \lya spectra.
+
+Our work confirms that rotation produces features in the \lya line
+strong enough that can be expected to be detected in observations and
+that should be taken into account into \lya modelling; the Doppler
+Boost offers an easy-to-implement approximation to explore such
+influence into already existing RT simulations.
 
@@
 
+\newpage
 \appendix

### 2018-03-31 paper/paper.tex
refs

@@
 
-Recent improvements in instrumentation have revolutionized the kind of
+Recent advances in instrumentation have revolutionized the kind of
 kinematic studies that can be performed on galaxies with small angular
 sizes on the sky.
-For instance, spatially resolved spectra on dwarf galaxies show signs
-of coherent rotation \citep{2009A&A...493..871S}.
-Some Compact Dwarf Galaxies actually show gas kinematics that range
-from pure rotation to high velocity dispersion without a clear
-rotation pattern.
-\citep{2015A&A...577A..21C,2017A&A...600A.125C}
-Some of those star forming galaxies are expected to have high neutral
-gas contents and low dust contents and thus show \lya\ emission with
-rotation imprints
-\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
+For instances, spatially resolved spectra on compact dwarf galaxies
+are now possible, and they confirm signs of gas kinematics showing
+pure rotation kinematics \citep{2015A&A...577A..21C,2017A&A...600A.125C}
+Recent mapping of high redshift circumgalactic regions are also able
+to resolve kinematic evicend of large scale rotation
+\citep{2018MNRAS.473.3907A} 
+These systems with star formation, some neutral gas and low dust
+contents are the kind of systems that are naturally expected to have
+\lya emission.
+What is the expected imprint of rotation on a resonant line such as
+the \lya line?
 
-The computational \lya modeling of systems showing rotation has only
+The computational \lya modeling of rotating systems has only
 been studied until recently by \cite{Garavito14}.
@@
 For instance \cite{Prescott14} presented observational results of a
-Doppler shift when taking spectra at two oposite sides of a LAE.
+Doppler shift when taking spectra at two oposite sides of a large
+($\approx 80$kpc) LAE.
 
@@
 
-\subsubsection{Relationship to MUSE observations}
+\subsubsection{Comparison against the MUSE-Wide Survey}
 
 In order to give the reader a feeling of the ranges of observational
-data that can be explained by the models presented here we perform a
-comparison against observational data from MUSE-Wide survey (MWS)
-\citep{2017A&A...606A..12H}.
+data that could be explained by the models presented here, 
+we perform a comparison against observational data from the MUSE-Wide
+Survey (MWS) \citep{2017A&A...606A..12H}.
 
 MWS provides public access to 831 spectra out of which 237 contain a
-detected \lya line as dominante. 
+detected \lya line as dominant. 
 The public dataset reports as the redshift of the galaxy the position
 of the dominant peak.
-We use this redshift to convert the spectra to rest-frame and compute
+We use this redshift to convert the spectra to a rest-frame and compute
 the standard deviation and skewness statitstics. 
@@
 valley/peak ratio. 
-We do not try to perform a full fit to all the spectra but only to
-explore whether the region of parameter space explored in our RT
-simulations shows any crossing with the results from the MWS.
+We do not try to perform a full fit to all the spectra
+\citep[e.g.][]{2017A&A...608A.139G} and wish only to
+know whether the region of parameter space explored in our RT
+simulations shows any crossing with the results from the MWS or
+shows any sign of atypical features.
+
+Figure \ref{fig:muse} shows the results of this experiment. 
+The small white circles correspond to the 217 \lya emitters in the MWS
+observations; the large black circles to the results of the
+simulations presented in this paper. 
+This shows that the systems modeled in this paper actually present an
+overlap with statistics derived from observations. 
+The region with standard deviation $>200$\kms could also be explained
+by systems with larger optical depths than we have used here. 
+However, the systems with a positive skewness, that is a line
+unbalanced towards the blue could be the sign of inflowing
+kinematics.
 
@@
 Boost offers an easy-to-implement approximation to explore such
-influence into already existing RT simulations.
+influence into already existing RT simulations and provide stronger
+footing to the interpretation of \lya kinematics maps
+\citep[e.g][]{2018MNRAS.473.3907A}.
 

### 2018-03-31 paper/paper.tex
edit

@@
 sizes on the sky.
-For instances, spatially resolved spectra on compact dwarf galaxies
-are now possible, and they confirm signs of gas kinematics showing
-pure rotation kinematics \citep{2015A&A...577A..21C,2017A&A...600A.125C}
-Recent mapping of high redshift circumgalactic regions are also able
-to resolve kinematic evicend of large scale rotation
-\citep{2018MNRAS.473.3907A} 
+For instance, spatially resolved spectra on compact dwarf galaxies
+have measured clear signs of gas showing pure rotation kinematics
+\citep{2015A&A...577A..21C,2017A&A...600A.125C}; recent mapping of
+high redshift circumgalactic regions have also revealed kinematic
+evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
 These systems with star formation, some neutral gas and low dust
-contents are the kind of systems that are naturally expected to have
-\lya emission.
-What is the expected imprint of rotation on a resonant line such as
-the \lya line?
+contents naturally show a \lya emission line \citep{PartridgePeebles},
+motivating the observational work to phenomenologically link
+kinematics traced by H$\alpha$ to \lya spectra \citep[e.g.][]{Herenz2016}.
 
-The computational \lya modeling of rotating systems has only
-been studied until recently by \cite{Garavito14}.
+Theoretically speaking, what is then the expected imprint of rotation
+on a resonant emission line such as the \lya line? To what extent is
+possible to constrain rotational kinematics from the \lya emission line? 
+Detailed radiative transfer (RT) \lya modeling of rotating systems
+started until recently by \cite{Garavito14}.
 In that work the authors studied the influence of pure solid body
-rotation on the \lya line's morpholgy.
+rotation on the \lya line's morphology.
 They found that rotation indeed introduces changes, the most
-noticeable change is the dependence of the spectra with the viewing
+noticeable being the dependence of the spectra with the viewing
 angle with respect to the rotation axis.
@@
 \end{figure*}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
+  \end{center}
+  \caption{\textbf{Spectra from receeding/approaching sides of a toy
+      model LAE.}  These results correspond to the RT simulation with
+    $\vout=25$\kms, $\vrot=50$\kms, $\tauh=10^5$. 
+    The spectra were computed for a viewing angle of
+    $\theta=90^{\circ}$. This toy model illustrates to what extent spectra from
+    opposite sides of a galaxy have an imprint of the rotational
+    kinematics. 
+    \label{fig:doppler}}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=0.48\textwidth]{./figures/results/muse_vs_simulations.pdf}
+  \end{center}
+  \caption{\textbf{Comparison between simulations and observations.}  
+    \label{fig:muse}}
+\end{figure}
 
@@
 
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
-  \end{center}
-  \caption{\textbf{Spectra from receeding/approaching sides of a toy
-      model LAE.}  These results correspond to the RT simulation with
-    $\vout=25$\kms, $\vrot=50$\kms, $\tauh=10^5$. 
-    The spectra were computed for a viewing angle of
-    $\theta=90^{\circ}$. 
-    \label{fig:doppler}}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-    \includegraphics[width=0.48\textwidth]{./figures/results/muse_vs_simulations.pdf}
-  \end{center}
-  \caption{\textbf{Comparison between observations and simulations.}  
-    \label{fig:muse}}
-\end{figure}
-
 \subsection{Theoretical Insights}
@@
 
-We find that the main effects of rotation on the \lya\ line morphology are:
+To address the first questionswe posed in the introduction (What is then
+the expected imprint of rotation on a resonant emission line such as
+the \lya line?)
+
+We can saythat the main effects of rotation on the \lya\ line morphology are:
 
@@
 
-As an application to observational data our results suggest that
+Addressing the second question we posed in the introduction (To what
+extent is possible to constrain rotational kinematics from the \lya
+emission line?)  we can say that as an application to observational data our results suggest that
 rotation should also be considered as an explanation of the
 non-vanishing \lya intensity at the line's center.
-We also show how the \lya spectra from two different sides of a galaxy
-can detect approaching/receeding gas motions We show that the
-distances between the peaks of these two spectra is smaller than the
-expected value of $2\times\vrot$ due to the differente weights at the
-surface of the emitting regions.  
-Finally we also use data from MUSE to demonstrate to what extent the
+
+A more straightfoward approach is takes into account how the
+\lya spectra from two different sides of a galaxy can detect
+approaching/receeding gas motions.
+In that case one has to be careful and take into account that the
+peaks of these two spectra is smaller than the expected value of
+$2\times\vrot$ due to the differente weights at the surface of the
+emitting regions.
+
+We also use data from MUSE to demonstrate to what extent the
 outputs of the RT results presented here can be expected to match
-hundreds of observed \lya spectra.
+hundreds of observed \lya spectra, but a full constraint of rotation
+properties certainly require a full Markov Chain Monte Carlo
+exploration of parameter space.
 

### 2018-03-31 paper/paper.tex
references

@@
 \begin{keywords}
-galaxies: dwarf --- radiative transfer --- Methods: numerical
+radiative transfer --- Methods: numerical
 \end{keywords}
@@
 Recent advances in instrumentation have revolutionized the kind of
-kinematic studies that can be performed on galaxies with small angular
-sizes on the sky.
+kinematic studies that can be performed on extra-galactic systems with small angular
+sizes on the sky, revealing the presence of gas rotation on vastly
+different physical scales.
 For instance, spatially resolved spectra on compact dwarf galaxies
@@
 motivating the observational work to phenomenologically link
-kinematics traced by H$\alpha$ to \lya spectra \citep[e.g.][]{Herenz2016}.
+H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.
 
@@
 more complex than pure rotation.
-The case for outflows as traced in the \lya\ line is much better
+The case for outflows as traced in the \lya line is much better
 stablished.
@@
 presented in the literature.
-Here we present for the first time such as study. 
-We study a simplified geometrical configuration corresponding to a
+Here we present such a study. 
+We investigate a simplified geometrical configuration corresponding to a
 spherical gas cloud with symmetrical radial outflows and solid body

### 2018-03-31 paper/paper.tex
references

@@
 
-Recent advances in instrumentation have revolutionized the kind of
-kinematic studies that can be performed on extra-galactic systems with small angular
-sizes on the sky, revealing the presence of gas rotation on vastly
-different physical scales.
+Recent advances in instrumentation have revealed the presence of gas
+rotation on vastly different physical scales.
 For instance, spatially resolved spectra on compact dwarf galaxies
 have measured clear signs of gas showing pure rotation kinematics
-\citep{2015A&A...577A..21C,2017A&A...600A.125C}; recent mapping of
+\citep{2015A&A...577A..21C,2017A&A...600A.125C} and the recent mapping of
 high redshift circumgalactic regions have also revealed kinematic
 evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
-These systems with star formation, some neutral gas and low dust
-contents naturally show a \lya emission line \citep{PartridgePeebles},
+These systems with star formation, neutral gas and low dust
+contents naturally show a \lya emission line \citep{PartridgePeebles}
 motivating the observational work to phenomenologically link
@@
 
-Theoretically speaking, what is then the expected imprint of rotation
-on a resonant emission line such as the \lya line? To what extent is
-possible to constrain rotational kinematics from the \lya emission line? 
+What is then the expected imprint of rotation on a resonant emission
+line such as the \lya line? To what extent is possible to constrain
+rotational kinematics from the \lya emission line? 
 Detailed radiative transfer (RT) \lya modeling of rotating systems

### 2018-03-31 paper/paper.tex
edit

@@
 \begin{center}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}\\
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}\\
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
-\includegraphics[height=0.27\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}\\
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}\\
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}\\
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}\\
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau5}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau6}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau7}
-\includegraphics[height=0.28\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau5}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau5}\\
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau6}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau6}\\
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau7}
+\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau7}
 \end{center}
@@
 \begin{figure}
-  \begin{center}
+\centering
     \includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
-  \end{center}
   \caption{\textbf{Spectra from receeding/approaching sides of a toy
@@
 \begin{figure}
-  \begin{center}
+\centering
 \includegraphics[width=0.48\textwidth]{./figures/results/muse_vs_simulations.pdf}
-  \end{center}
+
 \caption{\textbf{Comparison between simulations and observations.}  
@@
 Several authors have suggested that this central emission is caused by
-radiation that escapes the galaxy without scattering. 
+radiation that escapes the galaxy without scattering, mostly due to a
+clumpy interstellear medium \citep{Hansen06, 2016ApJ...833L..26G}.
 Rotation is an alternative that solves this issue.
@@
 
+A recent observational example is provided by
+\cite{2017A&A...608L...4R}.
+They report optical spectroscopy of a bright lensed galaxy at a
+redshift of $z=2.4$. 
+The object presents a triple peaked \lya profile that can be thought
+as the superposition of two components, a double asymmetric peak and a
+narrower symmetric peak around the line's center.
+One possible option for the narrow component is that it escaped the
+galaxy without substantial interaction with neutral Hydrogen.
+We especulate that the central feature could also be due to a purely 
+rotating compont that can produce such a feature as shown by
+\cite{tololo}. The triple peak could be then produced by two main
+decopled components one rotating the other outflowing.
+
+Another interesting example are the \lya lines in Green Pea galaxies \citep{2016ApJ...820..130Y}.
+\cite{2017ApJ...844..171Y} showed in a study of 43 Green Pea galaxies
+that 2/3 were strong \lya emitters. 
+Because the \lya escape fraction is estimated from the ration of
+H$\alpha$ to \lya at the respective line's center, the authors infer
+large escape fractions for galaxies with substantial \lya emission at
+the lines'center.
+We especulate that these features could also be the product of
+rotation in those galaxies.
+
 \subsubsection{\lya Kinematic Maps}
@@
 ($\approx 80$kpc) LAE.
+In more recent work \cite{2018MNRAS.473.3907A} mapped \lya emission
+around a quasar
 
@@
 rotation should also be considered as an explanation of the
-non-vanishing \lya intensity at the line's center.
+non-vanishing \lya intensity at the line's center, which could change
+the interpretation of recent observational results
+\citep[e.g.][]{2017ApJ...844..171Y,2017A&A...608L...4R} that claim
+instead that this flux constrains the escape fraction of \lya
+radiation.
 
-A more straightfoward approach is takes into account how the
+A more straightfoward approach takes into account how the
 \lya spectra from two different sides of a galaxy can detect

### 2018-03-31 paper/paper.tex
edit

@@
 line's center.
-We also develp a semi-analytic model that modifies the spectra of
+We also test a semi-analytic model that modifies the spectra of
 pure outflow simulations and manages to reproduce the line widths
@@
 precision, respectively.
+We argue that rotation is a required kinematic feature to take into
+account for a correct interpretation of \lya spectra in systems where
+rotation is evidences through kinematic maps but also in other
+systems, such as Green Pea galaxies, where a strong \lya emission at
+the line's center could be due to rotational kinematics. 
 \end{abstract}

### 2018-03-31 paper/paper.tex
casi listo

@@
 \end{center}
-\caption{\textbf{Valley Intensity:} We show for each \tauh the dependency that
-		the viewing angle $\theta$ has on the line's the valley intensity. The larger
-		$\cos{\theta}$ the lower is the valley intensity, so the higher $\theta$, the
-		higher the valley intensity. Also, the higher \tauh, the better is the fit RT-DS.
-		We fixed $\vrot=100$\kms.
+\caption{\textbf{Valley Intensity. } We show for each \tauh the dependency that
+  the viewing angle $\theta$ has on the line's the
+  valley intensity. $\vrot=100$\kms is fixed for all panels.
 		\label{fig:valley_intensity}}
@@
 \caption{\textbf{Comparison between simulations and observations.}  
+The observational results correspond to the MUSE-Wide Survey
+\citep{2017A&A...606A..12H}. The skewness and standard deviation from
+the simulations summarize the results presented in Figures
+\ref{fig:standard_deviation} and \ref{fig:skewness}.
   \label{fig:muse}}
@@
 
-The Monte Carlo code we use (CLARA)  follows the propagation of
+The Monte Carlo code we use (CLARA) \citep{CLARA} follows the propagation of
 individual photons through a neutral Hydrogen medium characterized by
@@
 \begin{equation}
-	v_{x}=\frac{x}{R}V_{\rm out}-\frac{y}{R}V_{\rm rot} ,
+	v_{x}=\frac{x}{R}v_{\rm out}-\frac{y}{R}v_{\rm rot} ,
 	\label{eq:vx}
@@
 \begin{equation}
-	v_{y}=\frac{y}{R}V_{\rm out}+\frac{x}{R}V_{\rm rot} ,
+	v_{y}=\frac{y}{R}v_{\rm out}+\frac{x}{R}v_{\rm rot} ,
 	\label{eq:vy}
@@
 \begin{equation}
-	v_{z}=\frac{z}{R}V_{\rm out},
+	v_{z}=\frac{z}{R}v_{\rm out},
 	\label{eq:vz}
@@
 \begin{equation}
-x' = x + \frac{\vec{V}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
+x' = x + \frac{\vec{v}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
 \label{eq:shift_x}
@@
 %
-where $x'$ is the photon's new frequency, $x$ is the photon's
+where $x'$ is the photon's new adimansional frequency, $x$ is the photon's
 frequency after being procesed by the outflow, $v_{\rm rot}$ is the 
@@
 
-\subsection{Theoretical Insights}
-
-A first approach to an analytical expression that returns the \lya spectrum
-from a rotating galaxy is presented by \cite{Garavito14}.
-This derivation is based on the assumption that the distribution of
-photons' propagation directions at the edge of the galaxy is
-anisotropic and depends of \tauh only. 
-So this approximation becomes more accurate with higher optical depth
-and it is also the reason why the Doppler Shift technique to induce
-rotational effects to an outflow \lya spectrum works.
-
-\subsection{Links to observations}
-
-There are at least three inmmediate ways the work we have presented
-here can be useful to observational efforts.
+In this section we discuss how the results we have presented can be
+connected to the interpretation of observational data.
 
@@
 
-To address the first questionswe posed in the introduction (What is then
+To address the first question we posed in the introduction (\emph{ehat is then
 the expected imprint of rotation on a resonant emission line such as
-the \lya line?)
-
-We can saythat the main effects of rotation on the \lya\ line morphology are:
+the \lya line?}) we can say that the main effects of rotation on the
+\lya\ line morphology are:
 
@@
 
-Addressing the second question we posed in the introduction (To what
+Addressing the second question we posed in the introduction (\emph{to what
 extent is possible to constrain rotational kinematics from the \lya
-emission line?)  we can say that as an application to observational data our results suggest that
+emission line?})  we can say that as an application to observational data our results suggest that
 rotation should also be considered as an explanation of the

### 2018-04-01 paper/paper.tex
latest

@@
 %=========================================================================
-\title[Outflows and rotation in LAEs]{Lyman-$\alpha$ photons through rotating outflows}
+\title[Lyman-α photons through rotating outflows]{Lyman-$\alpha$ photons through rotating outflows}
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
@@
 features impact the morphology of the Lyman-$\alpha$ emission line.
-We explore a range of Hydrogen optical depth of
+We explore a range of neutral Hydrogen optical depth of
 $10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
@@
 We also test a semi-analytic model that modifies the spectra of
-pure outflow simulations and manages to reproduce the line widths
-and flux change at the line's center within a $7\%$ and $50\%$
-precision, respectively.
+pure outflow simulations and manages to reproduce the radiative
+transfer results for the line widths and flux change at the line's
+center within a $7\%$ and $50\%$ precision, respectively, for an
+optical depth of $\tauh=10^5$ and within $2\%$ and $1\%$ for an
+optical depth of $\tauh=10^7$.   
 We argue that rotation is a required kinematic feature to take into
 account for a correct interpretation of \lya spectra in systems where
-rotation is evidences through kinematic maps but also in other
+rotation is already evident through kinematic maps but also in other
 systems, such as Green Pea galaxies, where a strong \lya emission at
-the line's center could be due to rotational kinematics. 
+the line's center could be explained by rotational kinematics. 
 \end{abstract}
@@
 \begin{keywords}
-radiative transfer --- Methods: numerical
+galaxies:ISM --- line:profiles --- radiative transfer --- methods: numerical
 \end{keywords}
@@
 \centering
-    \includegraphics[width=0.48\textwidth]{./figures/results/doppler_shift_logtau6_theta90}
-    \includegraphics[width=0.48\textwidth]{./figures/results/doppler_shift_logtau6_theta0}
+    \includegraphics[width=0.48\textwidth]{doppler_shift_logtau6_theta90}
+    \includegraphics[width=0.48\textwidth]{doppler_shift_logtau6_theta0}
   \caption{\textbf{Qualitative trends of changing outflow and
@@
 \begin{center}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout5_logtau5.pdf}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout25_logtau5.pdf}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout50_logtau5.pdf}\\
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout5_logtau6.pdf}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout25_logtau6.pdf}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout50_logtau6.pdf}\\
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout5_logtau7.pdf}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout25_logtau7.pdf}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_std_vout50_logtau7.pdf}
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout5_logtau5.pdf}
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout25_logtau5.pdf}
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout50_logtau5.pdf}\\
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout5_logtau6.pdf}
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout25_logtau6.pdf}
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout50_logtau6.pdf}\\
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout5_logtau7.pdf}
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout25_logtau7.pdf}
+\includegraphics[height=0.25\textwidth]{line_characterization_std_vout50_logtau7.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout5_logtau5}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout25_logtau5}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout50_logtau5}\\
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout5_logtau6}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout25_logtau6}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout50_logtau6}\\
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout5_logtau7}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout25_logtau7}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_skw_vout50_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout5_logtau5}
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout25_logtau5}
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout50_logtau5}\\
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout5_logtau6}
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout25_logtau6}
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout50_logtau6}\\
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout5_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout25_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_skw_vout50_logtau7}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout5_logtau5}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout25_logtau5}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout50_logtau5}\\
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout5_logtau6}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout25_logtau6}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout50_logtau6}\\
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout5_logtau7}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout25_logtau7}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_bi_vout50_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout5_logtau5}
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout25_logtau5}
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout50_logtau5}\\
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout5_logtau6}
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout25_logtau6}
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout50_logtau6}\\
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout5_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout25_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_bi_vout50_logtau7}
 \end{center}
@@
 \begin{center}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau5}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau5}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau5}\\
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau6}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau6}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau6}\\
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout5_vrot100_logtau7}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout25_vrot100_logtau7}
-\includegraphics[height=0.25\textwidth]{./figures/results/line_characterization_vi_vout50_vrot100_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout5_vrot100_logtau5}
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout25_vrot100_logtau5}
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout50_vrot100_logtau5}\\
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout5_vrot100_logtau6}
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout25_vrot100_logtau6}
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout50_vrot100_logtau6}\\
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout5_vrot100_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout25_vrot100_logtau7}
+\includegraphics[height=0.25\textwidth]{line_characterization_vi_vout50_vrot100_logtau7}
 \end{center}
@@
 \centering
-    \includegraphics[width=0.48\textwidth]{./figures/discussion/doppler}
+    \includegraphics[width=0.48\textwidth]{doppler}
   \caption{\textbf{Spectra from receeding/approaching sides of a toy
@@
 \centering
-\includegraphics[width=0.48\textwidth]{./figures/results/muse_vs_simulations.pdf}
+\includegraphics[width=0.44\textwidth]{muse_vs_simulations.pdf}
 
@@
 \cite{tololo}. The triple peak could be then produced by two main
-decopled components one rotating the other outflowing.
+decopled components: one rotating the other outflowing.
 
@@
 that 2/3 were strong \lya emitters. 
-Because the \lya escape fraction is estimated from the ration of
-H$\alpha$ to \lya at the respective line's center, the authors infer
-large escape fractions for galaxies with substantial \lya emission at
+There the \lya escape fraction is effectively estimated from the ratio
+of H$\alpha$ to \lya.
+Galaxies with large escape fractions also substantial \lya emission at
 the lines'center. 
-We especulate that these features could also be the product of
-rotation in those galaxies.
+We especulate that this feature could help to constrain rotation in
+Green Pea galaxies.
 
@@
   \begin{center}
-    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau5_theta90}
+    \includegraphics[width=0.49\textwidth]{doppler_shift_logtau5_theta90}
   \end{center}
@@
   \begin{center}
-    \includegraphics[width=0.49\textwidth]{./figures/results/doppler_shift_logtau7_theta90}
+    \includegraphics[width=0.49\textwidth]{doppler_shift_logtau7_theta90}
   \end{center}

### 2018-04-01 paper/paper.tex
minor typos

@@
 The case for outflows as traced in the \lya line is much better
-stablished.
+established.
 In many observations the \lya line profiles has a single peak
@@
     outflows and rotation.
-    The thick gray line shows the results of modifyin the pure outflow
+    The thick gray line shows the results of modifying the pure outflow
     solution by the Doppler shift presented in Equation \ref{eq:shift_x}
@@
     \includegraphics[width=0.48\textwidth]{doppler}
-  \caption{\textbf{Spectra from receeding/approaching sides of a toy
+  \caption{\textbf{Spectra from receding/approaching sides of a toy
       model LAE.}  These results correspond to the RT simulation with
@@
 %
-where $x'$ is the photon's new adimansional frequency, $x$ is the photon's
-frequency after being procesed by the outflow, $v_{\rm rot}$ is the 
+where $x'$ is the photon's new adimensional frequency, $x$ is the photon's
+frequency after being processed by the outflow, $v_{\rm rot}$ is the 
 rotational velocity at the point of escape of the photon, $\hat{k}$ is
@@
 
-The thick gray line correspods to the pure outflow solution
+The thick gray line corresponds to the pure outflow solution
 with the Doppler boost added to model rotation's influence.
@@
 We found by visual inspection of our spectra that BI$=2.5$ marks the
-transition between two peaks (however imbalanced) and a dominang
+transition between two peaks (however imbalanced) and a dominant
 single peak.
@@
 90^{\circ}$ (perpendicular to the rotation axis), which has the most
-extreme ifluence from rotation.
+extreme influence from rotation.
  The black triangles
@@
 
-The DS successfuly reproduces all trends with the optical depth,
+The DS successfully reproduces all trends with the optical depth,
 rotational velocity and outflow velocity.
@@
 
-Figure \ref{fig:skewness} presents the skwewness results for all the
+Figure \ref{fig:skewness} presents the skewness results for all the
 models together with the DS comparison following the same layout as
@@
 Skewness increases with rotational velocity and decreases with
-optical depth; rotation tries to smoothes the line diminishing the
-asymetries while a higher optical depth reinforces the line asymetries.
+optical depth; rotation tries to smooth the line diminishing the
+asymmetries while a higher optical depth reinforces the line asymmetries.
 The skewness does not have a monotonous trend with outflow velocity because
@@
 statistics. 
-As expected from the previous results the difference betwen RT and DS
+As expected from the previous results the difference between RT and DS
 decreases at higher optical depths and increases with increasing
@@
 the DS results. 
-The valley intensity is expresed as a fraction of the maximum peak
-intensty in the line, as such the valley/peak ratio is always less
+The valley intensity is expressed as a fraction of the maximum peak
+intensity in the line, as such the valley/peak ratio is always less
 than one. 
@@
 radiation that escapes the galaxy without scattering, mostly due to a
-clumpy interstellear medium \citep{Hansen06, 2016ApJ...833L..26G}.
+clumpy interstellar medium \citep{Hansen06, 2016ApJ...833L..26G}.
 Rotation is an alternative that solves this issue.
@@
 spectra is not modified by rotation, that asymmetry is solely
-controled by the outflow velocity.
+controlled by the outflow velocity.
 This means that results that have already derive a typical outflowing
@@
 galaxy without substantial interaction with neutral Hydrogen.
-We especulate that the central feature could also be due to a purely 
-rotating compont that can produce such a feature as shown by
+We speculate that the central feature could also be due to a purely 
+rotating component that can produce such a feature as shown by
 \cite{tololo}. The triple peak could be then produced by two main
@@
 Galaxies with large escape fractions also substantial \lya emission at
-the lines'center. 
-We especulate that this feature could help to constrain rotation in
+the line's center. 
+We speculate that this feature could help to constrain rotation in
 Green Pea galaxies.
@@
 In this toy model the distance between the peaks of the
-receeding/approaching spectra is close to $\sim 20\ \kms$, which is
+receding/approaching spectra is close to $\sim 20\ \kms$, which is
 less than half of the naively expected value of $2\vrot=100\ \kms$,
@@
 We use this redshift to convert the spectra to a rest-frame and compute
-the standard deviation and skewness statitstics. 
+the standard deviation and skewness statistics. 
 Given the uncertainty on the galaxy redshift we do not compute the
@@
 
-To address the first question we posed in the introduction (\emph{ehat is then
+To address the first question we posed in the introduction (\emph{what is then
 the expected imprint of rotation on a resonant emission line such as
@@
 These conclusions, specially the confirmation of the Doppler Boost as
-a good approximation to model solid body rotation, strenghten the
+a good approximation to model solid body rotation, strengthen the
 semi-analytic model reported by \cite{Garavito14} when they modeled
@@
 
-A more straightfoward approach takes into account how the
+A more straightforward approach takes into account how the
 \lya spectra from two different sides of a galaxy can detect
-approaching/receeding gas motions.
+approaching/receding gas motions.
 In that case one has to be careful and take into account that the
 peaks of these two spectra is smaller than the expected value of
-$2\times\vrot$ due to the differente weights at the surface of the
+$2\times\vrot$ due to the different weights at the surface of the
 emitting regions.
@@
 strong enough that can be expected to be detected in observations and
-that should be taken into account into \lya modelling; the Doppler
+that should be taken into account into \lya modeling; the Doppler
 Boost offers an easy-to-implement approximation to explore such

### 2018-04-02 paper/paper.tex
british package to break words in line ends
minor changes

@@
 \usepackage{bm}
+\usepackage[british]{babel}
 %Non reposionated tables
@@
 \newcommand{\vth}{\ifmmode v_{\mathrm{th}}\else $v_{\mathrm{th}}$~\fi}
+\newcommand{\hatk}{\ifmmode \hat{k}\else $\hat{k}$~\fi}
+\newcommand{\STD}{\ifmmode \mathrm{STD}\else $\mathrm{STD}$~\fi}
+\newcommand{\SKW}{\ifmmode \mathrm{SKW}\else $\mathrm{SKW}$~\fi}
+\newcommand{\BI}{\ifmmode \mathrm{BI}\else $\mathrm{BI}$~\fi}
 
@@
 What is then the expected imprint of rotation on a resonant emission
-line such as the \lya line? To what extent is possible to constrain
+line such as the \lya line? To what extent is it possible to constrain
 rotational kinematics from the \lya emission line? 
@@
 
-However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies is
+However, the gas dynamics in Lyman Alpha Emitter (LAE) galaxies are
 more complex than pure rotation.
-The case for outflows as traced in the \lya line is much better
+The case for outflows as traced in the \lya line is solidly
 established.  
-In many observations the \lya line profiles has a single peak
+In many observations the \lya line profile has a single peak
 redwards from the line's center, in other cases there is a double peak
@@
 \citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
-These features been readily explained as the consequence of multiple
-\lya photon scatterings through an homogeneous outflowing shell of
+These features have been explained as the consequence of multiple
+\lya photon scatterings through a homogeneous outflowing shell of
 neutral Hydrogen
@@
 
-An study of the combined effects of outflows and rotation has not been
-presented in the literature.
-Here we present such a study. 
+Nevertheless, an study of the combined effects of outflows and
+rotation has not been presented in the literature.
+Here we report on such a study. 
 We investigate a simplified geometrical configuration corresponding to a
@@
 Besides modeling the impact of joint rotation and outflows, we also
-check to what extent the analytical model presented by
-\cite{Garavito14} to explain the effects of rotation can also be
-applied in our case.
+check to what extent the semi-analytical model presented by
+\cite{Garavito14} an also be applied in our case.
 
@@
 \centering
-\includegraphics[width=0.44\textwidth]{muse_vs_simulations.pdf}
+\includegraphics[width=0.45\textwidth]{muse_vs_simulations.pdf}
 
@@
 its temperature, velocity field and global optical depth.
-The code assumes an homogeneous density throughout the simulated
+The code assumes a homogeneous density throughout the simulated
 volume.
@@
 coordinate; the outflows model is fully characterized by $v_{\rm out}$, the
-velocity at the sphere's surface.
+radial velocity at the sphere's surface.
 Rotation follows a solid body rotation profile, which is fully
-characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.
+characterized by $v_{\rm rot}$, the linear velocity at the sphere's surface.
 
@@
 \begin{equation}
-	v_{x}=\frac{x}{R}v_{\rm out}-\frac{y}{R}v_{\rm rot} ,
+	v_{x}=\frac{x}{R}\vout - \frac{y}{R}\vrot ,
 	\label{eq:vx}
@@
 \begin{equation}
-	v_{y}=\frac{y}{R}v_{\rm out}+\frac{x}{R}v_{\rm rot} ,
+	v_{y}=\frac{y}{R}\vout + \frac{x}{R}\vrot ,
 	\label{eq:vy}
@@
 \begin{equation}
-	v_{z}=\frac{z}{R}v_{\rm out},
+	v_{z}=\frac{z}{R}\vout,
 	\label{eq:vz}
@@
 origin at the sphere's center, $R$ is the radius of the sphere and the
-direction of the angular velocity vector corresponds to the $\hat{k}$
-unit vector.
+direction of the angular velocity vector is the $z$ axis.
 
 For each model setup we follow $10^5$ individual photons generated at
-the center of the sphere at the \lya\ line's center as they propagate
+the center of the sphere at the \lya line's center as they propagate
 through the volume and finally scape.
@@
 
-As input parameters we use \tauh$=\{10^5, 10^6, 10^7\}$,
-\vout$=\{5,25,50\}\kms$ and \vrot$=\{0,50,100\}\kms$, for a total of
+As input parameters we use $\tauh=\{10^5, 10^6, 10^7\}$,
+$\vout=\{5,25,50\}\;\kms$ and $\vrot=\{0,50,100\}\;\kms$, for a total of
 $27$ models with all the possible parameter combinations.
@@
 accounts for the effects of pure rotation on the
-\lya\ line morphology. 
+\lya line morphology. 
 The basic assumption of their analytical model is that each
@@
 \begin{equation}
-x' = x + \frac{\vec{v}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
+x' = x + \frac{\vec{v}_{\rm rot}\cdot \hatk}{\vth}
 \label{eq:shift_x}
@@
 where $x'$ is the photon's new adimensional frequency, $x$ is the photon's
-frequency after being processed by the outflow, $v_{\rm rot}$ is the 
-rotational velocity at the point of escape of the photon, $\hat{k}$ is
-the photon's direction of propagation and $v_{\rm th}$ is the thermal
+frequency after being processed only by the outflow, \vrot is the 
+rotational velocity at the point of escape of the photon, \hatk is
+the photon's direction of propagation and \vth is the thermal
 velocity of the sphere.
+We also define the viewing angle, $\theta$, as the angle between the
+rotation axis and the line of sight of a potential observer.
 
@@
 \section{Results}
-
 \label{sec:results}
@@
 \label{sec:qualitative}
-Figure \ref{fig:doppler_shift} summarizes the most important trends.
-The six panels correspond to $\tau=10^6$ and a viewing angle of
+
+Figure \ref{fig:doppler_shift} summarizes the most important trends
+from rhw RT simulations.
+In the left side, the six panels correspond to $\tau=10^6$ and a viewing angle of
 $\theta =90^{\circ}$, that is, perpendicular to the rotation axis of the
@@
 The thick gray line corresponds to the pure outflow solution
-with the Doppler boost added to model rotation's influence.
-At $\tauh=10^6$ the Doppler boost does a good job at capturing the broad
+with the Doppler shift added to model rotation's influence.
+At $\tauh=10^6$ the Doppler shift does a good job at capturing the broad
 morphological features introduced by rotation: the angle dependence,
@@
 
-In Figure \ref{fig:doppler_shift_theta_0} we show the same results as
-in Figure \ref{fig:doppler_shift} but for a viewing angle of $\theta =
-90^{\circ}$, that is parallel to the rotation axis. 
-In this case we confirmethe result presented by \cite{Garavito14},
+In the right side of Figure \ref{fig:doppler_shift} we show the same results as in the left one, but for a viewing angle of $\theta =
+0^{\circ}$, that is parallel to the rotation axis. 
+In this case we confirm the result presented by \cite{Garavito14},
 namely that pure rotation introduces a strong dependence with 
 viewing angle, a trend that we find also holds for rotation mixed with
-ouflows.
+outflows.
 
-The quality of the results from the Doppler boost improves for higher
+The quality of the results from the Doppler shift improves for higher
 \tauh values. 
-In the Appendix we show the same plots as Figures
-\ref{fig:doppler_shift} and \ref{fig:doppler_shift}, there it is
+In the Appendix we show the same plots as Figure \ref{fig:doppler_shift}, there it is
 evident that for $\tauh=10^5$ the results are not as good as they are
-for $\tauh=10^6$, and that for $\tauh=10^7$ the Doppler boos
+for $\tauh=10^6$, and that for $\tauh=10^7$ the Doppler shift
 provides a remarkable good approximation.
@@
 
-After finding the quantitative influence of the different parameters we
-move onto a qualitative study.
+After finding the qualitative influence of the different parameters we
+move onto a quantitative study.
 To do this we summarize the line morphology by four different
-characteristics: standard deviation (STD), skewness (SKW), bimodality
-(BI) and valley/peak ration.
-These quantities are defined by the following equations \citep{kokoska1999}.:
+scalars: standard deviation (STD), skewness (SKW), bimodality
+(BI) and valley/peak ratio.
+These quantities are defined by the following equations \citep{kokoska1999}:
 
@@
 \label{eq:std}
-\mathrm{STD} = \sqrt{m_2},
+\STD = \sqrt{m_2},
 \end{equation}
@@
 \label{eq:skw}
-\mathrm{SKW} = \frac{m_3}{m_2^{3/2}},
+\SKW = \frac{m_3}{m_2^{3/2}},
 \end{equation}
@@
 \label{eq:bi}
-\mathrm{BI} = \mathrm{KURTOSIS} - \mathrm{SKW}^2 = \frac{m_4}{m_2^{2}} - \frac{m_3^2}{m_2^{3}},
+\BI = \mathrm{KURTOSIS} - \SKW^2 = \frac{m_4}{m_2^{2}} - \frac{m_3^2}{m_2^{3}},
 \end{equation}
@@
 The SKW is adimensional and quantifies the peaks' asymmetry. 
-In the case of a bimodal distribution, $SKW>0$ means that the blue
-peak is taller and for $SWK<0$ the red peak is taller. 
+In the case of a bimodal distribution, $\SKW>0$ means that the blue
+peak is taller and for $\SKW<0$ the red peak is taller. 
 The BI is adimensional and quantifies whether the line has 1 or 2
-peaks: it is  always $\geq 1$ \citep{Pearson1929} and the 
+peaks: it is  always $\geq 1$ \citep{Pearson1929} and the closer 
 to 1, the more bimodal is the line (i.e. has 2 similar peaks). 
-We found by visual inspection of our spectra that BI$=2.5$ marks the
+We found by visual inspection of our spectra that $\BI=2.5$ marks the
 transition between two peaks (however imbalanced) and a dominant
@@
 depth are not monotonous.
-When the outflow velocity is low (\vout$<50\ \kms$), an increasing
+When the outflow velocity is low ($\vout<50\ \kms$), an increasing
 rotational velocity smears the two asymmetrical peaks pushing the line
@@
 increase. 
-On other situations (\vout$=50\ \kms$ and \tauh$\geq 10^6$) higher
+On other situations ($\vout=50\ \kms$ and $\tauh\geq 10^6$) higher
 rotational velocities the bimodality statistics decreases, which means
@@
 These results correspond to a fixed rotational velocity of
-\vrot$=100\ \kms$.
+$\vrot=100\ \kms$.
 The triangles correspond to the RT simulations and the line represents
@@
 The valley intensity is expressed as a fraction of the maximum peak
-intensity in the line, as such the valley/peak ratio is always less
-than one. 
+intensity in the line, as such the valley/peak ratio is always $<1$. 
 In every panel we see that the valley/peak ratio decreases as the
@@
 
-The valley/peak ratio at $\cos\theta=1$ matches results
+The valley/peak ratio at $\cos{\theta}=1$ matches results
 without rotation, this shows that for increasing rotational velocity
@@
 
-\subsubsection{\lya continuum leakage or rotation?}
+\subsection{\lya continuum leakage or rotation?}
 
-Another application of this new model to the current interpretation of \lya
+One application of this new model to the current interpretation of \lya
 spectra, is that the central ($V=0$) emission of the spectra seen in the
@@
 Only an additional exploration of rotational velocities and viewing
-angles need to be explored.
+angles needs to be explored.
 
@@
 \cite{tololo}. The triple peak could be then produced by two main
-decopled components: one rotating the other outflowing.
+decoupled components: one rotating the other outflowing.
 
@@
 We speculate that this feature could help to constrain rotation in
-Green Pea galaxies.
+Green Pea galaxies and find new correlations between \lya escape
+fraction and galaxy rotation.
 
-\subsubsection{\lya Kinematic Maps}
+\subsection{\lya Kinematic Maps}
 
@@
 For instance \cite{Prescott14} presented observational results of a
-Doppler shift when taking spectra at two oposite sides of a large
+Doppler shift when taking spectra at two opposite sides of a large
 ($\approx 80$kpc) LAE.
 In more recent work \cite{2018MNRAS.473.3907A} mapped \lya emission
-around a quasar
+around a quasar.
 
@@
 sum of Doppler shifts, the peak of the spectra is also weighted by the
-amount of mass with a given velocity. 
+amount of mass with a given line-of-sight velocity. 
 In this toy model the distance between the peaks of the
@@
 
-New spectrographs like MUSE should help obtain kinematic information
-from LAEs to build velocity maps in \lya. 
+Spectrographs like the Multi Unit Spectroscopic Explorer (MUSE) could
+obtain kinematic information from large samples of LAEs to build
+velocity maps in Ly$\alpha$.  
 This could be a natural extension of the work reported by
@@
 
-\subsubsection{Comparison against the MUSE-Wide Survey}
+\subsection{Comparison against the MUSE-Wide Survey}
 
@@
 In this paper we explore, for the first time in the literature,
-the results of a model for the emergent \lya\ line from rotating outflows.
+the results of a model for the emergent \lya line from rotating outflows.
 The main results for the model are computed from a Monte-Carlo
@@
 the \lya line?}) we can say that the main effects of rotation on the
-\lya\ line morphology are:
+\lya line morphology are:
 
@@
 Doppler shift computed as product of quantities at the surface of last
-scattering, namely  $\vec{v}_{\rm rot}\cdot \hat{k}$,  where
-$\vec{v}_{\rm rot}$ is the velocity due to rotation and $\hat{k}$ is
+scattering, namely  $\vec{v}_{\rm rot}\cdot \hatk$, where
+$\vec{v}_{\rm rot}$ is the velocity due to rotation and \hatk is
 the direction of the photon's propagation.  
-These conclusions, specially the confirmation of the Doppler Boost as
+These conclusions, specially the confirmation of the Doppler shift  as
 a good approximation to model solid body rotation, strengthen the
-semi-analytic model reported by \cite{Garavito14} when they modeled
-the effect of pure rotation on \lya\ spectra.
+semi-analytic model reported by \cite{Garavito14} where they modeled
+the effect of pure rotation on \lya spectra.
 
 Addressing the second question we posed in the introduction (\emph{to what
-extent is possible to constrain rotational kinematics from the \lya
+extent is it possible to constrain rotational kinematics from the \lya
 emission line?})  we can say that as an application to observational data our results suggest that
@@
 the interpretation of recent observational results
-\citep[e.g.][]{2017ApJ...844..171Y,2017A&A...608L...4R} that claim
-instead that this flux constrains the escape fraction of \lya
-radiation.
+\citep[e.g.][]{2017ApJ...844..171Y,2017A&A...608L...4R} and provide a
+test for rotation in the observed systems.
 
@@
 \lya spectra from two different sides of a galaxy can detect
-approaching/receding gas motions.
-In that case one has to be careful and take into account that the
+approaching/receding gas motions \citep{Prescott14,2018MNRAS.473.3907A}.
+In that case, one has to be careful and take into account that the
 peaks of these two spectra is smaller than the expected value of
@@
 
-We also use data from MUSE to demonstrate to what extent the
-outputs of the RT results presented here can be expected to match
-hundreds of observed \lya spectra, but a full constraint of rotation
-properties certainly require a full Markov Chain Monte Carlo
-exploration of parameter space.
+We also use data from MUSE \citep{2017A&A...606A..12H} to show
+to what extent the outputs of the RT results presented here can be
+expected to match hundreds of observed \lya spectra, but a full
+constraint of rotational properties certainly requires a full Markov
+Chain Monte Carlo exploration of parameter space.
 
@@
 that should be taken into account into \lya modeling; the Doppler
-Boost offers an easy-to-implement approximation to explore such
+Shift offers an easy-to-implement approximation to explore such
 influence into already existing RT simulations and provide stronger
-footing to the interpretation of \lya kinematics maps
-\citep[e.g][]{2018MNRAS.473.3907A}.
+footing to the interpretation of present and future \lya kinematics
+maps \citep[e.g][]{2018MNRAS.473.3907A}.
 

### 2018-04-02 paper/paper.tex
2 typos

@@
 
-Nevertheless, an study of the combined effects of outflows and
+Nevertheless, a study of the combined effects of outflows and
 rotation has not been presented in the literature.
@@
 the center of the sphere at the \lya line's center as they propagate
-through the volume and finally scape.
+through the volume and finally escape.
 We store the final frequency and propagation direction for each photon

### 2018-07-17 data/tau10E4/to_join_parts.py
data 10E4

@@ -1,0 +1,23 @@
+import os
+
+logtau = 4
+vrots = [0,50,100]
+vouts = [5,25,50]
+
+for vrot in vrots:
+    for vout in vouts:
+
+        path = './vrot'+str(vrot)+'/vout'+str(vout)+'/'
+        filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+
+        #os.system('rm '+path+filename+'_out.ascii')
+        os.system('cat '+path+filename+'_out* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_out.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_out.ascii')
+        os.system('rm temp.txt')
+
+        #os.system('rm '+path+filename+'_in.ascii')
+        os.system('cat '+path+filename+'_in* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_in.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_in.ascii')
+        os.system('rm temp.txt')

### 2018-07-17 data/tau10E4/vrot0/vout25/tau10E4_vrot0_vout25.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot0_vout25
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot0/vout25/tau10E4_vrot0_vout25.input

### 2018-07-17 data/tau10E4/vrot0/vout5/tau10E4_vrot0_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot0_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot0/vout5/tau10E4_vrot0_vout5.input

### 2018-07-17 data/tau10E4/vrot0/vout50/tau10E4_vrot0_vout50.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot0_vout50
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot0/vout50/tau10E4_vrot0_vout50.input

### 2018-07-17 data/tau10E4/vrot100/vout25/tau10E4_vrot100_vout25.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot100_vout25
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot100/vout25/tau10E4_vrot100_vout25.input

### 2018-07-17 data/tau10E4/vrot100/vout5/tau10E4_vrot100_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot100_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot100/vout5/tau10E4_vrot100_vout5.input

### 2018-07-17 data/tau10E4/vrot100/vout50/tau10E4_vrot100_vout50.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot100_vout50
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot100/vout50/tau10E4_vrot100_vout50.input

### 2018-07-17 data/tau10E4/vrot50/vout25/tau10E4_vrot50_vout25.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot50_vout25
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot50/vout25/tau10E4_vrot50_vout25.input

### 2018-07-17 data/tau10E4/vrot50/vout5/tau10E4_vrot50_vout5.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot50_vout5
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot50/vout5/tau10E4_vrot50_vout5.input

### 2018-07-17 data/tau10E4/vrot50/vout50/tau10E4_vrot50_vout50.sh

@@ -1,0 +1,11 @@
+#!/bin/sh
+#PBS -q batch
+#PBS -N tau10E4_vrot50_vout50
+#PBS -l mem=128mb
+#PBS -l nodes=1:ppn=6
+#PBS -M mc.remolina197@uniandes.edu.co
+#PBS -m abe
+
+module load openmpi/1.8.5
+cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/
+mpiexec -n 6 ./mine.x /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E4/vrot50/vout50/tau10E4_vrot50_vout50.input

### 2018-07-17 data/tau10E5/a.py

@@ -1,0 +1,23 @@
+import os
+
+logtau = 5
+vrots = [0]
+vouts = [25,50,100,200]
+
+for vrot in vrots:
+    for vout in vouts:
+
+        path = './vrot'+str(vrot)+'/vout'+str(vout)+'/'
+        filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+
+        #os.system('rm '+path+filename+'_out.ascii')
+        os.system('cat '+path+filename+'_out* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_out.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_out.ascii')
+        os.system('rm temp.txt')
+
+        #os.system('rm '+path+filename+'_in.ascii')
+        os.system('cat '+path+filename+'_in* >> temp.txt')
+        os.system('echo "x y z x_u y_u z_u x_frec escaped n_scatterings" > '+path+filename+'_in.ascii')
+        os.system('grep -v "#" temp.txt >> '+path+filename+'_in.ascii')
+        os.system('rm temp.txt')

### 2018-07-17 script/5th_run/to_make_inputs.py

@@ -1,0 +1,81 @@
+#Example of directory: /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot0/vout100
+
+logtaus = [4]
+vrots = [0, 50, 100]
+vouts = [5,25,50]
+
+for logtau in logtaus:
+    for vrot in vrots:
+        for vout in vouts:
+
+            filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+            path = '/hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E'+str(logtau)+'/vrot'+str(vrot)+'/vout'+str(vout)+'/'
+            input_file = open(path+filename+'.input', 'w')
+
+            input_file.write('% input files and formats*/\n')
+            input_file.write('InputDir\t'+path+'\n')
+            input_file.write('CubeName\tgrid_object_100\n')
+            input_file.write('\n')
+            input_file.write('% output files\n')
+            input_file.write('OutputDir\t'+path+'\n')
+            input_file.write('OutputFile\t'+filename+'\n')
+            input_file.write('\n')
+            input_file.write('\n')
+            input_file.write('% define the problem/test to solve\n')
+            input_file.write('NeufeldSlab\t0\n')
+            input_file.write('NeufeldCube\t0\n')
+            input_file.write('ExpandingSphere\t0\n')
+            input_file.write('RotatingSphere\t0\n')
+            input_file.write('OutflowRotationSphere\t1\n')
+            input_file.write('TestParallelVel\t0\n')
+            input_file.write('TestParallelVelFast\t0\n')
+            input_file.write('TestFirstScatter\t0\n')
+            input_file.write('TestRND\t0\n')
+            input_file.write('TestPerpVel\t0\n')
+            input_file.write('SimulationCube\t0\n')
+            input_file.write('HomogeneousInit\t0\n')
+            input_file.write('\n')
+            input_file.write('% parameters for the tests\n')
+            input_file.write('Test_a\t0.00014\n')
+            input_file.write('Test_x\t2.0\n')
+            input_file.write('OutputTestFile\ttest_b\n')
+            input_file.write('\n')
+            input_file.write('% define some physical parameters of the problem to solve\n')
+            input_file.write('Temperature\t10000.0 % in  Kelvin\n')
+            input_file.write('Tau\t1.0e'+str(logtau)+'\n')
[…]

### 2018-07-17 script/5th_run/to_make_scripts.py

@@ -1,0 +1,32 @@
+#Example of directory: /lustre/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E5/vrot0/vout100
+
+ppns = [6]
+logtaus = [4]
+vrots = [0, 50, 100]
+vouts = [5,25,50]
+
+for i in range(1):
+
+    logtau = logtaus[i]
+    ppn = ppns[i]
+
+    for vrot in vrots:
+        for vout in vouts:
+
+            filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+            path = '/hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E'+str(logtau)+'/vrot'+str(vrot)+'/vout'+str(vout)+'/'
+            script = open(path+filename+'.sh', 'w')
+
+            script.write('#!/bin/sh\n')
+            script.write('#PBS -q batch\n')
+            script.write('#PBS -N '+filename+'\n')
+            script.write('#PBS -l mem=128mb\n')
+            script.write('#PBS -l nodes=1:ppn='+str(ppn)+'\n')
+            script.write('#PBS -M mc.remolina197@uniandes.edu.co\n')
+            script.write('#PBS -m abe\n')
+            script.write('\n')
+            script.write('module load openmpi/1.8.5\n')
+            script.write('cd /hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/github/CLARA-MPI/src/\n')
+            script.write('mpiexec -n '+str(ppn)+' ./mine.x '+path+filename+'.input')
+
+            script.close()

### 2018-07-17 script/5th_run/to_run_scripts.py

@@ -1,0 +1,13 @@
+import os
+
+logtaus = [4]
+vrots = [0,  50, 100]
+vouts = [5,25,50]
+
+for logtau in logtaus:
+    for vrot in vrots:
+        for vout in vouts:
+            filename = 'tau10E'+str(logtau)+'_vrot'+str(vrot)+'_vout'+str(vout)
+            path = '/hpcfs/home/ciencias/fisica/pregrado/mc.remolina197/data/CLARA_RotationOutflowsData/tau10E'+str(logtau)+'/vrot'+str(vrot)+'/vout'+str(vout)+'/'
+            #os.system('qsub '+path+filename+'.sh')
+            print 'qsub '+path+filename+'.sh'

### 2018-07-17 paper/paper.tex
fixing running head
some fixes

@@
 %=========================================================================
-\title[Lyman-α photons through rotating outflows]{Lyman-$\alpha$ photons through rotating outflows}
+\title[Lyman-$\alpha$ photons through rotating outflows]{Lyman-$\alpha$ photons through rotating outflows}
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
@@
 These systems with star formation, neutral gas and low dust
-contents naturally show a \lya emission line \citep{PartridgePeebles}
+contents naturally produce a \lya emission line \citep{PartridgePeebles}
 motivating the observational work to phenomenologically link
@@
 neutral Hydrogen
-\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.
+\citep{2006A&A...460..397V,Orsi12,2015ApJ...812..123G}.
 
@@
 galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
-$M_{\odot}$.
+$M_{\odot}$. We also define the viewing angle, $\theta$, as the angle
+between the rotation axis and the line of sight of a potential observer.
 
@@
 the photon's direction of propagation and \vth is the thermal
-velocity of the sphere.
-We also define the viewing angle, $\theta$, as the angle between the
-rotation axis and the line of sight of a potential observer.
+velocity of the sphere. In this study, the latter is set to a fixed value
+of \vth $= 12.86$ \kms, which corresponds to a temperature of $T=10^4$ K.
+The reason of setting it fixed is so that the comparison between the simulations and the semianalitical model is clearer, because there is one less parameter space to explore.
+So, we set \vth to a constant and consistent value throughout the analysis.
+However we are aware of the smoothing effect it has over the Doppler shift, as seen in Eq.
+\ref{eq:shift_x}, and it should be an explicitly free parameter if this model is used to fit observations.
 
@@
 
-\subsection{Quantitative trends}
+\subsection{Quantitative Trends}
 \label{sec:quantitative}

### 2018-07-23 paper/paper.tex
abstract

@@
 %=========================================================================
-\title[Lyman-$\alpha$ photons through rotating outflows]{Lyman-$\alpha$ photons through rotating outflows}
+\title[Lyman-$\alpha$ photons through rotating
+  outflows]{Lyman-$\alpha$ photons through rotating outflows}  
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
@@
 kinematics of galaxies.
-Here we perform Monte Carlo radiative transfer simulations of outflowing
-gas with additional solid body rotation to understand how these kinematic
-features impact the morphology of the Lyman-$\alpha$ emission line.
+Here we introduce a semi-analytic model to quantify how rotating
+outflows impact the morphology of the Lyman-$\alpha$ emission line.   
+The model is contrasted against Monte Carlo radiative transfer
+simulations of outflowing gas with additional solid body rotation.
 We explore a range of neutral Hydrogen optical depth of
-$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
-outflow velocity $0\leq \vout/\kms\leq 50$.  
+$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq
+100$ and outflow velocity $0\leq \vout/\kms\leq 50$.  
 We find three important consequences of rotation.
@@
 line's center.
-We also test a semi-analytic model that modifies the spectra of
-pure outflow simulations and manages to reproduce the radiative
-transfer results for the line widths and flux change at the line's
-center within a $7\%$ and $50\%$ precision, respectively, for an
-optical depth of $\tauh=10^5$ and within $2\%$ and $1\%$ for an
-optical depth of $\tauh=10^7$.   
-We argue that rotation is a required kinematic feature to take into
-account for a correct interpretation of \lya spectra in systems where
-rotation is already evident through kinematic maps but also in other
-systems, such as Green Pea galaxies, where a strong \lya emission at
-the line's center could be explained by rotational kinematics. 
+The semi-analytic model reproduces the radiative transfer results for
+the line width and flux change at the line's center within a $7\%$
+and $50\%$ precision for an optical depth of
+$\tauh=10^5$, respectively, and within $2\%$ and $1\%$ for an optical
+depth of $\tauh=10^7$.
+The semi-analytic model is a convenient tool to introduce rotational
+kinematics as a post-processing step of idealized Monte Carlo
+simulations, providing and easy way to include rotation in models that
+seek to interpret \lya spectra in systems where rotation is evident
+through kinematic maps. 
 \end{abstract}

### 2018-07-23 paper/paper.tex
thermal

@@
 the photon's direction of propagation and \vth is the thermal
-velocity of the sphere. In this study, the latter is set to a fixed value
-of \vth $= 12.86$ \kms, which corresponds to a temperature of $T=10^4$ K.
-The reason of setting it fixed is so that the comparison between the simulations and the semianalitical model is clearer, because there is one less parameter space to explore.
-So, we set \vth to a constant and consistent value throughout the analysis.
-However we are aware of the smoothing effect it has over the Doppler shift, as seen in Eq.
-\ref{eq:shift_x}, and it should be an explicitly free parameter if this model is used to fit observations.
+velocity of the sphere. 
+In this study we use \vth $= 12.86$ \kms, which corresponds to a
+temperature of $T=10^4$ K.   
+We fix \vth throughout the paper, its influence is only
+of a multiplying constant that can be included without running new
+Monte-Carlo simulations, as the code works with the adimensional
+frequency $x$. 
+However, the thermal velocity is considered a free parameter 
+when comparing simulation results against observations.
 
-This allows us to produce new \lya spectra and compare them with the
-full radiative transfer solution including both outflows and
-rotation.
+The approach presented above allows us to produce new \lya spectra and
+compare them with the full radiative transfer solution including both
+outflows and rotation.
 

### 2018-07-26 paper/paper.tex
lw tau in paper

@@
 
+\subsection{Low optical thickness regimes}
+When located in the regimes of low optical thickness, \tauh $\lesssim 4$,  the agreement of the radiative transfer simulation with the semi-analytical solution is not as accurate.
+In this regime, only numerical simulations can help to predict the shape of the \lya profiles.
+We ran a simulation with low \tauh and compared it with the semi-analytical model.
+As seen in Figure \ref{fig:low_tau}, the model does not fit the simulation.
+This creates physical restrictions on the input of our model.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.48\textwidth]{low_tau_logtau4_theta90}
+	\caption{\textbf{Low optical thickness regimes}.  
+		Here we fix $\tauh=10^4$. 
+		The six panels correspond to $\theta=90^\circ$.
+		We vary \vrot increasing from left to right and \vout increasing
+		from top to bottom. 
+		The thin black line corresponds to the \lya line obtained with
+		CLARA without any rotation and the indicated outflow velocity.
+		The thick black line corresponds to the results including both
+		outflows and rotation.
+		The thick gray line shows the results of modifying the pure outflow
+		solution by the Doppler shift presented in Equation \ref{eq:shift_x}
+		(in thin line), if there is a radiative transfer of rotation and outflows
+		(thick and clear line), and if there is a radiative transfer of
+		only outflows, but also a Doppler shift from the rotational
+		velocity (thick and dark line).  
+		\label{fig:low_tau}}
+\end{figure}
+
 \section{Conclusions}

### 2018-08-06 paper/paper.tex
abstract y conclusiones en nueva perspectiva

@@
 line's center.
-The semi-analytic model reproduces the radiative transfer results for
-the line width and flux change at the line's center within a $7\%$
-and $50\%$ precision for an optical depth of
+For all viewing angles, the semi-analytic model reproduces the
+radiative transfer results for the line width and flux change at the
+line's center within a $7\%$ and $50\%$ precision for an optical depth of
 $\tauh=10^5$, respectively, and within $2\%$ and $1\%$ for an optical
@@
 kinematics as a post-processing step of idealized Monte Carlo
-simulations, providing and easy way to include rotation in models that
-seek to interpret \lya spectra in systems where rotation is evident
-through kinematic maps. 
+simulations, providing a framework to interpret \lya spectra in
+systems where rotation is expected or directly measured through
+kinematic maps.  
 \end{abstract}
@@
 In this paper we explore, for the first time in the literature,
-the results of a model for the emergent \lya line from rotating outflows.
-The main results for the model are computed from a Monte-Carlo
-radiative transfer simulation and confronted with a simple
-semi-analytic ansatz that adds the effects of rotation onto results of
-pure outflow kinematics.
+the results of a model for the emergent \lya line from rotating
+outflows. 
+We use a semi-analytic model, first presented in \citep{Garavito14}, that
+to capture the main effects of rotation and confront it against
+results from Monte-Carlo radiative transfer simulations. 
+The semi-analytic model only takes into account the Doppler shift computed
+as product of quantities at the surface of last scattering, namely
+$\vec{v}_{\rm rot}\cdot \hatk$, where $\vec{v}_{\rm rot}$ is the
+velocity due to rotation and \hatk is the direction of the photon's
+propagation.
 
@@
 \end{itemize}
-
-We find that these effects can be quantitatively explained by a
-Doppler shift computed as product of quantities at the surface of last
-scattering, namely  $\vec{v}_{\rm rot}\cdot \hatk$, where
-$\vec{v}_{\rm rot}$ is the velocity due to rotation and \hatk is
-the direction of the photon's propagation.  
-These conclusions, specially the confirmation of the Doppler shift  as
-a good approximation to model solid body rotation, strengthen the
-semi-analytic model reported by \cite{Garavito14} where they modeled
-the effect of pure rotation on \lya spectra.
+%
+All these effects can be quantitatively explained by the proposed
+semi-analytic model.
 
@@
 extent is it possible to constrain rotational kinematics from the \lya
-emission line?})  we can say that as an application to observational data our results suggest that
-rotation should also be considered as an explanation of the
-non-vanishing \lya intensity at the line's center, which could change
-the interpretation of recent observational results
-\citep[e.g.][]{2017ApJ...844..171Y,2017A&A...608L...4R} and provide a
-test for rotation in the observed systems.
+emission line?})  we can say that  the most straightforward approach
+takes into account how the \lya spectra from two different sides of a
+galaxy can detect approaching/receding gas motions
+\citep{Prescott14,2018MNRAS.473.3907A}. 
+In that case, we also show that one has to be careful and take into
+account that the  peaks of these two spectra is smaller than a naively
+expected value of $2\times\vrot$ due to the different weights at the
+surface of the emitting regions.
 
-A more straightforward approach takes into account how the
-\lya spectra from two different sides of a galaxy can detect
-approaching/receding gas motions \citep{Prescott14,2018MNRAS.473.3907A}.
-In that case, one has to be careful and take into account that the
-peaks of these two spectra is smaller than the expected value of
-$2\times\vrot$ due to the different weights at the surface of the
-emitting regions.
+As an application to observational data we use data from MUSE
+\citep{2017A&A...606A..12H} to show to what extent the outputs of the
+RT results presented here can be expected to change the results of
+pure outflow simulations and thus bettwe match hundreds of
+observed \lya spectra.
+We find that the effect on general line morphology statistics
+(standard deviation and skewness) is small, although a full constraint
+of rotational properties certainly requires a full Markov Chain Monte
+Carlo exploration of parameter space.
 
-We also use data from MUSE \citep{2017A&A...606A..12H} to show
-to what extent the outputs of the RT results presented here can be
-expected to match hundreds of observed \lya spectra, but a full
-constraint of rotational properties certainly requires a full Markov
-Chain Monte Carlo exploration of parameter space.
+Our results suggest that rotation should also be considered as an
+explanation of the non-vanishing \lya intensity at the line's center,
+and become a test for rotation in the observed systems, provided that
+the spectral resolution of the instrument is smaller than the expected
+rotational velocities.
 
-Our work confirms that rotation produces features in the \lya line
-strong enough that can be expected to be detected in observations and
-that should be taken into account into \lya modeling; the Doppler
-Shift offers an easy-to-implement approximation to explore such
-influence into already existing RT simulations and provide stronger
-footing to the interpretation of present and future \lya kinematics
-maps \citep[e.g][]{2018MNRAS.473.3907A}.
+To summarize, our work confirms that the semi-analytic model based on
+the Doppler Shift offers an easy-to-implement approximation to explore
+such influence into already existing RT simulations and provide
+stronger footing to the interpretation of present and future \lya
+kinematics maps \citep[e.g][]{2018MNRAS.473.3907A}.
 

### 2018-08-06 paper/paper.tex
typo

@@
 outflows. 
-We use a semi-analytic model, first presented in \citep{Garavito14}, that
+We use a semi-analytic model, first presented in \citep{Garavito14}, 
 to capture the main effects of rotation and confront it against

### 2018-08-08 paper/paper.tex
revised intro

@@
 evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
-These systems with star formation, neutral gas and low dust
-contents naturally produce a \lya emission line \citep{PartridgePeebles}
-motivating the observational work to phenomenologically link
-H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.
+Systems with star formation, neutral gas and low dust
+contents can produce a \lya emission line \citep{PartridgePeebles}
+motivating the observational work to phenomenologically link tracers
+of galaxy rotation such as H$\alpha$ to \lya spectra
+\citep[e.g.][]{Herenz2016}.
 
@@
 more complex than pure rotation.
-The case for outflows as traced in the \lya line is solidly
-established.  
 In many observations the \lya line profile has a single peak
@@
 rotation has not been presented in the literature.
-Here we report on such a study. 
+Here we report on such a study with the main aim of quantifying the
+validity of the semi-analytic approximation presented by
+\cite{Garavito14} in the case where outflows are also present. 
 We investigate a simplified geometrical configuration corresponding to a
 spherical gas cloud with symmetrical radial outflows and solid body
-rotation. 
-We base our modeling on a Monte-Carlo radiative transfer code.
-Besides modeling the impact of joint rotation and outflows, we also
-check to what extent the semi-analytical model presented by
-\cite{Garavito14} an also be applied in our case.
+rotation and contrast the semi-analytic model against the results of a
+Monte-Carlo radiative transfer code.
 

### 2018-08-09 paper/paper.tex
obs params

@@
 
-The Monte Carlo code we use (CLARA) \citep{CLARA} follows the propagation of
-individual photons through a neutral Hydrogen medium characterized by
-its temperature, velocity field and global optical depth.
-The code assumes a homogeneous density throughout the simulated
-volume.
+We use (CLARA) \citep{CLARA}, a Monte Carlo code that follows the
+propagation of individual photons through a neutral Hydrogen medium
+characterized by  its temperature, velocity field and global optical depth.
+The code assumes an homogeneous density throughout the simulated volume.
 In the current implementation we neglect the influence of dust.
@@
 the velocity magnitude increasing linearly with the radial
-coordinate; the outflows model is fully characterized by $v_{\rm out}$, the
+coordinate; this model is fully characterized by $v_{\rm out}$, the
 radial velocity at the sphere's surface.
@@
 
-For each model setup we follow $10^5$ individual photons generated at
+For each model setup we follow $10^5$ individual photons generated from
 the center of the sphere at the \lya line's center as they propagate
 through the volume and finally escape.
-We store the final frequency and propagation direction for each photon
+We store the frequency and propagation direction for each photon
 at its last scattering.
@@
 $27$ models with all the possible parameter combinations.
-The range of values have some overlap with the expectations from a
-galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
-$M_{\odot}$. We also define the viewing angle, $\theta$, as the angle
+The values for the outflow velocity are lower than other values used in the
+literature to allow for an interplay between the two kinematic features.
+Nevertheless, this range of parameters reproduces the main standard
+deviation and skewness  of high redshift LAES in recent observations
+(\label{subsec:kinematic}).
+
+We also define the viewing angle, $\theta$, as the angle
 between the rotation axis and the line of sight of a potential observer.
@@
 \subsection{\lya Kinematic Maps}
-
+\label{subsec:kinematic}
 Current observational facilities have the capability of spatially

### 2018-08-10 paper/paper.tex
edit

@@
 
+Throughout the paper we use a thermal velocity for a neutral Hydrogen
+gas of \vth $= 12.86$ \kms, which corresponds to a temperature of
+$T=10^4$ K.
+
 \begin{figure*}
@@
 
-We use (CLARA) \citep{CLARA}, a Monte Carlo code that follows the
+We use CLARA \citep{CLARA}, a Monte Carlo code that follows the
 propagation of individual photons through a neutral Hydrogen medium
@@
 Our basic model is an spherical distribution of neutral hydrogen,
-an approximation commonly used in the literature, as it explains a
+an approximation commonly used in the literature that explains a
 wide variety of observational features
@@
 
-The velocity field we use captures both outflows and rotation.
+We use a velocity field that captures both outflows and rotation.
 Outflows are described by a Hubble-like radial velocity profile with
-the velocity magnitude increasing linearly with the radial
-coordinate; this model is fully characterized by $v_{\rm out}$, the
-radial velocity at the sphere's surface.
-Rotation follows a solid body rotation profile, which is fully
-characterized by $v_{\rm rot}$, the linear velocity at the sphere's surface.
+the speed increasing linearly with the radial coordinate; this model
+is fully characterized by $v_{\rm out}$, the radial velocity at the
+sphere's surface. 
+Rotation follows a solid body rotation profile fully characterized by
+$v_{\rm rot}$ the linear velocity at the sphere's surface.
 
-The total velocity field corresponds to the superposition of rotation and
+The total velocity field is the superposition of rotation and
 outflows.
@@
 
-For each model setup we follow $10^5$ individual photons generated from
-the center of the sphere at the \lya line's center as they propagate
+For each run we follow $10^5$ individual photons generated from the
+origin at the \lya line's center as they propagate
 through the volume and finally escape.
@@
 $27$ models with all the possible parameter combinations.
-The values for the outflow velocity are lower than other values used in the
+The values for the outflow velocity are lower than values commonly used in the
 literature to allow for an interplay between the two kinematic features.
@@
 We also define the viewing angle, $\theta$, as the angle
-between the rotation axis and the line of sight of a potential observer.
+between the rotation axis and the line of sight of a potential
+observer.
 
@@
 \lya line morphology. 
-The basic assumption of their analytical model is that each
-differential surface element on the sphere Doppler shifts the photons
-that it emits.
+The basic assumption of the model is that each differential surface
+element on the sphere Doppler shifts the photons that it emits.
 In this paper we introduce this ansatz by post-processing the results
@@
 velocity of the sphere. 
-In this study we use \vth $= 12.86$ \kms, which corresponds to a
-temperature of $T=10^4$ K.   
-We fix \vth throughout the paper, its influence is only
-of a multiplying constant that can be included without running new
-Monte-Carlo simulations, as the code works with the adimensional
-frequency $x$. 
-However, the thermal velocity is considered a free parameter 
-when comparing simulation results against observations.
+We fix \vth throughout the paper. This factor is a multiplying
+constant that could be changed without running new Monte-Carlo
+simulations, as the code works with the adimensional frequency $x$.
 
-The approach presented above allows us to produce new \lya spectra and
-compare them with the full radiative transfer solution including both
-outflows and rotation.
+This semi-analytic model allows us to produce new \lya spectra from
+the outflow-only results and compare them with the full radiative
+transfer solution including both outflows and rotation.
 

### 2018-08-16 plots/paper/model.py
gaussian model

@@ -31,6 +31,14 @@ def doppler_shift(data, vrot=0.0, v_th=12.86):
 
     return new_data
 
+def gauss_shift(data, sigma=0.0):
+    '''Updates old_data adding doppler shifts. the result is stored in new_data.'''
+    new_data = data.copy()
+    new_data['x'], new_data['y'], new_data['z'] = get_unit_r_sphere(data)
+    n_points = len(new_data)
+    new_data['x_frec'] = data['x_frec'] + np.random.normal(loc=0, scale=sigma, size=n_points)
+    return new_data
+
 def read_data(vrot=0, vout=5, logtau=6, input_dir="../../data/"):
     '''Reads the data.'''
     tau_name = 'tau10E' + str(logtau)

### 2018-08-16 paper/paper.tex
gaussian smoothing

@@
 depth of $\tauh=10^7$.
+The results of the semi-analytic model for $\tauh=10^6$ and
+$\tauh=10^7$ can in turn be approximated by a gaussian smoothing with
+standard deviation $\sigma_v = \vrot\sin(\theta)/2$. 
 The semi-analytic model is a convenient tool to introduce rotational
@@
 
-\subsection{\lya continuum leakage or rotation?}
+\subsection{The semi-analytic model as gaussian smoothing}
 
-One application of this new model to the current interpretation of \lya
-spectra, is that the central ($V=0$) emission of the spectra seen in the
-\lya line, is a consequence of the viewing angle of the galaxy and can be
-controlled by it. 
-Several authors have suggested that this central emission is caused by
-radiation that escapes the galaxy without scattering, mostly due to a
-clumpy interstellar medium \citep{Hansen06, 2016ApJ...833L..26G}.
-Rotation is an alternative that solves this issue.
+The effects of the semi-analytic model on the pure outflowing spectra
+are similar to the expected results from a gaussian smoothing.
+this smoothing can also be a natural consequence of the intruments
+used to measure the LAE spectra.
+In other words, poor spectral resolution could be mistaken as an
+indication for rotation.
 
-We note that the relative intensity of the two peaks in the outflowing
-spectra is not modified by rotation, that asymmetry is solely
-controlled by the outflow velocity.
-This means that results that have already derive a typical outflowing
-velocity as to match the observational constraints do not have to be
-completely revised. 
-Only an additional exploration of rotational velocities and viewing
-angles needs to be explored.
+We model the effects of gaussian smoothing with a similar ansatz as
+the semi-analytic model by changing the frequency of each photon by a
+Gaussian random variable centered on zero and width $\sigma_x$:
+\begin{equation}
+x^{\prime} = x + \mathcal{N}(0,\sigma_x).
+\end{equation}
 
-A recent observational example is provided by
-\cite{2017A&A...608L...4R}.
-They report optical spectroscopy of a bright lensed galaxy at a
-redshift of $z=2.4$. 
-The object presents a triple peaked \lya profile that can be thought
-as the superposition of two components, a double asymmetric peak and a
-narrower symmetric peak around the line's center.
-One possible option for the narrow component is that it escaped the
-galaxy without substantial interaction with neutral Hydrogen.
-We speculate that the central feature could also be due to a purely 
-rotating component that can produce such a feature as shown by
-\cite{tololo}. The triple peak could be then produced by two main
-decoupled components: one rotating the other outflowing.
-
-Another interesting example are the \lya lines in Green Pea galaxies \citep{2016ApJ...820..130Y}.
-\cite{2017ApJ...844..171Y} showed in a study of 43 Green Pea galaxies
-that 2/3 were strong \lya emitters. 
-There the \lya escape fraction is effectively estimated from the ratio
-of H$\alpha$ to \lya.
-Galaxies with large escape fractions also substantial \lya emission at
-the line's center. 
-We speculate that this feature could help to constrain rotation in
-Green Pea galaxies and find new correlations between \lya escape
-fraction and galaxy rotation.
+We find that the results of the semi-analytic model on the total
+spectra at a given angle $\theta$ can be approximated using
+$\sigma_x = (1/2)\times(\vrot/\vth)\times \sin(\theta)$. 
+This approximation works well to reproduce the semi-analytic results
+for $\tau=10^6$ and $\tau=10^7$.
 
@@
 In this toy model the distance between the peaks of the
-receding/approaching spectra is close to $\sim 20\ \kms$, which is
-less than half of the naively expected value of $2\vrot=100\ \kms$,
+receding/approaching spectra is close to $\sim 25\ \kms$, which is
+a fourth of the naively expected value of $2\vrot=100\ \kms$,
 due to the fact that only a small fraction of the photons are emitted
@@
 
-\subsection{Low optical thickness regimes}
-When located in the regimes of low optical thickness, \tauh $\lesssim 4$,  the agreement of the radiative transfer simulation with the semi-analytical solution is not as accurate.
-In this regime, only numerical simulations can help to predict the shape of the \lya profiles.
-We ran a simulation with low \tauh and compared it with the semi-analytical model.
-As seen in Figure \ref{fig:low_tau}, the model does not fit the simulation.
-This creates physical restrictions on the input of our model.
+%\subsection{Low optical thickness regimes}
+%When located in the regimes of low optical thickness, \tauh $\lesssim
+%4$,  the agreement of the radiative transfer simulation with the
+%semi-analytical solution is not as accurate. In this regime, only
+%numerical simulations can help to predict the shape of the \lya
+%profiles. We ran a simulation with low \tauh and compared it with the
+%semi-analytical model. As seen in Figure \ref{fig:low_tau}, the model
+%does not fit the simulation. This creates physical restrictions on the
+%input of our model.
 
-\begin{figure}
-	\centering
-	\includegraphics[width=0.48\textwidth]{low_tau_logtau4_theta90}
-	\caption{\textbf{Low optical thickness regimes}.  
-		Here we fix $\tauh=10^4$. 
-		The six panels correspond to $\theta=90^\circ$.
-		We vary \vrot increasing from left to right and \vout increasing
-		from top to bottom. 
-		The thin black line corresponds to the \lya line obtained with
-		CLARA without any rotation and the indicated outflow velocity.
-		The thick black line corresponds to the results including both
-		outflows and rotation.
-		The thick gray line shows the results of modifying the pure outflow
-		solution by the Doppler shift presented in Equation \ref{eq:shift_x}
-		(in thin line), if there is a radiative transfer of rotation and outflows
-		(thick and clear line), and if there is a radiative transfer of
-		only outflows, but also a Doppler shift from the rotational
-		velocity (thick and dark line).  
-		\label{fig:low_tau}}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.48\textwidth]{low_tau_logtau4_theta90}
+%	\caption{\textbf{Low optical thickness regimes}.  
+%		Here we fix $\tauh=10^4$. 
+%		The six panels correspond to $\theta=90^\circ$.
+%		We vary \vrot increasing from left to right and \vout increasing
+%		from top to bottom. 
+%		The thin black line corresponds to the \lya line obtained with
+%		CLARA without any rotation and the indicated outflow velocity.
+%		The thick black line corresponds to the results including both
+%		outflows and rotation.
+%		The thick gray line shows the results of modifying the pure outflow
+%		solution by the Doppler shift presented in Equation \ref{eq:shift_x}
+%		(in thin line), if there is a radiative transfer of rotation and outflows
+%		(thick and clear line), and if there is a radiative transfer of
+%		only outflows, but also a Doppler shift from the rotational
+%		velocity (thick and dark line).  
+%		\label{fig:low_tau}}
+%\end{figure}
 

### 2018-08-16 paper/paper.tex
abstract update

@@
 depth of $\tauh=10^7$.
-The results of the semi-analytic model for $\tauh=10^6$ and
-$\tauh=10^7$ can in turn be approximated by a gaussian smoothing with
-standard deviation $\sigma_v = \vrot\sin(\theta)/2$. 
-The semi-analytic model is a convenient tool to introduce rotational
-kinematics as a post-processing step of idealized Monte Carlo
-simulations, providing a framework to interpret \lya spectra in
-systems where rotation is expected or directly measured through
-kinematic maps.  
+The semi-analytic model presented here is a convenient tool to
+introduce rotational kinematics as a post-processing step of idealized
+Monte Carlo simulations, providing a framework to interpret \lya
+spectra in systems where rotation is expected, or directly measured
+through kinematic maps.  
 \end{abstract}

### 2018-09-05 paper/paper.tex
gauss

@@
 100$ and outflow velocity $0\leq \vout/\kms\leq 50$.  
-We find three important consequences of rotation.
+We find three consequences of rotation.
 First, it introduces a dependency with viewing angle; second it
@@
 depth of $\tauh=10^7$.
+We also show that another, less precise, approximation is considering
+that rotation can be represented by a gaussian smoothing with standard
+deviation $\frac{1}{2} \vrot \sin \theta$, where $\theta$ is the angle between
+the line-of-sight and the rotation axis. 
 The semi-analytic model presented here is a convenient tool to
@@
 The effects of the semi-analytic model on the pure outflowing spectra
-are similar to the expected results from a gaussian smoothing.
-this smoothing can also be a natural consequence of the intruments
+are similar to the expected results from a gaussian smoothing, 
+such smoothing can also be a natural consequence of the intruments
 used to measure the LAE spectra.
-In other words, poor spectral resolution could be mistaken as an
-indication for rotation.
+As a consequence, natural experimental artifacts could be mistaken as
+an indication for rotation.
 
-We model the effects of gaussian smoothing with a similar ansatz as
-the semi-analytic model by changing the frequency of each photon by a
-Gaussian random variable centered on zero and width $\sigma_x$:
-\begin{equation}
-x^{\prime} = x + \mathcal{N}(0,\sigma_x).
+To understand to what extent the effects of rotation can be seen as a
+simple smoothing, we model the effects of gaussian smoothing with a
+similar ansatz as  the semi-analytic model by changing the frequency
+of each photon by a Gaussian random variable centered on zero and
+standard deviation $\sigma_x$: \begin{equation}
+x^{\prime} = x + \mathcal{N}(0,\sigma_x^2).
 \end{equation}
@@
 spectra at a given angle $\theta$ can be approximated using
-$\sigma_x = (1/2)\times(\vrot/\vth)\times \sin(\theta)$. 
-This approximation works well to reproduce the semi-analytic results
-for $\tau=10^6$ and $\tau=10^7$.
+$\sigma_x = (1/2)\times(\vrot/\vth)\times \sin\theta$. 
+This approximation reproduces within a few porcent all quantities
+measured in Section \ref{sec:quantitative} for the semi-analytic
+model, it only works well for $\tau=10^6$ and $\tau=10^7$.
+
+This result can be used as an order of magnitude estimate of the mimum
+espectral resolution required to observe the effects of a rotational
+velocity of $\vrot$. 
+The Space Telescope Imaging Spectrograph (STIS) on the Hubble Space
+Telescope (HST) can provide resolutions around 20 \kms
+\citep{2015PASA...32...27H}, which could allow for the detection of
+rotation features in nearby galaxies with at least $40$ \kms and a
+rotation axis perpendicular to the line of sight.  
+Other instruments such as the Cosmic Origins Spectrograph (COS) have
+varying resolutions between $20$\kms and $200$\kms according to the
+angular size of the source \citep{2018A&A...616A..60O}.
 

### 2018-09-05 paper/paper.tex
almost final version

@@
 depth of $\tauh=10^7$.
-We also show that another, less precise, approximation is considering
-that rotation can be represented by a gaussian smoothing with standard
-deviation $\frac{1}{2} \vrot \sin \theta$, where $\theta$ is the angle between
-the line-of-sight and the rotation axis. 
+Using this toy model we also show that the peaks of integrated spectra
+taken from opposite sides of an edge-on rotating gas distribution
+should have a separation of $\frac{1}{2}\vrot$. 
 The semi-analytic model presented here is a convenient tool to
@@
 neutral Hydrogen
-\citep{2006A&A...460..397V,Orsi12,2015ApJ...812..123G}.
+\citep{Verhamme06,Orsi12,2015ApJ...812..123G}.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.45\textwidth]{muse_vs_simulations.pdf}
-
-\caption{\textbf{Comparison between simulations and observations.}  
-The observational results correspond to the MUSE-Wide Survey
-\citep{2017A&A...606A..12H}. The skewness and standard deviation from
-the simulations summarize the results presented in Figures
-\ref{fig:standard_deviation} and \ref{fig:skewness}.
-  \label{fig:muse}}
-\end{figure}
-
 \section{Theoretical Models}
@@
 \tauh values. 
-In the Appendix we show the same plots as Figure \ref{fig:doppler_shift}, there it is
-evident that for $\tauh=10^5$ the results are not as good as they are
-for $\tauh=10^6$, and that for $\tauh=10^7$ the Doppler shift
-provides a remarkable good approximation.
+In the Appendix we show the same plots as Figure
+\ref{fig:doppler_shift}, there it is evident that for $\tauh=10^5$ the
+results are not as good as they are for $\tauh=10^6$, and that for
+$\tauh=10^7$ the Doppler shift provides a remarkable good
+approximation.
 
@@
 \label{subsec:kinematic}
+
 Current observational facilities have the capability of spatially
@@
 around a quasar.
+In this case one could measure the displacement between the
+peaks in the two different spectra. 
+The interest of this test is that the location of the peaks only
+depends on the kinematic properties of the sources and does not
+change by convolution with a gaussian window function due to limited
+instrumental resolution, as discussed in the previous section.
 
@@
 
-Although it is a good approximation to think the rotating spectra by a
-sum of Doppler shifts, the peak of the spectra is also weighted by the
-amount of mass with a given line-of-sight velocity. 
 In this toy model the distance between the peaks of the
@@
 $50\ \kms$. 
+Although it is a good approximation to think the rotating spectra by a
+sum of Doppler shifts, the peak of the spectra is also weighted by the
+amount of mass with a given line-of-sight velocity.
 
@@
 Reference Sample) galaxies.  
-The interpretation of such data should take into account the insights
+The interpretation of such data could take into account the insights
 and trends we have presented in this paper.
-
-\subsection{Comparison against the MUSE-Wide Survey}
-
-In order to give the reader a feeling of the ranges of observational
-data that could be explained by the models presented here, 
-we perform a comparison against observational data from the MUSE-Wide
-Survey (MWS) \citep{2017A&A...606A..12H}.
-
-MWS provides public access to 831 spectra out of which 237 contain a
-detected \lya line as dominant. 
-The public dataset reports as the redshift of the galaxy the position
-of the dominant peak.
-We use this redshift to convert the spectra to a rest-frame and compute
-the standard deviation and skewness statistics. 
-Given the uncertainty on the galaxy redshift we do not compute the
-valley/peak ratio. 
-We do not try to perform a full fit to all the spectra
-\citep[e.g.][]{2017A&A...608A.139G} and wish only to
-know whether the region of parameter space explored in our RT
-simulations shows any crossing with the results from the MWS or
-shows any sign of atypical features.
-
-Figure \ref{fig:muse} shows the results of this experiment. 
-The small white circles correspond to the 217 \lya emitters in the MWS
-observations; the large black circles to the results of the
-simulations presented in this paper. 
-This shows that the systems modeled in this paper actually present an
-overlap with statistics derived from observations. 
-The region with standard deviation $>200$\kms could also be explained
-by systems with larger optical depths than we have used here. 
-However, the systems with a positive skewness, that is a line
-unbalanced towards the blue could be the sign of inflowing
-kinematics.
-
-%\subsection{Low optical thickness regimes}
-%When located in the regimes of low optical thickness, \tauh $\lesssim
-%4$,  the agreement of the radiative transfer simulation with the
-%semi-analytical solution is not as accurate. In this regime, only
-%numerical simulations can help to predict the shape of the \lya
-%profiles. We ran a simulation with low \tauh and compared it with the
-%semi-analytical model. As seen in Figure \ref{fig:low_tau}, the model
-%does not fit the simulation. This creates physical restrictions on the
-%input of our model.
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.48\textwidth]{low_tau_logtau4_theta90}
-%	\caption{\textbf{Low optical thickness regimes}.  
-%		Here we fix $\tauh=10^4$. 
-%		The six panels correspond to $\theta=90^\circ$.
-%		We vary \vrot increasing from left to right and \vout increasing
-%		from top to bottom. 
-%		The thin black line corresponds to the \lya line obtained with
-%		CLARA without any rotation and the indicated outflow velocity.
-%		The thick black line corresponds to the results including both
-%		outflows and rotation.
-%		The thick gray line shows the results of modifying the pure outflow
-%		solution by the Doppler shift presented in Equation \ref{eq:shift_x}
-%		(in thin line), if there is a radiative transfer of rotation and outflows
-%		(thick and clear line), and if there is a radiative transfer of
-%		only outflows, but also a Doppler shift from the rotational
-%		velocity (thick and dark line).  
-%		\label{fig:low_tau}}
-%\end{figure}
 
@@
 surface of the emitting regions.   
-
-As an application to observational data we use data from MUSE
-\citep{2017A&A...606A..12H} to show to what extent the outputs of the
-RT results presented here can be expected to change the results of
-pure outflow simulations and thus bettwe match hundreds of
-observed \lya spectra.
-We find that the effect on general line morphology statistics
-(standard deviation and skewness) is small, although a full constraint
-of rotational properties certainly requires a full Markov Chain Monte
-Carlo exploration of parameter space.
-
-Our results suggest that rotation should also be considered as an
-explanation of the non-vanishing \lya intensity at the line's center,
-and become a test for rotation in the observed systems, provided that
-the spectral resolution of the instrument is smaller than the expected
-rotational velocities.
+Here the semi-analytic model also provides the correct quantitative
+insight.
 
@@
 the Doppler Shift offers an easy-to-implement approximation to explore
-such influence into already existing RT simulations and provide
-stronger footing to the interpretation of present and future \lya
-kinematics maps \citep[e.g][]{2018MNRAS.473.3907A}.
+such influence into already existing radiative transfer simulation
+simulations and provide stronger footing to the interpretation of
+present and future \lya kinematics maps
+\citep[e.g][]{2018MNRAS.473.3907A}.
 
@@
 
-\newpage
 \appendix
@@
 
-\begin{figure}
+\begin{figure*}
   \begin{center}
     \includegraphics[width=0.49\textwidth]{doppler_shift_logtau5_theta90}
-  \end{center}
-  \caption{\textbf{Qualitative trends of changing outflow and
-      rotational velocity.}
-    Same layout as Figure \ref{fig:doppler_shift},
-    this time  $\tauh=10^5$ and $\theta=90^\circ$.}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
     \includegraphics[width=0.49\textwidth]{doppler_shift_logtau7_theta90}
@@
       rotational velocity.}
-    Same layout as Figure \ref{fig:doppler_shift},
-    this time  $\tauh=10^7$ and $\theta=90^\circ$.}
-\end{figure}
+    Same layout as Figure \ref{fig:doppler_shift}. On the left:
+    $\tauh=10^5$ and $\theta=90^\circ$; on the right:     this time  $\tauh=10^7$ and $\theta=90^\circ$.}
+\end{figure*}
 

### 2018-09-10 paper/paper.tex
edit

@@
 First, it introduces a dependency with viewing angle; second it
-produces a line broadening and third it increases the flux at the
+broadens the line and third it increases the flux at the
 line's center.
@@
 depth of $\tauh=10^7$.
-Using this toy model we also show that the peaks of integrated spectra
+Using this model we also show that the peaks of integrated spectra
 taken from opposite sides of an edge-on rotating gas distribution
@@
 introduce rotational kinematics as a post-processing step of idealized
-Monte Carlo simulations, providing a framework to interpret \lya
-spectra in systems where rotation is expected, or directly measured
+Monte Carlo simulations; it provides a framework to interpret \lya
+spectra in systems where rotation is expected or directly measured
 through kinematic maps.  
@@
 outflows. 
-We use a semi-analytic model, first presented in \citep{Garavito14}, 
+We the a semi-analytic model first presented by \cite{Garavito14}
 to capture the main effects of rotation and confront it against
@@
 
-To address the first question we posed in the introduction (\emph{what is then
-the expected imprint of rotation on a resonant emission line such as
-the \lya line?}) we can say that the main effects of rotation on the
-\lya line morphology are:
+To address the first question we posed in the introduction (\emph{What
+  is the expected imprint of rotation on a resonant emission line such
+  as the \lya line?}) we find from the full RT simulations that the
+effects of rotation on the \lya line morphology are:
 
@@
 %
-All these effects can be quantitatively explained by the proposed
+All these effects can be qualitatively explained by the proposed
 semi-analytic model. 
+Quantitatively speaking the semi-analytic model provides a
+satisfactory answer for a neutral Hydrogen optical depth equal or
+larger than $10^6$.
 
-Addressing the second question we posed in the introduction (\emph{to what
+Addressing the second question we posed in the introduction (\emph{To what
 extent is it possible to constrain rotational kinematics from the \lya
-emission line?})  we can say that  the most straightforward approach
-takes into account how the \lya spectra from two different sides of a
-galaxy can detect approaching/receding gas motions
+emission line?})  we show that the most straightforward approach
+is measuring \lya spectra from two different regions of a
+galaxy to detect approaching/receeding gas motions
 \citep{Prescott14,2018MNRAS.473.3907A}. 
-In that case, we also show that one has to be careful and take into
-account that the  peaks of these two spectra is smaller than a naively
-expected value of $2\times\vrot$ due to the different weights at the
-surface of the emitting regions.   
+In that case, we also show that the distances between these two peaks 
+is four times shorter than the naively expected value of
+$2\times\vrot$.  
+This difference is produced by the different geometric weights at the
+emitting surface.    
 Here the semi-analytic model also provides the correct quantitative
@@
 
-To summarize, our work confirms that the semi-analytic model based on
-the Doppler Shift offers an easy-to-implement approximation to explore
-such influence into already existing radiative transfer simulation
-simulations and provide stronger footing to the interpretation of
-present and future \lya kinematics maps
-\citep[e.g][]{2018MNRAS.473.3907A}.
+To summarize, our works shows that the Doppler Shift offers an
+easy-to-implement approximation to explore such influence into already
+existing radiative transfer simulation simulations that include
+outflows, providing a versatile tool to interpret current and future \lya
+kinematics maps  \citep[e.g][]{2018MNRAS.473.3907A}.
 

### 2018-09-15 paper/paper.tex
update

@@
 literature to allow for an interplay between the two kinematic features.
-Nevertheless, this range of parameters reproduces the main standard
-deviation and skewness  of high redshift LAES in recent observations
-(\label{subsec:kinematic}).
+This range of parameters also produce emission lines with standard deviation and
+skewness in the same range to those of high redshift LAES in recent
+observations \citep{2017A&A...606A..12H}.
 

### 2018-09-15 paper/paper_old.tex
update

--- 2018-09-05 paper/paper.tex
@@
 %=========================================================================
-\title[Lyman-$\alpha$ photons through rotating
-  outflows]{Lyman-$\alpha$ photons through rotating outflows}  
+\title[Lyman-α photons through rotating outflows]{Lyman-$\alpha$ photons through rotating outflows}
 \author[M.C. Remolina-Gutierrez \& J.E. Forero-Romero]{
@@
 kinematics of galaxies.
-Here we introduce a semi-analytic model to quantify how rotating
-outflows impact the morphology of the Lyman-$\alpha$ emission line.   
-The model is contrasted against Monte Carlo radiative transfer
-simulations of outflowing gas with additional solid body rotation.
+Here we perform Monte Carlo radiative transfer simulations of outflowing
+gas with additional solid body rotation to understand how these kinematic
+features impact the morphology of the Lyman-$\alpha$ emission line.
 We explore a range of neutral Hydrogen optical depth of
-$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq
-100$ and outflow velocity $0\leq \vout/\kms\leq 50$.  
-We find three consequences of rotation.
+$10^5\leq\tauh\leq10^7$, rotational velocity $0\leq \vrot/\kms \leq 100$ and
+outflow velocity $0\leq \vout/\kms\leq 50$.  
+We find three important consequences of rotation.
 First, it introduces a dependency with viewing angle; second it
@@
 line's center.
-For all viewing angles, the semi-analytic model reproduces the
-radiative transfer results for the line width and flux change at the
-line's center within a $7\%$ and $50\%$ precision for an optical depth of
-$\tauh=10^5$, respectively, and within $2\%$ and $1\%$ for an optical
-depth of $\tauh=10^7$.
-Using this toy model we also show that the peaks of integrated spectra
-taken from opposite sides of an edge-on rotating gas distribution
-should have a separation of $\frac{1}{2}\vrot$. 
-The semi-analytic model presented here is a convenient tool to
-introduce rotational kinematics as a post-processing step of idealized
-Monte Carlo simulations, providing a framework to interpret \lya
-spectra in systems where rotation is expected, or directly measured
-through kinematic maps.  
+We also test a semi-analytic model that modifies the spectra of
+pure outflow simulations and manages to reproduce the radiative
+transfer results for the line widths and flux change at the line's
+center within a $7\%$ and $50\%$ precision, respectively, for an
+optical depth of $\tauh=10^5$ and within $2\%$ and $1\%$ for an
+optical depth of $\tauh=10^7$.   
+We argue that rotation is a required kinematic feature to take into
+account for a correct interpretation of \lya spectra in systems where
+rotation is already evident through kinematic maps but also in other
+systems, such as Green Pea galaxies, where a strong \lya emission at
+the line's center could be explained by rotational kinematics. 
 \end{abstract}
@@
 evidence for large scale rotation \citep{2018MNRAS.473.3907A}.
-Systems with star formation, neutral gas and low dust
-contents can produce a \lya emission line \citep{PartridgePeebles}
-motivating the observational work to phenomenologically link tracers
-of galaxy rotation such as H$\alpha$ to \lya spectra
-\citep[e.g.][]{Herenz2016}.
+These systems with star formation, neutral gas and low dust
+contents naturally show a \lya emission line \citep{PartridgePeebles}
+motivating the observational work to phenomenologically link
+H$\alpha$ kinematics to \lya spectra \citep[e.g.][]{Herenz2016}.
 
@@
 more complex than pure rotation.
+The case for outflows as traced in the \lya line is much better
+established.
 In many observations the \lya line profile has a single peak
@@
 \citep[e.g.][]{2010ApJ...717..289S,Erb14,Trainor16}.   
-These features have been explained as the consequence of multiple
+These features have been readily explained as the consequence of multiple
 \lya photon scatterings through a homogeneous outflowing shell of
 neutral Hydrogen
-\citep{Verhamme06,Orsi12,2015ApJ...812..123G}.
+\citep{2006A&A...460..397V,Orsi12,2012ApJ...751...29Y,2015ApJ...812..123G}.
 
-Nevertheless, a study of the combined effects of outflows and
-rotation has not been presented in the literature.
-Here we report on such a study with the main aim of quantifying the
-validity of the semi-analytic approximation presented by
-\cite{Garavito14} in the case where outflows are also present. 
+An study of the combined effects of outflows and rotation has not been
+presented in the literature.
+Here we present such a study. 
 We investigate a simplified geometrical configuration corresponding to a
 spherical gas cloud with symmetrical radial outflows and solid body
-rotation and contrast the semi-analytic model against the results of a
-Monte-Carlo radiative transfer code.
+rotation. 
+We base our modeling on a Monte-Carlo radiative transfer code.
+Besides modeling the impact of joint rotation and outflows, we also
+check to what extent the analytical model presented by
+\cite{Garavito14} to explain the effects of rotation can also be
+applied in our case.
 
@@
 our conclusions in Section \ref{sec:conclusions}.
-
-Throughout the paper we use a thermal velocity for a neutral Hydrogen
-gas of \vth $= 12.86$ \kms, which corresponds to a temperature of
-$T=10^4$ K.
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.45\textwidth]{muse_vs_simulations.pdf}
+
+\caption{\textbf{Comparison between simulations and observations.}  
+The observational results correspond to the MUSE-Wide Survey
+\citep{2017A&A...606A..12H}. The skewness and standard deviation from
+the simulations summarize the results presented in Figures
+\ref{fig:standard_deviation} and \ref{fig:skewness}.
+  \label{fig:muse}}
+\end{figure}
+
 \section{Theoretical Models}
@@
 
-We use CLARA \citep{CLARA}, a Monte Carlo code that follows the
-propagation of individual photons through a neutral Hydrogen medium
-characterized by its temperature, velocity field and global optical depth.
-The code assumes an homogeneous density throughout the simulated volume.
+The Monte Carlo code we use (CLARA) \citep{CLARA} follows the propagation of
+individual photons through a neutral Hydrogen medium characterized by
+its temperature, velocity field and global optical depth.
+The code assumes a homogeneous density throughout the simulated
+volume.
 In the current implementation we neglect the influence of dust.
 Our basic model is an spherical distribution of neutral hydrogen,
-an approximation commonly used in the literature that explains a
+an approximation commonly used in the literature, as it explains a
 wide variety of observational features
@@
 
-We use a velocity field that captures both outflows and rotation.
+The velocity field we use captures both outflows and rotation.
 Outflows are described by a Hubble-like radial velocity profile with
-the speed increasing linearly with the radial coordinate; this model
-is fully characterized by $v_{\rm out}$, the radial velocity at the
-sphere's surface. 
-Rotation follows a solid body rotation profile fully characterized by
-$v_{\rm rot}$ the linear velocity at the sphere's surface.
+the velocity magnitude increasing linearly with the radial
+coordinate; the outflows model is fully characterized by $v_{\rm out}$, the
+velocity at the sphere's surface.
+Rotation follows a solid body rotation profile, which is fully
+characterized by $v_{\rm rot}$, the linear velocity at the sphere's surface.
 
-The total velocity field is the superposition of rotation and
+The total velocity field corresponds to the superposition of rotation and
 outflows.
@@
 origin at the sphere's center, $R$ is the radius of the sphere and the
-direction of the angular velocity vector is the $z$ axis.
+direction of the angular velocity vector corresponds to the \hatk unit vector.
 
-For each run we follow $10^5$ individual photons generated from the
-origin at the \lya line's center as they propagate
-through the volume and finally escape.
-We store the frequency and propagation direction for each photon
+For each model setup we follow $10^5$ individual photons generated at
+the center of the sphere at the \lya line's center as they propagate
+through the volume and finally scape.
+We store the final frequency and propagation direction for each photon
 at its last scattering.
@@
 $27$ models with all the possible parameter combinations.
-The values for the outflow velocity are lower than values commonly used in the
-literature to allow for an interplay between the two kinematic features.
-Nevertheless, this range of parameters reproduces the main standard
-deviation and skewness  of high redshift LAES in recent observations
-(\label{subsec:kinematic}).
-
-We also define the viewing angle, $\theta$, as the angle
-between the rotation axis and the line of sight of a potential
-observer.
+The range of values have some overlap with the expectations from a
+galaxy with a total neutral hydrogen mass of $10^8$-$10^9$
+$M_{\odot}$.
 
@@
 \lya line morphology. 
-The basic assumption of the model is that each differential surface
-element on the sphere Doppler shifts the photons that it emits.
+The basic assumption of their analytical model is that each
+differential surface element on the sphere Doppler shifts the photons
+that it emits.
 In this paper we introduce this ansatz by post-processing the results
@@
 where $x'$ is the photon's new adimensional frequency, $x$ is the photon's
-frequency after being processed only by the outflow, \vrot is the 
+frequency after being processed by the outflow, \vrot is the 
 rotational velocity at the point of escape of the photon, \hatk is
@@
 velocity of the sphere.
-We fix \vth throughout the paper. This factor is a multiplying
-constant that could be changed without running new Monte-Carlo
-simulations, as the code works with the adimensional frequency $x$.
 
-This semi-analytic model allows us to produce new \lya spectra from
-the outflow-only results and compare them with the full radiative
-transfer solution including both outflows and rotation.
+This allows us to produce new \lya spectra and compare them with the
+full radiative transfer solution including both outflows and
+rotation.
 
@@
 
-Figure \ref{fig:doppler_shift} summarizes the most important trends
-from rhw RT simulations.
+Figure \ref{fig:doppler_shift} summarizes the most important trends.
 In the left side, the six panels correspond to $\tau=10^6$ and a viewing angle of
@@
 The thick gray line corresponds to the pure outflow solution
-with the Doppler shift added to model rotation's influence.
-At $\tauh=10^6$ the Doppler shift does a good job at capturing the broad
+with the Doppler boost added to model rotation's influence.
+At $\tauh=10^6$ the Doppler boost does a good job at capturing the broad
 morphological features introduced by rotation: the angle dependence,
@@
 In the right side of Figure \ref{fig:doppler_shift} we show the same results as in the left one, but for a viewing angle of $\theta =
-0^{\circ}$, that is parallel to the rotation axis. 
+90^{\circ}$, that is parallel to the rotation axis. 
 In this case we confirm the result presented by \cite{Garavito14},
@@
 
-The quality of the results from the Doppler shift improves for higher
+The quality of the results from the Doppler boost improves for higher
 \tauh values. 
-In the Appendix we show the same plots as Figure
-\ref{fig:doppler_shift}, there it is evident that for $\tauh=10^5$ the
-results are not as good as they are for $\tauh=10^6$, and that for
-$\tauh=10^7$ the Doppler shift provides a remarkable good
-approximation.
+In the Appendix we show the same plots as Figure \ref{fig:doppler_shift}, there it is
+evident that for $\tauh=10^5$ the results are not as good as they are
+for $\tauh=10^6$, and that for $\tauh=10^7$ the Doppler boost
+provides a remarkable good approximation.
 
-\subsection{Quantitative Trends}
+\subsection{Quantitative trends}
 \label{sec:quantitative}
@@
 To do this we summarize the line morphology by four different
-scalars: standard deviation (STD), skewness (SKW), bimodality
+characteristics: standard deviation (STD), skewness (SKW), bimodality
 (BI) and valley/peak ratio.
@@
 
-\subsection{The semi-analytic model as gaussian smoothing}
+\subsection{\lya continuum leakage or rotation?}
 
-The effects of the semi-analytic model on the pure outflowing spectra
-are similar to the expected results from a gaussian smoothing, 
-such smoothing can also be a natural consequence of the intruments
-used to measure the LAE spectra.
-As a consequence, natural experimental artifacts could be mistaken as
-an indication for rotation.
+One application of this new model to the current interpretation of \lya
+spectra, is that the central ($V=0$) emission of the spectra seen in the
+\lya line, is a consequence of the viewing angle of the galaxy and can be
+controlled by it. 
+Several authors have suggested that this central emission is caused by
+radiation that escapes the galaxy without scattering, mostly due to a
+clumpy interstellar medium \citep{Hansen06, 2016ApJ...833L..26G}.
+Rotation is an alternative that solves this issue.
 
-To understand to what extent the effects of rotation can be seen as a
-simple smoothing, we model the effects of gaussian smoothing with a
-similar ansatz as  the semi-analytic model by changing the frequency
-of each photon by a Gaussian random variable centered on zero and
-standard deviation $\sigma_x$: \begin{equation}
-x^{\prime} = x + \mathcal{N}(0,\sigma_x^2).
-\end{equation}
+We note that the relative intensity of the two peaks in the outflowing
+spectra is not modified by rotation, that asymmetry is solely
+controlled by the outflow velocity.
+This means that results that have already derive a typical outflowing
+velocity as to match the observational constraints do not have to be
+completely revised. 
+Only an additional exploration of rotational velocities and viewing
+angles needs to be explored.
 
-We find that the results of the semi-analytic model on the total
-spectra at a given angle $\theta$ can be approximated using
-$\sigma_x = (1/2)\times(\vrot/\vth)\times \sin\theta$. 
-This approximation reproduces within a few porcent all quantities
-measured in Section \ref{sec:quantitative} for the semi-analytic
-model, it only works well for $\tau=10^6$ and $\tau=10^7$.
+A recent observational example is provided by
+\cite{2017A&A...608L...4R}.
+They report optical spectroscopy of a bright lensed galaxy at a
+redshift of $z=2.4$. 
+The object presents a triple peaked \lya profile that can be thought
+as the superposition of two components, a double asymmetric peak and a
+narrower symmetric peak around the line's center.
+One possible option for the narrow component is that it escaped the
+galaxy without substantial interaction with neutral Hydrogen.
+We speculate that the central feature could also be due to a purely 
+rotating component that can produce such a feature as shown by
+\cite{tololo}. The triple peak could be then produced by two main
+decopled components: one rotating the other outflowing.
 
-This result can be used as an order of magnitude estimate of the mimum
-espectral resolution required to observe the effects of a rotational
-velocity of $\vrot$. 
-The Space Telescope Imaging Spectrograph (STIS) on the Hubble Space
-Telescope (HST) can provide resolutions around 20 \kms
-\citep{2015PASA...32...27H}, which could allow for the detection of
-rotation features in nearby galaxies with at least $40$ \kms and a
-rotation axis perpendicular to the line of sight.  
-Other instruments such as the Cosmic Origins Spectrograph (COS) have
-varying resolutions between $20$\kms and $200$\kms according to the
-angular size of the source \citep{2018A&A...616A..60O}.
+Another interesting example are the \lya lines in Green Pea galaxies \citep{2016ApJ...820..130Y}.
+\cite{2017ApJ...844..171Y} showed in a study of 43 Green Pea galaxies
+that 2/3 were strong \lya emitters. 
+There the \lya escape fraction is effectively estimated from the ratio
+of H$\alpha$ to \lya.
+Galaxies with large escape fractions also substantial \lya emission at
+the line's center. 
+We speculate that this feature could help to constrain rotation in
+Green Pea galaxies.
 
 \subsection{\lya Kinematic Maps}
-\label{subsec:kinematic}
 
@@
 around a quasar.
-In this case one could measure the displacement between the
-peaks in the two different spectra. 
-The interest of this test is that the location of the peaks only
-depends on the kinematic properties of the sources and does not
-change by convolution with a gaussian window function due to limited
-instrumental resolution, as discussed in the previous section.
 
@@
 
+Although it is a good approximation to think the rotating spectra by a
+sum of Doppler shifts, the peak of the spectra is also weighted by the
+amount of mass with a given velocity. 
 In this toy model the distance between the peaks of the
-receding/approaching spectra is close to $\sim 25\ \kms$, which is
-a fourth of the naively expected value of $2\vrot=100\ \kms$,
+receding/approaching spectra is close to $\sim 20\ \kms$, which is
+less than half of the naively expected value of $2\vrot=100\ \kms$,
 due to the fact that only a small fraction of the photons are emitted
@@
 $50\ \kms$.
-Although it is a good approximation to think the rotating spectra by a
-sum of Doppler shifts, the peak of the spectra is also weighted by the
-amount of mass with a given line-of-sight velocity.
 
-Spectrographs like the Multi Unit Spectroscopic Explorer (MUSE) could
-obtain kinematic information from large samples of LAEs to build
-velocity maps in Ly$\alpha$.  
+New spectrographs like MUSE should help obtain kinematic information
+from LAEs to build velocity maps in Ly$\alpha$. 
 This could be a natural extension of the work reported by
@@
 Reference Sample) galaxies.  
-The interpretation of such data could take into account the insights
+The interpretation of such data should take into account the insights
 and trends we have presented in this paper.
+
+\subsection{Comparison against the MUSE-Wide Survey}
+
+In order to give the reader a feeling of the ranges of observational
+data that could be explained by the models presented here, 
+we perform a comparison against observational data from the MUSE-Wide
+Survey (MWS) \citep{2017A&A...606A..12H}.
+
+MWS provides public access to 831 spectra out of which 237 contain a
+detected \lya line as dominant. 
+The public dataset reports as the redshift of the galaxy the position
+of the dominant peak.
+We use this redshift to convert the spectra to a rest-frame and compute
+the standard deviation and skewness statistics. 
+Given the uncertainty on the galaxy redshift we do not compute the
+valley/peak ratio. 
+We do not try to perform a full fit to all the spectra
+\citep[e.g.][]{2017A&A...608A.139G} and wish only to
+know whether the region of parameter space explored in our RT
+simulations shows any crossing with the results from the MWS or
+shows any sign of atypical features.
+
+Figure \ref{fig:muse} shows the results of this experiment. 
+The small white circles correspond to the 217 \lya emitters in the MWS
+observations; the large black circles to the results of the
+simulations presented in this paper. 
+This shows that the systems modeled in this paper actually present an
+overlap with statistics derived from observations. 
+The region with standard deviation $>200$\kms could also be explained
+by systems with larger optical depths than we have used here. 
+However, the systems with a positive skewness, that is a line
+unbalanced towards the blue could be the sign of inflowing
+kinematics.
 
@@
 In this paper we explore, for the first time in the literature,
-the results of a model for the emergent \lya line from rotating
-outflows. 
-We use a semi-analytic model, first presented in \citep{Garavito14}, 
-to capture the main effects of rotation and confront it against
-results from Monte-Carlo radiative transfer simulations. 
-The semi-analytic model only takes into account the Doppler shift computed
-as product of quantities at the surface of last scattering, namely
-$\vec{v}_{\rm rot}\cdot \hatk$, where $\vec{v}_{\rm rot}$ is the
-velocity due to rotation and \hatk is the direction of the photon's
-propagation.
+the results of a model for the emergent \lya line from rotating outflows.
+The main results for the model are computed from a Monte-Carlo
+radiative transfer simulation and confronted with a simple
+semi-analytic ansatz that adds the effects of rotation onto results of
+pure outflow kinematics.
 
@@
 \end{itemize}
-%
-All these effects can be quantitatively explained by the proposed
-semi-analytic model.
+
+We find that these effects can be quantitatively explained by a
+Doppler shift computed as product of quantities at the surface of last
+scattering, namely  $\vec{v}_{\rm rot}\cdot \hatk$, where
+$\vec{v}_{\rm rot}$ is the velocity due to rotation and \hatk is
+the direction of the photon's propagation.  
+These conclusions, specially the confirmation of the Doppler Boost as
+a good approximation to model solid body rotation, strengthen the
+semi-analytic model reported by \cite{Garavito14} where they modeled
+the effect of pure rotation on \lya spectra.
 
@@
 extent is it possible to constrain rotational kinematics from the \lya
-emission line?})  we can say that  the most straightforward approach
-takes into account how the \lya spectra from two different sides of a
-galaxy can detect approaching/receding gas motions
-\citep{Prescott14,2018MNRAS.473.3907A}. 
-In that case, we also show that one has to be careful and take into
-account that the  peaks of these two spectra is smaller than a naively
-expected value of $2\times\vrot$ due to the different weights at the
-surface of the emitting regions.   
-Here the semi-analytic model also provides the correct quantitative
-insight.
+emission line?})  we can say that as an application to observational data our results suggest that
+rotation should also be considered as an explanation of the
+non-vanishing \lya intensity at the line's center, which could change
+the interpretation of recent observational results
+\citep[e.g.][]{2017ApJ...844..171Y,2017A&A...608L...4R} that claim
+instead that this flux constrains the escape fraction of \lya
+radiation.
 
-To summarize, our work confirms that the semi-analytic model based on
-the Doppler Shift offers an easy-to-implement approximation to explore
-such influence into already existing radiative transfer simulation
-simulations and provide stronger footing to the interpretation of
-present and future \lya kinematics maps
+A more straightforward approach takes into account how the
+\lya spectra from two different sides of a galaxy can detect
+approaching/receding gas motions.
+In that case, one has to be careful and take into account that the
+peaks of these two spectra is smaller than the expected value of
+$2\times\vrot$ due to the different weights at the surface of the
+emitting regions.
+
+We also use data from MUSE to demonstrate to what extent the
+outputs of the RT results presented here can be expected to match
+hundreds of observed \lya spectra, but a full constraint of rotation
+properties certainly require a full Markov Chain Monte Carlo
+exploration of parameter space.
+
+Our work confirms that rotation produces features in the \lya line
+strong enough that can be expected to be detected in observations and
+that should be taken into account into \lya modeling; the Doppler
+Boost offers an easy-to-implement approximation to explore such
+influence into already existing RT simulations and provide stronger
+footing to the interpretation of \lya kinematics maps
 \citep[e.g][]{2018MNRAS.473.3907A}.
@@
 
-\begin{figure*}
+\begin{figure}
   \begin{center}
     \includegraphics[width=0.49\textwidth]{doppler_shift_logtau5_theta90}
+  \end{center}
+  \caption{\textbf{Qualitative trends of changing outflow and
+      rotational velocity.}
+    Same layout as Figure \ref{fig:doppler_shift},
+    this time  $\tauh=10^5$ and $\theta=90^\circ$.}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
     \includegraphics[width=0.49\textwidth]{doppler_shift_logtau7_theta90}
@@
       rotational velocity.}
-    Same layout as Figure \ref{fig:doppler_shift}. On the left:
-    $\tauh=10^5$ and $\theta=90^\circ$; on the right:     this time  $\tauh=10^7$ and $\theta=90^\circ$.}
-\end{figure*}
+    Same layout as Figure \ref{fig:doppler_shift},
+    this time  $\tauh=10^7$ and $\theta=90^\circ$.}
+\end{figure}
 

### 2018-09-21 paper/paper.tex
fix little typo

@@
 Figure \ref{fig:doppler_shift} summarizes the most important trends
-from rhw RT simulations.
+from the RT simulations.
 In the left side, the six panels correspond to $\tau=10^6$ and a viewing angle of
@@
 is measuring \lya spectra from two different regions of a
-galaxy to detect approaching/receeding gas motions
+galaxy to detect approaching/receding gas motions
 \citep{Prescott14,2018MNRAS.473.3907A}. 

### 2018-11-03 paper/paper.tex
last revision

@@
     CLARA without any rotation and the indicated outflow velocity.
-    The thick black line corresponds to the results including both
+    The thick black line corresponds to CLARA's results including both
     outflows and rotation.
     The thick gray line shows the results of modifying the pure outflow
-    solution by the Doppler shift presented in Equation \ref{eq:shift_x}
-    (in thin line), if there is a radiative transfer of rotation and outflows
-    (thick and clear line), and if there is a radiative transfer of
-    only outflows, but also a Doppler shift from the rotational
-    velocity (thick and dark line).  
+    solution (thin black line) by the Doppler shift presented in Equation \ref{eq:shift_x} using the respective \vrot. 
     \label{fig:doppler_shift}}
@@
 The basic assumption of the model is that each differential surface
-element on the sphere Doppler shifts the photons that it emits.
+element on the sphere Doppler shifts (DS) the photons that it emits.
 In this paper we introduce this ansatz by post-processing the results
@@
 outflows. 
-We the a semi-analytic model first presented by \cite{Garavito14}
+We use a semi-analytic model first presented by \cite{Garavito14}
 to capture the main effects of rotation and confront it against
@@
 easy-to-implement approximation to explore such influence into already
-existing radiative transfer simulation simulations that include
+existing radiative transfer simulations that include
 outflows, providing a versatile tool to interpret current and future \lya
-kinematics maps  \citep[e.g][]{2018MNRAS.473.3907A}.
+kinematics maps \citep[e.g][]{2018MNRAS.473.3907A,Erb18}.
 
